---
title: "ntw 2025 stats analyses"
date: "2025-12-02"
---

# # preamble & loading

stats analyses for ntw 2025

```{r load libraries and utils}
here::i_am("_ntw/scripts/analysis/stats-tents.Rmd")
library(here)

library(lme4) # glm, lm, glmer
#library(lmerTest) # lmer
#library(MuMIn) # dredge (nested model comparisons)
library(pscl) # zeroinfl
#library(MASS) # glm.nb

library(ggfortify) # autoplot (diagnosing)
library(tidyverse)

library(conflicted)
conflicts_prefer( 
  #lmerTest::lmer, # dont rlly need..
  dplyr::filter
)
# lmerTest::lmer overloads lme4::lmerMod summary() and anova() to provide p-vals for the fixed effs


# library(survival) # for fitting 
# library(survminer) # for plotting
# library(coxme) # mixed effects coxphs (but also frailty() might be enough...)

source(here("_ntw/scripts/analysis/filter_utils.R"))
source(here("_ntw/scripts/analysis/stats_utils.R"))
source(here("_ntw/scripts/wrangle-tents.R"))
```

notes: 
- `lmerTest` docs: https://www.jstatsoft.org/article/view/v082i13

```{r load data, message=FALSE}
source(here("_ntw/scripts/R/tidy-tents.R"))

# formatting
# imported_tents <- 
#   dfs_tents$n_pertent %>%
#   mutate(across(starts_with("trt"), as.factor))

#rm(dfs_pertent_ss)

# subset (all lab bugs only)
dfs <- list(
  # incl virgins; can use for laying but not hatching
  all = dfs_tidy$tents %>%
    filter(mate.pop != "field"),
  
  # use for hatching
  mated = dfs_tidy$tents %>% #dfs$tents
    FilterForLabEggs(),
  
  matedexptl = dfs_tidy$tents %>% #dfs$exptal
    FilterForLabEggs() %>%
    filter(mate.trt != 260)
)

# calc per tent SS and add cols/format
dfs <- dfs %>%
  lapply(., \(x){
    x %>%
      CalcTentCounts() %>%
      mutate(trt.minT = case_when(mate.trt == 419 ~ 19,
                                  mate.trt == 433 ~ 33,
                                  TRUE ~ 26),
             mate.ishs = factor(mate.ishs, levels = c("none", "both", "f", "m")),
             across(c("year", starts_with(c("trt", "mate"))), as.factor)
             )
  })

# make easier to interpret...
dfs$matedexptl$mate.trt <- dfs$matedexptl$mate.trt %>%
      recode_factor(`419` = "19", `426` = "26", `433` = "33")

```

# ### exptal only
- eggs/F: if the distribution is skewed right with mean greater than 1-3,  then round values to the nearest integer and use poisson

# # Z: eggs laid

## diagnosing

- tldr
  - offset... sometimes helps
  - need `n.f` as a predictor still.
  - rounding the rate is really nice!
  - so maybe use Z = round(eggs/F) for the models.

```{r warning=FALSE}
# Z: raw n.coll
# AICs: raw + offset = 4090.56; offset2 = 5013.118
mods <- list(
  raw = glm(n.coll ~ mate.trt*mate.ishs + n.f + year,
            data = dfs$matedexptl, family = "poisson"),
  # better QQ
  offset = glm(n.coll ~ mate.trt*mate.ishs + n.f + year, 
                   offset = n.f,
                   data = dfs$matedexptl, family = "poisson"),
  # bad QQ
  offset2 = glm(n.coll ~ mate.trt*mate.ishs + year, 
                    offset = n.f,
                    data = dfs$matedexptl, family = "poisson")
)

# Z: n.coll/f; sqrt(n.coll/f)
# AICs: none cuz poisonning non-integers
mods <- list(
  rawperf = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = dfs$matedexptl, family = "poisson"),
  # QQ looks like above
  offset =  glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
                        offset = n.f,
                        data = dfs$matedexptl, family = "poisson"),
  
  # (try the sqrt)
  # bad QQ
  sqrt_perf = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
             data = dfs$matedexptl, family = "poisson"),
  
  # QQ not that diff from 'offset'
  ## meh. this approach also doesnt feel right (using qb to deal with the non-integers)
  perf_qp = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
                   data = dfs$matedexptl, family = "quasipoisson")
)

# Z: round(n.coll/f) (since dist is right-skewed with mean > 3)
# AICs: roundperf/offset = 1430.344, offset2 = 2531.161
mods <- list(
  roundperf = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year, # use this?
                  data = dfs$matedexptl, family = "poisson"),
  # worse QQ (slightly)
  offset = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year,
                   offset = n.f, 
                   data = dfs$matedexptl, family = "poisson"),
  # uhh worse resids vs fitted? (slopes down)
  offset2 = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + year,
                   offset = n.f,
                   data = dfs$matedexptl, family = "poisson")
)

mods %>% DiagnoseModel()
mods %>% GetModelResults()
mods %>% lapply(., AIC)
```

# ### all bugs (ref only)

- 2026-02-10 after wkly mtg, just for ref only. but TLDR:
  - `n.f` and `year` should be inluded in the eggs laid mods
    - 2026-04-20: also i think `year` was being treated as a number lol (or a char? at least not a factor...)
  - Z = `n.coll.perf` is actually prob fine as the response var? (i.e. dont need to transform)
    - BUT this gets weird w poisson bc of the non-integers (so still need to address this)
  - need to run on a subset of the data without the ctrls (or else just as ctrls vs hot bc 26-26 mating group is confounded w/ `mate.ishs` and `mate.trt`)
  - do the hatching as `c(# hatch, # coll)` (not as success/failures)

# # Z: eggs laid

## diagnosing Z

trying out different response variables for the egg numbers.

- `mods_laid$sqrt` is the better transform (smaller resids, better QQ but still a little off..)
- `mods_laid$log` has a very ugly QQ plot

```{r eval=FALSE, warning=FALSE}
# trying out difference response variables
#(<mod>2 = code the trt as a more straightfwd interaction)
mods <- list(
  raw = glm(n.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
            data = dfs$mated, family = "poisson"),
  raw2 = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year, 
            data = dfs$mated, family = "poisson"),
  sqrt = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
             data = dfs$mated, family = "poisson"),
  sqrt2 = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year, 
             data = dfs$mated, family = "poisson"),
  log = glm(log(n.coll.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
            data = dfs$mated, family = "poisson"),
  log2 = glm(log(n.coll.perf + 1) ~ mate.trt*mate.ishs + n.f + year, 
            data = dfs$mated, family = "poisson"),
  sqrt_q = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
             data = dfs$mated, family = "quasipoisson")
)

mods %>% DiagnoseModel() # use sqrt -- more balanced
```

```{r eval=FALSE, warning=FALSE}
# trying out different groupings
mods <- list(
  raw = glm(n.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson"),
  sqrt = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson"),
  log = glm(log(n.coll.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson")
)

mods %>% DiagnoseModel() # use sqrt -- more balanced
```

## trying pred vars

figuring out which vars to include in the model and what model to use
- Z: eggs laid/collected
  - testing pred vars
    - [x] number f: include
    - [x] year: include (as fixed bc only 2 yrs)
    - [x] cage (random/fixed?): dont include bc breaks
    - [ ] fridge
    - [ ] colony
  - [x] testing interactions:
    - code as `trt.f*trt.m + ishs.f*ishs.m`?
    
### trts
- how to code the trts?

```{r eval=FALSE, warning=FALSE}
# lets give up on tents and look at the interactions instead
mods <- list(
  # which vars are needed?
  add = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m,
            data = dfs$mated, family = "poisson"),
  add_nf = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m + n.f,
            data = dfs$mated, family = "poisson"),
  add_yr = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m + n.f + year,
              data = dfs$mated, family = "poisson"),
  
  # testing ints
  ## incl year and n.f
  int = glm(sqrt.coll.perf ~ trt.f:trt.ishs.f + trt.m:trt.ishs.m + n.f + year,
            data = dfs$mated, family = "poisson"),
  # (this is the "add" model)
  # int2 = glm(sqrt.coll.perf ~ trt.f*trt.ishs.f + trt.m*trt.ishs.m + n.f + year,
  #                data = dfs$mated, family = "poisson"),
  int_is = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m:trt.ishs.f + n.f + year,
            data = dfs$mated, family = "poisson"),
  int_trt = glm(sqrt.coll.perf ~ trt.f:trt.m + trt.ishs.m:trt.ishs.f + n.f + year,
            data = dfs$mated, family = "poisson"),
  # (this is same as int_trt?)
  int_trt2 = glm(sqrt.coll.perf ~ trt.f*trt.m + trt.ishs.m*trt.ishs.f + n.f + year,
                 data = dfs$mated, family = "poisson"),
  int3 = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
             data = dfs$mated, family = "poisson")
)

# testing additional vars
anova(mods$add, mods$add_nf, test = "Chisq") # including n.f is better
anova(mods$add_nf, mods$add_yr, test = "Chisq") # including year is better
anova(mods$add, mods$add_yr, test = "Chisq")
drop1(mods$add_yr)
## so include yr and n.f

# testing add vs int?
anova(mods$add_yr, mods$int, test = "Chisq") # they are the same lol?

# testing coding interactions (trts + hs + sex separated)
anova(mods$int, mods$int_is, test = "Chisq") # int is better
anova(mods$int, mods$int_trt, test = "Chisq") # int_trt is marginally better
anova(mods$int, mods$int_trt2, test = "Chisq") # int_trt2 is marginally better

anova(mods$int_trt, mods$int_trt2, test = "Chisq") # these are the same? (so just use int_trt for the below)
anova(mods$int_is, mods$int_trt, test = "Chisq") # int_trt marginally better
anova(mods$add_yr, mods$int_trt, test = "Chisq") # int_trt marginally better
anova(mods$add_yr, mods$int3, test = "Chisq") # int3 marginally better...

# testing coding interactions (trts + hs separated)
anova(mods$int_trt, mods$int3, test = "Chisq") # these are the same?
AIC(mods$int_trt)
AIC(mods$int3)
drop1(mods$int_trt)
drop1(mods$int3)

# looking at the final model bc i think this is the better one?
mods$int3 %>% GetModelResults()

```


### cage    
- works when fixed (and sig) but not rly what we're trying to do here...

```{r try 2, eval=FALSE, warning=FALSE}
mods <- list(
  nocage = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
               data = dfs$mated, family = "poisson"),
  cage_fixed = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year + cage, # works but not what we want..
                   data = dfs$mated, family = "poisson")
  # breaks: bc of https://forum.posit.co/t/error-in-eval-f-x-downdated-vtv-is-not-positive-definite/73846/2?
  # cage_rand = glmer(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year + (1|cage),
  #                  data = dfs$mated, family = "poisson")
)

mods %>% GetModelResults()

```

- `cage` breaks things. not rly sure why though... (colinearity?)
```{r try 1 cages break, eval=FALSE, warning=FALSE, message=FALSE}
# drop tents with only 1 m/f trt combo (colinear?) for diagnosing
dfs <- list_modify(
  dfs,
  mod = dfs$mated %>%
  group_by(trt.f, trt.m) %>%
  mutate(n = n(),
         cage = case_when(n == 1 ~ NA_character_,
                          TRUE ~ cage))
  )

mods <- list(
  # diagnosing mods
  # full = glm(sqrt.coll.perf ~ mate.trt*trt.ishs.f*trt.ishs.m + n.f + cage + year,
  #            data = dfs$mated, family = "poisson"),
  # cage_fixed = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + cage,
  #              data = dfs$mod, na.action = na.omit,
  #              family = "poisson"),
  # cage_rand = glmer(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + (1|cage),
  #            data = dfs$mod, na.action = na.omit,
  #            family = "poisson"),
  test = glm(sqrt.coll.perf ~ trt.f + trt.m,
             data = dfs$mated, family = "poisson")
  
)
  
mods %>% GetModelResults()
mods %>% DiagnoseModel() # breaks (NA/missing data) 

resid(mods$test) # try -- when `cage` is included, resids are 0

```

### trying stuff w joel
- trying stuff with the mods:
  - use just a subset of the hot bugs only
  - try looking at the trt interaction + dropping different terms
  - trying trt as numeric vs factor
  - try other distributions

```{r trying stuff w joel 260210, eval=FALSE}
# how are the stats when the ctrl group is dropped?
df <- dfs$mated %>% filter(mate.trt != 260)

# if Z = n.coll these are amazing btw
mods <- list(
  # try dropping different terms
  m1 = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "poisson"),
  m2 = glm(n.coll.perf ~ mate.trt + mate.ishs + n.f + year,
            data = df, family = "poisson"),
  m3 = glm(n.coll.perf ~ mate.ishs + n.f + year,
            data = df, family = "poisson"),
  # try trt as numeric?
  m4 = glm(n.coll.perf ~ as.numeric(mate.trt)*mate.ishs + n.f + year,
            data = df, family = "poisson"),
  # fyi: different dist from m1-4 so cant anova it with the rest
  m5 = glm(n.coll.perf ~ mate.ishs + n.f + year,
            data = dfs$mated, family = "poisson")
)

mods %>% GetModelResults()

anova(mods$m1, mods$m2, mods$m3, mods$m4, test = "Chisq")
AIC(mods$m1, mods$m2, mods$m3, mods$m4)

anova(mods$m5, test = "Chisq")
summary(mods$m5)

# trying other transforms/dists
glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "gaussian") %>% DiagnoseModel()
glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "gaussian") %>% DiagnoseModel() # kinda bad
glm(n.coll.perf + 1 ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "Gamma") %>% DiagnoseModel()
```


# # Z: eggs hatched

## diagnosing Z

- dont need `n.f` in the model...? tho maybe need `n.coll.perf`?
- they all REALLY suck... need a different model? drop that one ctrl point?
  - try zero-inflated poissons!
  
### glms (bad: bc dist)
```{r , eval=FALSE, warning=FALSE, message=FALSE}
# i also just think this is bad bc of the preds LOL...
# regular glms:
mods <- list(
   raw = glm(n.hatch.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson"),
  # sqrt = glm(sqrt.hatch.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson"),
  # cbrt = glm((n.hatch.perf)^(1/3) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year,
  #            data = dfs$mated, family = "poisson"),
  log = glm(log(n.hatch.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson")
)

mods %>% DiagnoseModel() # they all really suck (cube rt is sorta okay)
qqnorm(mods$log$residuals); qqline(mods$log$residuals)
qqnorm(resid(mods$log)); qqline(resid(mods$log))
```

```{r eval=FALSE, warning=FALSE}
# more glms with different response var and diff preds
# (try on the whole number instead of hatch per f bc that doesn't rlly make sense... also try w the trt interaction)

# glm breaks vars are when factors
df <- dfs$mated %>%
  mutate(across(c("year", starts_with(c("mate", "trt"))), as.character))

mods <- list(
  raw = glm(n.hatch ~ mate.trt*mate.ishs + year,
            data = df, family = "poisson"),
  log = glm(log(n.hatch + 1) ~ mate.trt*mate.ishs + year, 
            data = df, family = "poisson")
  # raw_nb = MASS::glm.nb(n.hatch ~ mate.trt*mate.ishs + year,
  #              data = dfs$mated) # errors...
)

mods %>% DiagnoseModel()
```

### zips (try 1)
- need to code the trts differently.. otherwise warnings

```{r eval=FALSE}
# zip/zinb
mods <- list(
  raw_p = zeroinfl(n.hatch ~ trt.f|1, data = dfs$mated),
  raw_nb = zeroinfl(n.hatch ~ trt.f|1, data = dfs$mated, dist = "negbin")
  # raw_p2 = zeroinfl(n.hatch ~ mate.trt*mate.ishs + year, data = dfs$mated),
  # raw_nb2 = zeroinfl(n.hatch ~ mate.trt*mate.ishs + year, data = dfs$mated, dist = "negbin")
  
  # break: probably bc of how the trt info is handled
  # raw_p = zeroinfl(n.hatch ~ trt.f + trt.m + as.factor(trt.ishs.f) + as.factor(trt.ishs.m) + n.coll + year, 
  #                data = dfs$mated),
  # raw_nb = zeroinfl(n.hatch ~ trt.f + trt.m + as.factor(trt.ishs.f) + as.factor(trt.ishs.m) + n.coll + year, 
  #                data = dfs$mated, dist = "negbin")

)


mods %>% DiagnoseModel2()
#qqnorm(mods$raw_p$residuals); qqline(mods$raw_p$residuals)
#qqnorm(resid(mods$raw_p)); qqline(resid(mods$raw_p))
```

### zips (try 2)
- zip model wont fit if the trts are too piecemealy, so try just `mate.trt * mate.ishs`
  - (try `year` and `n.f` later)

```{r eval=FALSE}
df <- dfs$mated %>%
  mutate(mate.ishs = case_when(mate.ishs == "none" ~ NA_character_,
                               TRUE ~ as.character(mate.ishs)))

mods <- list(
  raw_p = zeroinfl(n.hatch ~ mate.trt*mate.ishs|1, 
                   data = df, dist = "poisson"),
  raw_nb = zeroinfl(n.hatch ~ mate.trt*mate.ishs|1, 
                    data = df, dist = "negbin")
)

mods %>% DiagnoseModel2() # resids are a little tighter w the negbin. does it need to be transformed tho ??

```

## trying pred vars

- where Z: eggs hatched
  - (glms are sorta wonky bc the distribution is sooo bad to begin w lol... (try ZIP model instead))
  - nts: dont need number of F for this.. just # of eggs coll???

# # Z: binomial (hatch, laid)

- [wip] Z: binomial (ie # hatch, # coll)
  - wonkkkyyy...

## trying pred vars
- seems like `trt` of the m/f is more impt? (everything sig except `trt.ishs`...)
- but go look at data transforming before continuing more lol

```{r, eval=FALSE}
# for simplicity, focus on the trts
dfs <- list_modify(
  dfs,
  mods = dfs$mated %>%
    mutate(n.fail = n.coll - n.hatch)
)

mods <- list(
  # as (# successes, # failures)
  raw_sf = glm(cbind(n.hatch, n.fail) ~
                 #trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + year,
                 #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
                 mate.trt*mate.ishs + n.f + year,
               data = dfs$mods, family = "binomial"),
  
  
  # as a prop
  raw_hc = glm(cbind(n.hatch, n.coll) ~
                 #trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + year,
                 #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
               mate.trt*mate.ishs + n.f + year,
               data = dfs$mods, family = "binomial"),
  
  raw_hprop = glm(n.hatch/n.coll ~ mate.trt*mate.ishs + n.f + year,
                  #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
               data = dfs$mods, family = "quasibinomial") # use qb for non-int binomials
)

#mods %>% DiagnoseModel() # breaks bc of cbind
mods %>% GetModelResults()
#anova(mods$raw_sf, mods$raw_hc, test = "Chisq") # breaks bc of cbind?
```
