---
title: "ntw 2025 stats analyses"
date: "2025-12-02"
---

# # preamble & loading

stats analyses for ntw 2025

```{r set pathfinding}
here::i_am("_ntw/scripts/analysis/stats-dev.Rmd")
```

```{r load libraries}
library(here)
library(lme4) # glm, lm, glmer
#library(lmerTest) # lmer
library(MuMIn) # dredge (nested model comparisons)
library(tidyverse)

library(conflicted)
conflicts_prefer( 
  #lmerTest::lmer, # dont rlly need..
  dplyr::filter
)
# lmerTest::lmer overloads lme4::lmerMod summary() and anova() to provide p-vals for the fixed effs

# library(survival) # for fitting 
# library(survminer) # for plotting
# library(coxme) # mixed effects coxphs (but also frailty() might be enough...)

source(here("_ntw/scripts/analysis/filter_utils.R"))
source(here("_ntw/scripts/analysis/stats_utils.R"))
```
notes

- `lmerTest` docs: https://www.jstatsoft.org/article/view/v082i13

## TODO
- keep year as a fixed effect! (i.e. no mixed effs for now LOL)

```{r load data, warning=FALSE}
source(here("_ntw/scripts/R/tidy-dev.R"))
```

# # pupal stats

## data prep

```{r dev data prep}
# prep exptal only + all bugs subsets. 
# further subset/mutate as needed closer to the codeblock (ie mutates, more filtering, etc etc)

dfs_imported <- list(
  all = dfs_tidy$wide,
  expt = dfs_tidy$wide %>% filter(trt != 260)
) %>%
  lapply(., \(x) {
    x %>%
      FilterOutLabTB() %>%
      FilterForNTWTrts() %>%
      filter(pop != "col",
             is.pup < 2, # drop culled bugs here
             ) %>%
      mutate(across(c("year", "trt.minT", "trt.type"), as.factor))
  })
```

## survival to pup

- 1 and 2-way ints are signif

```{r pup surv model generation}
mods_pupsurv <- list(
  expt_fullint = glm(is.pup ~ year*pop*trt.minT, 
                family = "binomial", data = dfs_imported$expt, 
                na.action = na.fail
                )
  )
```

```{r pup surv model results}
mods_pupsurv %>%
  lapply(., GetModelResults)

## dredging: print out is long asf so dont run this
# mods_pupsurv %>%
#   lapply(., \(x){
#     x %>% dredge() %>% subset(., weight > 0.05, recalc.weights = FALSE)
#   })
```


## pupal mass

### df prep
```{r}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% filter(!is.na(mass.pupa),
                 !is.na(sex))
  })
```

- 1-ways, `year:minT`, and `minT:pop` are sig

```{r pup mass model generation}
mods_pupmass <- list(
  expt_add = lm(mass.pupa ~ year + trt.minT + pop + sex,
           data = dfs$expt, na.action = na.fail),
  expt_fullint = lm(mass.pupa ~ year*trt.minT*pop*sex,
               data = dfs$expt, na.action = na.fail)
  
  # subsets of the full model are in a later chunk otherwise results r crazy
)
```

```{r pup mass model main results}
mods_pupmass %>%
  lapply(., GetModelResults)

## dredging. output is long so dont run this
# mods_pupmass %>%
#   lapply(., \(x){
#     x %>% dredge()
#   })
```
basically, per anovas, `lm(mass.pupa ~ year + minT + pop + sex + year:minT)` seems good...

other results: 

- bc the dredge of the full interactive model (`year*pop*minT*sex`) is crazy, how do the subsets compare?
  - doing `anova(simpler, complex)` per https://www.geeksforgeeks.org/machine-learning/comparing-two-linear-models-with-anova-in-r/ 


```{r pup mass model subsets, eval=FALSE}
# keep separate otherwise console blows up

mods_pupmass <- list_modify(
  mods_pupmass,
  # subsets of the full model
  expt_2way = lm(mass.pupa ~ (year + trt.minT + pop + sex)^2,
              data = dfs$expt, na.action = na.fail),
  expt_3way = lm(mass.pupa ~ year*trt.minT*pop + sex,
              data = dfs$expt, na.action = na.fail)
)

# testing anova(simplermod, complexmod)
anova(mods_pupmass$expt_add, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_2way, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_3way, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_2way, mods_pupmass$expt_3way, test = "Chisq")
```


## time to pup

- 1 ways and `year:minT`, `year:pop` are sig

notes
- checked that `sex` PROBABLY doesnt matter: bc sex is only determined after pupation anyway... (so omit from mods)

```{r pup time model generation}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      filter(!is.na(jdate.pupa)) %>%
      mutate(tt.pupa = jdate.pupa - jdate.hatch,
             devrate.pupa = 1/tt.pupa)
  })

mods_puptime <- list(
  expt_fullint = lm(devrate.pupa ~ year*trt.minT*pop, data = dfs$expt)
  )
```

```{r pup time model main results}
mods_puptime %>%
  lapply(., GetModelResults)
```
# # misc

## age vs instar

looking at dev trajectories: is `age ~ instar` linear?

```{r}
# df prep
dfs <- dfs_tidy$long %>%
  FilterOutLabTB() %>% 
  FilterForNTWTrts() %>%
  # pick out from 3rd-wander bc all have same trts
  filter(instar %in% c("3rd", "4th", "5th", 
                       "6th", "7th", "8th", 
                       "wander")) %>%
  mutate(across(c("year", "sup", starts_with(c("trt", "is."))),
                as.factor))
```

```{r}
# non-sups only
list(
  nosups = lm(tt ~ instar * trt, data = filter(dfs, is.sup == 0)),
  nosups2 = lm(tt ~ instar * trt * pop, data = filter(dfs, is.sup == 0))
  ) %>%
  GetModelResults()







```


## yr as random eff on pup dev

per 251217 wkly mtg, keep year as a fixed effect in the models (but here is some EDA for reference just in case.)

each chunk is self-contained (model fitting + tests) and overwrites the `mods_pup_randyr` object. just run `GetModelResults` on them to be reminded of what happened)

```{r pup surv rand yr, eval=FALSE}
### model generation (just focus on expt bugs)
dfs <-  dfs_imported$expt %>%
  mutate(year_n = as.numeric(year))

mods_pup_randyr <- list(
  surv = glmer(is.pup ~ pop*trt.minT + (1|year),
                     family = "binomial", data = dfs),
  
  surv_ctsyr = glmer(is.pup ~ pop*trt.minT + (1|year_n),
                     family = "binomial", data = dfs),
    # compare to `surv` to examine effs of yr as cts or cat
  
  surv_noyr = glm(is.pup ~ pop*trt.minT,
                     family = "binomial", data = dfs)
    # compare to `surv` to examine effs of year
)


### tests
# effect of cts vs cat yr
anova(mods_pup_randyr$surv, mods_pup_randyr$surv_ctsyr, test = "Chisq") 

# to get approx p-value for (1|year)?:
anova(mods_pup_randyr$surv, mods_pup_randyr$surv_noyr, test = "Chisq") 
  # i forget what this means...
```

notes about `(1|year)` on pup surv:

- `glmer(is.pup ~ pop*minT + (1|year))` results are same if `year` is numeric or factor
- can approximate the p-value of the random effect by using `anova` to compare models with and without the effect of interest (bc they're nested) -- see more https://stats.stackexchange.com/questions/381208/r-how-to-get-estimates-and-p-values-for-random-effects-in-glmer

```{r pup dev time rand yr, eval=FALSE}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      filter(!is.na(jdate.pupa)) %>%
      mutate(tt.pupa = jdate.pupa - jdate.hatch,
             devrate.pupa = 1/tt.pupa)
  })

# mostly reminding myself of how lmerTest works
mods_pup_randyr <- list(
  time_lmertest = lmerTest::lmer(tt.pupa ~ trt.minT*pop + (1|year), data = dfs$expt),
  time_lme4 = lme4::lmer(tt.pupa ~ trt.minT*pop + (1|year), data = dfs$expt)
)

mods_pup_randyr %>% lapply(., summary)
```

```{r pup mass rand yr, eval=FALSE}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      filter(!is.na(jdate.pupa),
             !is.na(sex),
             !is.na(mass.pupa)
      )
    })

mods_pup_randyr <- list(
  mass = lmer(mass.pupa ~ trt.minT*pop*sex + (1|year),
              data = dfs$expt)
)
```

