---
title: "ntw 2025 data viz"
date: "2025-12-01"
---

# # preamble & loading

data viz for ntw 2025

```{r set pathfinding}
here::i_am("_ntw/scripts/analysis/viz-dev.Rmd")
```

```{r load libraries and utils, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(patchwork) # for plot_layout
# library(survival) # for fitting surv
# library(survminer) # for plotting surv

# utils
source(here("_ntw/scripts/wrangle-dev.R"))
source(here("_ntw/scripts/analysis/filter_utils.R"))
source(here("_ntw/scripts/analysis/viz_utils.R"))
```

```{r load data, message=FALSE, warning=FALSE}
source(here("_ntw/scripts/R/tidy-dev.R"))

# add pre-filtered ntw dfs
dfs_tidy <- list_modify(
  dfs_tidy,
  widentw = dfs_tidy$wide %>% 
    FilterForNTWBugs() %>% filter(pop != "col"),
  longntw = dfs_tidy$long %>% 
    FilterForNTWBugs() %>% filter(pop != "col")
) %>%
  
  # make some factors for graphing ease
  lapply(., \(df) {
    df %>%
      mutate(trt.facet = case_when(trt == 260 ~ "ctrl",
                                   pop == "lab" ~ "lab",
                                   pop == "field" ~ "field")) %>%
      mutate(across(c(starts_with("trt"), "year"), as.factor))
  })
```

# ### larval dev


# # survival to pup
```{r}
.plotfn <- function(x){
  x %>%
  ggplot(aes(y = prop.pup,
             x = trt,
             col = year, #lty = year,
             group = year
             )) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = prop.pup + se.pup,
                    ymin = prop.pup - se.pup),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  # geom_text(aes(label = n),
  #           hjust = -0.2, vjust = -.2,
  #           show.legend = FALSE) +
  facet_wrap(~pop) +
  labs(y = "% survived to pupation",
       x = "treatment",
       col = "year", #lty = "year"
       ) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
}

# all trts
dfs_tidy$widentw %>%
  CalcSurvProps_def() %>%
  filter(!is.na(diet)) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

# expt trts only by minT
dfs_tidy$widentw %>%
  CalcSurvProps_def() %>%
  filter(trt != 260) %>% 
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_minT) +
    labs(x = "minimum T (°C)")

# ctrls vs expt
  # ctrls are too stacked bc of the x axis
dfs_tidy$widentw %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  facet_wrap(~trt.facet, scales = "free_x") +
  scale_x_discrete(labels = p_scales$labs_trt) +
  aes(shape = pop)
```

```{r}
# years sep (patchwork)
.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "% survived to pupation",
         lty = "population", shape = "population"
    ) +
    scale_color_brewer(palette = "Dark2") +
      scale_shape_manual(values = p_scales$shp_pop) +
  scale_linetype_manual(values = p_scales$lty_pop) +
    theme_bw()
}

p1 <- dfs_tidy$widentw %>%
  filter(trt == 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = year, 
      group = pop)

p2 <- dfs_tidy$widentw %>%
  filter(trt != 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = trt,
      group = year,
      col = year) +
  labs(x = "minimum temp (°C)") +
  scale_x_discrete(labels = p_scales$labs_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r}
# yrs combined patchwork
.prepfn <- function(x){
  x %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps()
}

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop,
               x = trt)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.type, labeller = as_labeller(p_scales$facs_trttype)) +
    labs(y = "% survived to pupation",
         lty = "population", shape = "population",
         x = "minimum temp (°C)"
    ) +
    scale_color_brewer(palette = "Dark2") +
      scale_shape_manual(values = p_scales$shp_pop) +
  scale_linetype_manual(values = p_scales$lty_pop) +
  scale_x_discrete(labels = p_scales$labs_minT) +
    theme_bw()
}

## TODO maybe add alpha'd out points in the back?
p1 <- dfs_tidy$widentw %>%
  filter(trt == 260) %>%
  .prepfn() %>%
  .plotfn()

p2 <- dfs_tidy$widentw %>%
  filter(trt != 260) %>%
  .prepfn() %>%
  .plotfn() +
  aes(group = pop)


p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r}
# yrs combined with points for individual yrs patchwork
dfs <- list(
  allyrs = dfs_tidy$widentw %>%
    CalcSurvProps_def() %>%
    MakeFactors(),
  
  noyr = dfs_tidy$widentw %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>%
    MakeFactors()
)


.plotfn <- function(x){
  x %>%
    ggplot(aes(y = prop.pup, lty = pop,
               x = trt, shape = pop,
               group = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.type, labeller = as_labeller(p_scales$facs_trttype)) +
    labs(y = "% survived to pupation",
         lty = "population", 
         shape = "population",
         #alpha = "population",
         x = "minimum temp (°C)",
    ) +
    scale_color_brewer(palette = "Dark2") +
    scale_shape_manual(values = #p_scales$shp_pop
                         c(`field` = 19, `lab` = 17)) +
    scale_linetype_manual(values = p_scales$lty_pop) +
    scale_x_discrete(labels = p_scales$labs_minT) +
    # scale_alpha_identity(values = pop, breaks = c(0.5, 1), labels = c("field", "lab")) +
    #scale_alpha_manual(values = c(`field` = 0.7, `lab` = 1)) +
    theme_bw() +
    scale_alpha_manual(guide = "none")
}

# trying alpha...
# dfs$noyr %>% 
#   filter(trt == 260) %>%
#   .plotfn() +
#   geom_point(aes(shape = pop), size = 2) +
#   geom_line() +
#   geom_point(data = dfs$allyrs %>% filter(trt == 260), size = 2,
#              aes(alpha = pop, col = year))


p1 <- dfs$noyr %>%
  filter(trt == 260) %>%
  .plotfn() +
  geom_point(filter(dfs$allyrs, trt == 260), 
             mapping = aes(col = year), size = 2, alpha = 0.5)

p2 <- dfs$noyr %>%
  filter(trt != 260) %>%
  .plotfn() +
  geom_point(filter(dfs$allyrs, trt != 260), 
             mapping = aes(col = year), size = 2, alpha = 0.5,
             position = position_dodge(-0.5))
  
p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1)) 
```


# # pup mass

### sexes combined
```{r}
.prepfn <- function(x){
  x %>%
    CalcDevSS_ad() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
  ggplot(aes(y = avg.mass, x = trt,
             #lty = year, 
             col = year,
             group = year,
             ))+
  geom_point() + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, 
                    ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") +
  # geom_text(aes(label = n),
  #           vjust = -0.2, hjust = -0.3,
  #           show.legend = FALSE) +
  facet_grid(~pop) +
  labs(y = "pupal mass (g)",
       caption = "mean±se; combined sexes") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
}
```

```{r}
dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
    scale_x_discrete(labels = p_scales$labs_trt)
```

```{r}
# patchwork
.prepfn <- function(x){
  x %>%
    CalcDevSS_ad() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.mass,
               shape = pop, lty = pop)) +
    geom_point(size = 2) + geom_line() +
    geom_errorbar(aes(ymax = avg.mass + se.mass, 
                      ymin = avg.mass - se.mass),
                  width = 0.25, lty = "solid") +
    # geom_text(aes(label = n),
    #           vjust = -0.2, hjust = -0.3,
    #           show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "pupal mass (g)",
         shape = "population", lty = "population") +
    scale_color_brewer(palette = "Dark2") +
    scale_linetype_manual(values = p_scales$lty_pop) +
    scale_shape_manual(values = p_scales$shp_pop) +
    theme_bw()
}

p1 <- dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, col = year,
      group = interaction(pop, year)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = "minimum temperature (°C)")

p1 + p2 +
  plot_layout(widths = c(1, 2),
              axes = "collect", guides = "collect") &
  ylim(c(1, 6.7)
```


### sexes parsed
```{r}
.prepfn <- function(x){
  x %>%
    filter(!is.na(sex)) %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}
```

```{r}
# 2x2 pop + sex (better)
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt,
               col = year,
               group = interaction(year)
               )) +
    geom_point() + geom_line() +
    # geom_text(aes(label = n),
    #           vjust = -0.5, hjust = -0.5, size = 3.5,
    #           position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) + # too busy
    geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                  width = 0.3, lty = "solid") + # remove if including ctrl
    labs(y = "pupal mass (g)", 
         caption = "mean±se") +
    facet_grid(sex~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260,
         #sex == "m"
         ) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_minT) +
    labs(x = "minimum T (°C)")

```

```{r}
# by sex (meh)
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt,
             col = year,
             group = interaction(year, pop), 
             shape = pop, lty = pop
             )) +
  geom_point(size = 2) + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") + # remove if including ctrl
  # geom_text(aes(label = n),
  #           vjust = -0.5, hjust = -0.5, size = 3.5,
  #           position = position_jitter(width = 0.2, height = 0.3, seed = 1),
  #           show.legend = FALSE) + # too busy
  labs(y = "pupal mass (g)", 
       caption = "mean±se") +
  facet_grid(~sex) +
    scale_color_brewer(palette = "Dark2") +
    scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  theme_bw()
}

dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

```{r}
# by pop (no)
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt,
             col = year,
             group = interaction(year, sex), 
             shape = sex, lty = sex
             )) +
  geom_point(size = 2) + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") +
  labs(y = "pupal mass (g)", 
       caption = "mean±se") +
  facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  theme_bw()
}

dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

# # pup time

### days to pup
```{r days to pup}
# sexes combined
.prepfn <- function(x){
  x %>%
    CalcDevSS_ad() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt, x = trt,
               col = year,
               group = year)) +
    geom_point(size = 2) + geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25) +
    # geom_text(aes(label = n),
    #           vjust = -0.75, hjust = -0.5, size = 3.5,
    #           #position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(y = "mean±se time to pupation (days)",
         x = "temperature trt",
         title = "time to pupation")
}

dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
# LOL the 260 lab every yr ...  i think its bc they take so long from 5 to wander

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

```

### pup rate
```{r pup rate}
# sexes combined
.prepfn <- function(x){
  x %>%
    CalcDevSS_ad() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.devrate, x = trt,
               col = year,
               group = year)) +
    geom_point(size = 2) + geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25) +
    # geom_text(aes(label = n),
    #           vjust = -0.75, hjust = -0.5, size = 3.5,
    #           #position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(y = "development rate (1/days)",
         x = "temperature trt",
         title = "time to pupation")
}

dfs_tidy$longntw %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$longntw %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

```{r}
# patchwork
.prepfn <- function(x){
  x %>%
    CalcDevSS_ad() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.devrate,
               shape = pop, lty = pop
               )) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "development rate (1/days)") +
    scale_color_brewer(palette = "Dark2") +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trttype)) +
    theme_bw() +
    labs(shape = "population", lty = "population") +
    scale_shape_manual(values = p_scales$shp_pop) +
    scale_linetype_manual(values = p_scales$lty_pop)
}


p1 <- dfs_tidy$longntw %>%
  .prepfn() %>% 
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- dfs_tidy$longntw %>%
  .prepfn() %>% 
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, group = interaction(pop, year), col = year) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = "minimum temperature (°C)")

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0.029, 0.054))
  
```



# # dev trajectories 

## df prep
```{r}
# stealing tidying code to include a tt.hatch date for later things
df <- dfs_tidy$wide %>% 
  # for dying bug attempts
  mutate(pmd.as = case_when(fate.dev == "pup" ~ "pupa",
                            fate.dev == "pmd" & !is.na(jdate.wander) & 
                              is.na(jdate.pupa) ~ "wander",
                            fate.dev == "pmd" & !is.na(jdate.6th) & is.na(jdate.pupa) ~ "sup",
                            fate.dev == "pmd" & !is.na(jdate.5th) ~ "5th",
                            fate.dev == "pmd" & !is.na(jdate.4th) ~ "4th",
                            fate.dev == "pmd" & !is.na(jdate.3rd) ~ "3rd",
                            fate.dev == "pmd" ~ "younger"),
         pmd.as = factor(pmd.as, levels = c("hatch", "pmd", "younger", 
                                            "3rd", "4th", "5th", "sup", "wander", "pupa"))
         ) %>%
  mutate(tt.hatch = 0,
         tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch,
         tt.8th = jdate.8th - jdate.hatch,
         tt.wander = jdate.wander - jdate.hatch,
         tt.pupa = jdate.pupa - jdate.hatch,
         tt.eclose = jdate.eclose - jdate.pupa,
         tt.long = jdate.long - jdate.eclose,
         #tt.exit = jdate.exit - jdate.hatch # breaks groups.. too many individ exits LOL
         tt.pmd = jdate.pmd - jdate.hatch
          ## hopefully this is better lol bc it should be only things that died have pmd..
  ) %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
  mutate(tt = case_when(is.na(tt) & !is.na(mass) ~ 999,
                        TRUE ~ as.numeric(tt))) %>%
  drop_na(tt) %>%
  mutate(tt = na_if(tt, 999)) %>%
  filter(!(instar == "long" & tt == 0)) %>%
  mutate(instar = factor(instar, 
                         levels = c("hatch", "pmd", "younger", "3rd", "4th", "5th", 
                                    "6th", "7th", "8th",
                                    "wander", "pupa", "eclose", "long"))) %>%
FilterForNTWBugs()
```

## y =  mass

```{r}
# make some variations of the df based off how the pupae are handled
  ## how does inclusion of culled bugs affect the trajectories?

.calcfn <- function(x){
  x %>%
    CalcDevSS() %>%
    filter(instar %in% c("3rd", "4th", "5th", "wander"))
}

dfs <- list(
  # default grouping
  all = df %>%
    PrepLarvalSS() %>%
    .calcfn(),
  
  # default grp with culled removed first
  all_wocull = df %>%
    filter(is.pup != 2) %>%
    PrepLarvalSS() %>%
    .calcfn(),
  
  # by pup, with survivors + culled combined together
  pups_wcull = df %>%
    mutate(is.pup = case_when(is.pup == 2 ~ 1,
                              TRUE ~ as.numeric(is.pup))) %>%
    PrepLarvalSS() %>% 
    group_by(is.pup, .add = TRUE) %>%
    .calcfn(),
  
  # by pup, with survivors + culled kept separate
  pups_sepcull = df %>%
    PrepLarvalSS() %>%
    group_by(is.pup, .add = TRUE) %>%
    .calcfn()
)

```

```{r}
# dev by year
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = year)) +
    geom_point() +
    geom_line() +
    # # hard to see with these lol bc theyre pretty close
    # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
    #                   ymin = avg.logmass - se.logmass,
    #                   width = CalcErrWd(avg.tt))) +
    # geom_errorbarh(aes(xmax = avg.tt + se.tt,
    #                    xmin = avg.tt - se.tt,
    #                    height = CalcErrHt(avg.logmass))) +
    facet_grid(pop ~ trt) +
    #scale_color_brewer(palette = "Greens") +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "year")
}

dfs$all %>%
  .plotfn() +
  labs(title = "development trajectories (incl pmd)")

dfs$all_wocull %>%
  .plotfn() +
  labs(title = "development trajectories (incl pmd)",
       caption = "culled omited")

dfs$pups_wcull %>%
  filter(is.pup != 0) %>%
  .plotfn() +
  labs(title = "development trajectories (pup'd only)")

dfs$pups_sepcull %>%
  filter(is.pup != 0) %>%
  .plotfn() +
  labs(title = "development trajectories (pup'd only)")
  # wonky lines prob bc of the unaddressed supernumeraries in the df prep rip

```

```{r}
# dev by year; parse out pmds more
.plotfn <- function(x){
  x %>%
    mutate(is.pup = as.factor(is.pup)) %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = year, lty = is.pup, 
               group = interaction(year, is.pup, trt, pop))) +
    geom_point() +
    geom_line() +
    facet_grid(pop ~ trt) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(title = "development trajectories",
         caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "year", lty = "fate"
    ) +
    scale_linetype_manual(values = c(`1` = "solid", `0` = "dashed"),
                          labels = c(`1` = "pupated", `0` = "died"))
    #scale_linetype_discrete(labels = c("pupated", "died"))
}

# surv, combined w culled
dfs$pups_wcull %>%
  .plotfn()

# surv, culled omitted
dfs$pups_sepcull %>%
  filter(is.pup != 2) %>%
  .plotfn()

```

```{r}
# dev by trt
.plotfn <- function(x){
  x %>%
  ggplot(aes(y = avg.logmass,
             x = avg.tt,
             col = trt)) +
  geom_point() +
  geom_line() +
  # hard to see with these lol bc theyre pretty close
  # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
  #                   ymin = avg.logmass - se.logmass,
  #                   width = CalcErrWd(avg.tt))) +
  # geom_errorbarh(aes(xmax = avg.tt + se.tt,
  #                    xmin = avg.tt - se.tt,
  #                    height = CalcErrHt(avg.logmass))) +
  facet_grid(pop ~ year) +
  #scale_color_brewer(palette = "Greens") +
  #scale_color_manual(values = hcl.colors(6, "Viridis", rev = TRUE)) +
    scale_color_manual(values = p_scales$cols_trt,
                       labels = p_scales$labs_trt) +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "treatment") +
    theme_bw()
}

dfs$pups_wcull %>%
  filter(is.pup != 0) %>%
  .plotfn() +
  labs(title = "development trajectories (pups only)")

dfs$all %>%
  .plotfn() +
  labs(title = "development trajectories (all)")

```

```{r}
# dev traj by trt + surv to pupation
.plotfn <- function(x){
  x %>%
    mutate(is.pup = case_when(is.pup == 0 ~ "died",
                              is.pup == 1 ~ "pupated")) %>%
    mutate(is.pup = factor(is.pup, levels = c("pupated", "died"))) %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = trt,
               group = interaction(year, pop, is.pup, trt))) +
    geom_point() +
    geom_line() +
    # hard to see with these lol bc theyre pretty close
    # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
    #                   ymin = avg.logmass - se.logmass,
    #                   width = CalcErrWd(avg.tt))) +
    # geom_errorbarh(aes(xmax = avg.tt + se.tt,
    #                    xmin = avg.tt - se.tt,
    #                    height = CalcErrHt(avg.logmass))) +
    facet_grid(is.pup ~ year) +
    #scale_color_brewer(palette = "Greens") +
    #scale_color_manual(values = hcl.colors(6, "Viridis", rev = TRUE)) +
    scale_color_manual(values = p_scales$cols_trt,
                       labels = p_scales$labs_trt) +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "treatment") +
    theme_bw()
}


dfs$pups_sepcull %>%
  filter(is.pup != 2,
         pop == "field") %>%
  .plotfn() +
  labs(title = "development trajectories (field)")
  
dfs$pups_sepcull %>%
  filter(is.pup != 2,
         pop == "lab") %>%
  .plotfn() +
  labs(title = "development trajectories (lab)")
```


```{r}
# wtf goin on w the lab controls
dfs$all %>%
  filter(trt == 260) %>%
  ggplot(aes(y = avg.logmass,
             x = avg.tt,
             col = year, lty = pop, shape = pop,
             group = interaction(year, pop))) +
  geom_errorbarh(aes(xmax = avg.tt + se.tt,
                     xmin = avg.tt - se.tt,
                     height = 0.3), lty = "solid") +
  geom_point() +
  geom_line() +
  facet_grid(#pop 
             ~ trt) +
  #scale_color_brewer(palette = "Greens") +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c(16, 1)) +
  labs(caption = "what's up w the controls") +
  theme_bw()
```

## y = instar

```{r}
# culled bugs ignored here lol. this is basically dfs$all i think
.prepfn <- function(x){
  x %>%
    filter(instar != "pmd") %>%
    PrepLarvalSS() %>%
    group_by(across(c("sup", "is.sup")), .add = TRUE) %>%
    CalcDevSS() %>%
    mutate(across(ends_with("sup"), replace_na, 0),
           across(ends_with("sup"), as.factor))
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = instar, x = avg.tt, 
               lty = sup,
               group = interaction(pop, trt, sup, year))) +
    geom_point() +
    geom_line() +
    # kinda messy, so omit..
    geom_errorbarh(aes(xmax = avg.tt + se.tt,
                       xmin = avg.tt - se.tt,
                       height = 0.25)) +
    labs(x = "days to instar",
         lty = "supernumerary?") +
    theme_bw()
}

# by year
df %>%
  .prepfn() %>% 
  filter(is.sup == 0) %>% # for non-sups
  #filter(is.sup != 0, sup != 8) %>% # for sups
  .plotfn() +
  aes(col = trt) +
  facet_grid(pop~year) +
  scale_color_manual(values = p_scales$cols_trt,
                     labels = p_scales$labs_trt)+
  labs(col = "treatment")

# by trt
df %>%
  .prepfn() %>%
  filter(is.sup == 0) %>% # for non-sups
  #filter(trt == 433) %>%
  #filter(is.sup != 0, sup != 8) %>% # for sups
  .plotfn() +
  aes(col = year) +
  facet_grid(pop~trt, labeller = labeller(trt = p_scales$labs_trt)) +
  scale_color_brewer(palette = "Dark2")

```

```{r attempts at pup version, eval=FALSE}
# this is too messy lol, at least for when looking at the non-sups
.prepfn <- function(x){
  x %>%
    filter(instar != "pmd") %>%
    mutate(is.pup = case_when(is.pup == 2 ~ 1, 
                              TRUE ~ is.pup)) %>% # need to do before grouping..
    PrepLarvalSS() %>%
    group_by(across(c("sup", "is.sup", "is.pup")), .add = TRUE) %>%
    CalcDevSS() %>%
    # (dont mutate is.pup here)
    mutate(across(ends_with("sup"), replace_na, 0),
           across(c(starts_with("is"), ends_with("sup")), as.factor))
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = instar, x = avg.tt, 
               lty = sup,
               shape = is.pup,
               group = interaction(pop, trt, sup, is.pup, year))) +
    geom_point() +
    geom_line() +
    scale_shape_manual(values = c(`1` = 19, `0` = 1),
                       labels = c(`1` = "survived", `0` = "died")) +
    labs(x = "days to instar",
         lty = "supernumerary?",
         shape = "survived to\npupation?") +
    theme_bw()
}

# by year
df %>%
  .prepfn() %>% 
  filter(is.sup == 0) %>% # for non-sups
  #filter(is.sup != 0, sup != 8) %>% # for sups
  .plotfn() +
  aes(col = trt) +
  facet_grid(pop~year) +
  scale_color_manual(values = p_scales$cols_trt,
                     labels = p_scales$labs_trt)+
  labs(col = "treatment")

# by trt
df %>%
  .prepfn() %>% 
  filter(is.sup == 0) %>% # for non-sups
  #filter(trt == 433) %>%
  #filter(is.sup != 0, sup != 8) %>% # for sups
  .plotfn() +
  aes(col = year) +
  facet_grid(pop~trt, labeller = labeller(trt = p_scales$labs_trt)) +
  scale_color_brewer(palette = "Dark2")
```

## dying bugs?

- when do bugs die? (esp those that make it to 5th/later instars but die before wandering at hot trts?)
  - (keep things simple for now and just focus on the non-sups)

- note: wrangling things that dont work lol
  - cant group by "exit" as an instar (too many unique exits)
  - no exit mass, so cant plot by avg mass

```{r try 3 df prep}
# pre:
  # want to get an avg line for the dev rate of survivors (+ non survivors?),
  # and plot death info separately

dfs <- list(
  ifpup = df %>%
    filter(is.sup == 0, is.pup != 2,
           #!(instar %in% c("hatch", "pupa")),
           instar != "pupa", # x-axis too long if include died as pups
           instar != "pmd",
           fate.dev != "pup") %>%
    PrepLarvalSS() %>%
    group_by(is.pup, .add = TRUE) %>%
    CalcDevSS() %>%
    mutate(is.pup = factor(is.pup)),
  
  # NOTE dfs are hardcoded into .plotfns below
  # lump < 3rds into "younger"
  pmdas = df %>%
    filter(is.sup == 0,  is.pup != 2,
           #!(instar %in% c("hatch", "pupa")),
           instar == "pmd",
           !is.na(pmd.as)) %>%
    mutate(pmd.as = case_when(pmd.as == "3rd" ~ "younger",
                              TRUE ~ as.character(pmd.as))) %>%
    PrepLarvalSS() %>%
    group_by(pmd.as, .add = TRUE) %>%
    CalcDevSS() %>%
    mutate(is.pup = 0) %>%
    mutate(is.pup = factor(is.pup)) %>%
    filter(instar == "pmd"),
  
  # leave < 3rds alone
  pmdas2 = df %>%
    filter(is.sup == 0, is.pup != 2,
           #!(instar %in% c("hatch", "pupa")),
           instar == "pmd",
           !is.na(pmd.as)) %>%
    PrepLarvalSS() %>%
    group_by(pmd.as, .add = TRUE) %>%
    CalcDevSS() %>%
    mutate(is.pup = 0) %>%
    mutate(is.pup = factor(is.pup))
  )
```

```{r try 3 timelines}
.plotfn <- function(x){
  x %>%
    #filter(trt == 433) %>%
    ggplot(aes(x = avg.tt,
               #y = year
               )) +
    
    # larval dev points only
    geom_point(data = . %>% filter(instar != "hatch"),
               aes(shape = instar, fill = instar),
               size = 2.5, stroke = NA) + 
    geom_line(aes(lty = is.pup),
              lwd = 0.25) +
    
    # add pmd points (hardcoded df)
    geom_point(data = dfs$pmdas %>% filter(pmd.as != "pupa", # gets super long
                                           #trt == 433
                                           ),
               aes(color = pmd.as), 
               shape = 4, size = 2.5, #show.legend = FALSE
               ) +
    geom_errorbarh(data = dfs$pmdas %>% filter(pmd.as != "pupa", # gets super long
                                               #trt == 433
                                               ),
                   aes(color = pmd.as,
                       xmax = avg.tt + se.tt,
                       xmin = avg.tt - se.tt),
                   lwd = 1, height = 0, alpha = 0.75, #show.legend = FALSE
                   ) +
    
    # scales
    scale_fill_manual(values = c(`pmd` = "black",
                                 `3rd` = "#F8766D", `4th` = "#A3A500", `5th` = "#00BF7D",
                                 `wander` = "#00B0F6")) +
    scale_color_manual(values = c(`younger` = "#F8766D",`4th` = "#A3A500", `5th` = "#00BF7D",
                                  `wander` = "#00B0F6", `pupa` = "#E76BF3")) +
    scale_shape_manual(values = c(`pmd` = 4, 
                                  `3rd` = 21, `4th` = 24, `5th` = 22, `wander` = 23)) +
    do.call(scale_linetype_manual, p_scales$lty_ispup) +
    
    # theme + legend
    theme_bw() +
    labs(y = "pop & developmental outcome", x = "days",
         col = "stage at death (± se)",
         fill = "instar",
         lty = "survived to pupation?") +
    guides(fill = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
           lty = guide_legend(title.position = "top", nrow = 1, byrow = TRUE),
           col = guide_legend(title.position = "top", nrow = 1, byrow = TRUE)) +
    theme(legend.position = "bottom",
          #legend.box = "vertical",
          legend.title = element_text(size = 9, hjust = 0.5),
          legend.margin = margin())
}

dfs$ifpup %>%
  .plotfn() +
  facet_grid(pop ~ trt, scales = "free_x", labeller = labeller(.cols = p_scales$labs_trt)) +
  aes(y = interaction(is.pup, year)) +
  scale_y_discrete(breaks = c("1.2025", "0.2025", "1.2024", "0.2024", "1.2023", "0.2023"),
                   labels = c("2025", "2025", "2024", "2024", "2023", "2023"),
                   # breaks = c(1:3),
                   # labels = levels(interaction(dfs$ifpup$is.pup, dfs$ifpup$year, sep = "&"))
                   )
dfs$ifpup %>%
  .plotfn() +
  facet_grid(year ~ trt, scales = "free_x", labeller = labeller(.cols = p_scales$labs_trt)) +
  aes(y = interaction(is.pup, pop)) +
  scale_y_discrete(breaks = c("1.lab", "0.lab", "1.field", "0.field"),
                   labels = c("lab", "lab", "field", "field"))
```
 
```{r try 3 trajectories}
# NOTE can plot either the individual pmds or the summary pmds -- just comment
# out the geom. but hard to see the ss errorbar if invidid pmds are included

# TODO try a version faceted by trt ~ pop instead

dfs$ifpup %>%
  filter(instar != "pupa",
         #year == 2025,
         #is.pup == 1 # some pup but die after
         ) %>%
  # set levels here so that y axis is in correct order later
  mutate(instar = factor(instar, levels = c("hatch", "younger", 
                                            "3rd", "4th", "5th", "wander", "pupa"))) %>%
  ggplot(aes(y = instar, 
             x = avg.tt,
             col = year,
             group = interaction(year, pop, trt, is.pup)
             )) +
  
  # dev info
  geom_point(shape = 16, size = 2,
             show.legend = FALSE
             ) +
  geom_line(aes(lty = is.pup)) +
  
  # invididual pmd info (needs new unsummarised df)
  geom_point(data = df %>%
               filter(pmd.as != "pupa",
                      is.sup == 0,
                      is.pup != 2,
                      #year == 2025
                      ) %>%
               mutate(is.pup == 0,
                      is.pup = factor(is.pup),
                      ),
             aes(group = interaction(pmd.as, year, pop),
                 #group = interaction(instar, trt, pop),
                 y = pmd.as, shape = instar,
                 x = tt),
             shape = 3, size = 0.5, alpha = 0.7,
             #show.legend = FALSE
             ) +


  # summary pmd info
  geom_point(data = dfs$pmdas2 %>% filter(pmd.as != "pupa"),
             aes(y = pmd.as, shape = instar,
                 group = interaction(pmd.as, year, pop)),
             #shape = 4,
              # NOTE assign value in scale_* instead of here to get
              # the shape to show in the legend on its own
             size = 2,
             #show.legend = FALSE
             ) +
  geom_errorbarh(data = dfs$pmdas2 %>% filter(pmd.as != "pupa"),
                 aes(y = pmd.as, lwd = instar,
                     group = interaction(pmd.as, year, pop),
                     xmax = avg.tt + se.tt,
                     xmin = avg.tt - se.tt),
                 #lwd = 1,
                 height = 0, alpha = 0.75,
                 #show.legend = FALSE
                 ) +
  
  # rest of the plot
  facet_grid(pop ~ trt, scales = "free_x", 
             labeller = labeller(.cols = p_scales$labs_trt)) +
  scale_y_discrete(limits = c("hatch", "younger", "3rd", "4th", "5th", "wander"
                              #"pupa"
                              )) +
  scale_color_brewer(palette = "Dark2") +
  do.call(scale_linetype_manual, p_scales$lty_ispup) +
  theme_bw() +
  scale_shape_manual(values = 4, breaks = "pmd", labels = "pmd") +
  scale_linewidth_manual(values = 1.5, breaks = "pmd", labels = "pmd") +
  guides(linewidth = guide_legend(override.aes = list(lwd = 1))) +
  labs(x = "days",
       lty = "survived to\n pupation?",
       shape = "if pmd", lwd = "if pmd", 
       col = "year")
```
### old attempts

```{r dying bugs try 2 eval=FALSE}
# pre: trying to use just 1 df lol
# post: 
  # still need 2 separate dfs (for dev stuff + pmd info.)
  # needs error bars.
# TOFIX 
  # is.pup == 1 bugs that die from pupa should be omitted from is.pup == 1 in these figs (see yr 2025, is.pup = 1)
  # grouping by is.pup and pmd.as simultaneously leads to double-plotting for pmd bugs

df %>%
  filter(is.sup == 0,
         !(instar %in% c("hatch", "pupa")),
         is.pup != 2) %>%
  PrepLarvalSS() %>%
  group_by(pmd.as, # TOFIX breaks the pmd stuff... (double plots)
           is.pup, .add = TRUE) %>%
  CalcDevSS() %>% #View()
  # mutate(is.pup = case_when(is.pup == 1 ~ "survived",
  #                           is.pup == 0 ~ "died")) %>%
  mutate(is.pup = as.factor(is.pup),
         # set same levels for instars so legends work properly
         pmd.as = factor(pmd.as, levels = c("hatch", "pmd", "younger", "3rd", "4th", "5th", "wander", "pupa")),
         instar = factor(instar, levels = c("hatch", "pmd", "younger", "3rd", "4th", "5th", "wander", "pupa"))) %>%
  filter(trt == 433) %>%
  ggplot(aes(y = interaction(is.pup, year),
             x = avg.tt,
             #col = instar,
             lty = is.pup
             )) +
  # dev points
  geom_point(#data = . %>% filter(is.na(pmd.as)),
             data = . %>% filter(instar != "pmd"),
             aes(shape = instar, fill = instar, col = instar),
             size = 2.5, #stroke = NA,
             #show.legend = FALSE
             ) +
  geom_line(#data = . %>% filter(is.na(pmd.as)),
            aes(#lty = interaction(is.pup, year),
                group = interaction(pop, trt, is.pup, year))
            )+
  
  # pmd points
  geom_point(data = . %>% filter(instar == "pmd",
                                 #pmd.as != "pupa", # too long
                                 is.pup != 2,
                                 !is.na(pmd.as)
                                 ),
             aes(col = pmd.as, #shape = pmd.as,
                 #shape = instar,
                 ),
             #shape = 1, 
             shape = 4,
             size = 2.5, lwd = 1.5,
             show.legend = FALSE
             ) +
  scale_shape_manual(values = c(`pmd` = 4, # gets lost when stroke = NA tho.
                                `3rd` = 21, `4th` = 24, `5th` = 22, `wander` = 23 # fill + col
                                #`3rd` = 16, `4th` = 17, `5th`= 15, `wander` = 7 # fill only
                                ),
                     #breaks = c("pmd", "younger", "3rd", "4th", "5th", "wander"),
                     limits = c("pmd", "3rd", "4th", "5th", "wander")) +
  scale_linetype_manual(labels = c(`1` = "survived", `0` = "died"),
                        values = c(`1` = "solid", `0` = "dashed")) +
  scale_color_manual(values = c(`pmd` = "black", `younger` = "#F8766D", `3rd` = "#B79F00",
                                `4th` = "#00BA38", `5th` = "#619Cff", `wander` = "#F564E3"),
                     limits = c(#"pmd",
                                "younger", "3rd", "4th", "5th", "wander")) +
  scale_fill_manual(values = c(`pmd` = "black", `younger` = "#F8766D", `3rd` = "#B79F00",
                                `4th` = "#00BA38", `5th` = "#619Cff", `wander` = "#F564E3")) +
  facet_grid(pop ~ trt) + 
  labs(shape = "instar", col = "instar", fill = "instar",
       caption = "timelines w/o err bars. needs fixes") +
  guides(fill = "none") +
  theme_bw()
```

```{r dying bugs try 1, eval=FALSE}
# post: thought i had to group the tt.pmd all by itself but actually i can just make seperate pups/pmd dfs!

dfs <- list(
  pups = df %>%
    filter(is.sup == 0, # keeping it simple for now...
           instar != "pmd") %>%
    mutate(is.pup = case_when(is.pup == 2 ~ 1, 
                              TRUE ~ is.pup)) %>%
    PrepLarvalSS() %>%
    group_by(is.pup, .add = TRUE) %>% # idk..
    CalcDevSS() %>%
    mutate(#across(ends_with("sup"), replace_na, 0),
           across(c(starts_with("is")), as.factor)) %>%
     filter(!is.na(is.pup),
            !(instar %in% c("pupa", "hatch"))) %>%
    MakeFactors(),
  
  pmd = df %>%
    filter(is.sup == 0,
           is.pup == 0, # TOFIX some things DO pup but die as pup
           instar == "pmd", # TOFIX should be pmd per instar, not across whole dev time
           !is.na(diet)) %>%
    group_by(across(c(year, pop, diet, starts_with("trt"), instar, is.pup))) %>%
    CalcDevSS() %>%
    mutate(across(starts_with("is"), as.factor)) %>%
    MakeFactors()
)

ggplot(dfs$pup %>% filter(trt == 433), 
       aes(y = interaction(is.pup, year), 
           x = avg.tt, 
           shape = instar,
           col = year,
           lty = is.pup,
           group = interaction(pop, trt, is.pup, year)
       )) +
  geom_point(size = 2) +
  geom_line() +
  geom_point(data = dfs$pmd %>% filter(trt == 433),
             size = 2) +
  scale_linetype_manual(values = c(`1` = "solid", `0` = "dashed"),
                        labels = c(`1` = "yes", `0` = "no")) +
  scale_color_brewer(palette = "Dark2") +
  facet_grid(pop ~ trt) +
  theme_bw() +
  labs(x = "days to instar",
       lty = "pupated?",
       #shape = "survived to\npupation?"
  )
```

# # misc larval dev

## supernumerary props

```{r}
dfs_tidy$widentw %>%
  CalcSupProps_def() %>%
  filter(status == "surv") %>%
  ggplot(aes(y = p,
             x = interaction(trt),
             fill = sup)) +
  geom_col() +
  geom_text(data = . %>% distinct(interaction(year, trt, pop), .keep_all = TRUE), 
            aes(label = N), y = 0.1) +
  facet_grid(pop~year) +
  labs(title = "# instars before pupation",
       y = "proportion survived", x = "trt",
       fill = "extra instars",
       caption = "N = total entered") +
  guides(x = guide_axis(angle = 45)) +
  scale_fill_brewer(palette = "Blues", direction = -1,
                    labels = c("0", "1", "2", "3", "died")) +
  scale_x_discrete(labels = c("26-26", "40-19", "40-26", "40-33")) +
  #scale_fill_discrete(labels = c("0", "1", "2", "3", "died")) +
  theme_bw()
```
# # EDA

## later larval performance

per KS 2021: Responses of MS to HW

- Z = 5th to wandering mass gain/dev time

```{r}
# df prep
df <- dfs_tidy$widentw %>%
  mutate(mass.last = case_when(!is.na(mass.8th) ~ mass.8th,
                               !is.na(mass.7th) ~ mass.7th,
                               !is.na(mass.6th) ~ mass.6th,
                               TRUE ~ mass.5th),
         jdate.last = case_when(!is.na(jdate.8th) ~ jdate.8th,
                                !is.na(jdate.7th) ~ jdate.7th,
                                !is.na(jdate.6th) ~ jdate.6th,
                                TRUE ~ jdate.5th),
         mass.late = mass.wander - mass.last,
         tt.late = jdate.wander - jdate.last,
         dev.late = 1/tt.late,
         
         mass.mid = mass.last - mass.3rd,
         tt.mid = jdate.last - jdate.3rd,
         dev.mid = 1/tt.mid,
         mass.mid2 = mass.wander - mass.3rd,
         tt.mid2 = jdate.wander - jdate.3rd,
         dev.mid2 = 1/tt.mid2
         ) %>%
  mutate(across(c(starts_with("trt"), "year"), as.factor)) %>%
  filter(is.sup == 0) %>%
  
  ## OPTIONAL: set grouping by year or not
  #group_by(across(c(starts_with("trt"), pop, year))) %>%
  group_by(across(c(starts_with("trt"), pop))) %>%
  summarise(n = n(),
            across(.cols = c("mass.late", "dev.late", "mass.mid", "dev.mid", "mass.mid2", "dev.mid2"),
                   .fns = list(avg = ~ mean(.x, na.rm = TRUE),
                               se = se),
                   .names = "{.fn}_{.col}")) %>%
  drop_na(avg_mass.late) %>%
  pivot_longer(cols = contains(c("mass", "dev")),
               names_to = c(".value", "stage"),
               names_sep = "\\."
               )
```

```{r eval=FALSE}
# separate late/mid stage performance plots
.plotfn <- function(x){
  x %>%
    ggplot(aes(x = trt, #col = year, group = interaction(trt.type, year)
               group = interaction(trt.type))) +
    geom_point(size = 2) +
    geom_line() +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

df %>%
  .plotfn() +
  aes(y = avg.mass.mid) +
  geom_errorbar(aes(ymax = avg.mass.mid + se.mass.mid,
                    ymin = avg.mass.mid - se.mass.mid,
                    width = 0.25)) +
  labs(y = "3rd-5th mass gain (mg)")

df %>%
  .plotfn() +
  aes(y = avg.mass.late) +
  geom_errorbar(aes(ymax = avg.mass.late + se.mass.late,
                    ymin = avg.mass.late - se.mass.late,
                    width = 0.25)) +
  labs(y = "5th-wander mass gain (mg)")

df %>%
  .plotfn() +
  aes(y = avg.dev.mid) +
  geom_errorbar(aes(ymax = avg.dev.mid + se.dev.mid,
                    ymin = avg.dev.mid - se.dev.mid,
                    width = 0.25)) +
  labs(y = "3rd-5th dev rate (1/days)")

df %>%
  .plotfn() +
  aes(y = avg.dev.late) +
  geom_errorbar(aes(ymax = avg.dev.late + se.dev.late,
                    ymin = avg.dev.late - se.dev.late,
                    width = 0.25)) +
  labs(y = "5th-wander dev rate (1/days)")

```

```{r}
# combined performance @ stage plots
.plotfn <- function(x){
  x %>%
    mutate(stage = case_when(stage == "mid" ~ "3rd - 5th",
                             stage == "mid2" ~ "3rd - wander",
                            stage == "late" ~ "5th - wander")) %>%
    ggplot(aes(x = trt, shape = stage, col = stage, 
               group = interaction(trt.type, stage))) +
    geom_point(size = 2) +
    geom_line() +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

df %>%
  .plotfn() +
  aes(y = avg_mass) +
  geom_errorbar(aes(ymax = avg_mass + se_mass,
                    ymin = avg_mass - se_mass),
                width = 0.25) +
  labs(y = "mass gain")

df %>%
  .plotfn() +
  aes(y = avg_dev) +
  geom_errorbar(aes(ymax = avg_dev + se_dev,
                    ymin = avg_dev - se_dev),
                width = 0.25) +
  labs(y = "dev rate (1/days)")
```

## alternate dev x axes explorations

(using dev rate)

- how best to show ctrl vs expt? level of mean/fluct?
- colors are good to show "ctrl" vs "expt" (hot bugs)
- some x-axes are a little confusing bc there's so many other points (in 2023).
  not sure what other `trt.` var to choose as another geom. 
  - (prob sth that emphasises the minTs or other aspect of the temp trt, but also dont want like
     3 additional levels max..)

### df prep  
```{r}
# making some dfs to test
.prepfn <- function(x){
  x %>%
    CalcDevSS_la() %>%
    filter(instar == "pupa")
}

dfs <- list(
  twobytwo = dfs_tidy$long %>%
    FilterOutLabTB() %>% 
    #filter(trt %in% c(260, 330, 419, 426, 433)) %>% # drops 267...
    .prepfn(),
  
  ntw = dfs_tidy$longntw %>%
    .prepfn()
)
```

```{r}
.plotfn <- function(x){
  list(
    by_minT = x %>%
      mutate(trt.flucT = as.numeric(as.character(trt.flucT)),
             trt.flucT = case_when(trt.flucT == 0 ~ "N",
                                   TRUE ~ "Y"),
      ) %>% 
      ggplot(aes(y = avg.devrate,
                 x = trt.minT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + 
      geom_point(#aes(shape = trt.flucT),
        aes(shape = trt.maxT),
        size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1),
    
    by_flucT = x %>%
      ggplot(aes(y = avg.devrate,
                 x = trt.flucT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + geom_point(aes(shape = trt.meanT), size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1),
    
    by_meanT = x %>%
      ggplot(aes(y = avg.devrate,
                 x = trt.meanT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + geom_point(aes(shape = trt.minT), size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1)
    )
}


dfs$twobytwo %>% .plotfn()

dfs$ntw %>% .plotfn()

```

# ### adult dev

## longevity

```{r EDA fridge bugs, eval = FALSE}
# EDA: effect of including fridge bugs?
.prepfn <- function(x){
  x %>%
    PrepAdultSS() %>%
    filter(instar == "long")
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt, x = trt,
               color = pop)
           )+
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid") +
    geom_point() + geom_line() +
    facet_grid(~ year) +
    labs(y = "longevity (days after eclosion)") 
}

# EDA: keep fridge bugs in the count
dfs_tidy$longntw %>%
  .prepfn() %>%
  CalcDevSS() %>%
  .plotfn() +
  aes(group = interaction(pop))

# EDA: parse out the fridge bugs
dfs_tidy$longntw %>%
  mutate(if.fridge = case_when(is.na(if.fridge) ~ "no",
                               TRUE ~ "yes")) %>%
  .prepfn() %>%
  group_by(if.fridge, .add = TRUE) %>%
  CalcDevSS() %>%
  #filter(is.na(if.fridge)) %>%
  .plotfn() +
  aes(group = interaction(pop, if.fridge),
      lty = if.fridge, shape = if.fridge)
```

```{r}
# patchwork
# final: remove the fridge bugs
.prepfn <- function(x){
  x %>%
    filter(is.na(if.fridge),
           instar == "long") %>%
    CalcDevSS_ad()
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               col = year, 
               shape = pop, lty = pop)
    ) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + geom_line() +
    facet_grid(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "longevity (days after eclosion)",
         shape = "population", lty = "population") +
    scale_shape_manual(values = p_scales$shp_pop) +
    scale_linetype_manual(values = p_scales$lty_pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    ylim(c(4, 17.5))
}

p1 <- dfs_tidy$longntw %>%
  filter(trt == 260) %>%
  .prepfn() %>%
  .plotfn() +
  aes(x = pop)

p2 <- dfs_tidy$longntw %>%
  filter(trt != 260) %>%
  .prepfn() %>%
  .plotfn() +
    aes(x = trt,
        group = interaction(year, pop)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = "minimum temperature (°C)")

p1 + p2 +
  plot_layout(widths = c(1,2),
              axes = "collect", guides = "collect")
```

