---
title: "ntw 2025 data viz"
date: "2025-12-01"
---

# # preamble & loading

data viz for ntw 2025

```{r set pathfinding}
here::i_am("_ntw/scripts/analysis/viz-dev.Rmd")
library(here)
```

```{r load libraries and utils, message=FALSE, warning=FALSE}
library(tidyverse)
library(patchwork) # for plot_layout
# library(survival) # for fitting surv
# library(survminer) # for plotting surv

# data
source(here("_ntw/scripts/R/tidy-dev.R"))

# utils
source(here("_ntw/scripts/utils/wrangle.R"))
source(here("_ntw/scripts/utils/analysis.R"))
source(here("_ntw/scripts/utils/analysis-viz.R"))
```

```{r convenience fns}
# convert things into factors for ease of graphing
MakeFactors <- function(ss_df){
  ss_df %>%
        mutate(trt.facet = case_when(trt == 260 ~ "ctrl",
                                     pop == "lab" ~ "lab",
                                     pop == "field" ~ "field")) %>%
        mutate(across(c("year", starts_with("trt")), as.factor))
}


#dfs_ldev <- dfs_ldev %>% MakeFactors()

#dfs_adev <- dfs_adev %>% MakeFactors()
```

# # larval dev

## survival to pup
```{r}
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    filter(pop != "col") %>%
    CalcSurvProps()
}

.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
  ggplot(aes(y = prop.pup,
             x = trt,
             col = year, #lty = year,
             group = year
             )) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = prop.pup + se.pup,
                    ymin = prop.pup - se.pup),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  # geom_text(aes(label = n),
  #           hjust = -0.2, vjust = -.2,
  #           show.legend = FALSE) +
  facet_wrap(~pop) +
  labs(y = "% survived to pupation",
       x = "treatment",
       col = "year", #lty = "year"
       ) +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
}
```

```{r}
# all trts
dfs_tidy$wide %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

# expt trts only by minT
dfs_tidy$wide %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_minT) +
    labs(x = "minimum T (°C)")

# ctrls vs expt
dfs_tidy$wide %>%
  .prepfn() %>%
  .plotfn() +
  facet_wrap(~trt.facet, scales = "free_x") +
  scale_x_discrete(labels = p_scales$labs_trt) +
  aes(shape = pop)

```

## pup mass

### sexes combined
```{r}
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    PrepAdultSS() %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
  ggplot(aes(y = avg.mass, x = trt,
             #lty = year, 
             col = year,
             group = year,
             ))+
  geom_point() + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, 
                    ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") +
  # geom_text(aes(label = n),
  #           vjust = -0.2, hjust = -0.3,
  #           show.legend = FALSE) +
  facet_grid(~pop) +
  labs(y = "pupal mass (g)",
       caption = "mean±se; combined sexes") +
  scale_color_brewer(palette = "Dark2") +
  theme_bw()
}

```

```{r}
dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
    scale_x_discrete(labels = p_scales$labs_trt)
```

### sexes parsed
```{r}
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    filter(!is.na(sex)) %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}
```

```{r}
# 2x2 pop + sex (better)
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.mass, x = trt,
               col = year,
               group = interaction(year)
               )) +
    geom_point() + geom_line() +
    # geom_text(aes(label = n),
    #           vjust = -0.5, hjust = -0.5, size = 3.5,
    #           position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) + # too busy
    geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                  width = 0.3, lty = "solid") + # remove if including ctrl
    labs(y = "pupal mass (g)", 
         caption = "mean±se") +
    facet_grid(sex~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260,
         #sex == "m"
         ) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_minT) +
    labs(x = "minimum T (°C)")

```

```{r}
# by sex (meh)
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.mass, x = trt,
             col = year,
             group = interaction(year, pop), 
             shape = pop, lty = pop
             )) +
  geom_point(size = 2) + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") + # remove if including ctrl
  # geom_text(aes(label = n),
  #           vjust = -0.5, hjust = -0.5, size = 3.5,
  #           position = position_jitter(width = 0.2, height = 0.3, seed = 1),
  #           show.legend = FALSE) + # too busy
  labs(y = "pupal mass (g)", 
       caption = "mean±se") +
  facet_grid(~sex) +
    scale_color_brewer(palette = "Dark2") +
    scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  theme_bw()
}

dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

```{r}
# by pop (no)
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.mass, x = trt,
             col = year,
             group = interaction(year, sex), 
             shape = sex, lty = sex
             )) +
  geom_point(size = 2) + geom_line() +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.3, lty = "solid") +
  labs(y = "pupal mass (g)", 
       caption = "mean±se") +
  facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    scale_shape_manual(values = c(1, 16)) +
  scale_linetype_manual(values = c("dashed", "solid")) +
  theme_bw()
}

dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

## pup time

### days to pup
```{r days to pup}
# sexes combined
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    PrepAdultSS() %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.tt, x = trt,
               col = year,
               group = year)) +
    geom_point(size = 2) + geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25) +
    # geom_text(aes(label = n),
    #           vjust = -0.75, hjust = -0.5, size = 3.5,
    #           #position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(y = "mean±se time to pupation (days)",
         x = "temperature trt",
         title = "time to pupation")
}

dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
# LOL the 260 lab every yr ...  i think its bc they take so long from 5 to wander

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

```

### pup rate
```{r pup rate}
# sexes combined
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    PrepAdultSS() %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}

.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.devrate, x = trt,
               col = year,
               group = year)) +
    geom_point(size = 2) + geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25) +
    # geom_text(aes(label = n),
    #           vjust = -0.75, hjust = -0.5, size = 3.5,
    #           #position = position_jitter(width = 0.2, height = 0.3, seed = 1),
    #           show.legend = FALSE) +
    facet_grid(~pop) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(y = "development rate (1/days)",
         x = "temperature trt",
         title = "time to pupation")
}

dfs_tidy$long %>%
  .prepfn() %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)

dfs_tidy$long %>%
  .prepfn() %>%
  filter(trt != 260) %>%
  .plotfn() +
  scale_x_discrete(labels = p_scales$labs_trt)
```

# # bonus larval dev

## dev trajectories 

### df prep
```{r}
# make some variations of the df based off how the pupae are handled
.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    PrepLarvalSS()
}

.calcfn <- function(x){
  x %>%
    CalcDevSS() %>%
    filter(instar %in% c("3rd", "4th", "5th", "wander"))
}

dfs <- list(
  # default grouping
  all = dfs_tidy$long %>%
    .prepfn() %>%
    .calcfn(),
  
  # by pup, with survivors + culled combined together
  pups_wcull = dfs_tidy$long %>%
    mutate(is.pup = case_when(is.pup == 2 ~ 1,
                              TRUE ~ as.numeric(is.pup))) %>%
    .prepfn() %>% 
    group_by(is.pup, .add = TRUE) %>%
    .calcfn(),
  
  # by pup, with survivors + culled kept separate
  pups_sepcull = dfs_tidy$long %>%
    .prepfn() %>%
    group_by(is.pup, .add = TRUE) %>%
    .calcfn()
)

```


### figs
```{r}
# dev by year
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = year)) +
    geom_point() +
    geom_line() +
    # hard to see with these lol bc theyre pretty close
    # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
    #                   ymin = avg.logmass - se.logmass,
    #                   width = CalcErrWd(avg.tt))) +
    # geom_errorbarh(aes(xmax = avg.tt + se.tt,
    #                    xmin = avg.tt - se.tt,
    #                    height = CalcErrHt(avg.logmass))) +
    facet_grid(pop ~ trt) +
    #scale_color_brewer(palette = "Greens") +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "year")
}

dfs$all %>%
  .plotfn() +
  labs(title = "development trajectories (incl pmd)")

dfs$pups_wcull %>%
  filter(is.pup != 0) %>%
  .plotfn() +
  labs(title = "development trajectories (pup'd only)")
```

```{r}
# dev by year; parse out pmds more
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    mutate(is.pup = as.factor(is.pup)) %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = year, lty = is.pup, 
               group = interaction(year, is.pup, trt, pop))) +
    geom_point() +
    geom_line() +
    facet_grid(pop ~ trt) +
    scale_color_brewer(palette = "Dark2") +
    theme_bw() +
    labs(title = "development trajectories",
         caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "year", lty = "fate"
    ) +
    scale_linetype_discrete(labels = c("pupated", "died"))
}

# surv, combined w culled
dfs$pups_wcull %>%
  .plotfn()

# surv, culled omitted
dfs$pups_sepcull %>%
  filter(is.pup != 2) %>%
  .plotfn()

```

```{r}
# dev by trt
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
  ggplot(aes(y = avg.logmass,
             x = avg.tt,
             col = trt)) +
  geom_point() +
  geom_line() +
  # hard to see with these lol bc theyre pretty close
  # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
  #                   ymin = avg.logmass - se.logmass,
  #                   width = CalcErrWd(avg.tt))) +
  # geom_errorbarh(aes(xmax = avg.tt + se.tt,
  #                    xmin = avg.tt - se.tt,
  #                    height = CalcErrHt(avg.logmass))) +
  facet_grid(pop ~ year) +
  #scale_color_brewer(palette = "Greens") +
  #scale_color_manual(values = hcl.colors(6, "Viridis", rev = TRUE)) +
    scale_color_manual(values = p_scales$cols_trt,
                       labels = p_scales$labs_trt) +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "treatment") +
    theme_bw()
}


dfs$pups_wcull %>%
  filter(is.pup != 0) %>%
  .plotfn() +
  labs(title = "development trajectories (pups only)")

dfs$all %>%
  .plotfn() +
  labs(title = "development trajectories (all)")

```
```{r}
# dev traj by trt + surv to pupation
.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    mutate(is.pup = case_when(is.pup == 0 ~ "died",
                              is.pup == 1 ~ "pupated")) %>%
    mutate(is.pup = factor(is.pup, levels = c("pupated", "died"))) %>%
    ggplot(aes(y = avg.logmass,
               x = avg.tt,
               col = trt,
               group = interaction(year, pop, is.pup, trt))) +
    geom_point() +
    geom_line() +
    # hard to see with these lol bc theyre pretty close
    # geom_errorbar(aes(ymax = avg.logmass + se.logmass,
    #                   ymin = avg.logmass - se.logmass,
    #                   width = CalcErrWd(avg.tt))) +
    # geom_errorbarh(aes(xmax = avg.tt + se.tt,
    #                    xmin = avg.tt - se.tt,
    #                    height = CalcErrHt(avg.logmass))) +
    facet_grid(is.pup ~ year) +
    #scale_color_brewer(palette = "Greens") +
    #scale_color_manual(values = hcl.colors(6, "Viridis", rev = TRUE)) +
    scale_color_manual(values = p_scales$cols_trt,
                       labels = p_scales$labs_trt) +
    labs(caption = "mean ± se",
         y = "log(mass mg)", x = "dev time (days)",
         col = "treatment") +
    theme_bw()
}


dfs$pups_sepcull %>%
  filter(is.pup != 2,
         pop == "field") %>%
  .plotfn() +
  labs(title = "development trajectories (field)")
  
dfs$pups_sepcull %>%
  filter(is.pup != 2,
         pop == "lab") %>%
  .plotfn() +
  labs(title = "development trajectories (lab)")
```


```{r}
# wtf goin on w the lab controls
dfs$all %>%
  filter(trt == 260) %>%
  MakeFactors() %>%
  ggplot(aes(y = avg.logmass,
             x = avg.tt,
             col = year, lty = pop, shape = pop,
             group = interaction(year, pop))) +
  geom_errorbarh(aes(xmax = avg.tt + se.tt,
                     xmin = avg.tt - se.tt,
                     height = 0.3), lty = "solid") +
  geom_point() +
  geom_line() +
  facet_grid(#pop 
             ~ trt) +
  #scale_color_brewer(palette = "Greens") +
  scale_color_brewer(palette = "Dark2") +
  scale_shape_manual(values = c(16, 1)) +
  labs(caption = "what's up w the controls") +
  theme_bw()
```


## supernumerary

```{r}
dfs_tidy$wide %>%
  FilterForNTWBugs() %>%
  CalcSupProps() %>%
  filter(status == "surv") %>%
  MakeFactors() %>%
  ggplot(aes(y = p,
             x = interaction(trt),
             fill = sup)) +
  geom_col() +
  geom_text(data = . %>% distinct(interaction(year, trt, pop), .keep_all = TRUE), 
            aes(label = N), y = 0.1) +
  facet_grid(pop~year) +
  labs(title = "# instars before pupation",
       y = "proportion survived", x = "trt",
       fill = "extra instars",
       caption = "N = total entered") +
  guides(x = guide_axis(angle = 45)) +
  scale_fill_brewer(palette = "Blues", direction = -1,
                    labels = c("0", "1", "2", "3", "died")) +
  scale_x_discrete(labels = c("26-26", "40-19", "40-26", "40-33")) +
  #scale_fill_discrete(labels = c("0", "1", "2", "3", "died")) +
  theme_bw()
```

## alternate x axes explorations

(using dev rate)

- how best to show ctrl vs expt? level of mean/fluct?
- colors are good to show "ctrl" vs "expt" (hot bugs)
- some x-axes are a little confusing bc there's so many other points (in 2023).
  not sure what other `trt.` var to choose as another geom. 
  - (prob sth that emphasises the minTs or other aspect of the temp trt, but also dont want like
     3 additional levels max..)
     
```{r}
# making some dfs to test
.prepfn <- function(x){
  x %>%
    PrepLarvalSS() %>%
    CalcDevSS() %>%
    filter(instar == "pupa")
}

dfs <- list(
  twobytwo = dfs_tidy$long %>%
    FilterOutLabTB() %>% 
    #filter(trt %in% c(260, 330, 419, 426, 433)) %>% # drops 267...
    .prepfn(),
  
  ntw = dfs_tidy$long %>%
    FilterForNTWBugs() %>%
    .prepfn()
)

```


```{r}
.plotfn <- function(x){
  list(
    by_minT = x %>%
      MakeFactors() %>%
      mutate(trt.flucT = as.numeric(as.character(trt.flucT)),
             trt.flucT = case_when(trt.flucT == 0 ~ "N",
                                   TRUE ~ "Y"),
      ) %>% 
      ggplot(aes(y = avg.devrate,
                 x = trt.minT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + 
      geom_point(#aes(shape = trt.flucT),
        aes(shape = trt.maxT),
        size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1),
    
    by_flucT = x %>%
      MakeFactors() %>%
      ggplot(aes(y = avg.devrate,
                 x = trt.flucT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + geom_point(aes(shape = trt.meanT), size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1),
    
    by_meanT = x %>%
      MakeFactors() %>% 
      ggplot(aes(y = avg.devrate,
                 x = trt.meanT, col = trt.type,
                 group = interaction(trt.type))) +
      geom_line() + geom_point(aes(shape = trt.minT), size = 2) +
      facet_grid(pop~year) +
      scale_color_discrete(direction = -1)
    )
}


dfs$twobytwo %>% .plotfn()

dfs$ntw %>% .plotfn()

```

# # adult dev

## longevity
```{r}

.prepfn <- function(x){
  x %>%
    FilterForNTWBugs() %>%
    filter(pop != "col") %>%
    PrepAdultSS() %>%
    filter(instar == "long")
}

.plotfn <- function(x){
  x %>%
    MakeFactors() %>%
    ggplot(aes(y = avg.tt, x = trt,
               color = pop)
           )+
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid") +
    geom_point() + geom_line() +
    facet_grid(~ year) +
    labs(y = "longevity (days after eclosion)") 
}

dfs_tidy$long %>%
  .prepfn() %>%
  CalcDevSS() %>%
  .plotfn() +
  aes(group = interaction(pop))

# this is silly...
# dfs_tidy$long %>%
#   .prepfn() %>%
#   group_by(if.fridge, .add = TRUE) %>%
#   CalcDevSS() %>%
#   mutate(if.fridge = case_when(!is.na(if.fridge) ~ "yes",
#                                TRUE ~ "no")) %>%
#   .plotfn() +
#   aes(group = interaction(pop, if.fridge),
#       lty = if.fridge)

dfs_tidy$long %>%
  .prepfn() %>%
  group_by(if.fridge, .add = TRUE) %>%
  CalcDevSS() %>% View()
  filter(is.na(if.fridge)) %>%
  .plotfn() +
  aes(group = interaction(pop))
  
```

