---
title: "tdt viz n stats"
date: "2025-06-16"
---

# preamble & loading

combined viz/stats for tdt stuff until i can be bothered to split them apart


```{r}
library(tidyverse)
#library(MuMIn) # for survival nested model comparisons
library(survival) # for fitting 
library(survminer) # for plotting
```

```{r}
here::i_am("2025/analysis-tdt.Rmd")
library(here)

source(here::here("set-paths.R"))
#source(here::here(bin_paths$y25, "compare-dev_p.R"))

#data <- read.csv(here(bin_paths$y25$data, "clean-working.csv"), header = TRUE) # can remove once wrangle script is done...
source(here(bin_paths$y25$root, "wrangle-tdt_p.R"))
```

# * setup (lists)

```{r}
# set up surv objects

survObj <- list()

# model fits

fits <- list()

```

# 2025-06-27 redoing cox fits lmao

cuz i love a fresh start!!!

- goals
  - [ ] set up km
    - [ ] which is the best time frame to use?
  - [ ] try a few cphs
  
## create fits

```{r}
survObj <- list_assign(survObj,
                       fromEnter = Surv(km$subsets$expts$exit.fromenter, km$subsets$expts$km.status),
                       fromRecovery = Surv(km$subsets$expts$exit.fromrecover, km$subsets$expts$km.status),
                       fromReturn = Surv(km$subsets$expts$exit.fromreturn, km$subsets$expts$km.status)
                       )
```

## visualise km

```{r}
# all exptal trts together
obj <- Surv(km$subsets$expts$exit.fromenter, km$subsets$expts$km.status)
fit <- survfit(obj ~ trt.recover + trt.duration, data = km$subsets$expts)

ggsurvplot(fit, data = km$subsets$expts,
           conf.int = TRUE,
           #xlim = c(45823, 45841),
           facet.by = "trt.recover",
           #linetype = "strata",
           palette = "strata"
           ) +
  geom_vline(xintercept = 2, color = "black") +
  geom_vline(xintercept = 3, color = "#F8766D") +
  geom_vline(xintercept = 4, color = "#00BFC4") +
  labs(caption = "black = entered recovery; colored = returned to 40") +
  theme(plot.caption = element_text(size = 10))

```

```{r}
# everyone together
obj <- Surv(km$data$exit.fromenter, km$data$km.status)
fit <- survfit(obj ~ trt.lab + trt.duration + trt.recover, data = km$data)

ggsurvplot(fit, data = km$data,
           #conf.int = TRUE,
           #xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #linetype = "strata",
           palette = "strata"
           ) +
  geom_vline(xintercept = 2, color = "black") +
  geom_vline(xintercept = 3, color = "#F8766D") +
  geom_vline(xintercept = 4, color = "#00BFC4")

```

## coxphing

```{r}
fits <- list_assign(fits, 
                    cph.enter = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration), data = km$subsets$expts),
                    cph.recovery = coxph(survObj$fromRecovery ~ trt.recover * as.factor(trt.duration), data = km$subsets$expts),
                    cph.return = coxph(survObj$fromReturn ~ trt.recover * as.factor(trt.duration), data = km$subsets$expts),
                    )

summary(fits$cph.enter)
summary(fits$cph.recovery)
summary(fits$cph.return)
```

```{r}
obj <- Surv(km$subsets$only40$exit.fromenter, km$subsets$only40$km.status)
obj2 <- Surv(km$subsets$only40$exit.fromreturn, km$subsets$only40$km.status)
fits <- list_assign(fits, 
                    cph.enter.addshelf = coxph(obj ~ trt.recover * as.factor(trt.duration) + shelf, data = km$subsets$only40),
                    cph.enter.intshelf = coxph(obj ~ trt.recover * as.factor(trt.duration) * shelf, data = km$subsets$only40),
                    cph.return.addshelf = coxph(obj2 ~ trt.recover * as.factor(trt.duration) + shelf, data = km$subsets$only40),
                    cph.return.intshelf = coxph(obj2 ~ trt.recover * as.factor(trt.duration) * shelf, data = km$subsets$only40),)

summary(fits$cph.enter.addshelf)
summary(fits$cph.enter.intshelf)
summary(fits$cph.return.addshelf)
summary(fits$cph.return.intshelf)
```









# 2025-06-22 coxph attempts

- todo
  - [x] shelf effects: look at starting exit-fromenter...

- notes:
  - also worked on: 2025-06-23, 2025-06-24
  - 2025-06-27: 
    - lowkey the fits arent that correct bc 1) the exit dates are calced wrong and 2) the `dh.<stage>` stuff was sorta assigned wrong (i think)
    - also i calced recovery time wrong lmao (used return time...)
  


```{r}
survObj <- list_assign(survObj, 
                       fromHatch = Surv(km$subsets$expts$exit.fromhatch, km$subsets$expts$km.status),
                       fromEnter = Surv(km$subsets$expts$exit.fromenter, km$subsets$expts$km.status),
                       fromRecovery = Surv(km$subsets$expts$exit.fromrecover, km$subsets$expts$km.status),
                       fromReturn = Surv(km$subsets$expts$exit.fromreturn, km$subsets$expts$km.status)
                       )
```


## survdiffing (**)

- NOTE this is where `kmdf` is

```{r trying survdiff interactions, eval=FALSE}

## so google ai says that survdiff is not intended to work w interaction terms LOL... sure!

##### trying to figure out how to get the interaction to show up but idky something about it breaks the formula
  # renaming to nicer vars doesnt work...

#survdiff(survObj$fromHatch ~ km$subsets$expts$trt.duration + km$subsets$expts$trt.recover)
# trying to pick out the columns names to get the interaction to work. see other attempts below
           # pluck(km$subsets, expts, trt.recover):pluck(km$subsets, expts, trt.duration))
           # 
           # hoist(as.data.frame(km$subsets), "trt.recover"):hoist(as.data.frame(km$subsets), "trt.duration"))
           # 
           # km$subsets$[expts][km$subsets$expts$,]:km$subsets[expts][km$subsets$expts$trt.recover,])


#### other attempts 

# kmdf <- km$subsets$expts %>%
#   rename(duration = trt.duration,
#          recover = trt.recover,
#          fromrec = exit.fromrecover,
#          stat = km.status)
# make a subset so naming is nicer, apparently u shouldnt use $ in formulas
# 
# survdiff(Surv(fromrec, stat) ~ 
#           duration + recover + interaction(duration, recover), data = kmdf) 
# this is bad (dont use interaction() cuz its ≠ to using :)
# 
# survdiff(Surv(fromrec, stat) ~ 
#           duration + recover + duration:recover, data = kmdf) # doesnt work
# 
#survdiff(formula = Surv(kmdf$fromrec, kmdf$stat) ~ duration + recover + duration:recover, data = kmdf) # doesnt work


### 3

# survdiff(survObj$fromHatch ~ poly(c(km$subsets$expts$trt.duration, km$subsets$expts$trt.recover), n = 2)) 
  # im totally using poly() wrong here but idr how to use it. but also def not what im looking for

### 4 am i going nuts

# anova(lm(fromrec ~ duration*recover, data = kmdf)) # i am not...

```


```{r eval=FALSE}
# survdiff starting from hatch time
## recovery temp + duration

survdiff(survObj$fromHatch ~ km$subsets$expts$trt.duration + km$subsets$expts$trt.recover 
         #+ `km$subsets$expts$trt.recover`:`km$subsets$expts$trt.duration`
         )

#                                                                    N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=19 40       40     42.8   0.17690   0.23620
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=26 45       45     39.2   0.86120   1.06911
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=33 45       45     40.6   0.46881   0.58530
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=19 44       44     42.0   0.09789   0.12282
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=26 45       45     54.9   1.79965   2.40490
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=33 45       45     44.5   0.00545   0.00693
#  Chisq= 3.7  on 5 degrees of freedom, p= 0.6 


#######

# survdiff starting from enter time
## recovery temp + duration

survdiff(survObj$fromEnter ~ km$subsets$expts$trt.duration + km$subsets$expts$trt.recover)

#                                                                    N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=19 40       40     47.7    1.2533     1.623
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=26 45       45     38.5    1.1097     1.365
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=33 45       45     32.3    5.0291     6.070
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=19 44       44     46.9    0.1740     0.223
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=26 45       45     55.6    2.0166     2.760
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=33 45       45     43.1    0.0843     0.106
#  Chisq= 10.4  on 5 degrees of freedom, p= 0.06 


#######


# looking for diffs starting at recovery
## based on recovery temp
survdiff(survObj$fromRecovery ~ km$subsets$expts$trt.recover)

#                                  N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.recover=19 84       84     84.7   0.00617   0.00976
# km$subsets$expts$trt.recover=26 90       90     94.1   0.18157   0.29812
# km$subsets$expts$trt.recover=33 90       90     85.1   0.27711   0.43458
#  Chisq= 0.5  on 2 degrees of freedom, p= 0.8 

survdiff(survObj$fromRecovery ~ km$subsets$expts$trt.duration)
## based on recovery period
#                                    N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.duration=24 130      130      123      0.45     0.894
# km$subsets$expts$trt.duration=48 134      134      141      0.39     0.894
#  Chisq= 0.9  on 1 degrees of freedom, p= 0.3 


survdiff(survObj$fromRecovery ~ km$subsets$expts$trt.duration + km$subsets$expts$trt.recover)
# (cant do the interaction)

#                                                                    N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=19 40       40     42.8   0.17690   0.23620
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=26 45       45     39.2   0.86120   1.06911
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=33 45       45     40.6   0.46881   0.58530
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=19 44       44     42.0   0.09789   0.12282
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=26 45       45     54.9   1.79965   2.40490
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=33 45       45     44.5   0.00545   0.00693
#  Chisq= 3.7  on 5 degrees of freedom, p= 0.6


########

# survdiff starting from returning time
## recovery temp + duration
survdiff(survObj$fromReturn ~ km$subsets$expts$trt.duration + km$subsets$expts$trt.recover)

#                                                                    N Observed Expected (O-E)^2/E (O-E)^2/V
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=19 40       40     42.8   0.17690   0.23620
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=26 45       45     39.2   0.86120   1.06911
# km$subsets$expts$trt.duration=24, km$subsets$expts$trt.recover=33 45       45     40.6   0.46881   0.58530
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=19 44       44     42.0   0.09789   0.12282
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=26 45       45     54.9   1.79965   2.40490
# km$subsets$expts$trt.duration=48, km$subsets$expts$trt.recover=33 45       45     44.5   0.00545   0.00693
#  Chisq= 3.7  on 5 degrees of freedom, p= 0.6 

```
## cphing

### fitting cph

#### effects of recovery temp and duration
```{r}
# for all:
# n= 264, number of events= 264 

fits <- list_assign(fits, cph.enter = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration), 
                                               data = km$subsets$expts))
summary(fits$cph.enter)

#                                coef  exp(coef)   se(coef)      z      p
# trt.recover               0.0671178  1.0694215  0.0347183  1.933   0.0532 .
# trt.duration              0.0235347  1.0238138  0.0246043  0.957   0.3388  
# trt.recover:trt.duration -0.0012162  0.9987846  0.0009202 -1.322   0.1863  
#
# Concordance= 0.562  (se = 0.018 )
# Likelihood ratio test= 8.4  on 3 df,   p=0.04
# Wald test            = 8.92  on 3 df,   p=0.03
# Score (logrank) test = 9.02  on 3 df,   p=0.03

fits <- list_assign(fits, cph.recovery = coxph(survObj$fromRecovery ~ trt.recover * as.factor(trt.duration), 
                                               data = km$subsets$expts))
summary(fits$cph.recovery)
#                                coef  exp(coef)   se(coef)      z Pr(>|z|)
# trt.recover               0.0279399  1.0283338  0.0350318  0.798    0.425
# trt.duration              0.0116693  1.0117376  0.0248779  0.469    0.639
# trt.recover:trt.duration -0.0006269  0.9993733  0.0009280 -0.676    0.499
# 
# Concordance= 0.535  (se = 0.019 )
# Likelihood ratio test= 1.54  on 3 df,   p=0.7
# Wald test            = 1.57  on 3 df,   p=0.7
# Score (logrank) test = 1.58  on 3 df,   p=0.7


fits <- list_assign(fits, cph.return = coxph(survObj$fromReturn ~ trt.recover * as.factor(trt.duration), 
                                             data = km$subsets$expts))
summary(fits$cph.return)
#                                coef  exp(coef)   se(coef)      z Pr(>|z|)
# trt.recover               0.0279399  1.0283338  0.0350318  0.798    0.425
# trt.duration              0.0116693  1.0117376  0.0248779  0.469    0.639
# trt.recover:trt.duration -0.0006269  0.9993733  0.0009280 -0.676    0.499
# 
# Concordance= 0.535  (se = 0.019 )
# Likelihood ratio test= 1.54  on 3 df,   p=0.7
# Wald test            = 1.57  on 3 df,   p=0.7
# Score (logrank) test = 1.58  on 3 df,   p=0.7

fits <- list_assign(fits, cph.return2 = coxph(survObj$fromReturn ~ trt.recover * trt.duration, 
                                             data = km$subsets$expts))
summary(fits$cph.return2)
# same results as above

fits <- list_assign(fits, cph.return3 = coxph(survObj$fromReturn ~ as.factor(trt.recover) * trt.duration, 
                                             data = km$subsets$expts))
summary(fits$cph.return3)

#                                            coef exp(coef)  se(coef)      z Pr(>|z|)
# as.factor(trt.recover)26               0.711290  2.036617  0.497724  1.429    0.153
# as.factor(trt.recover)33               0.419369  1.521002  0.496156  0.845    0.398
# trt.duration                           0.005547  1.005563  0.009358  0.593    0.553
# as.factor(trt.recover)26:trt.duration -0.020138  0.980063  0.012938 -1.557    0.120
# as.factor(trt.recover)33:trt.duration -0.009401  0.990643  0.012886 -0.730    0.466
# 
# Concordance= 0.536  (se = 0.023 )
# Likelihood ratio test= 3.75  on 5 df,   p=0.6
# Wald test            = 3.65  on 5 df,   p=0.6
# Score (logrank) test = 3.67  on 5 df,   p=0.6

```

so the cph fits only really matter starting from the enter time?

#### shelf effects

```{r}
# looking at shelf effects

fits <- list_assign(fits, cph.shelf = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration) * shelf, 
                                               data = km$subsets$expts))
summary(fits$cph.shelf)


fits <- list_assign(fits, cph.shelf2 = coxph(survObj$fromEnter ~ shelf + trt.recover * as.factor(trt.duration), 
                                               data = km$subsets$expts))
summary(fits$cph.shelf2)

fits <- list_assign(fits, cph.shelf3 = coxph(survObj$fromEnter ~ shelf + trt.recover * trt.duration, 
                                               data = km$subsets$expts))
summary(fits$cph.shelf3)
```

tldr in all cases, `trt.recover` becomes signif once u throw in the shelf LOL.. (generally p < 0.5)

### plotting cph

```{r eval=FALSE}
# plot cphs...
# plot(x = fits$cph.recovery$linear.predictors,
#      y = fits$cph.recovery$residuals,
#      data = km$subsets$expts, 
#      #facet.by = trt.recovery
#      )
  # this is not how to use survival::plot.coxph lolol (cph doesnt give u a survival curve per se. its a regression)

autoplot(survfit(coxph(survObj$fromReturn ~ trt.recover * trt.duration, 
                                             data = km$subsets$expts)),
         facets = TRUE,
         )
```

```{r eval=FALSE}
# per google:
# check if prop hazards assumption is met (is HR betwn groups constant over time?)
  # https://stats.stackexchange.com/questions/560975/how-to-interpret-schoenfield-residual-plot
ggcoxzph(cox.zph(fits$cph.enter), ggtheme = theme_minimal())
ggcoxzph(cox.zph(fits$cph.recovery), ggtheme = theme_minimal())
ggcoxzph(cox.zph(fits$cph.return), ggtheme = theme_minimal())

# TODO idk what im doing here so... lets revisit later
```

```{r}
fit <- survfit(Surv(kmdf$exit.fromenter, kmdf$stat) ~ recover + duration, data = kmdf) 
# ok survfit super doesnt like "subsets" being in the df name lol

# kmdf_summ <- kmdf %>%
#   group_by(trt)
#   summarise()

ggsurvplot(fit, 
           data = kmdf,
           conf.int = TRUE,
           #xlim = c(45823, 45841),
           #facet.by = "trt.duration",
           facet.by = "duration",
           #linetype = "strata",
           palette = "strata") +
  
  # i think the stuff for vline needs to come from a separate df LOL...
  # also, lowkey not sure how this works considering the time depends on the hatch day... (avg them??)
  
  geom_vline(xintercept = kmdf$dh.recover - kmdf$dh.exit, #tried w dh.enter, kinda wonky
             #color = "red"
             #aes(xintercept = kmdf$dh.recover - kmdf$dh.enter, color = kmdf$recover)
             ) +
  geom_vline(xintercept = kmdf$return - kmdf$df.exit,
             color = "blue")

# TODO idk gotta think about wtf im plotting for this one a bit... draw it out
  
```








# initial attempts

## km curves

```{r eval=FALSE, include=FALSE}
# fit all
fit <- survfit(Surv(dh.exit, km.status) ~ trt.enter + trt.recover + trt.duration, data = km$data)
fit <- survfit(Surv(dh.exit, km.status) ~ trt, data = km$data)

           # linetype = c("1", "1", "1", "1",
           #              "2", "2", "2",
           #              "3", "3", "3"),
           palette = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")
           )  


  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_linetype_manual(values = c("1", "1", "1", "1",
                                   "2", "2", "2",
                                   "3", "3", "3"))




ggsurvplot(fit, data = km$data,
           #conf.int = TRUE,
           xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #linetype = "strata",
           #palette = "strata")
           )

+ 
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_linetype_manual(values = c("1", "1", "1", "1",
                                   "2", "2", "2",
                                   "3", "3", "3"))



fit <- survfit(Surv(dh.exit, km.status) ~ trt.enter + trt.recover + trt.duration, data = km$data)
#fit <- survfit(coxph(Surv(km.exit, km.status) ~ trt, data = km.data))


ggsurvplot(fit, data = km$data, 
           #conf.int = TRUE,
           xlim = c(45823, 45839),
           facet.by = "trt.duration",
           #group.by = c("trt.recover", "trt.duration"),
           #palette = "trt.recover",
           palette = "strata"
           #linetype = "trt.duration",
           ) +
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", 
                                "deepskyblue", "skyblue1", "goldenrod1", "darkorange1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", 
                                "deepskyblue", "skyblue1", "goldenrod1", "darkorange1"))

  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1"),
                     labels = c(19, 26, 33, 40)) +
  scale_linetype_manual(#values = c("0" = 1, "24" = 2, "48" = 3),
                        values = c(1, 2, 3),
                        labels = c("0", "24", "48"))
  
  
  
  
  
# colorRampPalette(c(startcol, endcol))(ngroups)
  
# autoplot(fit, data)+ scale_color_manual, full
  
  
  
  
  
ggsurvplot(fit, data = km$data,
           #conf.int = TRUE,
           xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #palette = viridis_pal(option = "H", direction = -1)(10)
           palette = hcl.colors(10, palette = "Roma")
           #linetype = "strata",
           #palette = "strata")
           ) +
  labs(caption = "2025-06-18 15:57 data; starting from hatch")
```

```{r eval=FALSE, include=FALSE}
# try subsets
subset <- km$data %>% filter(,
                             #trt > 100 & trt < 200,
                             is.40 == "Y",
                             #is.ctrl == "Y"
                             #trt %in% c(133, 233)
                             )

fit <- survfit(Surv(dh.exit, km.status) ~ trt + trt.duration + trt.recover, data = subset)
#fit <- survfit(Surv(km.exit, km.status) ~ trt.recover + trt.duration, data = subset)


#summary(fit)
#autoplot(fit)

ggsurvplot(fit, data = subset, 
           conf.int = TRUE,
           xlim = c(45823, 45839),
           #facet.by = "trt.recover",
           group.by = c("trt.recover", "trt.duration"),
           #palette = "trt.recover",
           #linetype = "trt.duration",
           ) +
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1"),
                     labels = c(19, 26, 33, 40)) +
  scale_linetype_manual(#values = c("0" = 1, "24" = 2, "48" = 3),
                        values = c(1, 2, 3),
                        labels = c("0", "24", "48"))
  
  
  # scale_color_manual(values = c(c("19", "119", "219") = "deepskyblue", 
  #                               c("26", "126", "226") = "skyblue1",
  #                               c("33", "133", "233") = "goldenrod1",
  #                               "40" = "darkorange1")) +
  # scale_linetype_manual(values = c(c("19", "26", "33", "40") = 1,
  #                                  c("119", "126", "133") = 2,
  #                                  c("219", "226", "233") = 3))

  # scale_linetype_manual(values = c("19" = 1, "26" = 1, "33" = 1, "40" = 1,
  #                                  "119" = 2, "126" = 2, "133" = 2,
  #                                  "219" = 3, "226" = 3, "233" = 3))
```

```{r eval=FALSE, include=FALSE}
fit <- survfit(coxph(Surv(dh.exit, km.status) ~ trt, data = subset))
#library(ggplot2)
autoplot(fit)

```





