---
title: "2025 TDT data cleaning"
date: "2025-06-14"
---

# 0. roadmap

1. load data & packages; pre-filtering
2. column formatting + rearrangement
3. save output into `2025/data/`

re-run script whenever the gsheets gets updated


# 1. load data & packages

```{r message = FALSE}

library(tidyr)
library(purrr)
library(dplyr)
library(lubridate) # for handling time stuff
library(googlesheets4)

here::i_am("2025/cleaning-tdt.Rmd")
library(here)

source(here::here("set-paths.R"))
```

googlesheets import

```{r}
# need to correct dates (add 2025)
raw <- read_sheet("https://docs.google.com/spreadsheets/d/1rxZSI-8ubhMkeM2Sh1xNi5IKF0Vt_BDFItAg_Ak476M/edit?gid=1826657546#gid=1826657546", 
                  sheet = "jun 2025 pilot",
                  col_types = "c",
                  na = c("#VALUE!")) %>%
  drop_na() %>%
  na_if("--") %>% na_if("")

data <- raw %>% 
  rename(trt = treatment) %>%
  mutate(trt.recover = case_when(trt.type == "ctrl" ~ NA_character_,
                                 TRUE ~ trt.recover)) %>%
  select(-c(trt.type, starts_with("dv"))) %>%
  mutate(across(starts_with("date."), ~ stringr::str_c("2025/", .)))
```

# 2. content fixing and standardising

filtering, value fixing

```{r}
# pull out accidentally culled before 3rd (fate = 2, instar.exit < 3)
data <- data %>%
  filter(!(fate == 2 & instar.exit < 3) | is.na(fate)) # retain is.na(fate) for things still developing

data[, 'notes'] <- lapply(data[, 'notes'], gsub, pattern=",", replacement = "&") 
  # replaces commas w/ &'s to avoid csv breakage after export

```

column formatting, date-time fixing

```{r}
data <- data %>%
    mutate(across(c(trt, trt.enter, trt.recover, trt.duration, id, instar.exit, fate,
                    #starts_with("mass.", "dv."))
                    starts_with("mass")), as.numeric))
    #mutate(across(starts_with("date."), as.Date, format = "%m/%d")) #%>%
    #mutate(across(starts_with("time."), hm)) # breaks parsing bc converts to "HH MM SS"

# add '2025' to yrs bc i didnt do it in gsheets, convert to D-T objects, calculate the actual decimal hrs
# bc gsheets does it differently from R
data <- data %>%
  #mutate(across(starts_with("date."), ~ stringr::str_c("2025/", .))) %>%
  mutate(dh.hatch = as_datetime(paste(date.hatch, time.hatch), format = "%Y/%m/%d %H:%M"),
         dh.enter = as_datetime(paste(date.enter, time.enter), format = "%Y/%m/%d %H:%M"),
         dh.recover = as_datetime(paste(date.recover, time.recover), format = "%Y/%m/%d %H:%M"),
         dh.return = as_datetime(paste(date.return, time.return), format = "%Y/%m/%d %H:%M"),
         dh.exit = case_when(!is.na(date.culled) ~ as_datetime(paste(date.culled, time.died), format = "%Y/%m/%d %H:%M"),
                             !is.na(date.died) ~ as_datetime(paste(date.died, time.died), format = "%Y/%m/%d %H:%M"),)
         ) %>%
  mutate(across(starts_with("dh."), ~ as.numeric(as.Date(., format = "%j") + hour(.) + minute(.)/60)))

# TODO might need to keep "expt type" to separate out 40 ctrl from other "ctrls" ?

```

```{r eval=FALSE, include=FALSE}
## TODO archive later, but refer back to for date-time futzing lol

# fix dv.<stage> bc gsheets doesnt do it the same as R lol --> convert to numeric
  # as.Date = date only
data %>%
  mutate(dv.test = as_datetime(paste(date.hatch, time.hatch), 
                               format = "%Y-%m-%d %H:%M"),
         time.test2 = hour(dv.test),
         time.test = hour(as_datetime(time.hatch, format = "%H:%M")),
         j.test = as.numeric(as.Date(dv.test, format = "%j") + 
                               as.numeric(as.Date(time.hatch, format = "%H:%M"))/60),
         # j.test2 = as.numeric(as.Date(dv.test, format = "%j") + 
         #                       as.numeric(hour(time.hatch)) +
         #                        as.numeric(minute(time.hatch)/60)),
         .before = 1)


data %>% # combine the <stages> and then convert
  mutate(across(starts_with("date"), as.Date, format = "%m/%d"),
         across(starts_with("time"), as_datetime, format = "%H:%M"),
         .before = 1)


data %>%
  mutate(across(starts_with("date."), ~ stringr::str_c("2025/", .))) %>%
  mutate(dt.hatch = as_datetime(paste(date.hatch, time.hatch), format = "%Y/%m/%d %H:%M"),
         dt.enter = as_datetime(paste(date.enter, time.enter), format = "%Y/%m/%d %H:%M"),
         dt.recover = as_datetime(paste(date.return, time.recover), format = "%Y/%m/%d %H:%M"),
         dt.return = as_datetime(paste(date.return, time.return), format = "%Y/%m/%d %H:%M"),
         dt.died = case_when(!is.na(date.culled) ~ as_datetime(paste(date.culled, time.died), format = "%Y/%m/%d %H:%M"),
                             !is.na(date.died) ~ as_datetime(paste(date.died, time.died), format = "%Y/%m/%d %H:%M"),),
         )


```


```{r}
# other filters

# data <- data %>%
#   dplyr::filter(!(notes.ignore %in% c("early pmd", "lost"))) # dropping things that died in 1 day or got lost
# 
# data <- select(data, -c("date.LPI", "date.15")) # not rly needed for analyses/LPI is empty lol


# value fixing
# data[, -ncol(data)] <- lapply(data[, -ncol(data)], gsub, pattern="\\?", replacement = "") 
#   # removes `?`s from all columns excepts notes column (assuming its the last col in the df)


# data[, 'id'] <- lapply(data[, 'id'], gsub, pattern="\\.\\d", replacement = "") 
#   # turns the repeat entries into whole numbers so theres no decimals going on after as.numeric(id)
```

julian date conversions

```{r}
data <- data %>%
  mutate(across(starts_with("date."), as.Date, "%Y/%m/%d")) %>%
  mutate(across(starts_with("date."), as.Date, "%j", .names = "j{.col}")) %>%
  mutate(across(starts_with("jdate."), as.numeric)) %>%
  mutate(jdate.exit = case_when(!is.na(jdate.culled) ~ jdate.culled,
                                !is.na(jdate.died) ~ jdate.died)) %>%
  select(-c(jdate.died, jdate.culled))
```


# 3. saving output & tidying up

```{r}
# setup
today <- format(Sys.time(), "%y%m%d-%H%M")

#data <- mutate(data, across(starts_with("time."), as.character))
  # breaks write_csv if left as a `lubridate` object
```

```{r}
# export
# TODO convert these to here() later
write.csv(raw, file = paste0("~/Documents/repos/ntw/2025/data/archive/", today, "_raw.csv"), row.names = FALSE)
write.csv(data, file = paste0("~/Documents/repos/ntw/2025/data/archive/", today, "_clean-full.csv"), row.names = FALSE)

# extract the useful stuff
data <- data %>%
  select(c("cohort", "shelf", starts_with("trt"), "id", starts_with(c("jdate", "mass", "dh")), "instar.exit", "fate"))

write.csv(data, file = paste0(here(bin_paths$y25$data), "/clean-working.csv"), row.names = FALSE)
```

remove troubleshooting objects

```{r}
rm(today, data, raw)
```



