---
title: "ntw_data_cleaning"
date: "2023-11-07"
---

# roadmap

1. load data & packages
2. cleaning: prefiltering, fix columns, add columns
3. save output

# 1. load data & packages
```{r message = FALSE}
library(conflicted)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(tidyr)
library(purrr)

# to load data from gsheets
library(googlesheets4)
```

## - data import from gsheets

note: 2024-07-02 updated gsheet with the field diapausers(?); should be the final update. as of now, has yet to be incorporated into cleaned data used for analyses (should look into "tossing" criteria before deleting this note)

```{r}
# before importing,
  # run cleaning macros (larva sheets)
  # check column/row alignment + proper sorting (adult sheets)

# larva data
data_field <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet = "june field", col_types = "c")
data_labsu <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="june 2023 lab", col_types = "c")
data_labF1 <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="lab F1", col_types = "c")
data_labsp <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="consolidated + tidyed", col_types = "c")
```

## - consolidation

```{r}
# function to standardise column format
fix.columns <- function(data) {
  cleaned_data <- data %>%
    mutate(across(starts_with("date."),  as.Date, format = "%m/%d/%y")) %>%
    mutate(mutate(across(starts_with("mass."), as.numeric))
    )
  
  return(cleaned_data)
}

# add IDs to imported data sheets and merge
data_field <- mutate(data_field, src = "field")
data_labF1 <- mutate(data_labF1, src = "F1")
data_labsp <- mutate(data_labsp, src = "sp")
data_labsu <- mutate(data_labsu, src = "su")

data_all <- data.table::rbindlist(lapply(list(data_field, data_labF1, data_labsp, data_labsu), fix.columns), fill = T)

check_raw <- data_all #for troubleshooting lol
```


# 2. clean data

## - prefiltering
```{r}
# filter out individuals that didnt hatch at 26C / ctrl temps
data_all <- data_all %>%
    filter(!(src == "sp" & (from.E == 2 | (from.E == 0 & (treatment == 337 | treatment == 330 | treatment == 267)))))

# filter random deaths (got squished/empty)
data_all <- subset(data_all, !(ID %in% c(1002, 1007, 1017)))
#data_all <- data_all %>% filter(is.na(reason.ignore))
```

## - column/value formatting, other fixes
```{r}
# julian date formatting
data_all <- data_all %>%
  mutate(across(starts_with("date."), as.Date, "%j", .names = "j{.col}")) %>%
  mutate(across(starts_with("jdate."), as.numeric))

# value standardising
data_all$treatment[data_all$treatment=="426"] <- "337" # match for consistency
data_all$instar.enter[data_all$trt.stage == "1st"] <- "hatch" # change field 1sts to hatchlings
data_all$final.fate[data_all$final.fate == "LP"] <- "LPI"

# add instar.enter levels
data_all$instar.enter <- factor(data_all$instar.enter, levels = c("hatch", "1st", "3rd", "4th"))

# add in late dead pupa
data_all <- data_all %>%
  mutate(final.fate = case_when(!is.na(date.exit) & is.na(final.fate) ~ "pupa",
                                TRUE ~ as.character(final.fate)))
```

## - column creation for stats
```{r}
# survival/development binaries
data_all <- data_all %>% mutate(
  if.stuck = case_when(is.na(date.stuck) ~ "N", TRUE ~ "Y"),
  if.pupa = case_when(is.na(date.LP) & is.na(date.pupa) ~ "N", TRUE ~ "Y"),
  if.sup = case_when(is.na(date.6th) ~ "N", TRUE ~ "Y"),
  sup = case_when(!is.na(date.7th) ~ 7, !is.na(date.6th) ~ 6, TRUE ~ 0) # type of sup
  )

# subdivide treatments
data_all <- data_all %>%
  mutate(treatment = as.character(treatment),
         trt.stage = as.character(paste(treatment,instar.enter, sep = "-")), # treatment @ instar
         grp.trt = paste(expt.group, trt.stage, sep = "-")) # cohort & treatment

# mean/fluc Ts; correction for diurnal trts
data_all <- data_all %>%
  mutate(meanT = case_when(treatment == 260 | treatment == 267 | treatment == "ctrl" ~ 26,
                           treatment == 337 | treatment == 330 ~ 33,
                           treatment == 433 ~ 36.5,
                           treatment == 419 ~ 29.5),
         flucT = case_when(treatment == 260 | treatment == 330 ~ 0,
                           treatment == 267 | treatment == 337 ~ 7,
                           treatment == 433 ~ 3.5,
                           treatment == 419 ~ 10.5),
         maxT = case_when(treatment == 260 | treatment == "ctrl" ~ 26,
                          treatment == 267 | treatment == 330 ~ 33,
                          treatment > 331 ~ 40),
         minT = case_when(treatment == 260 | treatment == "ctrl" ~ 26,
                          treatment == 267 | treatment == 419 ~ 19,
                          treatment == 337 ~ 26,
                          treatment == 433 | treatment == 330 ~ 33)) %>%
  mutate(flucT = case_when((treatment == 260 & 
                              (expt.group == "C" | expt.group == "D" | expt.group == "E" | expt.group == "F" | expt.group == "G" | expt.group == "H")) ~ 2.5,
                           TRUE ~ as.numeric(flucT)))
```

## - column creation for notes
```{r}
# indicate field/lab pops, diet, parents
data_all <- data_all %>%
  mutate(pop = case_when(location == "CC" ~ "field",
                         src == "F1" ~ "F1",
                         TRUE ~ "lab"),
         diet = case_when((expt.group == "F" | pop == "field") ~ "TB",
                          TRUE ~ "LD"),
         parent.tent = case_when(src == "F1" ~ as.character(parent.tent),
                                 TRUE ~ "parent"))

# fill in reason.ignore (optional reasons to ignore individs later when graphing/stats/etc)
  # indicate overly long-lived individuals: things that took > 25 days to die
data_all <- data_all %>% 
  mutate(reason.ignore = case_when(jdate.pmd-jdate.hatch > 25 ~ "slow larva",
                                   #jdate.eclose - jdate.pupa > 25 ~ "diapaused?",
                                   src == "F1" & ID < 3036 ~ "early F1",
                                   expt.group == "E" & treatment == "337" & ID < 247 ~ "early field",
                                   #final.fate == "culled" & instar.culled == "pupa" ~ "winter culling",
                                   final.fate == "culled" & instar.culled != "pupa" ~ "culled larva",
                                   reason.pmd == "toast" ~ "hot larva",
                                   TRUE ~ as.character(reason.ignore)))

# check: table(data_all$reason.ignore)

check_newcols <- data_all
```

# 3. save the output

```{r}
today <- format(Sys.time(), "%y%m%d")

# MQs only
mqs <- data_all %>% filter(species == "MQ")

write.csv(mqs, file = "~/Documents/repos/ntw/2023/data/clean-MQs.csv", row.names = FALSE)
write.csv(mqs, file = paste0("~/Documents/repos/ntw/2023/data/archive/", today, "_clean-MQs.csv" ), row.names = FALSE)


# wordier notes
notes <- data_all %>% select(c("treatment", "ID", "expt.group", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "mass.died", "pupa.deformities", "notes", "reason.pmd", "reason.ignore", ends_with(".stuck"), ends_with(".culled")))

write.csv(notes, file = "~/Documents/repos/ntw/2023/data/notes_ntw.csv", row.names = FALSE)
write.csv(notes, file = paste0("~/Documents/repos/ntw/2023/data/archive/", today, "_notes_ntw.csv" ), row.names = FALSE)
  
# cleaned up sexta
check_final <- data_all %>%
  filter(is.na(species)) %>% 
  select(-c("species", "from.E", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "pupa.deformities", "mass.died", "toss.if", "old.date.pmd", "reason.pmd", "fate.code", "notes", "mass.stuck", "date.stuck", ends_with(".culled"), "eclose-3"))

write.csv(check_final, file = "~/Documents/repos/ntw/2023/data/clean-ntw.csv", row.names = FALSE)
write.csv(check_final, file = paste0("~/Documents/repos/ntw/2023/data/archive/", today, "_clean_ntw.csv" ), row.names = FALSE)
```

## - troubleshooting

```{r}
# data_all <- check_raw
# data_all <- check_newcols
# data_all <- check_final
```

```{r}
# if all goes well!
rm(check_raw, check_newcols)
rm(today, mqs, notes, check_final, data_all)
rm(data_field, data_labF1, data_labsp, data_labsu, fix.columns)
```
