---
title: "entsoc 2024 figs and analyses"
date: "2024-10-22"
---

# 0. roadmap

goal: making figs/running stats for entsoc poster, focusing on lab bugs

# 1. load data & packages

```{r}
source("./helpers_entsoc.R")

source("../2023/helpers_tents.R")

dtents <- data_tstats 
#%>% select(-c("trt.type", "trt.pair"))

dad <- data_longevity

rm(data_tstats, data_hatch, data_longevity,
   labels.alltrts, labels.exptrts, RYB,
   x_err_ncoll, x_err_ncollf, y_err_hrate)
```

convenience functions

```{r}
# calc std err
se <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
```



value/columns renaming and filtering

```{r}
# adult data
dad <- dad %>%
  rename(jdate.eclose = jdate.ec,
         jdate.surv = jdate.died) %>%
  mutate(mass.eclose = mass.eclose/1000,
         for.surv = case_when(track.reason %in% c("surv-fert", "surv") ~ 1),
         for.fert = case_when(track.reason %in% c("surv-fert", "fert") ~ 1),
         trt.type = case_when(maxT == minT ~ "ctrl",
                              TRUE ~ "expt")) %>%
  select(-c(track.reason, starts_with("date")))


# tent data
dtents <- dtents %>%
  filter(pop == "lab") %>%
  #rename(trt.typeold = trt.type) %>%
  mutate(trt.mateto = case_when(trt.m != trt.f | trt.m + trt.f == 520 ~ "ctrl",
                                TRUE ~ "hs"),
         trt.sex = case_when(trt.m > trt.f ~ "m",
                             trt.f > trt.m ~ "f",
                             trt.m == trt.f ~ "both"),
         trt.hs = case_when(trt.m > trt.f ~ as.numeric(trt.m),
                            trt.f > trt.m ~ as.numeric(trt.f),
                            trt.m + trt.f > 600 ~ as.numeric(trt.m),
                            TRUE ~ 260),
         trt.isctrl = case_when(trt.m + trt.f == 520 ~ 1,
                              TRUE ~ 0),
         trt.ref = case_when(trt.m + trt.f == 520 ~ "ctrl",
                             trt.m != trt.f ~ "btwn",
                             trt.m == trt.f ~ "win")
         ) %>%
  select(-pop)


# change mass to g
d23 <- d23 %>%
  mutate(mass.pupa = mass.pupa/1000,
         mass.eclose = mass.eclose/1000)
```


# 2. adult, larval dev

## a. math

```{r}
# larval stats
ss_la <- d23 %>%
  group_by(trt.type, minT, 
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n.all - n.pmd, 
            prop.survpup = round(1-(n.pmd/n.all), digits=2)
            ) #%>% mutate_all(~ifelse(is.nan(.), NA, .))

# adult stats
ss_ad <- dad %>%
  filter(!is.na(for.surv), pop == "lab") %>%
  group_by(#diet, # gets too parsed out w/ diet lol
           trt.type, minT, sex
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.surv - jdate.eclose)),
            se.tt = sd(na.omit(jdate.surv - jdate.eclose))/sqrt(length(na.omit(jdate.surv - jdate.eclose))),
            avg.mass = mean(na.omit(mass.eclose)),
            se.mass = sd(na.omit(mass.eclose)/sqrt(length(na.omit(mass.eclose)))),
            n = sum(na.omit(!is.na(jdate.eclose)))
            )
```

# 3. tent stats

## a. pre-math (fixing #s)

these calcs are copy/pasted/tweaked from `/2023/analyses_tents.Rmd`

```{r}
# determine actual # of ovipositing females
dtents <- dtents %>%
  mutate_at(c(6:14), ~replace_na(., 0)) %>% 
  group_by(id.tent) %>%
  mutate(jdate.rel = jdate - first(jdate)) %>%
  ungroup() %>%
  group_by(id.tent, trt.f, trt.m, trt.sex, jdate.rel) %>%
  arrange(jdate.rel, .by_group = TRUE) %>%
  mutate(#n.laying = n.curr.f + n.died.f, .after = "n.curr.f",
         n.ovi = dplyr::lag(n.curr.f, default = first(n.curr.f)) + n.died.f, .after = "n.curr.f") %>%
  ungroup()


# where n.ovipositing females = 0 but n.collected > 0, add the # eggs collected to the day before
# (or last date of ovipositing females)
# first, figure out the rows
cond <- (dtents$n.coll > 0 & dtents$n.ovi == 0)

ovifix <- dtents %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

negovi <- ovifix[ovifix$females.laying < 0, ]

lastovi <- ovifix[ovifix$n.curr.f > 0 & ovifix$n.ovi >= 1, ]

test2 <- c(negovi$index, lastovi$index)

# then, edit values
#summ_rate$n.coll[negovi$index] <- 0
dtents <- dtents %>% 
  mutate(n.coll = case_when(rownames(.) %in% negovi$index ~ 0,
                            rownames(.) %in% lastovi$index ~ n.coll + 1,
                            !(rownames(.) %in% c(negovi$index, lastovi$index)) ~ as.numeric(n.coll)
                            )) %>%
  select(-females.laying)

rm(cond, ovifix, negovi, lastovi)

# removing extra 0's
# dtents <- dtents %>%
#   filter(!(n.coll == 0 & n.ovi == 0))
  # tbh not sure why i did this in the old version. this drops a few #s here and there

```

## b. fecundity rates (# laid)

ref: `summ_rates_all` calcs in `analyses_tents.Rmd`

```{r}
# this calcs how # of tents, # eggs laid, # ovi f changes daily per trt
ss_ovirates <- dtents %>%
  group_by(trt.ref, trt.sex, trt.hs, jdate.rel) %>%
  summarise(n.tents = n(), #idk why i use mutate in the orig lol
         eggs.perovi = sum(n.coll)/sum(n.ovi),
         eggsovi.pertent = eggs.perovi / n.tents) %>%
  mutate_all(~ifelse(is.nan(.), NA, .))

```


## c. fertility rates (# hatched)

ref: `summ_hatch` calcs in `analyses_tents.Rmd`

```{r}
ss_hatchrates <- dtents %>%
  group_by(trt.ref, trt.sex, trt.hs) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.laid = sum(n.coll, na.rm = TRUE),
            se.laid = se(n.coll),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            se.hatched = se(n.tothatch),
            n.fs = sum(n.new.f, na.rm = TRUE),
            prop.hatched = n.hatched/n.laid,
            se.hatchprop = sqrt(prop.hatched*(1-prop.hatched)/n.laid),
            prop.hatchf = prop.hatched/n.fs,
            se.hatchfprop = sqrt(prop.hatchf*(1-prop.hatchf)/n.laid)
            )
```


