---
title: "2024 dev analyses"
date: "2024-07-16"
---

# roadmap & loading

## - **questions**

- (pilot) what is the average time spent in each instar after transferring to constant temps?
  - where do bugs "line up" in their dev at 4th and 5th?
- does getting stuck make a difference on survival?
- are the avg growth outcomes different if you keep the survivors vs not?
- (ntw) how 2024 ctrls compare to 2023 ctrls?


```{r message = FALSE}
source("./helpers_ox.R") # need run cleaning.Rmd beforehand if gsheet has been updated

# INCASE: bring this back if u need it and are breaking things
  # library(tidyverse)
  # conflicted::conflicts_prefer(dplyr::filter)
  # all_wide <- read.csv("~/Documents/repos/ntw/2024/data/ox.csv", header = TRUE)
```

# july pilot 

## = time to dev stages

2024-09-12: taking out the 6ths bc there's just not rly enough of them to do anything interesting w/ and they mess things up lol

### - wrangling

```{r}
# doing this math separately from the helpers stuff bc this is more specific to the pilot lol

# relative to time entering trt
trttime_wide <-  wide_all %>%
  mutate(labs.trt = factor(labs.trt, c("26±7°C", "26±0°C", "33±0°C", "36±0°C", "38±0°C"))) %>%
  filter(sup == 0 | is.na(sup)) %>%
  mutate(t3.3rd = jdate.3rd - jdate.3rd,
         t3.4th = jdate.4th - jdate.3rd,
         t3.5th = jdate.5th - jdate.3rd,
         t3.6th = jdate.6th - jdate.3rd,
         t3.wander = jdate.wander - jdate.3rd,
         t3.pmd = jdate.pmd - jdate.3rd,
         tt.intrt = case_when(is.na(jdate.pmd) ~ jdate.pupa - jdate.3rd,
                              TRUE ~ jdate.pmd - jdate.3rd)
         )

# stats of major dev pts (long)
trttime_long <- trttime_wide %>%
  select(-(contains(c("4d", "5d", "6d")))) %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "t3")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
  filter(!is.na(instar) & !(instar %in% c("2nd", "15", "culled", "")))
```


### - stats

2024-09-12: need to take handle the culled somehow

```{r}
# includes pmd and all trts
trttime_lstats <- trttime_long %>%
  #filter(is.pmd == 0) %>%
  filter(instar %in% c("3rd", "4th", "5th", "6th", "wander")) %>%
  group_by(labs.trt, instar) %>% # can add extra filters...
  mutate(t3h = t3*24,
         fate.pmd = case_when(fate == "pmd" ~ 1, TRUE ~ 0),
         fate.culled = case_when(grepl("culled", fate) ~ 1, TRUE ~ 0),
         #fate.surv = case_when(fate %in% c("pupa", "eclose") ~ 1, TRUE ~ 0)
         ) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.t3 = mean(na.omit(t3)),
            se.t3 = sd(na.omit(t3))/sqrt(length(na.omit(t3))),
            avg.t3h = mean(na.omit(t3h)),
            se.t3h = sd(na.omit(t3h))/sqrt(length(na.omit(t3h))),
            n=n(), # includes culled #s
            n.pmd = sum(fate.pmd),
            n.culled = sum(fate.culled),
            pct.surv = round((n - n.culled - n.pmd)/(n - n.culled) * 100, digits = 1), # excludes culled from pctg
            ) %>%
  #mutate(labs.surv = str_replace(labs.surv, " \\(.0*\\.0\\)", "")) %>%
  mutate(pct.surv = case_when(pct.surv == 0 | pct.surv == 100 ~ NA_character_,
                              TRUE ~ sprintf("%.1f", pct.surv)),
         labs.surv = case_when(is.na(pct.surv) ~ as.character(n),
                               TRUE ~ paste0(n, "\n(", pct.surv, ")")))
# oops math
  # mutate(n.culled = dplyr::lag(n.culled, default = 0),
  #        pct.surv =  round(n - n.pmd/(n - n.culled) * 100, digits = 1))
```

### - plot

TODO: 
- [x] reorder trts?
- how to handle the culled bugs

```{r}
## major dev points


# by hours
trttime_lstats %>%
  #filter(labs.trt != "38±0°C") %>%
  ggplot(aes(y = labs.trt, x = avg.t3h, shape = instar, color = labs.trt)) +
  geom_point(size = 2) +
  geom_line(aes(group = labs.trt, lty = labs.trt), lwd = 0.8, show.legend = FALSE) +
  geom_errorbarh(aes(xmin = avg.t3h - se.t3h, xmax = avg.t3h + se.t3h), height = 0.15) +
  geom_text(aes(label = labs.surv,
                  #paste0(n, "\n(", pct.surv, ")"), vjust = -0.5
                  vjust = -0.95
                ),
            size = 3.5, lineheight = 0.9,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(minor_breaks = seq(0, 250, by = 24),
                     breaks = seq(0, 400, by = 48)) +
  theme_bw() +
  labs(title = "avg instar times (hrs)",
       y = "treatment", x = "hrs since 3rd",
       color = "treatment",
       caption = "n (% surv)")

# by days
trttime_lstats %>%
  #filter(labs.trt != "38±0°C") %>%
  ggplot(aes(y = labs.trt, x = avg.t3, shape = instar, color = labs.trt)) +
  geom_point(size = 2) +
  geom_line(aes(group = labs.trt, lty = labs.trt), lwd = 0.8, show.legend = FALSE) +
  geom_errorbarh(aes(xmin = avg.t3 - se.t3, xmax = avg.t3 + se.t3), height = 0.15) +
  geom_text(aes(label = labs.surv, vjust = -0.9),
            size = 3.5, lineheight = 0.9,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(0, 18, by = 2)) +
  theme_bw() +
  labs(title = "avg instar times (days)",
       y = "treatment", x = "days since 3rd",
       color = "treatment",
       caption = "n (% surv)")
```

## = ctrl: dev stages by cohort

wrangle (recalc stats by cohort)

```{r}
trttime_ctrlstats <- trttime_long %>%
  filter(!(labs.trt %in% c("36±0°C", "38±0°C"))) %>%
  filter(instar %in% c("3rd", "4th", "5th", "6th", "wander")) %>%
  group_by(cohort, labs.trt, instar) %>%
  mutate(t3h = t3*24,
         fate.pmd = case_when(fate == "pmd" ~ 1, TRUE ~ 0),
         fate.culled = case_when(grepl("culled", fate) ~ 1, TRUE ~ 0),
         #fate.surv = case_when(fate %in% c("pupa", "eclose") ~ 1, TRUE ~ 0)
         ) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.t3 = mean(na.omit(t3)),
            se.t3 = sd(na.omit(t3))/sqrt(length(na.omit(t3))),
            avg.t3h = mean(na.omit(t3h)),
            se.t3h = sd(na.omit(t3h))/sqrt(length(na.omit(t3h))),
            n=n(), # includes culled #s
            n.pmd = sum(fate.pmd),
            n.culled = sum(fate.culled),
            pct.surv = round((n - n.culled - n.pmd)/(n - n.culled) * 100, digits = 1), # excludes culled from pctg
            ) %>%
  #mutate(labs.surv = str_replace(labs.surv, " \\(.0*\\.0\\)", "")) %>%
  mutate(pct.surv = case_when(pct.surv == 0 | pct.surv == 100 ~ NA_character_,
                              TRUE ~ sprintf("%.1f", pct.surv)),
         labs.surv = case_when(is.na(pct.surv) ~ as.character(n),
                               TRUE ~ paste0(n, "\n(", pct.surv, ")")))
# oops math
  # mutate(n.culled = dplyr::lag(n.culled, default = 0),
  #        pct.surv =  round(n - n.pmd/(n - n.culled) * 100, digits = 1))
```

plot

```{r}
trttime_ctrlstats %>%
  ggplot(aes(y = cohort, x = avg.t3, shape = instar, color = labs.trt)) +
  geom_point(size = 2) +
  geom_line(aes(group = cohort, lty = cohort), lwd = 0.8, show.legend = FALSE) +
  facet_wrap(~labs.trt) +
  geom_errorbarh(aes(xmin = avg.t3 - se.t3, xmax = avg.t3 + se.t3), height = 0.15) +
  geom_text(aes(label = labs.surv, vjust = -0.9),
            size = 3.5, lineheight = 0.9,
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(0, 18, by = 2)) +
  theme_bw() +
  labs(title = "avg instar times (days)",
       y = "cohort", x = "days since 3rd",
       color = "treatment",
       caption = "n (% surv); ctrl trts")
```



## = fine-scale development

want to see how mass changes day-by-day in the later instars up to wandering

### - stats

```{r}
fine_stats <- long_all %>% # todo: 2024-09-22: theres some unexpected NAs in here?
  filter(instar %in% c("3rd", "4th", "5th", "6th", "wander")) %>%
  group_by(labs.trt, #instar, # does nawt work w instar lol
           t3) %>%
  mutate(logmass = log(na.omit(mass))) %>% #filter(t3 == 0) %>% View
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n()) %>%
  filter(n > 1) # drops the lonely things bc they fudge the graph up


majortimes_stats <- trttime_long %>%
  filter(instar %in% c("4th", "5th", "6th", "wander")) %>%
  group_by(labs.trt, 
           instar) %>%
  summarise(avg.t3 = mean(na.omit(t3)),
            se.t3 = sd(na.omit(t3)/sqrt(length(na.omit(t3)))),
            max.t3 = avg.t3 + se.t3, 
            min.t3 = avg.t3 - se.t3)
```

### - plot

TODO:

- need to split the 6ths from the rest bc they have different dev trajectories (everything is fine if u omit it tho 👍)

```{r}
# faceted ver: run majortime_stats grouped by labs.trt
ggplot(fine_stats, aes(y = avg.logmass, x = t3, color = labs.trt, #shape = instar
                       )) +
  geom_point() + geom_line(lwd = 0.8) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass,
                    width = 0.5)) +
  # geom_line(data = long_all, 
  #           aes(y = log(mass), group = interaction(id, labs.trt)),
  #           alpha = 0.25) +
  geom_vline(majortimes_stats, mapping = aes(xintercept = avg.t3, lty = instar, color = labs.trt), # meh could be better
             #show.legend = FALSE
             ) + 
  facet_grid(labs.trt~.) +
  theme_bw() +
  scale_color_manual(values = c("#2166ac", "#92c5de", "#f4a582", "#d6604d", "#b2182b")) +
  labs(title = "avg daily growth after entering trt",
       x = "days since 3rd", y = "avg log(mass (mg))",  
       color = "treatment")




# combined: run majortime_stats w/o grouping by labs.trt
# for rectangle fills, see:
  # https://stackoverflow.com/questions/71733503/background-bars-in-ggplot2-using-geom-rect
  # https://stackoverflow.com/questions/62590533/how-to-shade-an-area-in-ggplot-mixing-geom-rect-and-geom-line

ggplot(fine_stats, aes(y = avg.logmass, x = t3, color = labs.trt, #shape = instar
                       )) +
  geom_point() + geom_line(lwd = 0.8) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass,
                    width = 0.5)) +
  # geom_line(data = long_all, 
  #           aes(y = log(mass), group = interaction(id, labs.trt)),
  #           alpha = 0.25) +
  # geom_ribbon(data = majortimes_stats, mapping = aes(y = 0, x = avg.t3, xmin = avg.t3 - se.t3, xmax = avg.t3 + se.t3,
  #                                             lty = instar), # meh could be better
  #             inherit.aes = FALSE
  #            #show.legend = FALSE
  #            ) + 
  geom_rect(majortimes_stats, mapping = aes(xmin = min.t3, xmax = max.t3,
                                            ymin = -Inf, ymax = Inf),
            alpha = 0.6, fill = "grey70",
            inherit.aes = FALSE) +
  geom_vline(majortimes_stats, mapping = aes(xintercept = avg.t3, lty = instar)) +
  theme_bw() +
  scale_color_manual(values = c("#2166ac", "#92c5de", "#f4a582", "#d6604d", "#b2182b")) +
  labs(title = "avg daily growth after entering trt",
       x = "days since 3rd", y = "avg log(mass (mg))",  
       color = "treatment",
       caption = "vlines = avg±sd instar duration")
```


# bradford plate data i dont understand how to use excel

##- fit std curve

```{r}
df <- read.csv("~/Documents/repos/ntw/2024/data/plates/241009_bradford-1.csv", header = TRUE)

# blank correct
df %>%
  filter(type == "std") %>%
  group_by(cxn) %>%
  summarise(avg= mean(abs)) # blank = 0.721

df <- df %>% 
  mutate(blank = 0.721,
         bc = abs - 0.721) 

# take avgs
std <- df %>%
  filter(type == "std") %>%
  group_by(cxn) %>%
  summarise(bc.avg = mean(bc),
            bca.wt = bc.avg)

unk <- df %>%
  filter(type == "unk") %>%
  group_by(time, temp, rep) %>%
  summarise(bc.avg = mean(bc),
            bca.wt = bc.avg/log(mass)) # mass-corrected

# fit std curve and unks
stdc <- lm(cxn ~ poly(bc.avg, 2, raw = TRUE), data = std)
stdc.wt <- lm(cxn ~ poly(bca.wt, 2, raw = TRUE), data = std)

unk$temp[is.na(unk$temp)] <- "ctrl"
unk$temp <- factor(unk$temp, levels = c("ctrl", "26", "33", "36"))


std <- cbind(std, fit = predict(stdc, type = "response"))
unk <- cbind(unk, fit = predict(stdc, unk))


std <- cbind(std, fit.wt = predict(stdc.wt, type = "response"))
unk <- cbind(unk, fit.wt = predict(stdc.wt, unk))

```

## - plots

```{r}
# plot
#library(viridis)

ggplot(data = std, aes(y = cxn, x = bc.avg)) +
  geom_point(aes()) + 
  geom_line(data = std, aes(y = fit))

# raw abs
ggplot(data = std, aes(y = cxn, x = bc.avg)) +
  geom_point(aes()) + 
  #geom_line(mapping = aes(x = cxn)) +
  geom_line(data = std, aes(y = fit)) +
  geom_point(data = unk, aes(y = fit, #shape = time, 
                             color = temp), size = 2) +
  facet_wrap(~time) +
  labs(#title = "non-mass corrected abs",
       title = "protein concentrations (Bradford)",
       y = "protein concentration (ug/mL)",
       x = "avg abs @ 595 nm",
       caption = "T0 = pre-trt; T1 = 100 hrs in trt; T2 = 5th d1",
       color = "treatment") +
  theme_bw() +
  scale_shape_manual(values = c(4, 1, 16)) +
  #scale_color_brewer(palette = "BrBG", direction = -1)
  scale_color_manual(values = c("seagreen4", "cyan3", "purple4", "darkorange"))
  #scale_color_viridis(discrete = TRUE, option = "mako")


# mass-corrected abs
ggplot(data = std, aes(y = cxn, x = bca.wt)) +
  geom_point() + 
  #geom_line(mapping = aes(x = cxn)) +
  geom_line(data = std, aes(y = fit.wt)) +
  geom_point(data = unk, aes(y = fit.wt, #shape = time, 
                             color = temp), size = 2) +
  facet_wrap(~time) +
  labs(#title = "mass corrected abs",
       title = "protein concentrations (Bradford)",
       y = "protein concentration (ug/mL)",
       x = "avg abs @ 595 nm",
       caption = "T0 = pre-trt; T1 = 100 hrs in trt; T2 = 5th d1",
       color = "treatment") +
  #theme_bw() +  scale_shape_manual(values = c(4, 1, 16)) +
  scale_color_brewer(palette = "BrBG", direction = -1) +
  ylim(0, 11) + xlim(0, 0.2)
  #scale_color_viridis(discrete = TRUE, option = "mako")

```


```{r}
# does mass correlate with abs ????
df %>%
  filter(type == "unk") %>%
  mutate(temp = case_when(is.na(temp) ~ "ctrl",
                          TRUE ~ as.character(temp)),
         temp = factor(temp, levels = c("ctrl", "26", "33", "36"))) %>%
  group_by(time, temp, rep) %>%
  summarise(avg.abs = mean(abs),
            mass = mass) %>%
  filter(time != "T0") %>%
  ggplot(aes(x = log(mass), y = avg.abs,
             color = temp, shape = time)) +
  geom_point(size = 2) +
  facet_wrap(~time) +
  #scale_shape_manual(values = c(4, 1, 16)) +
  #scale_color_brewer(palette = "BrBG", direction = -1) +
  scale_color_manual(values = c("cyan3", "purple4", "darkorange")) +
  #theme_bw() +
  labs(title = "abs vs mass",
       y = "avg abs @ 595 nm") +
  ylim(0, 2) +
  xlim(5.5, 8.5)

```

# sod assay (try 1)

2025-01-20: need to fix this per 240115 joel mtg lol

##- fit std curve

```{r}
df <- read.csv("~/Documents/repos/ntw/2024/data/plates/241212_sod.csv", header = TRUE)

df <- mutate(df,
             bc = abs.post - abs.blank)

# take avgs
std <- df %>%
  filter(type == "std") %>%
  group_by(cxn) %>%
  summarise(bc.avg = mean(bc),
            bca.wt = bc.avg)

unk <- df %>%
  filter(type == "unk") %>%
  group_by(time, temp, rep) %>%
  summarise(bc.avg = mean(bc),
            bca.wt = bc.avg/log(mass)) # mass-corrected

# fit std curve and unks
stdc <- lm(cxn ~ poly(bc.avg, 2, raw = TRUE), data = std)
stdc.wt <- lm(cxn ~ poly(bca.wt, 2, raw = TRUE), data = std)

unk$temp[is.na(unk$temp)] <- "ctrl"
unk$temp <- factor(unk$temp, levels = c("ctrl", "26", "33", "36"))


std <- cbind(std, fit = predict(stdc, type = "response"))
unk <- cbind(unk, fit = predict(stdc, unk))


std <- cbind(std, fit.wt = predict(stdc.wt, type = "response"))
unk <- cbind(unk, fit.wt = predict(stdc.wt, unk))

```

## - plots

```{r}
# plot
#library(viridis)

ggplot(data = std, aes(x = cxn, y = bc.avg)) +
  geom_point(aes()) + 
  geom_line(data = std, aes(x = fit))

# raw abs
ggplot(data = std, aes(x = cxn, y = bc.avg)) +
  geom_point(aes()) + 
  #geom_line(mapping = aes(x = cxn)) +
  geom_line(data = std, aes(x = fit)) +
  geom_point(data = unk, aes(x = fit, #shape = time, 
                             color = temp), size = 2) +
  facet_wrap(~time) +
  labs(#title = "non-mass corrected abs",
       title = "SOD activity",
       x = "SOD cxn (U/mL)",
       y = "avg abs @ 450 nm",
       caption = "T0 = pre-trt; T1 = 100 hrs in trt; T2 = 5th d1",
       color = "treatment") +
  theme_bw() +
  scale_shape_manual(values = c(4, 1, 16)) +
  #scale_color_brewer(palette = "BrBG", direction = -1)
  scale_color_manual(values = c("seagreen4", "cyan3", "purple4", "darkorange"))
  #scale_color_viridis(discrete = TRUE, option = "mako")


# mass-corrected abs
ggplot(data = std, aes(y = cxn, x = bca.wt)) +
  geom_point() + 
  #geom_line(mapping = aes(x = cxn)) +
  geom_line(data = std, aes(y = fit.wt)) +
  geom_point(data = unk, aes(y = fit.wt, #shape = time, 
                             color = temp), size = 2) +
  facet_wrap(~time) +
  labs(#title = "mass corrected abs",
       title = "SOD activity",
       x = "SOD cxn (U/mL)",
       y = "avg abs @ 450 nm",
       caption = "T0 = pre-trt; T1 = 100 hrs in trt; T2 = 5th d1",
       color = "treatment") +
  #theme_bw() +  scale_shape_manual(values = c(4, 1, 16)) +
  scale_color_brewer(palette = "BrBG", direction = -1) +
  ylim(0, 11) + xlim(0, 0.2)
  #scale_color_viridis(discrete = TRUE, option = "mako")

```


```{r}
# does mass correlate with abs ????
df %>%
  filter(type == "unk") %>%
  mutate(temp = case_when(is.na(temp) ~ "ctrl",
                          TRUE ~ as.character(temp)),
         temp = factor(temp, levels = c("ctrl", "26", "33", "36"))) %>%
  group_by(time, temp, rep) %>%
  summarise(avg.abs = mean(abs),
            mass = mass) %>%
  filter(time != "T0") %>%
  ggplot(aes(x = log(mass), y = avg.abs,
             color = temp, shape = time)) +
  geom_point(size = 2) +
  facet_wrap(~time) +
  #scale_shape_manual(values = c(4, 1, 16)) +
  #scale_color_brewer(palette = "BrBG", direction = -1) +
  scale_color_manual(values = c("cyan3", "purple4", "darkorange")) +
  #theme_bw() +
  labs(title = "abs vs mass",
       y = "avg abs @ 595 nm") +
  ylim(0, 2) +
  xlim(5.5, 8.5)

```

# sod assay (try 2)

## - fit std curve

- using 4PLC: https://janalin.github.io/analyse-ELISA/calibration.html

```{r}
# minpack.lm::nlsLM
df <- read.csv("~/Documents/repos/ntw/2024/data/plates/prelim.csv", header = TRUE)

sod <- filter(df, assay == "sod")
```


