---
title: "entsoc 2024 figs and analyses"
date: "2024-10-22"
---

# 0. roadmap

goal: making figs/running stats for entsoc poster, focusing on lab bugs

# 1. load data & packages

```{r}
source("./helpers_entsoc.R")

source("../2023/helpers_tents.R")

dtents <- data_tstats 
#%>% select(-c("trt.type", "trt.pair"))

dad <- data_longevity

rm(data_tstats, data_hatch, data_longevity,
   labels.alltrts, labels.exptrts, RYB,
   x_err_ncoll, x_err_ncollf, y_err_hrate)
```

convenience functions

```{r}
# calc std err
se <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
```



value/columns renaming and filtering

```{r dev data formatting}
# change mass to g
d23 <- d23 %>%
  mutate(mass.pupa = mass.pupa/1000,
         mass.eclose = mass.eclose/1000)

# adult data
dad <- dad %>%
  rename(jdate.eclose = jdate.ec,
         jdate.surv = jdate.died) %>%
  mutate(mass.eclose = mass.eclose/1000,
         for.surv = case_when(track.reason %in% c("surv-fert", "surv") ~ 1),
         for.fert = case_when(track.reason %in% c("surv-fert", "fert") ~ 1),
         trt.type = case_when(maxT == minT ~ "ctrl",
                              TRUE ~ "expt")) %>%
  select(-c(track.reason, starts_with("date")))
```

```{r tent data formatting}
# tent data
dtents <- dtents %>%
  filter(pop == "lab") %>%
  #rename(trt.typeold = trt.type) %>%
  mutate(trt.mateto = case_when(trt.m != trt.f | trt.m + trt.f == 520 ~ "ctrl",
                                TRUE ~ "hs"),
         trt.sex = case_when(trt.m > trt.f ~ "m",
                             trt.f > trt.m ~ "f",
                             trt.m == trt.f ~ "both"),
         trt.hs = case_when(trt.m > trt.f ~ as.numeric(trt.m),
                            trt.f > trt.m ~ as.numeric(trt.f),
                            trt.m + trt.f > 600 ~ as.numeric(trt.m),
                            TRUE ~ 260),
         trt.isctrl = case_when(trt.m + trt.f == 520 ~ 1,
                              TRUE ~ 0),
         trt.ref = case_when(trt.m + trt.f == 520 ~ "ctrl",
                             trt.m != trt.f ~ "btwn",
                             trt.m == trt.f ~ "win")
         ) %>%
  select(-pop)
```


# 2. adult, larval dev

## a. df prep & summary stats

```{r dev stats}
# larval stats
ss_la <- d23 %>%
  group_by(trt.type, minT, 
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            # avg.mass = mean(na.omit(mass.pupa)),
            # se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            #n = n(),
            n = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n - n.pmd, 
            prop.survpup = round(1-(n.pmd/n), digits=2)
            ) %>% 
  #mutate_all(~ifelse(is.nan(.), NA, .))
  mutate(stage = "la")

# pupal dev (eclosion time)
ss_pu <- d23 %>%
  group_by(trt.type, minT) %>%
  summarise(avg.tt = mean(na.omit(jdate.eclose - jdate.pupa)),
            se.tt = sd(na.omit(jdate.eclose - jdate.pupa))/sqrt(length(na.omit(jdate.eclose - jdate.pupa))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),) %>%
  mutate(stage = "pu")

# adult stats
ss_ad.sex <- dad %>%
  filter(!is.na(for.surv), pop == "lab") %>%
  group_by(#diet, # gets too parsed out w/ diet lol
           trt.type, minT, sex
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.surv - jdate.eclose)),
            se.tt = sd(na.omit(jdate.surv - jdate.eclose))/sqrt(length(na.omit(jdate.surv - jdate.eclose))),
            avg.mass = mean(na.omit(mass.eclose)),
            se.mass = sd(na.omit(mass.eclose)/sqrt(length(na.omit(mass.eclose)))),
            n = sum(na.omit(!is.na(jdate.eclose)))
            ) %>%
  mutate(stage = "ad")

ss_ad <- dad %>%
  filter(!is.na(for.surv), pop == "lab") %>%
  group_by(
           trt.type, minT
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.surv - jdate.eclose)),
            se.tt = sd(na.omit(jdate.surv - jdate.eclose))/sqrt(length(na.omit(jdate.surv - jdate.eclose))),
            avg.mass = mean(na.omit(mass.eclose)),
            se.mass = sd(na.omit(mass.eclose)/sqrt(length(na.omit(mass.eclose)))),
            n = sum(na.omit(!is.na(jdate.eclose)))
            ) %>%
  mutate(stage = "ad")
```

```{r final dev df}
# combine dfs
ss_dev <- full_join(ss_la, ss_pu)
ss_dev <- full_join(ss_dev, ss_ad)
ss_dev <- full_join(ss_dev, ss_ad.sex)

ss_dev$stage <- factor(ss_dev$stage, levels = c("la", "pu", "ad"))

ss_dev.expt <- filter(ss_dev, trt.type == "expt")
ss_dev.ctrl <- filter(ss_dev, trt.type == "ctrl")
```

```{r final stats df}
# df for exptal bugs only (for modeling)
d23.expt <- filter(d23, trt.type == "expt")
```


## b. figs & modeling

### - survival

```{r surv fig}
plot.surv <- #ss_dev.expt %>%
  ss_dev %>%
  filter(trt.type == "expt") %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), group = stage,
             #color = trt.type
             )) +
  geom_point(size = 2, shape = 16) +
  geom_line(lty = 2) +
  #geom_text(aes(label = n.surv), vjust = 2.25) +
  theme_bw() +
  labs(y = "proportion survived to pupation",
       x = "larval nighttime temperature (°C)") +
  ylim(c(0, 1))

plot.surv
```


```{r surv stats}
# new df for survival data: drops accidental deaths/na outcomes (0 = eclose, 1 = death)
d_surv <- d23.expt %>%
  rename(status = surv.outcome) %>%
  filter(status < 2)

mod_surv <- glm(status ~ as.factor(minT), data = d_surv, family = "binomial")
mod_surv2 <- glm(status ~ as.factor(minT)*sex, data = d_surv, family = "binomial")

summary(mod_surv)
anova(mod_surv, test = "Chisq") # minT ***
```



### - mass

```{r mass fig}
# after running stats: sex matters for adults lol

plot.devmass <- #ss_dev.expt %>%
  ss_dev %>%
  filter(trt.type == "expt") %>%
  #filter(!(stage %in% c("ad.sex", "pu"))) %>% # m/f aren't that different
  #filter(stage %in% c("ad.sex", "pu")) %>%
  filter(stage == "ad" & !is.na(sex) | stage == "pu") %>%
  # mutate(stage = case_when(stage == "la" ~ "pu",
  #                          TRUE ~ as.character(stage)),
  #        stage = factor(stage, levels = c("pu", "ad.sex"))) %>%
  # mutate(sex = case_when(is.na(sex) ~ "etc",
  #                        TRUE ~ sex)) %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), group = interaction(stage, sex),
             #group = stage,
             color = sex, shape = sex,
             #color = trt.type
             )) +
  geom_point(size = 2) +
  geom_line(aes(lty = stage)) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.1) +
  theme_bw() +
  labs(y = "mass (g)",
       x = "larval nighttime temperature (°C)") +
  scale_linetype_manual(labels = c("pupa", "adult"), values = c(2, 1)) +
  scale_shape_manual(#breaks = ~ .x[!is.na(.x)],
                       values = c(17, 15), limits = c("f", "m"),
                       na.translate = TRUE, na.value = 16
                       ) +
  scale_color_discrete(#breaks = ~ .x[!is.na(.x)],
                       limits = c("f", "m"),
                       na.translate = TRUE, na.value = "black"
                       ) +
  ylim(c(1.15, 4.6)) # 1.25-4.6 if omitting sex
  
plot.devmass
```

stats

```{r pupal mass stats}
# trying other possible influencing factors first
mod_pmass1 <- lm(mass.pupa ~ as.factor(minT)*sex, data = d23)
mod_pmass2 <- lm(mass.pupa ~ as.factor(minT), data = d23) # better per AIC

anova(mod_pmass1) # minT ***, sex ns (0.19), int ns (0.42)
anova(mod_pmass2) # minT ***

AIC(mod_pmass1) # 351.1274
AIC(mod_pmass2) # 348.6995
```

```{r adult mass stats}
mod_amass1 <- lm(mass.eclose ~ as.factor(minT)*sex, data = d23) # better per AIC
mod_amass2 <- lm(mass.eclose ~ as.factor(minT), data = d23)

anova(mod_amass1) # minT ***, sex **
anova(mod_amass2) # minT ***

AIC(mod_amass1) # 214.7664
AIC(mod_amass2) # 220.7273
```



### - time

```{r dev time fig} 
plot.devtime <- #ss_dev.expt %>% 
  ss_dev %>%
  filter(trt.type == "expt") %>%
  #filter(!(stage %in% c("ad.sex", "pu"))) %>% # m/f not that different
  filter(stage %in% c("ad", "la") & is.na(sex)) %>% 
  ggplot(aes(y = avg.tt, x = as.factor(minT), group = stage,
             #color = trt.type
             )) +
  geom_point(size = 2, shape = 16) +
  geom_line(aes(lty = stage)) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt),
                width = 0.08) +
  theme_bw() +
  scale_linetype_manual(labels = c(#"adult\n(eclosion → death)", "larva\n(hatching → pupation)",
                                   "larva", "adult"
                                   #"pupa (pupation → eclosion)"
                                   ),
                        values = c(4, 1#, 3
                                   )
                        ) +
  labs(y = "development time / longevity (days)",
       x = "larval nighttime temperature (°C)")

plot.devtime
```

stats

```{r larval dtime stats}
mod_ltime1 <- lm((jdate.pupa - jdate.hatch) ~ as.factor(minT)*as.factor(sup), data = d23) # better per AIC
mod_ltime2 <- lm((jdate.pupa - jdate.hatch) ~ as.factor(minT), data = d23)

anova(mod_ltime1) # minT ***, sup ***, int ns (0.75)
anova(mod_ltime2) # minT ***

AIC(mod_ltime1) # 480.7616
AIC(mod_ltime2) # 576.1871
```


```{r adult dtime stats}
mod_atime1 <- lm((jdate.surv - jdate.eclose) ~ as.factor(minT)*sex, data = d23) 
mod_atime2 <- lm((jdate.surv - jdate.eclose) ~ as.factor(minT), data = d23) # better per AIC

anova(mod_atime1) # minT ns (0.55), sex ns (0.81)
anova(mod_atime2) # minT ns

AIC(mod_atime1) # 158.7356
AIC(mod_atime2) # 156.8043
```


### - combined plot

```{r}
# library(patchwork)
# check that lty/shape values match across plots for merge to work


(plot.surv | plot.devmass | plot.devtime) +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(1:1)) &
  theme(legend.position = "top") &
  scale_linetype_manual(labels = c("larva", "pupa", "adult", "adult"), values = c(4,2,1,1), 
                        drop = FALSE) # good enough man idk

#ggsave("./figs/entsoc/big 3.png", dpi = "retina")
```




# 3. tent stats

## a. hatch rate calcs

these calcs are copy/pasted/tweaked from `/2023/analyses_tents.Rmd`

### - pre-math (fixing #s)

```{r fixing calcs}
# determine actual # of ovipositing females
dtents <- dtents %>%
  mutate_at(c(6:14), ~replace_na(., 0)) %>% 
  group_by(id.tent) %>%
  mutate(jdate.rel = jdate - first(jdate)) %>%
  ungroup() %>%
  group_by(id.tent, trt.f, trt.m, trt.sex, jdate.rel) %>%
  arrange(jdate.rel, .by_group = TRUE) %>%
  mutate(#n.laying = n.curr.f + n.died.f, .after = "n.curr.f",
         n.ovi = dplyr::lag(n.curr.f, default = first(n.curr.f)) + n.died.f, .after = "n.curr.f") %>%
  ungroup()


# where n.ovipositing females = 0 but n.collected > 0, add the # eggs collected to the day before
# (or last date of ovipositing females)
# first, figure out the rows
cond <- (dtents$n.coll > 0 & dtents$n.ovi == 0)

ovifix <- dtents %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

negovi <- ovifix[ovifix$females.laying < 0, ]

lastovi <- ovifix[ovifix$n.curr.f > 0 & ovifix$n.ovi >= 1, ]

test2 <- c(negovi$index, lastovi$index)

# then, edit values
#summ_rate$n.coll[negovi$index] <- 0
dtents <- dtents %>% 
  mutate(n.coll = case_when(rownames(.) %in% negovi$index ~ 0,
                            rownames(.) %in% lastovi$index ~ n.coll + 1,
                            !(rownames(.) %in% c(negovi$index, lastovi$index)) ~ as.numeric(n.coll)
                            )) %>%
  select(-females.laying)

rm(cond, ovifix, negovi, lastovi)

# removing extra 0's
# dtents <- dtents %>%
#   filter(!(n.coll == 0 & n.ovi == 0))
  # tbh not sure why i did this in the old version. this drops a few #s here and there

```


### - fecundity rates (# laid)

ref: `summ_rates_all` calcs in `analyses_tents.Rmd`

```{r}
# i dont need this lol

# # this calcs how # of tents, # eggs laid, # ovi f changes daily per trt
# ss_ovirates <- dtents %>%
#   group_by(trt.ref, trt.sex, trt.hs, jdate.rel) %>%
#   summarise(n.tents = n(), #idk why i use mutate in the orig lol
#          eggs.perovi = sum(n.coll)/sum(n.ovi),
#          eggsovi.pertent = eggs.perovi / n.tents) %>%
#   mutate_all(~ifelse(is.nan(.), NA, .))

```


### - fertility rates (# hatched)

ref: `summ_hatch` calcs in `analyses_tents.Rmd`

```{r}
# want: # eggs laid per female (overall), prop hatched in tent
ss_hatchrates <- dtents %>%
  group_by(trt.ref, trt.sex, trt.hs) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.laid = sum(n.coll, na.rm = TRUE),
            #se.laid = se(n.coll),
            n.fs = sum(n.new.f, na.rm = TRUE), # = total # fs in a tent
            n.laidf = round(n.laid/n.fs, digits = 2),
            #se.laidf = se(sum(n.coll, na.rm = TRUE)/sum(n.new.f)), # doesnt work
            se.laidf = sqrt(n.laidf*(1-n.laidf/n.laid)), # hmmmm.
            n.hatched = sum(n.tothatch, na.rm = TRUE), # needed downstream
            #se.hatched = se(n.tothatch),
            prop.hatched = n.hatched/n.laid,
            se.hatchprop = sqrt(prop.hatched*(1-prop.hatched)/n.laid),
            # prop.hatchf = prop.hatched/n.fs,
            # se.hatchfprop = sqrt(prop.hatchf*(1-prop.hatchf)/n.laid)
            )
```


## b. figs

```{r}
# a note that colors were formerly
# c("#f15a29", "#b2325c", "#4d276a")
```


### - win hatch

```{r win hatch rates plot}
plot.whatch <- ss_hatchrates %>%
  filter(trt.ref %in% c("win", "ctrl")) %>%
  ggplot(aes(y = prop.hatched, x = n.laidf, color = as.factor(trt.hs))) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin= prop.hatched - se.hatchprop, ymax = prop.hatched + se.hatchprop), 
                width = 8
                ) +
  geom_errorbarh(aes(xmin = n.laidf - se.laidf, xmax = n.laidf + se.laidf), 
                 height = 0.005
                 ) +
  #geom_errorbarh(aes(xmin = n.laid - se.laid, xmax = n.laid + se.laid), height = 0.01) +
  geom_text(aes(label = n.fs), vjust = -0.7, hjust = -0.7, show.legend = FALSE) +
  theme_bw() +
  labs(color = "larval temperature treatment",
       y = "proportion eggs hatched",
       x = "# eggs laid per female")+
  guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
  scale_color_manual(labels = c("26-26°C", "40-19°C", "40-26°C", "40-33°C"),
                     #values = c("#0b809f", "#4d266b", "#b12256", "#f88e28")
                     #values = c("#4393c3", "#f4a582", "#d6604d", "#b2182b")
                     values = c("#4393ce", "#fdae61", "#d73027", "#a50026")
                     )

plot.whatch # y: 0-0.18?, x: 0-200
```


### - btwn hatch

```{r btwn hatch rates plot}
plot.bhatch <- ss_hatchrates %>%
  filter(trt.ref == "btwn") %>%
  ggplot(aes(y = prop.hatched, x = n.laidf, color = as.factor(trt.hs), shape = trt.sex)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin= prop.hatched - se.hatchprop, ymax = prop.hatched + se.hatchprop), 
                width = 8
                ) +
  geom_errorbarh(aes(xmin = n.laidf - se.laidf, xmax = n.laidf + se.laidf), 
                 height = 0.005
                 ) +
  geom_text(aes(label = n.fs), vjust = -0.9, hjust = -0.8, show.legend = FALSE) +
  theme_bw() +
  labs(color = "larval temperature\ntreatment",
       shape = "treated sex",
       y = "proportion eggs hatched",
       x = "# eggs laid per female") +
  scale_color_manual(labels = c("40-19°C", "40-26°C", "40-33°C"),
                     #values = c("#4d266b", "#b12256", "#f88e28"),
                     #values = c("#f4a582", "#d6604d", "#b2182b"),
                     #values = c("#fdae61", "#f46d43", "#a50026"),
                     values = c("#fdae61", "#d73027", "#a50026")) +
  guides(color = "none", # omit for plot_layout to work
         shape = guide_legend(nrow = 2)
         ) + 
  scale_shape_manual(values = c(17, 15))

plot.bhatch # y: 0-0.12, x = 0-350
```
### - combined plot

```{r}
(plot.whatch | plot.bhatch) +
  plot_layout(guides = "collect", 
              axes = "collect", 
              widths = c(1:1)
              ) &
  theme(legend.position = "bottom",
        legend.justification = "left"
        ) &
  # scale_color_manual(labels = c("26-26°C", "40-19°C", "40-26°C", "40-33°C"),
  #                    values = c("#4393ce", "#fdae61", "#d73027", "#a50026")
  #                    )
  ylim(c(-0.005, 0.18)) &
  xlim(c(0, 340))

ggsave("./figs/entsoc/tents-2.png",dpi = "retina")
```


## c. modeling

### - df prep & value-checking

```{r}
# calc avg eggs/F (fecundity) and overall prop hatch PER TENT
d_repro <- dtents %>%
  filter(trt.ref != "ctrl") %>%
  group_by(id.tent, trt.ref, trt.sex, trt.hs) %>%
  summarise(n.laid = sum(n.coll, na.rm = TRUE),
            n.fs = sum(n.new.f, na.rm = TRUE), # = total # fs in a tent
            avg.laidf = round(n.laid/n.fs, digits = 2), # fecundity stat
            log.fec = log(avg.laidf + 1),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            prop.hatched = n.hatched/n.laid, # fert stat
            log.fert = log(prop.hatched + 1) 
            )

# imo need log.fert and avg.laidf is ok... just looking at 0's. but should check qqplots later

d_win <- d_repro %>%
  filter(trt.ref == "win")
d_btwn <- d_repro %>%
  filter(trt.ref == "btwn")
```

### - btwn trts
```{r modeling btwn}
# no random effs bc of tent grouping lol

# fecundity 
mod_bfec <- lm(avg.laidf ~ trt.hs*trt.sex, data = d_btwn)
mod_bfec.log <- lm(log.fec ~ trt.hs*trt.sex, data = d_btwn)

# fertility
mod_bfert <- lm(prop.hatched ~ trt.hs*trt.sex, data = d_btwn)
mod_bfert.log <- lm(log.fert ~ trt.hs*trt.sex, data = d_btwn)
```

```{r checking btwn mods}
anova(mod_bfec) # trt (0.95), sex (.17), int (0.82) ns 
anova(mod_bfec.log) # trt (0.79), sex (.97), int (0.64) ns 
AIC(mod_bfec) # 177.0551
AIC(mod_bfec.log) # 61.76643

anova(mod_bfert) 
anova(mod_bfert.log)  
AIC(mod_bfert) # better?? not rly tho - they are pretty close anyway lol
AIC(mod_bfert.log)
```

```{r qq btwn mods}
# fecundity
qqnorm(resid(mod_bfec), main = "QQ btwn fecundity")
qqline(resid(mod_bfec))

qqnorm(resid(mod_bfec.log), main = "QQ log(btwn fecundity)") # this is better (sorta)
qqline(resid(mod_bfec.log))

# fertility 
# they both rly suck bc of a shitton of 0s lol
qqnorm(resid(mod_bfert), main = "QQ btwn fertility")
qqline(resid(mod_bfert))

qqnorm(resid(mod_bfert.log), main = "QQ log(btwn fertility)")
qqline(resid(mod_bfert.log))
```


Analysis of Variance Table

Response: avg.laidf
               Df Sum Sq Mean Sq F value Pr(>F)
trt.hs          1     48    48.0  0.0039 0.9517
trt.sex         1  27487 27487.0  2.2042 0.1685
trt.hs:trt.sex  1    680   680.4  0.0546 0.8200
Residuals      10 124702 12470.2               
Analysis of Variance Table

Response: log.fec
               Df Sum Sq Mean Sq F value Pr(>F)
trt.hs          1  0.242  0.2421  0.0732 0.7922
trt.sex         1  0.005  0.0052  0.0016 0.9693
trt.hs:trt.sex  1  0.784  0.7844  0.2372 0.6368
Residuals      10 33.075  3.3075               
[1] 177.0551
[1] 61.76643


Analysis of Variance Table

Response: prop.hatched
               Df   Sum Sq   Mean Sq F value Pr(>F)
trt.hs          1 0.002134 0.0021337  0.3753 0.5553
trt.sex         1 0.007761 0.0077610  1.3653 0.2726
trt.hs:trt.sex  1 0.000566 0.0005664  0.0996 0.7595
Residuals       9 0.051161 0.0056845               
Analysis of Variance Table

Response: log.fert
               Df   Sum Sq   Mean Sq F value Pr(>F)
trt.hs          1 0.001633 0.0016333  0.3449 0.5715
trt.sex         1 0.006554 0.0065543  1.3841 0.2696
trt.hs:trt.sex  1 0.000402 0.0004023  0.0849 0.7773
Residuals       9 0.042620 0.0047355               
[1] -25.0981
[1] -27.47261



### - win trts
```{r modeling win}
# fecundity 
mod_wfec.log <- lm(log.fec ~ trt.hs, data = d_win)

# fertility
mod_wfert <- lm(prop.hatched ~ trt.hs, data = d_win)
```


```{r checking win mods}
anova(mod_wfec.log) # trt (0.94) ns

anova(mod_wfert) # trt (0.21) ns

```


```{r qqplots win}
# fecundity
qqnorm(resid(mod_wfec.log)) 
qqline(resid(mod_wfec.log))

# fertility 
qqnorm(resid(mod_wfert))
qqline(resid(mod_wfert))

```



# 4. some archived stuff

- archived 2025-01-20 from `2024/analyses_ntw-compare.Rmd` when i first tried to combine 23v24 for entsoc lol (before I decided to just focus on 2023 data)
  - moved code into this doc 2025-02-13
- needs the dfs (all ntw bugs and 2023 ntw bugs) to run properly lolol, so see thw wrangling from there.


## - summary stats

// todo 
- 2024-10-17: keeping bugs separate by year for now, will see how they look separate before trying to combine
- 2025-01-20: filter out bugs on their correct diet types only for now

```{r}
# sep by sups (bc the sups have their own dev trajectory...)
ss_wide.sup <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT, sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) # `minT` is ok here

# all dev trajectories together
ss_wide <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT,
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) #%>% 
  #mutate_all(~ifelse(is.nan(.), NA, .)) # breaks `minT`

```


prereqs: a cleaned up `d23`

## - stage-specific stats

```{r}
# slapping sth on for entsocing
ss_larva <- d23 %>%
  filter(pop == "lab") %>%
  group_by(year, pop, diet, trt.type, minT, 
           #sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n.all - n.pmd, 
            prop.survpup = round(1-(n.pmd/n.all), digits=2)
            ) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  mutate(minT = factor(minT, levels = c("19", "26", "33"))) %>%
  filter(!(pop == "lab" & diet == "TB") &
           trt.type == "expt")
```




## - figs


*notes:* 

- use lab bugs on LD + field on TB only!!!! (this is why u cant just reuse ur LB figs lol)
- 2025-01-20: below figs assume all combos of diet types and pops are included in the `all` df.

```{r}
# (messing around w 2023 data only for entsoc,,)
# using LB code lol

# survival
p1 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = prop.survpup, x = minT, shape = pop, color = pop,
             )) +
  geom_point(size=3) + 
  geom_line(aes(lty = pop, group = pop)) +
  geom_text(aes(label = n.all), hjust = -0.5, show.legend = FALSE) +
  theme_bw() +
  labs(title = "23v24 ntw pupal survival",
       y = "proportion survived to pupation", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  ylim(c(0, 1))

p1
```

```{r}
# avg tt pup
pd <- position_dodge(0.3)

p2 <-
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.tt, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3, #position = pd
             ) +  
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop), #position = pd
            ) +
  theme_bw() +
  ylim(c(18, 34)) +
  labs(title = "23v24 ntw larval development",
       y = "time to pupation (days)", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year)

p2
```

```{r}
# avg mass
p3 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.mass, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3) +  
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop)) +
  theme_bw() +
  #ylim(c(1750, 4750)) +
  labs(title = "23v24 ntw pupal mass",
       y = "mass at pupation (mg)", 
       x = "minimum temperature (°C)",
       #caption = "2023 data only; lab+LD and field+TB"
       ) +
  facet_wrap(~year)

p3
```

```{r}
# merged plot
p1 + p2 + p3 +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(1:1)) &
  theme(legend.position = "top") &
  labs(lty = "population", shape = "population",
       title = NULL) &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16))
  #guides(shape = "none", lty = "none")
```




