---
title: "ntw 2023 v 2024 data comparisons"
date: "2024-10-08"
---

# 0. roadmap

goal: compare 2023 and 2024 ntw development data. needs some wrangling lol (esp 2023)

- 2025-02-13: commenting out/moved out some of the 26/33 vs ±0/7 stuff. moved some of it to a draft doc

wip: combine fertility data...

# 1. load data & packages

```{r}
source("./helpers_ntw-compare.R")

test <- d23
#d23 <- test
```


# 2. wrangling dfs

standardising some data values and column names in `d23`

```{r}
# pick relevant data and rename columns/values to match 2024

# compare column names
# setdiff(names(d24), names(d23)) # differences
# setdiff(names(d23), names(d24))
# intersect(names(d24), names(d23)) # same

# filter rows and rename columns
d23 <- d23 %>%
  filter(reason.ignore != "lost" | src != "F1" #| !is.na(final.fate)
           ) %>% 
  filter(trt.stage %in% c("260-hatch", "267-hatch", "330-hatch", "337-hatch", "419-hatch","433-hatch")) %>%
  filter(expt.group %in% c("A", "B", "C", "D", "E", "F", "H")) %>%
  rename(trt = treatment,
         id = ID,
         cohort = expt.group,
         notes.ignore = reason.ignore)

# standardise values
d23$trt[d23$trt == 337] <- 426

# d23 <- d23 %>%
#   mutate(final.fate = case_when(notes.ignore %in% c("hot larva", "cut", "culled larva", "wet diet") | final.fate == "culled" ~ "other",
#                                 TRUE ~ as.character(final.fate))
#          )

d23 <- d23 %>%
  mutate(fate = case_when(notes.ignore %in% c("culled larva", "cut", "hot larva", "lost", "wet diet") ~ "other"))

# match values to d24
d23 <- d23 %>%
  mutate(sup = case_when(!is.na(jdate.7th) ~ 7,
                         !is.na(jdate.6th) ~ 6,
                         is.na(jdate.pmd) & !is.na(jdate.pupa) ~ 0),
         # case_when(sup = 0 & !is.na(jdate.5th) ~ NA_real_,
         #                 TRUE ~ as.numeric(sup)),
         surv.outcome = case_when(fate == "other" | (is.na(jdate.pmd) & is.na(jdate.pupa)) ~ 2,
                                  !is.na(date.pupa) | !is.na(date.LP) ~ 1,
                                  !is.na(date.pmd) ~ 0,
                                  #TRUE ~ 2
                                  ),
         # surv.outcome = case_when(!is.na(date.pupa) | !is.na(date.LP) ~ 0,
         #                          final.fate == "pmd" ~ 1,
         #                          final.fate == "other" ~ 2), # treat injuries the same as this
         flucT = case_when(flucT == 2.5 ~ 0,
                           TRUE ~ as.numeric(flucT)),
         trt.type = case_when(meanT == 26 & flucT == 0 ~ "ctrl",
                              TRUE ~ "expt"),
         trt = as.numeric(trt),
         #id = paste(id, year, sep = "-")
         )

# troubleshooting
# filter(d23, is.na(surv.outcome)) %>% View()
#test %>% group_by(final.fate) %>% summarise(n = n())
#test %>% group_by(reason.ignore) %>% summarise(n = n())
#filter(test, is.na(final.fate)) %>% View()

test2 <- d23
#d23<-test2

# how to handle LPIs? (in 23 and 24?)

d23 <- d23 %>%
  select(-c("jdate.collected", "jdate.15", "jdate.stuck",
            "jdate.exit", "jdate.LP", "jdate.died")) %>% # d23 exclusives
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "instar.enter", starts_with(c("jdate", "mass")), 
           "sex",
           ends_with("T", ignore.case = FALSE), "trt.type",
           "sup", "surv.outcome", "year"
          ))

# append longevity data to d23
d23_longevity[d23_longevity$id == 1415, "sex"] <- "f" # fix data

d23 <- merge(d23, d23_longevity, all = TRUE)
```


fine-tuning `d24`

```{r}
d24 <- d24 %>%
  #mutate(id = paste(id, year, sep = "-")) %>%
  select(-"jdate.culled") %>%
  filter(instar.enter == "hatch")
```



merge data

```{r}
all <- merge(d24, d23, all = TRUE
             ) %>%
  drop_na(id) %>%
  select(-c("instar.enter",
            ))

# drop things that died in 1 day
  # 2025-01-23: can i do this??????
all <- filter(all, jdate.pmd - jdate.enter > 1 | is.na(jdate.pmd - jdate.enter))

# add in exit dates to more easily parse out things still developing (for 2024 data)
# and other convenience things
all <- all %>% 
  mutate(
        # jdate.exit = case_when(!is.na(jdate.pmd) ~ jdate.pmd,
        #                         !is.na(jdate.pupa) ~ jdate.pupa,
        #                         TRUE ~ NA_integer_),
         minT = factor(minT, levels = c(19, 26, 33)),
         year = factor(year, levels = c(2023, 2024)),
         for.ntw = case_when(#year == 2023 & cohort %in% c("A", "B") ~ "N",
                             trt %in% c(260, 419, 426, 433) ~ "Y",
                             TRUE ~ "N"),
         for.const = case_when(#flucT %in% c(0, 7) & meanT %in% c(26, 33) ~ "Y",
                               trt %in% c(260, 267, 330, 426) ~ "Y",
                               TRUE ~ "N")
                               # trt.type == "ctrl" | (meanT == 33 & flucT %in% c(0, 7)) ~ "Y",
                               # TRUE ~ "N")
         )

```


# 3. df prep (math)

## - pivot prep

2025-01-20: filtering out diet wonkiness here lol. we can revisit it later if needed

```{r}
all_wide <- all %>%
  filter(!(pop == "lab" & diet == "TB")) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch,
         tt.8th = jdate.8th - jdate.hatch, # omit if only d23
         tt.wander = jdate.wander - jdate.hatch,
         tt.pupa = jdate.pupa - jdate.hatch,
         tt.eclose = jdate.eclose - jdate.pupa,
         tt.surv = jdate.surv - jdate.eclose,
         tt.lsurv = jdate.lsurv - jdate.lec, # for d23 longevity
         tt.pmd = jdate.pmd - jdate.hatch
         )  %>%
  # drop extra d23 longevity columns
  mutate(tt.surv = case_when(is.na(tt.surv) ~ as.numeric(tt.lsurv),
                             TRUE ~ as.numeric(tt.surv))) %>%
  #filter(year == 2023 & sex == "f" & !is.na(jdate.lsurv)) %>% View()
  select(-c("jdate.lsurv", "jdate.lec", "tt.lsurv"))

all_long <- all_wide %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
  drop_na(jdate) %>%
  drop_na(tt) # drops NA's if an individual didnt reach a certain stage


# filtering by expt type
ntw_wide <- filter(all_wide, for.ntw == "Y")
#cvf_wide <- filter(all_wide, for.const == "Y")
```

## - stage-specific summary stats

2025-01-20 troubleshooting:
- https://stackoverflow.com/questions/73533852/create-multiple-dataframes-based-on-the-dates-of-another-dataframe-in-r
- https://stackoverflow.com/a/24376207/3358227


```{r}
# apply calcs to data
ss_all <- lapply(list(ntw_wide, cvf_wide), calc.devstats)

ss_ntw <- Reduce(full_join, ss_all[[1]]) %>% 
  calc.ssadj()

# ss_cvf <- Reduce(full_join, ss_all[[2]]) %>%
#   calc.ssadj()
```

## - dev ss variations

```{r}
ss_moredev <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab")) %>% 
  calc.ssmoredev()

```


## - dev outcomes 

//todo
- 2025-02-04: try to group by surv outcome maybe... (instead of splitting into 2 dfs)

```{r}
# pmds + sup of those that pup'ed
ss_devall <- ntw_wide %>%
  mutate(trt = as.factor(trt)) %>%
  filter(!(diet == "TB" & pop == "lab")) %>%
  group_by(year, diet, pop, trt) %>%
  summarise(n.tot = n(),
            n.pmd = sum(na.omit(surv.outcome == 0)),
            n.misc = sum(na.omit(surv.outcome == 2)),
            n.pup = sum(na.omit(surv.outcome == 1)),
            n.sup = sum(na.omit(surv.outcome == 1 & sup > 0)),
            n.dev = n.tot - n.misc, # = pmd + pups of any sup
            n.5th = sum(na.omit(sup == 0 & surv.outcome == 1)),
            n.6th = sum(na.omit(sup == 6 & surv.outcome == 1)),
            n.7th = sum(na.omit(sup == 7 & surv.outcome == 1)),
            n.8th = sum(na.omit(sup == 8 & surv.outcome == 1)),
            p.pmd = round(n.pmd/n.dev, digits = 3),
            p.sup = round(n.sup/n.dev, digits = 3),
            p.5th = round(n.5th/n.dev, digits = 3),
            p.6th = round(n.6th/n.dev, digits = 2),
            p.7th = round(n.7th/n.dev, digits = 2),
            p.8th = round(n.8th/n.dev, digits = 2)
            )


# sup of those that pmd'd
ss_devpmd <- ntw_wide %>%
  mutate(trt = as.factor(trt)) %>%
  filter(!(diet == "TB" & pop == "lab")) %>%
  filter(surv.outcome == 1) %>%
  group_by(year, diet, pop, trt) %>%
  summarise(n.tot = n(),
            n.5th = sum(na.omit(sup == 0)),
            n.6th = sum(na.omit(sup == 6)),
            n.7th = sum(na.omit(sup == 7)),
            n.8th = sum(na.omit(sup == 8)),
            p.5th = round(n.5th/n.tot, digits = 2),
            p.6th = round(n.6th/n.tot, digits = 2),
            p.7th = round(n.7th/n.tot, digits = 2),
            p.8th = round(n.8th/n.tot, digits = 2)
            )

# pivots

ss_devall2 <- ss_devall %>%
  pivot_longer(cols = starts_with(c("n.", "p.")),
               names_to = c(".value", "stage"),
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  filter(stage %in% c("pmd", "5th", "6th", "7th"))

ss_devpmd2 <- ss_devpmd %>%
  pivot_longer(cols = starts_with(c("n.", "p.")),
               names_to = c(".value", "stage"),
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
    filter(stage %in% c("5th", "6th", "7th"))
```


# 4. figs & analyses prep

## - figs ref

larval "#1B9E77" 
pup "#D95F02" 
adult "#7570B3"

```{r}
# fig troubleshooting
# filter(all_wide, year == 2023 & sex == "f" & !is.na(tt.surv)) %>% View()
# filter(ntw_wide, year == 2023 & sex == "f" & !is.na(tt.surv)) %>% View()
# filter(ntw_wide, year == 2023 & sex == "f" & pop == "field" & minT == 33) %>% View()
# filter(ss_ntw, year == 2023 & sex == "f" & stage == "ad") %>% View()
```

## - stats prep

```{r}
ntw_expt <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab") & trt.type == "expt") %>%
  mutate(dmass = mass.pupa - mass.eclose,
         rate.pup = mass.pupa/tt.pupa)

```

# - pupal survival

## - figs

// todo
2025-01-23: 
- make these functions also LOL

```{r}
# NTs on pupal survival 

p1 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "expt") %>% 
  ggplot(aes(y = prop.survpup, x = minT,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  geom_text(aes(label = n.dev), hjust = -0.25, vjust = -0.5) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = prop.survpup, x = year,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  geom_text(aes(label = n.dev), hjust = -0.25, vjust = -0.5) +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs increase survival to pup?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       y = "proportion survived to pupation") &
  guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16)) &
  scale_color_manual(values = "#1B9E77") &
  ylim(c(0, 1))

```
## - stats

```{r}
ntw_expt.surv <- ntw_expt %>%
  filter(surv.outcome != 2)
```

```{r}
mod.surv_minT <- glm(surv.outcome ~ minT*pop, data = ntw_expt.surv, family = "binomial")
# mod.surv_meanT <- glm(surv.outcome ~ meanT*pop, data = ntw_expt.surv, family = "binomial")
  # ignore meanT bc it's confounded w minT

anova(mod.surv_minT, mod.surv_meanT, test = "Chisq") 

```


# - dev outcomes (sup)

## - figs: pmds + pups

```{r}
# not stacking stages/grouping by pop as desired
ss_devall2 %>%
  ggplot(aes(y = p, x = trt, fill = stage)) +
  #geom_col(position = "dodge") + facet_wrap(~year) +
  #geom_col() + facet_wrap(year~pop) +
  geom_col(position = "dodge", stat = "identity") + facet_wrap(~year) +
  labs(caption = "pmds + dev fates of pups") +
  theme_bw()

# okay apparently u just cant stack AND dodge at the same time lol says google
  # https://stackoverflow.com/questions/12715635/ggplot2-bar-plot-with-both-stack-and-dodge
  # https://stackoverflow.com/questions/46597278/how-to-plot-a-stacked-and-grouped-bar-chart-in-ggplot
ss_devall2 %>%
  ggplot(aes(y = p, x = interaction(pop, trt), fill = stage)) +
  geom_bar(position = "stack", stat = "identity") + facet_grid(~year) +
  #geom_col() + facet_wrap(year~pop) +
  labs(caption = "pmds + dev fates of pups") +
  theme_bw()

# this is the one!!!!!
ss_devall2 %>%
  ggplot(aes(y = p, x = interaction(pop, trt), fill = stage)) +
  geom_bar(position = "stack", 
           #position = position_fill(reverse = TRUE),
           stat = "identity") + 
  facet_grid(~interaction(pop,year), scale = "free", 
             #labeller = as_labeller(interaction(pop,year) = c("field 2023", "lab 2023", "field 2024", "lab 2024"))
             ) +
  #geom_col() + facet_wrap(year~pop) +
  labs(caption = "pmds + instars of pups",
       fill = "stage",
       x = "D-N treatment (°C)",
       y = "proportion") +
  scale_x_discrete(labels = c("26-26", "40-19", "40-26", "40-33"), guide = guide_axis(angle = 45)) +
  #theme(axis.text.x.bottom = element_text(angle = 90)) +
  scale_fill_brewer(palette = "Blues", direction = -1) +
  theme_bw()
  

# trying w/ alpha.. tldr i think i need to combine some of these variables tgt lol. but also redo the math
# (its not adding to 1 bc need x=interaction)
ss_devall2 %>%
  ggplot(aes(y = p, x = pop, fill = stage, group = trt, alpha = trt)) +
  geom_bar(position = "dodge", stat = "identity") + facet_grid(~year) +
  #geom_col() + facet_wrap(year~pop) +
  labs(caption = "pmds + dev fates of pups") +
  theme_bw()
```

## - figs: pups only

```{r}
# looking only at survivors
ss_devpmd2 %>%
  ggplot(aes(y = p, x = interaction(pop, trt), fill = stage)) +
  geom_bar(position = "stack", 
           #position = position_fill(reverse = TRUE),
           stat = "identity") + 
  facet_grid(~interaction(pop,year), scale = "free", 
             #labeller = as_labeller(interaction(pop,year) = c("field 2023", "lab 2023", "field 2024", "lab 2024"))
             ) +
  #geom_col() + facet_wrap(year~pop) +
  labs(caption = "instars of things that survived to pupation",
       fill = "stage",
       x = "D-N treatment (°C)",
       y = "proportion") +
  scale_x_discrete(labels = c("26-26", "40-19", "40-26", "40-33"), guide = guide_axis(angle = 45)) +
  #theme(axis.text.x.bottom = element_text(angle = 90)) +
  scale_fill_brewer(palette = "Blues", 
                    #direction = -1
                    ) +
  theme_bw()
```




# - dev time & longevity

// todo
- 2025-01-30: look into forcing y-axes limits to be the same w `ggh4x`
  - https://teunbrand.github.io/ggh4x/articles/Facets.html
  
## - figs: by sex

```{r}
# NTs on development time/longevity
# see this
  # https://stackoverflow.com/questions/78193007/how-do-i-control-the-alpha-of-a-geom-point-fill

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = stage, #fill = stage, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = stage), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  #facet_wrap(~year) + # orig
  facet_grid(stage~year, #scales = "free_y"
             ) +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.tt, x = year,
             color = stage, 
             #fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  #geom_point(aes(alpha = year), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  geom_point(#fill = NA, 
             size = 2) +
  labs(x = "year") +
  #facet_wrap(~trt.type) + # orig
  facet_grid(stage~trt.type, #scales = "free_y"
             ) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs reduce dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "development time / longevity (days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(21, 24, 22)
                     #c(16, 17, 15)
                     #c(1, 2, 0)
                     ) &
  scale_color_manual(values = c("#1B9E77", "#7570B3")) &
  scale_fill_manual(values = c("#1B9E77", "#7570B3")) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank()) #&
  #expand_limits(y = 37)
  #scale_alpha_manual(values = c(0.1,1)) &
  #ylim(c(2.5, 35))

```

# - mass

## - figs: by sex

```{r}
# NTs on pupal/adult mass

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = stage, #fill = stage, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = stage), size = 2) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(stage~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.mass, x = year,
             color = stage, #fill = stage, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(#aes(alpha = year), 
             size = 2) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_grid(stage~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  scale_fill_manual(values = c("#D95F02", "#7570B3")) &
  #scale_alpha_manual(values = c(0.1,1)) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank()) #&
  #ylim(c(0.75, 7.25))

```


# - dmass

## - figs: by sex

```{r}
# separated sexes
p1 <- ss_moredev %>% 
  filter(sex != "both" & 
           trt.type == "expt") %>%
  ggplot(aes(y = avg.dmass, x = minT,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.dmass - se.dmass, ymax = avg.dmass + se.dmass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(sex~year) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p1

p2 <- ss_moredev %>% 
  filter(sex != "both" & 
           trt.type == "ctrl") %>%
  ggplot(aes(y = avg.dmass, x = year,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.dmass - se.dmass, ymax = avg.dmass + se.dmass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "year") +
  facet_grid(sex~trt.type) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p2


p1 + p2 +
  labs(caption = "lower NTs x change in mass by sex") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass pupa - mass eclosion (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())
```

## - figs: both sexes

```{r}
# sexes combined
p1 <- ss_moredev %>% 
  filter(sex == "both" & 
           trt.type == "expt") %>%
  ggplot(aes(y = avg.dmass, x = minT,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.dmass - se.dmass, ymax = avg.dmass + se.dmass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(sex~year) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p1

p2 <- ss_moredev %>% 
  filter(sex == "both" & 
           trt.type == "ctrl") %>%
  ggplot(aes(y = avg.dmass, x = year,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.dmass - se.dmass, ymax = avg.dmass + se.dmass), width = 0.25) +
  geom_line(aes(lty = pop)) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "year") +
  facet_grid(~trt.type) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p2


p1 + p2 +
  labs(caption = "lower NTs x change in mass") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass pupa - mass eclosion (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())
```

# - pup growth rate 

## - figs: both sexes

```{r}
# sexes combined
p1 <- ss_moredev %>% 
  filter(sex == "both" & 
           trt.type == "expt") %>%
  ggplot(aes(y = avg.ratepup, x = minT,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.ratepup - se.ratepup, ymax = avg.ratepup + se.ratepup), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(sex~year) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p1

p2 <- ss_moredev %>% 
  filter(sex == "both" & 
           trt.type == "ctrl") %>%
  ggplot(aes(y = avg.ratepup, x = year,
             shape = sex, color = sex,
             group = interaction(pop, sex)
             )) +
  geom_errorbar(aes(ymin = avg.ratepup - se.ratepup, ymax = avg.ratepup + se.ratepup), width = 0.25) +
  geom_line(aes(lty = pop)) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  labs(x = "year") +
  facet_grid(~trt.type) +
  #scale_color_manual(values = aes_2col) +
  theme_bw()

p2


p1 + p2 +
  labs(caption = "lower NTs x pup growth rate") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "pup growth rate (g/days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())
```








