---
title: "ntw data comparisons"
date: "2024-10-08"
---

# 0. roadmap

goal: compare 2023 and 2024 ntw development data. needs some wrangling lol (esp 2023)

wip: combine fertility data...

# 1. load data & packages

```{r}
source("./helpers_ntw-compare.R")

test <- d23
#d23 <- test
```


# 2. wrangling dfs

standardising some data values and column names in `d23`

```{r}
# pick relevant data and rename columns/values to match 2024

# compare column names
setdiff(names(d24), names(d23)) # differences
setdiff(names(d23), names(d24))
intersect(names(d24), names(d23)) # same

# filter rows and rename columns
d23 <- d23 %>%
  filter(reason.ignore != "lost" | src != "F1" #| !is.na(final.fate)
           ) %>% 
  filter(trt.stage %in% c("260-hatch", "267-hatch", "330-hatch", "337-hatch", "419-hatch","433-hatch")) %>%
  filter(expt.group %in% c("A", "B", "C", "D", "E", "F", "H")) %>%
  rename(trt = treatment,
         id = ID,
         cohort = expt.group,
         notes.ignore = reason.ignore)

# standardise values
d23$trt[d23$trt == 337] <- 426

d23 <- d23 %>%
  mutate(final.fate = case_when(notes.ignore %in% c("hot larva", "cut", "culled larva", "wet diet") ~ "other",
                                TRUE ~ as.character(final.fate)))

# match values to d24
d23 <- d23 %>%
  mutate(sup = case_when(sup = 0 & !is.na(jdate.5th) ~ NA_real_,
                         TRUE ~ as.numeric(sup)),
         surv.outcome = case_when(!is.na(date.pupa) | !is.na(date.LP) ~ 0,
                                  final.fate == "pmd" ~ 1,
                                  final.fate == "other" ~ 2), # treat injuries the same as this
         flucT = case_when(flucT == 2.5 ~ 0,
                           TRUE ~ as.numeric(flucT)),
         trt.type = case_when(meanT == 26 & flucT == 0 ~ "ctrl",
                              TRUE ~ "expt"),
         trt = as.numeric(trt),
         #id = paste(id, year, sep = "-")
         )

test2 <- d23
#d23<-test2

# how to handle LPIs? (in 23 and 24?)

d23 <- d23 %>%
  select(-c("jdate.collected", "jdate.15", "jdate.stuck",
            "jdate.exit", "jdate.LP", "jdate.died")) %>% # d23 exclusives
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "instar.enter", starts_with(c("jdate", "mass")), 
           "sex",
           ends_with("T", ignore.case = FALSE), "trt.type",
           "sup", "surv.outcome", "year"
          ))
```


fine-tuning `d24`

```{r}
d24 <- d24 %>%
  #mutate(id = paste(id, year, sep = "-")) %>%
  select(-"jdate.culled") %>%
  filter(instar.enter == "hatch")
```



merge data

```{r}
all <- merge(d24, d23, all = TRUE
             ) %>%
  drop_na(id) %>%
  select(-c("instar.enter",
            ))

# drop things that died in 1 day
  # 2025-01-23: can i do this??????
all <- filter(all, jdate.pmd - jdate.enter > 1 | is.na(jdate.pmd - jdate.enter))

# add in exit dates to more easily parse out things still developing (for 2024 data)
# and other convenience things
all <- all %>% 
  mutate(
        # jdate.exit = case_when(!is.na(jdate.pmd) ~ jdate.pmd,
        #                         !is.na(jdate.pupa) ~ jdate.pupa,
        #                         TRUE ~ NA_integer_),
         minT = factor(minT, levels = c(19, 26, 33)),
         year = factor(year, levels = c(2023, 2024)),
         for.ntw = case_when(#year == 2023 & cohort %in% c("A", "B") ~ "N",
                             trt %in% c(260, 419, 426, 433) ~ "Y",
                             TRUE ~ "N"),
         for.const = case_when(#flucT %in% c(0, 7) & meanT %in% c(26, 33) ~ "Y",
                               trt %in% c(260, 267, 330, 426) ~ "Y",
                               TRUE ~ "N")
                               # trt.type == "ctrl" | (meanT == 33 & flucT %in% c(0, 7)) ~ "Y",
                               # TRUE ~ "N")
         )

```


# 3. df prep (math)

generic math

2025-01-20: filtering out diet wonkiness here lol. we can revisit it later if needed

```{r}
all_wide <- all %>%
  filter(!(pop == "lab" & diet == "TB")) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch,
         tt.8th = jdate.8th - jdate.hatch, # omit if only d23
         tt.wander = jdate.wander - jdate.hatch,
         tt.pupa = jdate.pupa - jdate.hatch,
         tt.eclose = jdate.eclose - jdate.pupa,
         tt.surv = jdate.surv - jdate.eclose
         )

all_long <- all_wide %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
  drop_na(jdate) %>%
  drop_na(tt) # drops NA's if an individual didnt reach a certain stage


# filtering by expt type
ntw_wide <- filter(all, for.ntw == "Y")
cvf_wide <- filter(all, for.const == "Y")
```

stage-specific summary stats

2025-01-20 troubleshooting:
- https://stackoverflow.com/questions/73533852/create-multiple-dataframes-based-on-the-dates-of-another-dataframe-in-r
- https://stackoverflow.com/a/24376207/3358227


```{r}
# apply to data
ss_all <- lapply(list(ntw_wide, cvf_wide), calc.devstats)

ss_ntw <- Reduce(full_join, ss_all[[1]]) %>% 
  calc.ssadj()

ss_cvf <- Reduce(full_join, ss_all[[2]]) %>%
  calc.ssadj()
```


# 4. NT figs!

larval "#1B9E77" 
pup "#D95F02" 
adult "#7570B3"

## - survival

// todo
2025-01-23: 
- make these functions also LOL

```{r}
# NTs on pupal survival 

p1 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "expt") %>% 
  ggplot(aes(y = prop.survpup, x = minT,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = prop.survpup, x = year,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs increase survival to pup?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       y = "proportion survived to pupation") &
  guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16)) &
  scale_color_manual(values = "#1B9E77") &
  ylim(c(0, 1))

```
## - dev time/longevity

```{r}
# NTs on development time/longevity
# see this
  # https://stackoverflow.com/questions/78193007/how-do-i-control-the-alpha-of-a-geom-point-fill

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.tt, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs reduce dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "development time / longevity (days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(21, 24, 22)
                     #c(16, 17, 15)
                     #c(1, 2, 0)
                     ) &
  scale_color_manual(values = c("#1B9E77", "#7570B3")) &
  scale_fill_manual(values = c("#1B9E77", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(5, 35))

```

## - mass

```{r}
# NTs on pupal/adult mass

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.mass, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  scale_fill_manual(values = c("#D95F02", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(1.1, 7.25))

```

# 5. const figs

## - survival

```{r}
# NTs on pupal survival 

p1 <- ss_cvf %>% 
  filter(stage == "la") %>% View()
  ggplot(aes(y = prop.survpup, x = trt.type,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type)
             )) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1
```

```{r}
p2 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = prop.survpup, x = year,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs increase survival to pup?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       y = "proportion survived to pupation") &
  guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16)) &
  scale_color_manual(values = "#1B9E77") &
  ylim(c(0, 1))

```
## - dev time/longevity

```{r}
# NTs on development time/longevity
# see this
  # https://stackoverflow.com/questions/78193007/how-do-i-control-the-alpha-of-a-geom-point-fill

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "ctrl") %>% 
  ggplot(aes(y = avg.tt, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs reduce dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "development time / longevity (days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(21, 24, 22)
                     #c(16, 17, 15)
                     #c(1, 2, 0)
                     ) &
  scale_color_manual(values = c("#1B9E77", "#7570B3")) &
  scale_fill_manual(values = c("#1B9E77", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(5, 35))

```

## - mass

```{r}
# NTs on pupal/adult mass

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  ggplot(aes(y = avg.mass, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  scale_fill_manual(values = c("#D95F02", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(1.1, 7.25))

```





# zzz archive


## 4. entsoc figs

archived 2025-01-20


summary stats

// todo 
- 2024-10-17: keeping bugs separate by year for now, will see how they look separate before trying to combine
- 2025-01-20: filter out bugs on their correct diet types only for now

```{r}
# sep by sups (bc the sups have their own dev trajectory...)
ss_wide.sup <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT, sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) # `minT` is ok here

# all dev trajectories together
ss_wide <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT,
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) #%>% 
  #mutate_all(~ifelse(is.nan(.), NA, .)) # breaks `minT`

```


prereqs: a cleaned up `d23`

- stats

```{r}
# slapping sth on for entsocing
ss_larva <- d23 %>%
  filter(pop == "lab") %>%
  group_by(year, pop, diet, trt.type, minT, 
           #sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n.all - n.pmd, 
            prop.survpup = round(1-(n.pmd/n.all), digits=2)
            ) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  mutate(minT = factor(minT, levels = c("19", "26", "33"))) %>%
  filter(!(pop == "lab" & diet == "TB") &
           trt.type == "expt")
```




- figs


*notes:* 

- use lab bugs on LD + field on TB only!!!! (this is why u cant just reuse ur LB figs lol)
- 2025-01-20: below figs assume all combos of diet types and pops are included in the `all` df.

```{r}
# (messing around w 2023 data only for entsoc,,)
# using LB code lol

# survival
p1 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = prop.survpup, x = minT, shape = pop, color = pop,
             )) +
  geom_point(size=3) + 
  geom_line(aes(lty = pop, group = pop)) +
  geom_text(aes(label = n.all), hjust = -0.5, show.legend = FALSE) +
  theme_bw() +
  labs(title = "23v24 ntw pupal survival",
       y = "proportion survived to pupation", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  ylim(c(0, 1))

p1
```

```{r}
# avg tt pup
pd <- position_dodge(0.3)

p2 <-
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.tt, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3, #position = pd
             ) +  
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop), #position = pd
            ) +
  theme_bw() +
  ylim(c(18, 34)) +
  labs(title = "23v24 ntw larval development",
       y = "time to pupation (days)", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year)

p2
```

```{r}
# avg mass
p3 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.mass, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3) +  
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop)) +
  theme_bw() +
  #ylim(c(1750, 4750)) +
  labs(title = "23v24 ntw pupal mass",
       y = "mass at pupation (mg)", 
       x = "minimum temperature (°C)",
       #caption = "2023 data only; lab+LD and field+TB"
       ) +
  facet_wrap(~year)

p3
```

```{r}
# merged plot
p1 + p2 + p3 +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(1:1)) &
  theme(legend.position = "top") &
  labs(lty = "population", shape = "population",
       title = NULL) &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16))
  #guides(shape = "none", lty = "none")
```




