---
title: "2024 dev analyses"
date: "2024-07-16"
---

# roadmap & loading

## - **questions**

- (pilot) what is the average time spent in each instar after transferring to constant temps?
  - where do bugs "line up" in their dev at 4th and 5th?
- does getting stuck make a difference on survival?
- are the avg growth outcomes different if you keep the survivors vs not?
- (ntw) how 2024 ctrls compare to 2023 ctrls?


```{r message = FALSE}
source("./helpers_ox.R") # need run cleaning.Rmd beforehand if gsheet has been updated

# INCASE: bring this back if u need it and are breaking things
  # library(tidyverse)
  # conflicted::conflicts_prefer(dplyr::filter)
  # all_wide <- read.csv("~/Documents/repos/ntw/2024/data/data-ox.csv", header = TRUE)
```

# july pilot 

## = controls




## = time to dev stages

2024-09-12: taking out the 6ths bc there's just not rly enough of them to do anything interesting w/ and they mess things up lol

### - wrangling

```{r}
# doing this separately from the helpers stuff bc this is more specific to the pilot lol
trttime_wide <-  wide_all %>%
  filter(sup == 0 | is.na(sup)) %>%
  mutate(t3.3rd = jdate.3rd - jdate.3rd,
         t3.4th = jdate.4th - jdate.3rd,
         t3.5th = jdate.5th - jdate.3rd,
         t3.6th = jdate.6th - jdate.3rd,
         t3.wander = jdate.wander - jdate.3rd,
         t3.died = jdate.pmd - jdate.3rd,
         tt.intrt = case_when(is.na(jdate.pmd) ~ jdate.pupa - jdate.3rd,
                              TRUE ~ jdate.pmd - jdate.3rd)
         )

# stats of major dev pts
trttime_long <- trttime_wide %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "t3")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)"))
```


### - stats

2024-09-12: need to take handle the culled somehow

```{r}
# includes pmd
trttime_lstats <- trttime_long %>%
  #filter(is.pmd == 0) %>%
  filter(instar %in% c("3rd", "4th", "5th", "6th", "wander")) %>%
  group_by(labs.trt, instar) %>%
  mutate(t3h = t3*24) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.t3 = mean(na.omit(t3)),
            se.t3 = sd(na.omit(t3))/sqrt(length(na.omit(t3))),
            avg.t3h = mean(na.omit(t3h)),
            se.t3h = sd(na.omit(t3h))/sqrt(length(na.omit(t3h))),
            n=n())
```

### - plot

TODO: 

- pull out 6ths
- reorder trts?
- how to handle the culled bugs

```{r}
# by hours
trttime_lstats %>%
  ggplot(aes(y = labs.trt, x = avg.t3h, shape = instar, color = labs.trt)) +
  geom_point(size = 2) +
  geom_line(aes(group = labs.trt, lty = labs.trt), lwd = 0.8, show.legend = FALSE) +
  geom_errorbarh(aes(xmin = avg.t3h - se.t3h, xmax = avg.t3h + se.t3h), height = 0.15) +
  geom_text(aes(label = n, vjust = -1.5),
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(minor_breaks = seq(0, 250, by = 24),
                     breaks = seq(0, 400, by = 48)) +
  theme_bw() +
  labs(title = "avg instar times (hrs)",
       y = "treatment", x = "hrs since 3rd",
       color = "treatment")

# by days
trttime_lstats %>%
  filter(labs.trt != "38±0°C") %>%
  ggplot(aes(y = labs.trt, x = avg.t3, shape = instar, color = labs.trt)) +
  geom_point(size = 2) +
  geom_line(aes(group = labs.trt, lty = labs.trt), lwd = 0.8, show.legend = FALSE) +
  geom_errorbarh(aes(xmin = avg.t3 - se.t3, xmax = avg.t3 + se.t3), height = 0.15) +
  geom_text(aes(label = n, vjust = -1.5),
            check_overlap = TRUE, show.legend = FALSE) +
  scale_x_continuous(breaks = seq(0, 18, by = 2)) +
  theme_bw() +
  labs(title = "avg instar times (days)",
       y = "treatment", x = "days since 3rd",
       color = "treatment")
```

## = fine-scale development

want to see how mass changes day-by-day in the later instars up to wandering

### - stats

```{r}
fine_stats <- long_all %>%
  filter(instar %in% c("3rd", "4th", "5th", "6th", "wander")) %>%
  group_by(labs.trt, #instar, # does nawt work w instar lol
           t3) %>%
  mutate(logmass = log(na.omit(mass))) %>% 
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

majortimes_stats <- trttime_long %>%
  filter(instar %in% c("4th", "5th", "6th", "wander")) %>%
  group_by(instar) %>%
  summarise(avg.t3 = mean(na.omit(t3)))
```

### - plot

TODO:

- need to split the 6ths from the rest bc they have different dev trajectories

```{r}
ggplot(fine_stats, aes(y = avg.logmass, x = t3, color = labs.trt, #shape = instar
                       )) +
  geom_point() + geom_line(lwd = 0.8) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass,
                    width = 0.5)) +
  # geom_line(data = long_all, 
  #           aes(y = log(mass), group = interaction(id, labs.trt)),
  #           alpha = 0.25) +
  geom_vline(majortimes_stats, mapping = aes(xintercept = avg.t3, lty = instar), # meh could be better
             #show.legend = FALSE
             ) + 
  #facet_wrap(~labs.trt) +
  theme_bw() +
  labs(x = "days since 3rd", y = "avg log(mass (mg))",  
       color = "treatment")
```



