---
title: "2024 ntw dev data cleaning"
date: "2024-09-23"
---

# 0. roadmap

1. load data & packages; pre-filtering
2. column formatting + rearrangement
3. save output into `2024/data/`

*notes:* this is mostly copied over from `cleaning_ox.Rmd` so some comments/notes have been taken out if u compare versions

# 1. load data & packages

```{r message = FALSE}

##### if doing a major update of the sheet, archive the old one first

library(tidyr)
library(purrr)
library(dplyr)
library(lubridate) # for handling time stuff

library(googlesheets4)
```

googlesheets import

```{r}
data <- read_sheet("https://docs.google.com/spreadsheets/d/1TOUsH7zNT6jF7SltYsD_8pK5tI7yR-Gqx-kDrodmJ00/edit?gid=1256964024", sheet = "ntw repeat", col_types = "c")

incase <- data
data <- incase
```

pre-filtering/value fixing

```{r}
# filtering

data <- data %>%
  dplyr::filter(!(notes.ignore %in% c("early pmd", "lost"))) # dropping things that died in 1 day or got lost

data <- select(data, -c("date.LPI", "date.15")) # not rly needed for analyses/LPI is empty lol


# value fixing

data[, -ncol(data)] <- lapply(data[, -ncol(data)], gsub, pattern="\\?", replacement = "") 
  # removes `?`s from all columns excepts notes column (assuming its the last col in the df)

data[, 'notes'] <- lapply(data[, 'notes'], gsub, pattern=",", replacement = ";") 
  # replaces commas w/ ;'s to avoid csv breaking

data[, 'id'] <- lapply(data[, 'id'], gsub, pattern="\\.\\d", replacement = "") 
  # turns the repeat entries into whole numbers so theres no decimals going on after as.numeric(id)

```

column type standardising

```{r}
fix.columns <- function(data) {
  cleaned_data <- data %>%
    mutate(across(starts_with(c("date.", "stuck.")),  as.Date, format = "%m/%d"))%>%
    mutate(time.enter = lubridate::hm(time.enter)) %>%
    mutate(across(starts_with("mass."), as.numeric)) %>%
    mutate(treatment = as.numeric(treatment),
           id = as.numeric(id)) %>%
    rename(trt = treatment,
           pop = population)
  
  return(cleaned_data)
}

data <- fix.columns(data)
```


# 2. column reformatting + creation

## - impt ones

julian date conversions

```{r}
data <- data %>%
  mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
  mutate(across(starts_with("jdate."), as.numeric))
```

treating field 1sts as hatchlings

```{r}
# data$jdate.hatch[(data$pop == "field" & data$instar.enter == "1st")] <- as.numeric(data$jdate.collected - 1)
  # assume they hatched the day before collecting ### this is breaking as of 2024-10-08

data <- data %>%
  mutate(jdate.hatch = case_when(pop == "field" & instar.enter == "1st" ~ as.numeric(jdate.collected - 1),
                                 pop == "field" & instar.enter == "hatch" & is.na(jdate.hatch) ~ as.numeric(jdate.collected),
                                 TRUE ~ as.numeric(jdate.hatch)))
  # assume hatch date for late/early field hatchlings

# data %>% filter(instar.enter %in% c("1st", "hatch") & pop == "field") %>% select(c("id", "instar.enter", "jdate.collected", "jdate.hatch")) %>% View() # troubleshooting

data$instar.enter[data$instar.enter == "1st"] <- "hatch" 
  # change instar.enter


```


adding in temperature treatment info

```{r}
data <- data %>%
  mutate(meanT = case_when(trt == 260 ~ 26,
                           trt == 419 ~ 29.5,
                           trt == 426 ~ 33,
                           trt == 433 ~ 36.5),
         flucT = case_when(trt == 260 ~ 0,
                           trt == 419 ~ 10.5,
                           trt == 426 ~ 7,
                           trt == 433 ~ 3.5),
         maxT = case_when(trt == 260 ~ 26,
                          TRUE ~ 40),
         minT = case_when(trt == 260 | trt == 426 ~ 26,
                          trt == 419 ~ 19,
                          trt == 433 ~ 33),
         trt.type = case_when(trt == 260 ~ "ctrl",
                              TRUE ~ "expt")
         )
```


## - useful ones

parse out dev trajectories

// TODO: 2024-09-23 revisit the eclosion thing w/ a dedicated `fate` column

```{r}
data <- data %>%
  mutate(sup = case_when(!is.na(date.8th) ~ 8,
                         !is.na(date.7th) ~ 7,
                         !is.na(date.6th) ~ 6,
                         (is.na(date.pmd) & !is.na(date.pupa)) ~ 0)) 

```


survival outcomes

```{r}
# 0 = pupated, 1 = pmd, 2 = died of other reason (i.e. injury)

data <- data %>%
  mutate(surv.outcome = case_when(!is.na(date.pmd) & notes.ignore == "injured" ~ 2,
                                  !is.na(date.pmd) ~ 1,
                                  !is.na(date.pupa) ~ 0))
```




## - fun ones

add in stuck indicators per instar

```{r}
# 0 if no, 1 if yes

data <- data %>%
  mutate(is.stuck3rd = case_when(is.na(stuck.3rd) ~ 0, TRUE ~ 1),
         is.stuck4th = case_when(is.na(stuck.4th) ~ 0, TRUE ~ 1),
         is.stuck5th = case_when(is.na(stuck.5th) ~ 0, TRUE ~ 1),
         is.stuck6th = case_when(is.na(stuck.6th) ~ 0, TRUE ~ 1),
         is.stuck7th = case_when(is.na(stuck.7th) ~ 0, TRUE ~ 1))

```


# 3. saving output & tidying up

maybe keep?
  - instar.enter (idk bc they're all sposed to start at hatching lol)
  - notes.ignore


```{r}
data$time.enter <- as.character(data$time.enter) 
  # breaks write_csv if left as a `lubridate` object

today <- format(Sys.time(), "%y%m%d")

# wordier notes/less-informative columns
notes <- data %>%
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "species", "location", "date.collected", # collection/demog info
           "instar.enter", "time.enter", # prob dont need
           starts_with(c("date.", "h.")), contains(c("stuck", "notes")), # dev stuff that is jdated
           "sup", "surv.outcome"
           )) #%>% View()

write.csv(notes, file = "~/Documents/repos/ntw/2024/data/ntw-notes.csv", row.names = FALSE)
write.csv(notes, file = paste0("~/Documents/repos/ntw/2024/data/archive/", today, "_ntw-notes.csv" ), row.names = FALSE)
  
# MQs only
mqs <- data %>%
  filter(species == "MQ")

write.csv(mqs, file = "~/Documents/repos/ntw/2024/data/ntw-MQ.csv", row.names = FALSE)
write.csv(mqs, file = paste0("~/Documents/repos/ntw/2024/data/archive/", today, "_ntw-MQ.csv" ), row.names = FALSE)

  
# useful columns only
export <- data %>%
  filter(is.na(species)) %>%
  select(-"jdate.collected") %>%
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "instar.enter", starts_with(c("jdate", "mass")), 
           "sex",
           ends_with("T", ignore.case = FALSE), "trt.type",
           "sup", "surv.outcome"
          )) #%>% View()

write.csv(export, file = "~/Documents/repos/ntw/2024/data/ntw.csv", row.names = FALSE)
write.csv(export, file = paste0("~/Documents/repos/ntw/2024/data/archive/", today, "_ntw.csv" ), row.names = FALSE)

```

remove troubleshooting objects

```{r}
#rm(incase, data, fix.columns)
rm(today, notes, mqs, export)
```



