---
title: "fertility analyses"
date: "2024-02-15"
---

purpose: looking at fertility stats

*roadmap:*
0. load data, pckgs
1. ...

*questions:*
- fertility
  - overall hatch rates per trt/tent? factors contributing to hatch rates?
  - which trts lead to hatching?
  - controls: how do expt compare to controls?
  - daily avg hatch rate
  - running avg of egg production
  - time btwn coll vs hatching?
  - sex diffs?

- longevity
  - time to ec? what are contributing factors?
  - survival post ec? what factors? (affected by time in 4C?)

# 0. load data, helper fns
```{r}
source("./helpers_tents.R")
# map(other_helpers, ~discard(., .p = ~all))
  # working on it: https://stackoverflow.com/questions/63533428/in-r-remove-list-of-list-based-on-name
```
# 1. analyses hodgepodge

## - overall hatch rates
```{r}
hrate_summ <- data_tstats %>%
  group_by(#id.tent,
           trt.f, trt.m, pop
           ) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.collected = sum(n.coll, na.rm = TRUE),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            #start = first(jdate), end = last(jdate),
            t.duration = max(jdate) - min(jdate)
            ) %>%
  drop_na(trt.m) # drop tents w/o males

```

