---
title: "acc_analyses"
date: "2023-12-07"
---

# 0.1 load data & packages
```{r message = FALSE}
source("./helpers_ntw.R")
rm(list=c(temps_helpers, NT_helpers))
```

# 1.0 super simple stats
```{r}
# acc stats
wide_all %>%
  filter.acc2() %>% filter( pop == "lab") %>%
  group_by(trt.stage) %>%
  summarise(n=n(),
            pct.pmd = round(sum(final.fate == "pmd")/n*100, digits = 1),
            #pct.slow = round(sum(ignore.reason == "slow")/n*100, digits = 1),
            pct.sup = round(sum(if.sup == "Y")/n*100, digits = 1),
            pct.6th = round(sum(sup == 6)/sum(if.sup == "Y")*100, digits = 1),
            pct.7th = round(sum(sup == 7)/sum(if.sup == "Y")*100, digits = 1))
            #n.pmd = sum(final.fate == "pmd"))
```

## - sup barplots
```{r}
# make these into barplots instead

wide_all %>%
  filter.acc2() %>% filter(pop == "lab") %>%
  mutate(sup = factor(sup)) %>%
  ggplot(aes(x = trt.stage, fill = sup)) +
  geom_bar(stat = "count", position = position_fill(reverse = TRUE)) +
  labs(title = "acc: supernumerary fates", fill = "supernumerary stage", x = "treatment", y = "percent") +
  scale_x_discrete(labels = acc_labels) +
  theme_bw() + scale_fill_brewer(palette = "Greens")

```

# 2.0 (old) growth / surv stuff

## overall growth trends; end points

240201: just for ref rn

compare overall development of all individuals compared to average growth

<!-- ## - set up the df (**) -->
```{r}
# pull out the lab bugs
long_lf <- long_all %>% filter(pop == "lab" | pop == "field") 

# individual dev stats
dev_L <- long_lf %>%
  filter(!(final.fate == "pmd" | final.fate == "misc")) %>%
  filter.ins.topup() %>%
  mutate(logmass = log(mass))

devsumm_L <- dev_L %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

# for pulling out supernumeraries.. need to tweak
# devsumm_Lsups <- dev_L %>%
#   group_by(pop, sup, trt.stage, instar) %>%
#   summarise(avg.mass = mean(na.omit(mass)),
#             se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
#             avg.tt = mean(na.omit(tt)),
#             se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
#             avg.logmass = mean(na.omit(logmass)),
#             se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
#             n=n())

# see 'plot NTs' under 'overall growth'
  
```

<!-- ## - plot acc -->
```{r}
acc_dev <- dev_L %>% filter(pop == "lab") %>% filter.acc2()
#acc_devsumm <- devsumm_L %>% filter(pop == "lab") %>% filter.acc2()
acc_devsumm <- dev_L %>% filter.acc2() %>% calc.devsumm.trtstg()

acc_devsumm %>%
  filter(instar != "pupa") %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line() +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=(acc_dev %>% filter(instar != "pupa")), aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  acc_aes() +
  labs(title = "accumulation: average development from 3rd - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "lab, no pmd, at 26 +/- 0 until enter trt")

acc_devP <- acc_dev %>% filter(instar == "pupa")
acc_devsummP <- dev_L %>%
  filter.acc2() %>%
  group_by(trt.stage, sex, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt)))) %>% 
  filter(instar == "pupa")

acc_devsummP %>%
  ggplot(aes(x=avg.tt, y=avg.mass, color=trt.stage)) +
  geom_point(size = 2) +
  geom_point(data=acc_devP, aes(x=tt, y=mass, color=trt.stage), alpha = 0.275) +
  y_err_mass(0.5) + x_err_tt(1) +
  acc_aes() + facet_wrap(~sex) +
  labs(title = "accumulation: average development to pupa", color = "treatment", y = "mass (mg)", x = "days to instar")
```

## survival to pupation

<!-- ## - censoring + subsetting -->
```{r}
# censor data
surv_P <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

acc_surv <- surv_P %>% filter.acc2()
```

<!-- ## - create survival objects and fits -->
```{r}
acc_sobj <- Surv(acc_surv$timeto.exit, acc_surv$censor)
acc_sfit <- survfit(acc_sobj ~ trt.stage, data = acc_surv)
```

<!-- ## - plot survival curves -->
```{r}
ggsurvplot(fit = acc_sfit, data=acc_surv,
           conf.int = T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "accumulation: survival from hatch-pup", 
           palette = c("#00C2D1", "#F3922B", "#C23C1E", "#710A36"),
           legend.title = "treatment-stage entered", 
           legend.labs = c("26°C @ hatch", "337-hatch"="33±7°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th" )
           ) +
  labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")

```

## cox-ph survival stuff?

- check as variables: lab/TB diet; if.sup; if.stuck

<!-- ##  - if sup x to pup/eclosion -->

```{r}
#check_suppup <- wide_all %>% filter(if.pupa == "Y")

#mod_suppup <- lm(if.pupa ~ if.sup + treatment + pop, data = check_suppup, na.action=na.omit)
  # does this not work bc Y needs to be a # LOL

# i think this shoudl be part of survival stuff

```

# 3.0 redoing growth stats

per the new analyses. aka im just copying over all the code from `temps_analyses` and replacing stuff where needed (just doing this for completion lol)

## a. calc stats
```{r}
# filter out pupal stuff & the correct bugs lol
acc_all <- wide_all %>%
  ## relevel instars
  mutate(instar.intrt = case_when(instar.enter == "hatch" ~ 5,
                                  instar.enter == "3rd" ~ 3,
                                  instar.enter == "4th" ~ 2)) %>%
  filter.acc2() %>%
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled") %>%
  ## code survival and labeling
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0), 
         trt = case_when(treatment == 260 ~ "control (26±0°C)",
                           treatment == 337 ~ "exptal (33±7°C)")) %>%
  ## add sample sizes
  group_by(trt, instar.enter) %>%
  mutate(n = n(),
         n_pmd = sum(final.fate == "pmd"),
         n_surv = n - n_pmd) %>%
  ungroup()

# calculate the pup stats we're interested in
summary_acc <- acc_all %>%
  group_by(treatment, trt, instar.enter, instar.intrt) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            n_surv = n_all - n_pmd,
            prop.survpup = round(1-(n_pmd/n_all), digits=2))
```

## b. plot
```{r mass and time together}
# visualise time to pup, pupal mass across expt groups
summary_acc %>% ggplot(aes(y = avg.mass, x = avg.tt, color = as.factor(instar.intrt), shape = trt)) +
  geom_point(size = 2.5) +
  y_err_mass(err = 0.1) + x_err_tt(err = 100) + theme_bw() +
  #scale_color_manual(values = c("#619Cff", "#00BA38", "#F8766D")) +
  scale_color_manual(values = RYB) +
  labs(title = "acc: avg time to and mass at pupation", 
       y = "average pupal mass (mg)", 
       x = "average time to pupa from hatching (days)", 
       caption = "all lab; trt = 33±7°C, 6 hr soaks",
       color = "instars spent in treatment", shape = "treatment")
```

generally, control conditions = biggest and happiest bugs (big and fast). exptal trts are all slower and smaller. the less time spent at exptl trt, the biiiiger they are. time is generally about the same tho (not quite a hatch > 3rd > 4th tho, interestingly for the time. maybe bc hatch has a lot of time to acclimate and 4th is too much of a shock?)

that said, pup time still isn't that much bigger (only about 2 more days), considering how ~stressful~ these conds are... (aka the bugs wre still healthy atp lol)

```{r plot 1 response var at a time}
# avg pup mass due to acc
summary_acc %>% ggplot(aes(y = avg.mass, x = as.factor(instar.intrt), shape = as.factor(trt))) +
  geom_point(size = 2) + geom_text(aes(label = n_surv), vjust = -1.1, hjust = -0.5) +
  geom_line(aes(lty = trt, group = trt)) +
  y_err_mass(err = 0.04) + theme_bw() +
  labs(title = "acc: avg mass at pupation", 
       y = "average pupal mass (mg)", 
       x = "instars spent in treatment", 
       caption = "all lab",
       shape = "treatment", lty = "treatment")

# avg tt pupa due to acc
summary_acc %>% ggplot(aes(y = avg.tt, x=as.factor(instar.intrt), shape = as.factor(trt))) +
  geom_point(size = 2) + geom_text(aes(label = n_surv), vjust = -1.1, hjust = -0.5) +
  geom_line(aes(lty = trt, group = trt)) +
  y_err_tt(err = 0.04) + theme_bw() +
  labs(title = "acc: avg time to pupation", 
       y = "avg time to pupation (days)", 
       x = "instars spent in treatment", 
       caption = "all lab",
       shape = "treatment", lty = "treatment")

# avg surv due to acc
summary_acc %>% 
  filter(treatment == 337) %>%
  ggplot(aes(y = prop.survpup, x = as.factor(instar.intrt))) +
  geom_point(size = 2) + 
  geom_text(aes(label = n_all), vjust = -1, hjust = -1.1, check_overlap = TRUE) +
  geom_line(aes(lty = trt, group = trt), lty = "dashed") +
  theme_bw() + 
  #geom_blank(aes(y=1.0095)) +
  # expand_limits(y = c(0.87, 1.025)) +
  labs(title = "acc: proportion survived to pupation", 
       y = "prop. survived to pupation", 
       x = "instars spent in treatment", 
       caption = "all lab; ctrl (n=16) all survived",
       shape = "treatment", lty = "treatment") +
  ylim(c(0.7, 1.05))
```

```{r KM curves}
# fit models
kmfit.trttime <- survfit(Surv(acc_all$tt.trt, acc_all$status) ~ trt.stage, data = acc_all)
kmfit.devtime <- survfit(Surv(acc_all$tt.exit, acc_all$status) ~ trt.stage, data = acc_all)

# base plot
ggsurvplot(fit = kmfit.trttime, data = acc_all,
           conf.int = TRUE,
           palette = RYB,
           xlab = "days since entering treatment")

ggsurvplot(fit = kmfit.devtime, data = acc_all,
           conf.int = TRUE,
           palette = RYB,
           xlab = "days since hatching")

```

# 4.0 models,, ?

240201: gunna skip modeling for now lolll




# followups

1. 240527: looking for the avg length of time in an instar at diff trts

```{r}
acc_dev %>%
  filter(pop == "lab",
         #trt.stage %in% c("260-hatch", "330-hatch"),
         !(instar %in% c("pupa"))
         ) %>%
  mutate(tthrs = 24*tt) %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.ttdays = mean(na.omit(tt)),
            se.ttdays = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
            avg.tthrs = mean(na.omit(tthrs)),
            se.tthrs = sd(na.omit(tthrs))/sqrt(length(na.omit(tthrs))),
            n=n()) %>% #View()
  # mutate(trt.stage = factor(trt.stage, level = c("337-hatch", "260-hatch", "337-3rd", "337-4th")),
  #        trt.stage = case_when(trt.stage == "260-hatch" ~ "26±0°C @hatch",
  #                              trt.stage == "337-hatch" ~ "33±7°C @hatch",
  #                              trt.stage == "337-3rd" ~ "33±7°C @3rd",
  #                              trt.stage == "337-4th" ~ "33±7°C @4th")) %>%
  ggplot(aes(y=trt.stage, x = avg.tthrs, color = trt.stage)) +
  geom_line() + geom_point(aes(shape=instar), size = 3) +
  labs(y = "treatment", x = "avg hrs post-hatch", title = "instar duration (trt @ various instars) - lab") +
  xlim(c(65, 375))
```





