---
title: "temps_analyses"
date: "2023-12-07"
---

# 0. load data & packages
```{r}
source("./helpers_ntw.R")
rm(list=c(acc_helpers, NT_helpers))
```

# 1. pct sup
```{r}
wide_all %>%
  filter.temps2() %>% filter(pop == "lab") %>%
  group_by(treatment) %>%
  summarise(n=n(),
            pct.pmd = round(sum(final.fate == "pmd")/n*100, digits = 1),
            #pct.slow = round(sum(ignore.reason == "slow")/n*100, digits = 1),
            pct.sup = round(sum(if.sup == "Y")/n*100, digits = 1),
            pct.6th = round(sum(sup == 6)/sum(if.sup == "Y")*100, digits = 1),
            pct.7th = round(sum(sup == 7)/sum(if.sup == "Y")*100, digits = 1))
            #n.pmd = sum(final.fate == "pmd"),)

```

## a. barplots
```{r}
# make these into barplots instead

wide_all %>%
  filter.temps2() %>% filter(pop == "lab") %>%
  mutate(sup = factor(sup)) %>%
  ggplot(aes(x = trt.stage, fill = sup)) +
  geom_bar(stat = "count", position = position_fill(reverse = TRUE)) +
  labs(title = "temps: supernumerary fates", fill = "supernumerary stage", x = "treatment", y = "proportion supernumerary") +
  scale_x_discrete(labels = temp_labels) +
  theme_bw() + scale_fill_brewer(palette = "Greens")
```

# *** lowkey ignore ***

i sorta dislike everything under here so we're kinda ignoring it rn (maybe picking + choosing tho)

# overall growth trends; end points

compare overall development of all individuals compared to average growth

<!-- ## - set up the df (**) -->
```{r}
# pull out the lab bugs
long_lf <- long_all %>% filter(pop == "lab" | pop == "field") 

# individual dev stats
dev_L <- long_lf %>%
  filter(!(final.fate == "pmd" | final.fate == "misc")) %>%
  filter.ins.topup() %>%
  mutate(logmass = log(mass))

devsumm_L <- dev_L %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

# for pulling out supernumeraries.. need to tweak
# devsumm_Lsups <- dev_L %>%
#   group_by(pop, sup, trt.stage, instar) %>%
#   summarise(avg.mass = mean(na.omit(mass)),
#             se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
#             avg.tt = mean(na.omit(tt)),
#             se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
#             avg.logmass = mean(na.omit(logmass)),
#             se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
#             n=n())

# see 'plot NTs' under 'overall growth'
  
```

<!-- ## - plot temps -->
```{r}
temps_dev <- dev_L %>% filter(pop == "lab") %>% filter.temps2()
#temps_devsumm <- devsumm_L %>% filter(pop == "lab") %>% filter.temps2()
temps_devsumm <- dev_L %>% filter.temps2() %>% calc.devsumm.trtstg()

temps_devsumm %>%
  filter(instar != "pupa") %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line() +
  geom_point(data=(temps_dev %>% filter(instar != "pupa")), aes(x=tt, y=logmass, color=trt.stage), alpha = 0.15) + 
  geom_point(aes(shape = instar), size = 3) +
  #y_err_logmass(0.5) + x_err_tt(1) +
  temp_aes() +
  labs(title = "temps: average development from 3rd - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "lab, no pmd, enter at hatch")

temps_devP <- temps_dev %>% filter(instar == "pupa")
  # tbh can drop lol

temps_devsummP <- dev_L %>%
  filter.temps2() %>%
  group_by(trt.stage, sex, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt)))) %>% 
  filter(instar == "pupa")

temps_devsummP %>%
  ggplot(aes(x=avg.tt, y=avg.mass, color=trt.stage)) +
  geom_point(size = 2) +
  geom_point(data=temps_devP, aes(x=tt, y=mass, color=trt.stage), alpha = 0.275) +
  y_err_mass(0.5) + x_err_tt(1) +
  temp_aes() + facet_wrap(~sex) +
  labs(title = "temps: average development to pupa", color = "treatment", y = "mass (mg)", x = "days to instar")


### plot with response on the y (for the pupated only)
# survival vs mean/fluc T
temps_dev
  


# dev rate vs mean/fluc T
temps_devP %>%
  mutate(devrate = 1/tt) %>%
  group_by(meanT, flucT, sex) %>%
  summarise(avg.devrate = mean(devrate)) %>%
  ggplot(aes(y=avg.devrate, x = meanT, shape = as.factor(flucT), lty = as.factor(flucT))) +
  geom_point() + geom_line() +
  facet_wrap(~sex) + theme_bw() +
  labs(y = "average development rate (to pupa)", x = "mean temperature (°C)", shape = "fluctuation (°C)", lty = "fluctuation (°C)")
```

# survival to pupation

<!-- ## - censoring + subsetting -->
```{r}
# censor data
surv_P <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

temps_surv <- surv_P %>% filter.temps2()
```

<!-- ## - create survival objects and fits -->
```{r}
temps_sobj <- Surv(temps_surv$timeto.exit, temps_surv$censor)
temps_sfit <- survfit(temps_sobj ~ treatment, data = temps_surv)
```

<!-- ## - plot survival curves -->
```{r}
ggsurvplot(fit = temps_sfit, data=temps_surv,
           conf.int = T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp: survival from hatch-pup",
           #palette = temp_colors,
           palette = c("#00C2D1","#1929B3", "#F9C639", "#710A36"),
           legend.title = "treatment", 
           legend.labs = temp_labels
           ) +
  labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")
```

# cox-ph survival stuff?

- check as variables: lab/TB diet; if.sup; if.stuck

<!-- ##  - if sup x to pup/eclosion -->

```{r}
#check_suppup <- wide_all %>% filter(if.pupa == "Y")

#mod_suppup <- lm(if.pupa ~ if.sup + treatment + pop, data = check_suppup, na.action=na.omit)
  # does this not work bc Y needs to be a # LOL

# i think this shoudl be part of survival stuff

```

<!-- # x=trt, y = response graphs -->

```{r}
# subset data
temps_all <- wide_all %>% filter.temps2()

#ggplot()

```


** (ok restarting here lol) **

(this mostly takes the code from `overall_controls.Rmd` LOL)

# 2. visualise growth, dev, surv

## a. calc stats
```{r calculate stats}
# filter out pupal stuff & the correct bugs lol
temps_all <- wide_all %>%
  filter.temps2() %>%
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled") %>%
  mutate(tt.pupa = jdate.pupa - jdate.hatch)

# calculate the pup stats we're interested in
summary_temps <- temps_all %>%
  group_by(meanT, flucT) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            n_surv = n_all - n_pmd, 
            prop.survpup = round(1-(n_pmd/n_all), digits=2))
  # mutate(meanT = as.factor(meanT),
  #        flucT = as.factor(flucT))
  # # nvm lets NOT do this bc i'll forget to fix when we do the surv stuff
```


## b. plots

```{r plot}
# visualise time to pup, pupal mass across expt groups
summary_temps %>% ggplot(aes(y = avg.mass, x = avg.tt, color = as.factor(meanT), shape = as.factor(flucT))) +
  geom_point(size = 2.5) +
  y_err_mass(err= 0.1) + x_err_tt(err = 100) + theme_bw() +
  labs(title = "2x2: avg time to and mass at pupation", 
       y = "average pupal mass (mg)", 
       x = "average time to pupa from hatching (days)", 
       caption = "all lab",
       shape = "fluctuation temp", color = "mean temp") +
  scale_color_manual(labels = c("26°C", "33°C"), 
                     values = c("#00BFC4", "#F8766D")) +
  scale_shape_discrete(labels = c("± 0°C", "± 7°C"))

# eyeeee dk what this means honestly LOLLL (not sure how to make sense of it)
```
so... hotter mean = lower mass. higher fluct = longer tt pup...

let's try to not plot 80 things at once and just put 1 response var on the y lol

```{r plot 1 response var at a time}
# avg pup mass due to temp
summary_temps %>% ggplot(aes(y = avg.mass, x = as.factor(meanT), shape = as.factor(flucT))) +
  geom_point(size = 2) + geom_text(aes(label = n_surv), vjust = -1.1, hjust = -0.5) +
  geom_line(aes(lty = as.factor(flucT), group = flucT)) +
  y_err_mass(err = 0.04) + theme_bw() +
  labs(title = "2x2: avg mass at pupation", 
       y = "average pupal mass (mg)", 
       x = "mean temp (°C)", 
       caption = "all lab; N",
       shape = "fluctuation temp", lty = "fluctuation temp") +
  scale_shape_discrete(labels = c("± 0°C", "± 7°C")) +
  scale_linetype_discrete(labels = c("± 0°C", "± 7°C"))
  
# avg tt pupa due to temp
summary_temps %>% ggplot(aes(y = avg.tt, x = as.factor(meanT), shape = as.factor(flucT))) +
  geom_point(size = 2) + geom_text(aes(label = n_surv), vjust = -1.1, hjust = -0.5) +
  geom_line(aes(lty = as.factor(flucT), group = flucT)) +
  y_err_tt(err = 0.04) + theme_bw() +
  labs(title = "2x2: avg time to pupation", 
       y = "average time to pupation from hatching (days)", 
       x = "mean temp (°C)", 
       caption = "all lab; N",
       shape = "fluctuation temp", lty = "fluctuation temp") +
  scale_shape_discrete(labels = c("± 0°C", "± 7°C")) +
  scale_linetype_discrete(labels = c("± 0°C", "± 7°C"))

# avg surv due to temp
summary_temps %>% ggplot(aes(y = prop.survpup, x = as.factor(meanT), #color = as.factor(flucT)
                             )) +
  geom_point(size = 2) + 
  geom_text(aes(label = n_all), vjust = -1.1, hjust = -0.5,
            position = position_dodge(0.1)
            #position = position_jitter(height = 0.001)
            ) +
  #geom_text(aes(label = n_all), vjust = -1.1, hjust = -0.5, position = position_jitter()) +
  geom_line(aes(lty = as.factor(flucT), group = flucT)) +
  theme_bw() + 
  #geom_blank(aes(y=1.0095)) +
  # expand_limits(y = c(0.87, 1.025)) +
  labs(title = "2x2: proportion survived to pupation", 
       y = "prop. survived to pupation", 
       x = "mean temp (°C)", 
       caption = "all lab; N",
       # color = "fluctuation temp", 
       lty = "fluctuation temp") +
  # scale_color_manual(values = c( "#00BFC4", "#F8766D"),
  #                      labels = c("± 0°C", "± 7°C")) +
  scale_linetype_discrete(labels = c("± 0°C", "± 7°C")) +
  ylim(c(0.87, 1.03))

```


```{r KM curve}
# KM curve

# add binaries and fit model
temps_all <- temps_all %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# kmfit.temps <- survfit(Surv(temps_all$tt.exit, temps_all$status) ~ meanT*flucT, data = temps_all)
  # this doesnt work bc u cant do int terms. til

kmfit.temps <- survfit(Surv(temps_all$tt.exit, temps_all$status) ~ treatment, data = temps_all)

# plot
ggsurvplot(fit = kmfit.temps, data = temps_all,
           conf.int = TRUE,
           palette = RYB,
           xlab = "days since hatching")
```

- this KM curve is not really interesting (everything overlaps + surv is high overall) so we'll just leave it here!
  - altho probably of note is that the mean=33 does separate out from mean=26 (lower survival). but again the probabilies/CIs are close/overlap so.
  
let's move into models! (240129: uhh after some followup lol)

## // followup 01: reviz by minT

bc im having a hard time seeing how the 2x2 results connect to the NTs results, which are all grouped by minT on the x... so we'll also just redo these viz lol

```{r sum stats}
# add in a missing minT... and other stats/things
temps_all$minT[temps_all$treatment == 330] <- 33

temps_all <- temps_all %>%
  group_by(meanT, flucT, minT) %>%
  mutate(n = n(),
         n_pmd = sum(final.fate == "pmd"),
         n_surv = n - n_pmd) %>%
  ungroup()
  
# resummarise w min T as a grouping variable
summary_temps2 <- temps_all %>%
  group_by(meanT, flucT, minT) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"),
            n_surv = n_all - n_pmd, 
            prop.survpup = round(1-(n_pmd/n_all), digits=2))
```

```{r ttpup plots}
# tt pup graph v1
temps_all %>% ggplot(aes(y = tt.pupa, x = as.factor(minT), fill = as.factor(flucT))) +
  geom_boxplot() +
  #y_err_tt(err= 0.25) + 
  theme_bw() +
  scale_fill_manual(values = RYB) + 
  labs(title = "2x2: avg time to pupation", 
       x = "minimum T",
       y = "average days to pupation from hatching", 
       caption = "faceted by mean temp",
       fill = "fluctuation temp (°C)") +
  facet_grid(~meanT, space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 0), check_overlap = TRUE)

# tt pup graph v2
temps_all %>% ggplot(aes(y = tt.pupa, x = as.factor(minT), fill = as.factor(meanT))) +
  geom_boxplot() +
  #y_err_tt(err= 0.9) + 
  theme_bw() +
  scale_fill_manual(values = RYB) + 
  labs(title = "2x2: avg time to pupation", 
       x = "minimum T",
       y = "average days to pupation from hatching", 
       caption = "faceted by fluct temp",
       fill = "mean temp (°C)") +
  facet_grid(~flucT, space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 0), check_overlap = TRUE)

# the blank spaces are sooo confusing lol. let's try a point plot?

# tt pup graph v3
summary_temps2 %>% 
  ggplot(aes(y = avg.tt, x = as.factor(minT), shape = as.factor(flucT), color = as.factor(meanT))) +
  geom_point(aes(), size = 2.5) +
  geom_line(aes(group = meanT, lty = as.factor(meanT))) +
  y_err_tt(err = 0.25) + 
  theme_bw() +
  scale_color_manual(values = RYB) + 
  labs(title = "2x2: avg time to pupation", 
       x = "minimum T",
       y = "average days to pupation from hatching", 
       #caption = "faceted by fluct temp",
       color = "mean temp (°C)",
       shape = "fluct temp (°C)",
       lty = "mean temp (°C)"
       )
  #facet_grid(~flucT, space = "free") +
  #geom_text(aes(label = paste0("n = ", n_surv), y = 0), check_overlap = TRUE)

# still... confusing lol idk how to group. lets try some faceting

# tt pup graph v4 - the only saved one lol
summary_temps2 %>% 
  ggplot(aes(y = avg.tt, x = as.factor(minT), 
             shape = as.factor(flucT)
             #color = as.factor(meanT))
         )) +
  geom_point(aes(), size = 2.5) +
  #geom_line(aes(group = meanT, lty = as.factor(meanT))) +
  y_err_tt(err = 0.25) + 
  theme_bw() +
  scale_color_manual(values = RYB) + 
  facet_wrap(~meanT) +
  labs(title = "2x2: avg time to pupation", 
       x = "minimum T",
       y = "average days to pupation from hatching", 
       caption = "faceted by mean temp",
       #color = "fluct temp (°C)",
       shape = "fluct temp (°C)"
       ) +
  #facet_grid(~flucT, space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 12), check_overlap = TRUE) +
  ylim(c(11.75, 20))

# errr i guess this works. hard to visualise this many temperature variables lol

# tt pup graph v5 - also saved
summary_temps2 %>% 
  ggplot(aes(y = avg.tt, x = as.factor(meanT), 
             #shape = as.factor(flucT)
             color = as.factor(flucT))
         ) +
  geom_point(aes(), size = 2.5) +
  #geom_line(aes(group = meanT, lty = as.factor(meanT))) +
  y_err_tt(err = 0.25) + 
  theme_bw() +
  scale_color_manual(values = RYB) + 
  facet_wrap(~minT) +
  labs(title = "2x2: avg time to pupation", 
       x = "mean T",
       y = "average days to pupation from hatching", 
       caption = "faceted by min temp",
       color = "fluct temp (°C)"
       #shape = "fluct temp (°C)"
       ) +
  #facet_grid(~flucT, space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 12), check_overlap = TRUE, color = "black") +
  ylim(c(11.75, 20))
```

```{r}
# this df is from analyses_ntw-ctrls

# visualise % survival
summary_lab260.2 %>% 
  mutate(x.label = case_when(x.label == "0°C" ~ "26±0°C",
                             x.label == "2.5°C"~ "26±2.5°C (diurnal)"),
         diet = case_when(diet == "LD" ~ "lab diet",
                           diet == "TB" ~ "tobacco"),
         expt.group = case_when(expt.group == "A" ~ "Feb '23",
                                expt.group == "B" ~ "Mar '23",
                                expt.group == "D" ~ "Jul '23",
                                expt.group == "F" ~ "Aug '23"),
         expt.group = factor(expt.group, levels = c("Feb '23", "Mar '23", "Jul '23", "Aug '23")),
         ) %>%
  ggplot(aes(y = pct.survpup, fill = diet, x = expt.group)) +
  geom_col() + theme_bw() +
  scale_fill_manual(values = RYB) +
  #y_err(avg.mass, se.mass) + #SIGH
  # scale_fill_hue(labels = c("A (n = 21)",
  #                            "B (n = 16)",
  #                            "D (n = 18)",
  #                            "F (n = 19)")) +
  geom_text(aes(label = paste0("n = ", n_all), y = 20), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt cohort", 
       #caption = "reared during: Jan-Mar '23 (A, B), Jul '23 (D), Aug '23 (F)"
       ) +
  facet_grid(~x.label, scales = "free_x", space = "free")

```

# 3. modeling

lets confirm what our graphs suggest + look for interactions/or lack thereof.

## a. pup mass

graphs suggest there is no mean/fluct int: let's confirm

```{r}
mod.mass <- lm(mass.pupa ~ meanT*flucT, data = temps_all)
anova(mod.mass)
anova(mod.mass, test = "Chisq")
summary(mod.mass)
  # meanT + flucT are sig on their own (***) and (**) but not the interaction
  # lowkey why is SS sooo big for mean tho LOL

# add sex as an additional check. sex shoooould be signif
mod.mass2 <- lm(mass.pupa ~ meanT*flucT*sex, data = temps_all[temps_all$sex %in% c("m", "f"),])
anova(mod.mass2) # (it is)
anova(mod.mass2, test = "Chisq")
summary(mod.mass2)

# since the MxF interaction isnt signif, drop it from the model
  # 240308: idk why i said that bc sex IS significant lmfaoo do NOT use this model. use # 2
mod.mass4 <- lm(mass.pupa ~ meanT + flucT, data = temps_all)
anova(mod.mass4)
summary(mod.mass4) ## use this!

# is this different for those that survived to pupae?
mod.mass5 <- lm(mass.pupa ~ meanT*flucT, 
                data = temps_all[temps_all$if.pupa == "Y", ])
anova(mod.mass5)
summary(mod.mass5)
  # not really
```
240308: use mod.2


## b. tt pup

confirm there there is a signif interaction btwn mean/fluct

```{r}
mod.tt <- lm(tt.pupa ~ meanT*flucT, data = temps_all)
anova(mod.tt) # fuller model is better
summary(mod.tt) # all the things matter

# let's add sex? (there shouldnt be anything)
mod.tt2 <- lm(tt.pupa ~ meanT*flucT*sex, data = temps_all)
summary(mod.tt2) # a little but not rly an effect of sex so w/e
anova(mod.tt2)

```


## c. survival

probably nothing of note will arise but we can add the glms in here anyway

```{r}
# what affects overall survival?
mod.surv <- glm(status ~ meanT*flucT, data = temps_all, family = binomial)
summary(mod.surv)
anova(mod.surv, test="Chisq") # actually test if diffrential inclusion of terms are meaningful(?)
anova(mod.surv) # tests overall variance
  # in this expt, neither of these LOL
```

# 4. list of random followups

(look for followup XX in the headers)

1. 240129: i just wanna reframe these graphs in terms of NTs... do we notice any new patterns? (mostly bc im having a hard time translating these results into the NTs (sweats))


