---
title: "overall_controls"
date: "2024-01-15"
---

purpose: checking controls over all groups. this is adding visualisations/taking some code from `other_analyses.Rmd`

also doing some other overall analyses, e.g:

- adult masses

# 0. load data, helper fns
```{r message = FALSE}
source("./helpers_ntw.R")
```

# 1. checking if lab 260s across all trts are the same to each other
```{r}
# filter lab 260 trts
data_lab260 <- wide_all %>%
  filter(treatment == 260 & pop == "lab") %>%
  filter(expt.group !="C") %>% # N is too small bc most of them culled
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled")

# calculate pup summary stats
summary_lab260 <- data_lab260 %>%
  group_by(expt.group) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))
```

## a. visualise
```{r}
# visualise time to pup, pupal mass across expt groups
summary_lab260 %>% ggplot(aes(y = avg.mass, x = avg.tt, color = expt.group)) +
  geom_point() +
  y_err_mass(err = 0.9) + x_err_tt(err = 0.9) + theme_bw() +
  scale_color_manual(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)"),
                     values = RYB) +
  labs(title = "avg time to and mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# let's isolate more... just masspup
summary_lab260 %>% ggplot(aes(y = avg.mass, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_mass(err = 0.9) + theme_bw() +
  scale_fill_manual(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)"),
                 values = RYB) +
  labs(title = "avg mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "expt group", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# just tt pup
summary_lab260 %>% ggplot(aes(y = avg.tt, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_tt(err = 0.9) + theme_bw() +
  scale_fill_manual(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)"),
                 values = RYB) +
  labs(title = "avg time to pupation for lab controls", 
       y = "average time to pupation from hatching", 
       x = "expt group", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# visualise % survival
summary_lab260 %>% ggplot(aes(y = pct.survpup, fill = expt.group, x = expt.group)) +
  geom_col() + theme_bw() +
  scale_fill_manual(values = RYB) +
  #y_err(avg.mass, se.mass) + #SIGH
  # scale_fill_hue(labels = c("A (n = 21)",
  #                            "B (n = 16)",
  #                            "D (n = 18)",
  #                            "F (n = 19)")) +
  geom_text(aes(label = paste0("n = ", n_all), y = 10), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")
```

### // followup 01: per 240117-8 meetings, let's facet these graphs a bit differently (+ try a bargraph?)

```{r data prep again}
data_lab260 <- data_lab260 %>%
    mutate(expt.group = case_when(expt.group == "A" ~ "Feb '23",
                                expt.group == "B" ~ "Mar '23",
                                expt.group == "D" ~ "Jul '23",
                                expt.group == "F" ~ "Aug '23"),
         expt.group = factor(expt.group, levels = c("Feb '23", "Mar '23", "Jul '23", "Aug '23")),
         diet = case_when(diet == "LD" ~ "lab",
                          diet == "TB" ~ "tobacco"))

# let's test on mass pup first.

# redo the summary with other groupings
summary_lab260.2 <- data_lab260 %>%
  mutate(flucT = case_when((expt.group == "D" | expt.group == "F") ~ 2.5,
                           TRUE ~ 0)) %>% # add in the correct fluct lol
  group_by(expt.group, diet, flucT) %>% # nts: dont need src
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass =  sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            n_surv = n_all - n_pmd,
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1)) %>%
  mutate(x.label = case_when(flucT == 0 ~ "0°C", # add better names for the facet labels
                             TRUE ~ "2.5°C"))

# column adjustments for boxplotting
data_lab260.2 <- data_lab260 %>%
    mutate(flucT = case_when((expt.group == "D" | expt.group == "F") ~ 2.5,
                           TRUE ~ 0)) %>% # add in the correct fluct lol
    mutate(x.label = case_when(flucT == 0 ~ "0°C",
                             TRUE ~ "2.5°C"),
           tt.pupa = jdate.pupa-jdate.hatch) %>%
  group_by(expt.group) %>%
  mutate(n = n(),
         n_pmd = sum(final.fate == "pmd"),
         n_surv = n - n_pmd) %>%
  ungroup()
  
```

```{r graphing again}
# mass.pup graph
data_lab260.2 %>% ggplot(aes(y = mass.pupa, x = expt.group, fill = diet)) +
  geom_boxplot() +
  #y_err_mass(err = 0.9) + 
  theme_bw() +
  scale_fill_manual(labels = c("lab", "tobacco"),
                    values = RYB) +
  labs(title = "avg mass at pupation for lab controls", 
       y = "average pupal mass (mg)", 
       x = "expt group", 
       #caption = "reared during: Jan-Mar '23 (A, B), Jul '23 (D), Aug '23 (F)"
       ) +
  facet_grid(~x.label, scales = "free_x", space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 15), check_overlap = TRUE)
  # this is kinda interesting. let's look at the others. and not bar graphs lol

# tt pup graph
data_lab260.2 %>% 
  ggplot(aes(y = tt.pupa, x = expt.group, fill = diet)) +
  geom_boxplot() +
  #y_err_tt(err = 0.9) + 
  theme_bw() +
  scale_fill_manual(labels = c("lab", "tobacco"),
                      values = RYB) + 
  labs(title = "avg time to pupation for lab controls", 
       x = "expt group",
       y = "average days to pupation from hatching", 
       #caption = "reared during: Jan-Mar '23 (A, B), Jul '23 (D), Aug '23 (F)"
       ) +
  facet_grid(~x.label, scales = "free_x", space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 15), check_overlap = TRUE)

# visualise % survival
summary_lab260.2 %>% 
  ggplot(aes(y = pct.survpup, fill = diet, x = expt.group)) +
  geom_col() + theme_bw() +
  scale_fill_manual(values = RYB) +
  #y_err(avg.mass, se.mass) + #SIGH
  # scale_fill_hue(labels = c("A (n = 21)",
  #                            "B (n = 16)",
  #                            "D (n = 18)",
  #                            "F (n = 19)")) +
  geom_text(aes(label = paste0("n = ", n_all), y = 10), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       #caption = "reared during: Jan-Mar '23 (A, B), Jul '23 (D), Aug '23 (F)"
       ) +
  facet_grid(~x.label, scales = "free_x", space = "free")
  
```



```{r KM plot}
# visualise survival dynamics w/ KM
  # add survival binary & time
data_lab260 <- data_lab260 %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# fit model
kmfit.lab260 <- survfit(Surv(data_lab260$tt.exit, data_lab260$status) ~ expt.group, data = data_lab260)

# plot
ggsurvplot(fit = kmfit.lab260, data = data_lab260,
           conf.int=T,
           palette = RYB)
  # well ... this generally fits as expected thus far lol.
  # seems like death time is a bit more wonky under diet
```


LOL okay rip (re: pt plot) let's model it then... (km plot for surv lol)

## b. modeling
```{r mass models}
# mass model
mod.mass <- lm(mass.pupa ~ expt.group, data = data_lab260)
anova(mod.mass)

mod.mass2 <- lm(mass.pupa ~ expt.group*sex, data = data_lab260)
anova(mod.mass2)
  # adding sex doesnt help

mod.mass3 <- lm(mass.pupa ~ sex, data = data_lab260)
anova(mod.mass3)
  # but sex alone is signif, as expected


# quick visualisation

# before that, lets add the diurnal temps back in lol
data2_lab260 <- data_lab260 %>%
  mutate(flucT = case_when((expt.group == "D" | expt.group == "F") ~ 2.5,
                           TRUE ~ 0))

plot.design(mass.pupa ~ as.factor(expt.group), data = data2_lab260)
  # B is really big? mostly bc it seems like there were 9 f and 7 m though
plot.design(mass.pupa ~ as.factor(sex), data = data2_lab260)
  # as expected
```
word no major effect. sex somewhat confirms it

```{r ttpup models}
# ttpup model
# a note that only grp F ate TB diet lol

mod.tt <- lm((jdate.pupa - jdate.hatch) ~ expt.group, data = data_lab260)
anova(mod.tt)
  # oops (EG is signif by ***)
summary(mod.tt)

mod.tt2 <- lm((jdate.pupa - jdate.hatch) ~ expt.group*sex, data = data_lab260)
anova(mod.tt2)
  # OOPS.. (only EG is signif by ***)

mod.tt3 <- lm((jdate.pupa - jdate.hatch) ~ sex, data = data_lab260)
anova(mod.tt3)
  # no sig

# # lets add the diurnal temps back in lol
# data2_lab260 <- data_lab260 %>%
#   mutate(flucT = case_when((expt.group == "D" | expt.group == "F") ~ 2.5,
#                            TRUE ~ 0))
  # moved up

mod.tt4 <- lm((jdate.pupa - jdate.hatch) ~ flucT, data = data2_lab260)
anova(mod.tt4)
  # diff SS as the expt.group but same level of sig

mod.tt5 <- lm((jdate.pupa - jdate.hatch) ~ flucT*diet, data = data2_lab260)
anova(mod.tt5)
  # flucT and diet both signif
summary(mod.tt5)

mod.tt6 <- lm((jdate.pupa - jdate.hatch) ~ flucT*diet*sex, data = data2_lab260)
anova(mod.tt6)
  # flucT and diet both signif; diet:sex a little sig. sex is not

mod.tt7 <- lm((jdate.pupa - jdate.hatch) ~ diet, data = data2_lab260)
anova(mod.tt7)
summary(mod.tt7)
  # diet and intercept model are both sig lol?

# quick visualisation
plot.design((jdate.pupa - jdate.hatch) ~ as.factor(expt.group), data = data2_lab260)
  # ... lol
```
errrm so group does have a big effect on ttpup, probably due to the diet + diurnal (sweats;;). see mod 5

let's look at survival...

```{r survival glms prep}
# survival glms

# add some more binaries
data2_lab260 <- data2_lab260 %>%
  mutate(if.pupa2 = case_when(if.pupa == "Y" ~ 0, TRUE ~ 1),
         tt.exit = jdate.exit - jdate.hatch)

# does survival depend on time? mass? diet? temp trt?

```

```{r survival glm-ing}
# modeling

# not sure if i should just use KM for anything with time in it...
mod.surv1a <- glm(if.pupa2 ~ mass.pupa*tt.exit, data = data2_lab260, family = binomial(link="logit"))
mod.surv1b <- glm(if.pupa2 ~ mass.pupa*tt.exit, data = data2_lab260, family = binomial)
  # default link for binomial is the logit lol

anova(mod.surv1a)
anova(mod.surv1b)
summary(mod.surv1b)
  # seems like mass + tt dont matter... (but they depend on temp/diet tho?)

mod.surv2 <- glm(if.pupa2 ~ flucT*diet, data = data2_lab260, family = binomial)
anova(mod.surv2)
summary(mod.surv2)
  # diet has a barely signif contribution to overall survival?
  # well that's good, actually

mod.surv3 <- glm(if.pupa2 ~ sex, data = data2_lab260, family = binomial)
anova(mod.surv3)
  # also seems like sex doesnt matter which is good!

mod.surv4 <- glm(if.pupa2 ~ expt.group, data = data2_lab260, family = binomial)
anova(mod.surv4)
  # this is also a good sign i think

# one more quick visualisation
data2_lab260$expt.group <- as.factor(data2_lab260$expt.group)
plot.design(if.pupa2 ~ expt.group, data = data2_lab260)
  # you can see F is the most different and D is also a little different.
  # idk what the numbers mean tho (ranges 0.0 to 0.4. avging 1's and 0's?)

# revisit some of the old models
anova(mod.surv1a, test = "Chisq")
  # u get nothin. thats good i think

# the only model with any signif
anova(mod.surv2, test = "Chisq") # see `?anova.glm`
anova(glm(if.pupa2 ~ diet*flucT, data = data2_lab260, family = binomial), test = "Chisq")
  # so a simpler model is better?

# deciding to skip the cox rn bc it's pretty obvious that survival is different w/in each lab bug cohort. debateable if grp D (diurnal + LD) is diff from A/B though (const + LD)

```

## c. summary...

how did cohort affect lab animals? mostly bc of the temp fluct/diet w/in each cohort ("cohort-related").

- mass wasnt really rlly affected by cohort-related stuff. was significantly affected by sex, as assumed

- tt.pup was significantly affected by fluct + diet in combination? maybe more fluct tho,,, see `mod.tt7`?

- surv seems to be very minorly affected by diet,,, see `mod.surv2`. doesnt super seem to be bc of mass/time (that's good?)

- OVERALL i think this is like. an okay sign of the controls? (i.e. the mean temp wasn't messing things up.) mostly ttpup seems to be affected by all the random stuff i did -- so this'll be important to keep in mind.

# 2. checking field 260s

```{r}
# filter field 260 trts
data_field260 <- wide_all %>%
  filter(treatment == 260 & pop == "field") %>%
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled")

# calculate pup summary stats
summary_field260 <- data_field260 %>%
  group_by(expt.group) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))
```

## a. visualise

stealing the code from above `lab` stuff lol + adjusting variable names as needed

```{r main plots}
# visualise time to pup, pupal mass across expt groups
summary_field260 %>% ggplot(aes(y = avg.mass, x = avg.tt, color = expt.group)) +
  geom_point() +
  y_err_mass(err = 0.9) + x_err_tt(err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg time to and mass at pupation for field controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "H = 426 topoff")

# let's isolate more... just masspup
summary_field260 %>% ggplot(aes(y = avg.mass, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_mass(err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg mass at pupation for field controls", 
       y = "average pupal mass", 
       x = "expt group", 
       caption = "H = 426 topoff")

# just tt pup
summary_field260 %>% ggplot(aes(y = avg.tt, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_tt(err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg time to pupation for field controls", 
       y = "average time to pupation from hatching", 
       x = "expt group", 
       caption = "H = 426 topoff")

# visualise % survival
summary_field260 %>% ggplot(aes(y = pct.survpup, fill = expt.group, x = expt.group)) +
  geom_col() + theme_bw() +
  geom_text(aes(label = paste0("n = ", n_all), y = 5), check_overlap = TRUE) +
  labs(title = "% field bugs survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "H = 426 topoff")
```

overall i care most abt group E and F matching. this generally seems to be the case which is good. don't think we'll pursue modeling these for the most part

```{r KM plot}
# visualise survival dynamics w/ KM
  # add survival binary & time
data_field260 <- data_field260 %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# fit model
kmfit.field260 <- survfit(Surv(data_field260$tt.exit, data_field260$status) ~ expt.group, data = data_field260)

# plot
ggsurvplot(fit = kmfit.field260, data = data_field260,
           conf.int=T)
  # they all have huuuuge CIs...
```
should followup on some of those freakishly long lived things in grp E (probably sups?) but otherwise i'm also reasonably content with this

## b. summary

basically nothing is that surprising to me here. they seem to be similar (which is good)

# 3. checking lab 337s

(population effects will be addressed in NTs only!)

same qns as before: do they look the "same" across trts? (also need to check removing that extra grp lol)

(and once again we are just gunna copy a lot of code lol)

```{r}
# filter lab 337 trts
data_lab337 <- wide_all %>%
  filter(treatment == 337 & pop == "lab" & instar.enter == "hatch") %>%
  filter(expt.group !="C") %>% # N is too small bc most of them culled
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled")

# calculate pup summary stats
summary_lab337 <- data_lab337 %>%
  group_by(expt.group) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))
```

## a. visualise
```{r main plots}
# visualise time to pup, pupal mass across expt groups
summary_lab337 %>% ggplot(aes(y = avg.mass, x = avg.tt, color = expt.group)) +
  geom_point() +
  y_err_mass(err = 0.9) + x_err_tt(err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("A (n = 4)",
                                "B (n = 27)",
                                "F (n = 27)")) +
  labs(title = "avg time to and mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "A = 24 but accidental deaths; C omitted bc of culling")

# let's isolate more... just masspup
summary_lab337 %>% ggplot(aes(y = avg.mass, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_mass(err = 0.9) + theme_bw() +
  scale_fill_hue(labels = c("A (n = 4)",
                                "B (n = 27)",
                                "F (n = 27)")) +
  labs(title = "avg time to and mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "A = 24 but accidental deaths; C omitted bc of culling")

# just tt pup
summary_lab337 %>% ggplot(aes(y = avg.tt, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_tt(err = 0.9) + theme_bw() +
  scale_fill_manual(labels = c("A (n = 4)",
                                "B (n = 27)",
                                "F (n = 27)"),
                    values = RYB) +
  labs(title = "avg time to pupation for lab controls (33±7)", 
       y = "average time to pupation (days)", 
       x = "expt group/cohort", 
       caption = "A = 24 but accidental deaths; C omitted bc of culling")

# visualise % survival
summary_lab337 %>% ggplot(aes(y = pct.survpup, fill = expt.group, x = expt.group)) +
  geom_col() + theme_bw() +
  geom_text(aes(label = paste0("n = ", n_all), y = 10), check_overlap = TRUE) +
  scale_fill_manual(values = RYB) +
  labs(title = "% lab ctrls (33±7) survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "A = 24 but accidental deaths; C omitted bc of culling")
```

```{r KM plot}
# visualise survival dynamics w/ KM
  # add survival binary & time
data_lab337 <- data_lab337 %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# fit model
kmfit.lab337 <- survfit(Surv(data_lab337$tt.exit, data_lab337$status) ~ expt.group, data = data_lab337)

# plot
ggsurvplot(fit = kmfit.lab337, data = data_lab337,
           conf.int=T, palette = RYB)

```
## b. summary

let's ignore grp A lol but not really sure what's going on in grp B and F re: tt pup and survival... tbh i think it's hard to say just bc they're really far apart in time lol soooo... bookmark this and we'll revisit later.. but basically B did a lot better than F for some reason :p (shorter dev time and bette survival overall)

## // followup 02: also look at diet stuff

240129: based on the 260s (followup 1), there's some pretty big diffs btwn the bugs as a result of diet + temp + pop etc. -- want to followup on these before going thru with more of the NTs analyses on the NT bugs.

`temps_analyses` (only on lab bugs) suggests that temp/fluct dont matter tooo much in determining ttpup/survival. 

```{r}
# recalc summary stats with diet included in grouping
summary_lab337.2 <- data_lab337 %>%
  filter(expt.group != "A") %>% # drop the toast group
  group_by(expt.group, diet) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            n_surv = n_all - n_pmd,
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))

# add extra columns for boxplot
data_lab337 <- data_lab337 %>%
  mutate(tt.pupa = jdate.pupa - jdate.hatch) %>%
  group_by(expt.group) %>%
  mutate(n = n(),
         n_pmd = sum(final.fate == "pmd"),
         n_surv = n - n_pmd) %>%
  ungroup()
  
```

```{r ttpup and surv plots}
# (mass is similar. not including the code)

# tt pup graph
data_lab337 %>% 
  filter(expt.group != "A") %>% #drop the toast group
  ggplot(aes(y = tt.pupa, x = expt.group, fill = diet)) +
  geom_boxplot() +
  #y_err_tt(err = 0.9) + 
  theme_bw() +
  scale_fill_manual(labels = c("lab diet", "tobacco diet"),
                      values = RYB) + 
  labs(title = "avg time to pupation for lab controls", 
       x = "expt group",
       y = "average days to pupation from hatching", 
       caption = "reared in Mar '23 (B), Aug '23 (F)") +
  #facet_grid(~x.label, scales = "free_x", space = "free") +
  geom_text(aes(label = paste0("n = ", n_surv), y = 15), check_overlap = TRUE)

# visualise % survival
summary_lab337.2 %>% ggplot(aes(y = pct.survpup, fill = diet, x = expt.group)) +
  geom_col() + theme_bw() +
  scale_fill_manual(values = RYB) +
  geom_text(aes(label = paste0("n = ", n_all), y = 10), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "reared in Mar '23 (B), Aug '23 (F)")

# see above for the KM plots

```

# // followup 03: smush plots together

want to put 2x2 and NT data on the same panels

```{r filter data}
# refilter NTs without the lab bugs on lab diet
NTs <- wide_all %>%
  filter.NTs2() %>%
  filter(!(pop == "lab" & diet == "LD")) %>%
  filter(pop == "lab") %>% # only keep lab for comparison to 2x2
  mutate(expt = "NTs") # for survival filtering

# filter temps
temps <- wide_all %>%
  filter.temps2() %>%
  mutate(expt = "temps")

# combine
smush <- rbind(NTs, temps) %>%
  mutate(status = case_when(final.fate == "pmd" ~ 1, 
                            TRUE ~ 0),
         if.fluct = case_when(treatment == 260 | treatment == 330 ~ "N",
                              TRUE ~ "Y"),
         treatment2 = paste(minT, if.fluct, sep = "-")) %>%
  filter((expt == "NTs" & if.fluct == "Y") | expt == "temps")


# calc summ stats we're interested in
summary_smush <- smush %>%
  group_by(expt, meanT, flucT, minT, maxT, if.fluct, .add = TRUE) %>% 
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n(),
            n.pmd = sum(final.fate == "pmd"),
            n.pmd = case_when(is.na(n.pmd) ~ 0, TRUE ~ as.numeric(n.pmd)),
            n.surv = n.all - n.pmd, 
            prop.survpup = round(1-(n.pmd/n.all), digits=2)) %>%
  mutate(label = case_when(expt == "temps" ~ "2x2",
                           expt == "NTs" ~ "min temps (max 40C)"))


```
```{r test fig}
summary_smush %>%
  ggplot(aes(y = avg.mass, x = as.factor(meanT), color = as.factor(minT), shape = label)) +
  geom_point(size = 2) + y_err_mass(err = 0.25) +
  geom_text(aes(label = n.surv), hjust = -0.5) +
  #geom_line(aes(group = interaction(expt, meanT, flucT))) +
  facet_wrap(~flucT, nrow = 1) +
  labs(color = "min temp (C)", shape = "expt", 
       y = "average pupal mass (mg)", x = "mean T",
       title = "pupal mass in response to varying mean/fluct temps",
       caption = "facet = flucts; lab bugs only; NTs on TB") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D")) +
  theme_bw()

summary_smush %>%
  ggplot(aes(y = avg.tt, x = as.factor(meanT), color = as.factor(minT), shape = label)) +
  geom_point(size = 2) + y_err_tt(err = 0.25) +
  geom_text(aes(label = n.surv), hjust = -0.5) +
  facet_wrap(~flucT, nrow = 1) +
  labs(color = "min temp (C)", shape = "expt", 
       y = "avg time to pupation (days since hatching)", x = "mean T",
       title = "time to pupation in response to varying mean/fluct temps",
       caption = "facet = flucts; lab bugs only; NTs on TB") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D")) +
  theme_bw()

summary_smush %>%
  ggplot(aes(y = prop.survpup, x = as.factor(meanT), color = as.factor(minT), shape = label)) +
  geom_point(size = 2) +
  geom_text(aes(label = n.all), hjust = -0.5) +
  facet_wrap(~flucT, nrow = 1) +
  labs(color = "min temp (C)", shape = "expt", 
       y = "proportion survival", x = "mean T",
       title = "survival to pupation in response to varying mean/fluct temps",
       caption = "facet = flucts; lab bugs only; NTs on TB") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D")) +
  #scale_color_manual(values = RYB) +
  theme_bw()

```



```{r make lists}
plots.mass <- summary_smush %>% 
  ungroup() %>%
  group_split(expt) %>%
  map(~ggplot(data = ., aes(y = avg.mass, x = as.factor(meanT), color = as.factor(minT), shape = as.factor(expt))) +
        geom_point(size = 2) +  
        y_err_mass(err = 0.25) +
        theme_bw() +
        facet_wrap(~flucT) +
        theme(legend.position="none") +
        geom_text(aes(label = n.surv), vjust = -1.1, hjust = -0.5) +
        labs(
             y = "average pupal mass (mg)", 
             x = "minimum temperature (°C)" 
             )
        # ylim(c(2100, 4750))
      )

plots.mass[[1]]

```

# 4. adult masses

honestly we're just going to bulk calculate them here lol bc i don't think they're really the central analyses to their respective analysis scripts, so!

```{r calc stats}
# add labels per expt; calc stats
# wide_all <- wide_all %>%
#   mutate(expt.temps = case_when(expt.group == "A" | expt.group == "B" ~ "Y"),
#          expt.acc = case_when(expt.group == "B" ~ "Y"),
#          expt.ntw = case_when(expt.group == "C" | expt.group == "D" | expt.group == "E" | expt.group == "F" | expt.group == "H" ~ "Y"),
#          expt.F1 = case_when(src == "F1" | ID %in% c(1186, 1188, 1195, 1192, 1262, 1290, 1264, 1248,
#                                                      1193, 1184, 1287, 1180,
#                                                      1353, 1344, 1349, 1345, 1389) ~ "Y"))
summ.adults <- function(data){
  summ_data <- data %>%
  filter(sex == "f" | sex == "m") %>%
  #group_by(treatment, trt.stage, pop, src, sex) %>%
  summarise(avg.mass = mean(na.omit(mass.eclose)),
         se.mass = sd(na.omit(mass.eclose))/sqrt(length(na.omit(mass.eclose))),
         avg.tt = mean(na.omit(jdate.eclose - jdate.pupa)),
         se.tt = sd(na.omit(jdate.eclose - jdate.pupa))/sqrt(length(na.omit(jdate.eclose - jdate.pupa))),
         n = n())
  
  return(summ_data)
}
```

```{r plots}
# temps
wide_all %>%
  filter.temps2() %>% 
  group_by(treatment, sex) %>% summ.adults() %>%
  ggplot(aes(y = avg.mass, x = treatment, color = sex)) +
  geom_point(aes(shape = sex), size = 2) +
  #annotate("text", label = n) +
  geom_text(aes(label = n), hjust = -2, check_overlap = T, show.legend = FALSE) +
  y_err_mass(err = 0.3) +
  theme_bw() +
  labs(title = "temps: avg adult mass")

# acc
wide_all %>%
  filter.acc2() %>% 
  group_by(trt.stage, sex) %>% summ.adults() %>%
  ggplot(aes(y = avg.mass, x = trt.stage, shape = sex, color = sex)) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -2, check_overlap = T, show.legend = FALSE) +
  y_err_mass(err = 0.3) +
  theme_bw() +
  labs(title = "acc: avg adult mass")

# NTs
wide_all %>%
  filter.NTs2() %>% 
  group_by(treatment, pop, sex) %>% summ.adults() %>%
  mutate(treatment = factor(treatment, levels = c("260", "419", "337", "433"))) %>%
  ggplot(aes(y = avg.mass, x = treatment, shape = sex, color = sex)) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -1, check_overlap = T, show.legend = FALSE) +
  y_err_mass(err = 0.3) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "NTs: avg adult mass")

# F1s
  # TODO: need to revisit the filtering lol bc its a bit all over the place...
wide_all %>%
  filter.F1s() %>% filter(treatment != 337) %>%
  mutate(tent = case_when(parent.tent == "301-I" | ID %in% c(1186, 1188, 1195, 1192, 1262, 1290, 1264, 1248) ~ "301-I",
                          parent.tent == "301-F" | ID %in% c(1193, 1184, 1287, 1180) ~ "301-F",
                          parent.tent == "107-C" | ID %in% c(1353, 1344, 1349, 1345, 1389) ~ "107-C"),
         tent = factor(tent, levels = c("301-I", "301-F", "107-C")),
         parent.trt = case_when(tent == "301-I" ~ "m/f @ 40-19C",
                                tent == "301-F" ~ "f @ 40-19C",
                                tent == "107-C" ~ "f @ 40-33C"),
         treatment2 = case_when(treatment == 260 ~ "control",
                                treatment != 260 ~ "exptl")) %>%
  drop_na(tent) %>%
  group_by(tent, parent.trt, src, treatment, treatment2, 
           sex) %>% summ.adults() %>%
  mutate(
         # treatment = factor(treatment, levels = c("260", "419", "433")),
         # treatment = case_when(treatment == 260 ~ "26-26",
         #                       treatment == 419 ~ "40-19",
         #                       treatment == 433 ~ "40-33"),
         src = case_when(src == "su" ~ "parent", T ~ as.character(src)),
         src = factor(src, levels = c("parent", "F1"))) %>%
  ggplot(aes(y = avg.mass, x = src, color = treatment2, shape = sex), position = position_dodge(2)) +
  geom_point(size = 2) +
  geom_text(aes(label = n), hjust = -1, check_overlap = T, show.legend = FALSE) +
  y_err_mass(err = 0.3) +
  facet_wrap(~parent.trt, nrow = 1) +
  theme_bw() +
  scale_color_manual(values = c("#00BFC4", "#F8766D")) +
  labs(title = "F1s: avg adult mass",
       color = "temperature treatment")
```

# // lunch bunch

```{r stats}
data_LB <- wide_all %>%
  mutate(labeled.treatment = case_when(treatment == 260 ~ "26±0",
                               treatment == 337 ~ "33±7",
                               TRUE ~ as.character(treatment)),
         labeled.group = case_when(expt.group == "A" ~ "Feb '23",
                                 expt.group == "B" ~ "Mar '23",
                                 expt.group == "D" ~ "Jul '23",
                                 expt.group == "F" ~ "Aug '23"),
         labeled.group = factor(labeled.group, levels = c("Feb '23", "Mar '23", "Jul '23", "Aug '23")),
         labeled.diet = case_when(diet == "LD" ~ "lab",
                                  diet == "TB" ~ "tobacco"),
         labeled.supfate = case_when(!is.na(jdate.pmd) ~ "pmd",
                                     !is.na(jdate.pmd) & if.pupa == "N" ~ "pmd",
                                     is.na(jdate.pmd) & if.pupa == "Y" & if.sup == "N" & sup == 0 ~ "normal",
                                     is.na(jdate.pmd) & if.pupa == "Y" & if.sup == "Y" & sup == 6 ~ "6",
                                     is.na(jdate.pmd) & if.pupa == "Y" & if.sup == "Y" & sup == 7 ~ "7"),
         labeled.supfate = factor(labeled.supfate, levels = c("pmd", "7", "6", "normal")))
```

### sup
```{r}
data_LB %>%
  filter(!(final.fate == "culled" | final.fate == "accidental")) %>%
  filter(expt.group !="C") %>%
  filter(treatment == 260 | treatment == 337) %>%
  filter(instar.enter == "hatch" & pop == "lab") %>%
  mutate(treatment = case_when(treatment == 260 ~ "26±0",
                               treatment == 337 ~ "33±7",
                               TRUE ~ as.character(treatment))) %>%
  # group_by(labeled.group, treatment) %>%
  # summarise(n_all = n(),
  #           n_pmd = count(labeled.supfate == "pmd"),
  #           n_6 = count(labeled.supfate == "6"),
  #           n_norm = count(labeled.supfate == "normal"),
  #           pct.pmd = n_pmd/n_all,
  #           pct.6 = n_6/n_all,
  #           pct.norm = n_norm/n_all) %>%
  ggplot(aes(fill = labeled.supfate, x = treatment)) +
  geom_bar(stat = "count", position = "fill") +
  facet_wrap(~labeled.group, nrow = 1) +
  labs(y = "proportion",
       x = "temperature treatment (°C)",
       fill = "development outcome") +
  scale_fill_brewer(palette = "Spectral", labels = c("failed to pupate", "pupated, supernumerary instars", "pupated, normal instars")) +
  theme_bw()
```


### big 3
```{r stats}
# lots of changes made specifically for labeling the 2nd grp of figs lol

summ_ctrlsLB <- data_LB %>%
  filter(final.fate != "culled" | final.fate != "accidental") %>%
  filter(expt.group !="C") %>% # N is too small bc most of them culled
  filter(treatment == 260 #| treatment == 337
         ) %>%
  filter(instar.enter == "hatch" & pop == "lab") %>%
  mutate(labeled.group = case_when(expt.group == "A" ~ "Feb",
                                 expt.group == "B" ~ "Mar",
                                 expt.group == "D" ~ "Jul",
                                 expt.group == "F" ~ "Aug"),
         labeled.group = factor(labeled.group, levels = c("Feb", "Mar", "Jul", "Aug")),
         flucT = case_when(flucT == 0 ~ "±0°C",
                           flucT == 2.5 ~ "±2.5°C",
                           TRUE ~ as.character(flucT))) %>%
  group_by(labeled.group, labeled.treatment, meanT, flucT, labeled.diet) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd", na.rm = TRUE), 
            prop.survpup = 1 - n_pmd/n_all,
            #pct.survpup = round(100-(n_pmd/n_all*100), digits=1),
            n_sup = sum(if.sup == "Y", na.rm = TRUE),
            pct.sup = round(n_sup/n_all*100, digits = 1), # just geom_bar this
            pct.6th = round(sum(sup == 6)/n_all*100, digits = 1),
            n_normal = n_all-n_pmd-n_sup,
            pct.normal = round(n_normal/n_all*100, digits = 1))

```

```{r facet by time}
# surv
summ_ctrlsLB %>%
  ggplot(aes(y = prop.survpup, x = labeled.treatment, color = labeled.diet)) +
  geom_point() +
  facet_wrap(~labeled.group, nrow = 1) +
  theme_bw()

# mass
summ_ctrlsLB %>%
  ggplot(aes(y = avg.mass, x = labeled.treatment, color = labeled.diet)) +
  y_err_mass() +
  geom_point() +
  facet_wrap(~labeled.group, nrow = 1) +
  theme_bw()


# time
summ_ctrlsLB %>%
  ggplot(aes(y = avg.tt, x = labeled.treatment, color = labeled.diet)) +
  y_err_tt() +
  geom_point() +
  facet_wrap(~labeled.group, nrow = 1) +
  theme_bw()
```


```{r facet by temp+diet}
# for 260s only bc these were the expt wide ctrls (337 were just to have a condition to compare to w/in each trt...)

# surv
p1 <- 
summ_ctrlsLB %>%
  ggplot(aes(y = prop.survpup, x = labeled.group, shape = labeled.diet)) +
  geom_point(size = 2.5, show.legend = FALSE) +
  facet_wrap(~flucT, scales = "free_x") +
  labs(y = "proportion surviving to pupation",
       x = "cohort") +
  ylim(c(0, 1))

# mass
p2 <- 
summ_ctrlsLB %>%
  ggplot(aes(y = avg.mass, x = labeled.group, shape = labeled.diet)) +
  y_err_mass(err = 0.5) +
  geom_point(size = 2.5) +
  facet_wrap(~flucT, scales = "free_x") +
  labs(y = "mass at pupation (mg)",
       x = "cohort") +
  ylim(c(0, 7000))


# time
p3 <- 
summ_ctrlsLB %>%
  ggplot(aes(y = avg.tt, x = labeled.group, shape = labeled.diet)) +
  y_err_tt(err = 0.5) +
  geom_point(size = 2.5) +
  facet_wrap(~flucT, scales = "free_x") +
  labs(y = "time to pupation (days)",
       x = "cohort") +
  ylim(c(0, 32))

# merge
p1 + p2 + p3 +
  plot_layout(nrow = 1, guides = "collect", axes = "collect",
              widths = c(1:1)) &
  theme_bw() &
  labs(shape = "diet type") &
  scale_shape_manual(values = c(19, 1)) &
  theme(legend.position = "top")
```



# chronological list of followups
 
 so i can keep track of these things better: they're inserted where relevant/applicable but kept track of chronologically here.
 
 1. 240118: trying different lab ctrl visualisations
 2. 240129: same but in the 337s
 3. smushing plots together (idk dont know where to put this rn)
 4. 240310: adult masses


