---
title: "NTs_analyses"
date: "2023-12-07"
---

# 0.1 load data & packages
```{r message = FALSE}
library(conflicted)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(tidyr)
library(purrr)
library(ggplot2)

# survival curve stuff
library(survival) # for computing
library(survminer) # for visualising

wide_all <- read.csv("~/Documents/repos/_not-public/1_data/ntw_data/clean-gsheets.csv", header = TRUE)

```

## - pivot to long
```{r}
long_all <- wide_all %>% select(-(starts_with("date."))) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch, 
         tt.wander = jdate.wander - jdate.hatch,
         #tt.pupa = jdate.pupa - jdate.wander,
         tt.pupa = jdate.pupa-jdate.hatch,
         tt.15 = jdate.15-jdate.hatch,
         tt.eclose = jdate.eclose-jdate.pupa,
         tt.exit = jdate.exit - jdate.enter,
         tt.surv = jdate.surv - jdate.hatch) %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "h", "tt")),
               names_to = c(".value", "instar"),
               #names_sep = ".",
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  rename(molt.status = h) %>%
  drop_na(jdate) %>% drop_na(tt) %>% # drops NA's if an individual didnt reach a certain stage
  filter(instar != "15")

# add instar factor levels
long_all <- long_all %>% mutate(instar = factor(instar, levels=c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit")))
#long_all$instar <- factor(long_all$instar, c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit"))

#rm(data_all)

```

# 0.2 define helpful functions & aesthetics

## - filtering groups
```{r}
# separate experiment groups
filter.NTs2 <-function(data){
  filtered_data <- data %>% 
    filter(trt.stage == "260-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch") %>% mutate(trt.stage = factor(trt.stage, levels = c("260-hatch", "419-hatch", "337-hatch", "433-hatch"))) %>%
    filter(expt.group == "D" | expt.group == "E" | expt.group == "F" | expt.group == "H")
  
    return(filtered_data)
}

# separate instars
filter.ins.topup <- function(data) {
  filtered_data <- data %>%
      filter(instar == "4th" | instar == "5th" | instar == "6th" | instar == "7th" | instar == "wander" | instar == "pupa")
  
  return(filtered_data)
}
```

## - doing math
```{r}
# calc dev summ stats

calc.devsumm <- function(long_data) {
  summary <- long_data %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              logmass = log(mass),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  return(summary)
}

calc.devsumm.trtstg <- function(long_data) {
  summary <- long_data %>%
    group_by(pop, trt.stage, instar) %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              logmass = log(mass),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  
  return(summary)
}

calc.devsumm.trtstgsex <- function(long_data) {
  summary <- long_data %>%
    group_by(pop, sex, trt.stage, instar) %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              logmass = log(mass),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  
  return(summary)
}

```


## - adding geoms
```{r}
# error bars
y_err_logmass <- function(x) {
  list(geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass), width = x))
}

y_err_mass <- function(x) {
  list(geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = x))
}

y_err_tt <- function(x){
  list(geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = x))
}

x_err_tt <- function(x){
  list(geom_errorbarh(aes(xmin = avg.tt - se.tt, xmax = avg.tt + se.tt), width = x))
}

# recoloring theme + legend
NT_aes <- function(x){
  list(theme_bw(), scale_color_manual(values=NT_colors, labels=NT_labels))
}
```


## - define aesthetics
```{r}
# NTs: same DTs, different NTs 
NT_trts = c("260-hatch", "419-hatch", "337-hatch", "433-hatch")
NT_labels = c("260-hatch"="26/26 (26±0°C)", "419-hatch"="40/19 (29.5±10.5°C)", "337-hatch"="40/26 (33±7°C)", "433-hatch"="40/33 (36.5±3.5°C)")
NT_colors = c("260-hatch"="#F4B942", "419-hatch"="#4059AD", "337-hatch"="#6B9AC4", "433-hatch"="#97D8C4")
  # although i think 267 is the better comparison, i have more 260s
```

 - calc timeto (3rd-pup)
```{r}
pivot.ttpup <- function(wide_data) {
  pivoted_data <- wide_data %>%
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         "6th" = jdate.6th - jdate.hatch,
         "7th" = jdate.7th - jdate.hatch, 
         wander = jdate.wander - jdate.hatch,
         pupa = jdate.pupa - jdate.wander) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("3rd", "4th", "5th", "wander", "pupa"), # pivot differences by stage
                names_to = "instar",
                values_to = "days.to") %>%
  drop_na(days.to) # drop whoever didnt hit a certain stage
  
  return(pivoted_data)
}
```

# 2.0 super simple stats
```{r}
# NT stats
wide_all %>%
  filter.NTs2() %>% filter(pop == "field") %>%
  group_by(treatment) %>%
  summarise(field_n=n(),
            pct.pmd = round(sum(final.fate == "pmd")/field_n*100, digits = 1),
            #pct.slow = round(sum(ignore.reason == "slow")/n*100, digits = 1),
            pct.sup = round(sum(if.sup == "Y")/field_n*100, digits = 1),
            pct.6th = round(sum(sup == 6)/sum(if.sup == "Y")*100, digits = 1),
            pct.7th = round(sum(sup == 7)/sum(if.sup == "Y")*100, digits = 1))
            #n.pmd = sum(final.fate == "pmd"),

wide_all %>%
  filter.NTs2() %>% filter(pop == "lab") %>%
  group_by(treatment) %>%
  summarise(lab_n=n(),
            pct.pmd = round(sum(final.fate == "pmd")/lab_n*100, digits = 1),
            #pct.slow = round(sum(ignore.reason == "slow")/n*100, digits = 1),
            pct.sup = round(sum(if.sup == "Y")/lab_n*100, digits = 1),
            pct.6th = round(sum(sup == 6)/sum(if.sup == "Y")*100, digits = 1),
            pct.7th = round(sum(sup == 7)/sum(if.sup == "Y")*100, digits = 1))


```

## - sup barplots
```{r}
# make these into barplots instead

wide_all %>%
  filter.NTs2() %>% 
  #filter(!(pop == "lab" & diet == "TB")) %>%
  filter(pop != "F1") %>%
  mutate(sup = factor(sup)) %>%
  ggplot(aes(x = trt.stage, fill = sup)) +
  geom_bar(stat = "count", position = position_fill(reverse = TRUE)) +
  labs(title = "NTs: supernumerary fates", fill = "supernumerary stage", x = "treatment", y = "percent") +
  scale_x_discrete(labels = NT_labels, guide = guide_axis(angle = 45)) +
  facet_wrap(~pop) +
  theme_bw() + scale_fill_brewer(palette = "Greens") 
  #theme(axis.text.x = element_text(angle = 30, vjust = 0.5, hjust=1))

# did i filter out the diet stuff...
  
```

# 3.0 overall growth trends; end points

compare overall development of all individuals compared to average growth

## - set up the df (**)
```{r}
# pull out the lab bugs
long_lf <- long_all %>% filter(pop == "lab" | pop == "field") 

# individual dev stats
dev_L <- long_lf %>%
  filter(!(final.fate == "pmd" | final.fate == "misc")) %>%
  filter.ins.topup() %>%
  mutate(logmass = log(mass))

devsumm_L <- dev_L %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

# for pulling out supernumeraries.. need to tweak
# devsumm_Lsups <- dev_L %>%
#   group_by(pop, sup, trt.stage, instar) %>%
#   summarise(avg.mass = mean(na.omit(mass)),
#             se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
#             avg.tt = mean(na.omit(tt)),
#             se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
#             avg.logmass = mean(na.omit(logmass)),
#             se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
#             n=n())

# see 'plot NTs' under 'overall growth'
  
```

## - plot NTs
```{r}
NT_dev <- dev_L %>% filter.NTs2()
#NT_devsumm <- devsumm_L %>% filter.NTs()
NT_devsumm <- dev_L %>% filter.NTs2() %>% calc.devsumm.trtstg()

#NTsups_devsumm <- devsumm_Lsups %>% filter.NTs2()
NTsups_devsumm <- NT_dev %>%
  group_by(pop, sup, trt.stage, instar) %>%
  calc.devsumm() %>%
  mutate(if.sup = case_when(sup == 0 ~ "N", TRUE ~ "Y"))

NT_devsumm %>%
  #filter(!(instar == "3rd" | instar == "pupa")) %>%
  filter(instar != "pupa") %>%
  #filter(sup == 0) %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line() +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=(NT_dev %>% filter(instar != "pupa")), aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(~pop) +
  labs(title = "NTs: average development from 3rd - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch")

NTsups_devsumm %>%
  filter(!(instar == "3rd" | instar == "pupa")) %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line(aes(lty = as.factor(sup))) +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=NT_dev, aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(if.sup~pop) +
  labs(title = "NTs: average development from 4th - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch")

# trying this again adding "if.sup" as Y/N
  # a little better but still hard to compare
NTsups_devsumm %>%
  filter(!(instar == "3rd" | instar == "pupa")) %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line(aes(lty = as.factor(sup))) +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=NT_dev, aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(~pop) +
  labs(title = "NTs: average development from 4th - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch")

# try again, focusing on trt.stage's producing sups instead instead
NTsups_devsumm %>%
  filter(!(instar == "3rd" | instar == "pupa")) %>%
  #filter(if.sup == "Y") %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line(aes(lty = as.factor(sup))) +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=NT_dev, aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(if.sup~pop) +
  labs(title = "NTs: average development from 4th - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch") +
  scale_x_continuous(expand = c(0.001, 1))

NTsups_devsumm %>%
  filter(!(instar == "3rd" | instar == "pupa")) %>%
  #filter(if.sup == "Y") %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line(aes(lty = as.factor(sup))) +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=NT_dev, aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(sup~pop, nrow=3) +
  labs(title = "NTs: average development from 4th - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch") +
    scale_x_continuous(expand = c(0.001, 1))

# this is it...
  #but need to filter out the lab bugs on TB/regular diet i think
NTsups_devsumm %>%
  filter(!(instar == "3rd" | instar == "pupa")) %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line(aes(lty = as.factor(sup))) +
  geom_point(aes(shape = instar), size = 3) +
  geom_point(data=NT_dev, aes(x=tt, y=logmass, color=trt.stage), alpha = 0.25) + 
  NT_aes() + facet_wrap(pop~trt.stage, ncol=4) +
  labs(title = "NTs: average development from 4th - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "L+F, no pmd, enter at hatch") 


NT_devP <- NT_dev %>% filter(instar == "pupa")
NT_devsummP <- dev_L %>%
  filter.NTs2() %>%
  group_by(pop, sex, trt.stage, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt)))) %>% 
  filter(instar == "pupa")

NT_devsummP %>%
  ggplot(aes(x=avg.tt, y=avg.mass, color=trt.stage)) +
  geom_point(size = 2) +
  geom_point(data=NT_devP, aes(x=tt, y=mass, color=trt.stage), alpha = 0.275) +
  y_err_mass(0.5) + x_err_tt(1) +
  NT_aes() + facet_wrap(sex~pop) +
  labs(title = "NTs: average development to pupa", color = "treatment", y = "mass (mg)", x = "days to instar")



```

# 4.0 survival to pupation

## - censoring + subsetting
```{r}
# censor data
surv_P <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

NTl_surv <- surv_P %>% filter(pop == "lab") %>% filter.NTs2()
NTf_surv <- surv_P %>% filter(pop == "field") %>% filter.NTs2()
```

## - create survival objects and fits
```{r}
NTf_sobj <- Surv(NTf_surv$timeto.exit, NTf_surv$censor)
NTf_sfit <- survfit(NTf_sobj ~ treatment, data = NTf_surv)

NTl_sobj <- Surv(NTl_surv$timeto.exit, NTl_surv$censor)
NTl_sfit <- survfit(NTl_sobj ~ treatment, data = NTl_surv)
```

## - plot survival curves
```{r}
# how to combine these into one?
ggsurvplot(fit = NTl_sfit, data=NTl_surv,
           conf.int=T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NTs: lab survival from hatch-pup", 
           palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)"),
           #strata = pop
           ) +
  labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")

ggsurvplot(fit = NTf_sfit, data=NTf_surv,
           conf.int=T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NTs: field survival from hatch-pup", 
           palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)"),
           #strata = pop
           )

```

```{r}
## need to redo with some of the earlier field ones yeeted too LOL
```

# 4.1 NT stats

```{r}
# subset NTs

check_NTstats <- wide_all %>% filter.NTs2()
```

## - trt x tt.pup
```{r}
# model
mod_checkNT_ttpup <- lm((jdate.pupa - jdate.hatch) ~ pop + treatment, data=check_NTstats)
  # look into the sigif when expt.group is added...
  # try adding pop*trt?

anova(mod_checkNT_ttpup)

summary(mod_checkNT_ttpup)

# plot
#devsumm_L %>% # should redo this with calcing dev only for the group...
long_all %>%
  filter.NTs2() %>% 
  calc.devsumm.trtstg() %>%
  filter(instar == "pupa") %>%
  ggplot(aes(x = trt.stage, y = avg.tt, color = pop)) +
  geom_point() + theme_bw() + y_err_tt(0.5) +
  labs(title = "avg time to pupation")

```

## - trt x mass.pup
```{r}
# model
mod_checkNT_masspup <- lm((mass.pupa) ~ pop + treatment + sex, data=check_NTstats)

anova(mod_checkNT_masspup)

summary(mod_checkNT_masspup)

# plot
long_all %>%
  filter.NTs2() %>% 
  calc.devsumm.trtstgsex() %>%
  filter(instar == "pupa") %>%
  ggplot(aes(x = trt.stage, y = avg.mass, color = pop)) +
  geom_point() + theme_bw() + y_err_mass(0.5) +
  facet_wrap(~sex) +
  labs(title = "avg mass at pupation")
```

# 5.0 cox-ph survival stuff?

- check as variables: lab/TB diet; if.sup; if.stuck

##  - if sup x to pup/eclosion

```{r}
#check_suppup <- wide_all %>% filter(if.pupa == "Y")

#mod_suppup <- lm(if.pupa ~ if.sup + treatment + pop, data = check_suppup, na.action=na.omit)
  # does this not work bc Y needs to be a # LOL

# i think this shoudl be part of survival stuff

```

## - 231130 followups: glms

```{r}
# binomial survival to pup

check_NTstats <- check_NTstats %>% 
  mutate(if.pupa2 = case_when(if.pupa == "Y" ~ 1, TRUE ~ 0),
         tt.pupa = jdate.pupa - jdate.hatch)

# going to try a bunch of things
mod_checkNT_survpup1 <-  glm(if.pupa2 ~ trt.stage + pop, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup1)
  # dont rlly know how to interpret...

mod_checkNT_survpup2 <-  glm(if.pupa2 ~ trt.stage + mass.pupa + tt.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup2)
  # algo did not converge

mod_checkNT_survpup3 <- glm(if.pupa2 ~ trt.stage + mass.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup3)
  # algo did not converge

mod_checkNT_survpup4 <- glm(if.pupa2 ~ trt.stage*mass.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup4)
  # algo did not converge

mod_checkNT_survpup5 <- glm(if.pupa2 ~ meanT*flucT + pop, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup5)
  # this works!

# redo this with taking out the 260 group
check_NTstats2 <- check_NTstats %>% 
  filter(meanT != 26) %>%
  mutate(dayT = case_when(TRUE ~ 40),
         nightT = case_when(treatment == 419 ~ 19,
                            treatment == 337 ~ 26,
                            treatment == 433 ~ 33))

mod_checkNT_survpup5a <- glm(if.pupa2 ~ meanT*flucT + pop, data = check_NTstats2, family = "binomial")
summary(mod_checkNT_survpup5a)

mod_checkNT_survpup5b <- glm(if.pupa2 ~ nightT + pop, data = check_NTstats2, family = "binomial")
summary(mod_checkNT_survpup5b)

# trying things w/ joel... (see that one slide)

anova(mod_checkNT_survpup1, test = "Chisq") # analysis of deviance (not looking at coef; rather than variance)
  # interp the same as an anova

plot.design(if.pupa2 ~ trt.stage + pop, data = check_NTstats)
  # coefs: how diff are they from the intercept
  # anova: are they signif different (compare all tgt simultaneously)
  # this is an easy way to visualise results



mod_checkNT_survpup6 <- glm(if.pupa2 ~ mass.pupa, data = check_NTstats, family = "binomial")
  # breaks. also breaks when u try to throw sex in here so let's stop here

```

# 6.0 x=trt, y = response graphs

```{r}
# subset data
long_NTs <- long_all %>% 
  #filter(is.NA(ignore.reason)) %>%
  filter.NTs2()
  # "slows" are messing things up

long_NTs %>%
  #filter(ignore.reason == "NA") %>%
  filter(instar == "pupa") %>%
  group_by(pop, sex, meanT) %>%
  calc.devsumm() %>%
  ggplot(aes(x = meanT, y = avg.mass)) +
  geom_point() +
  geom_line(aes(lty = pop)) +
  y_err_mass(x=0.5) +
  facet_wrap(~sex) +
  theme_bw() +
  labs(title = "NT: mean temperatures on pupal mass", caption = "NT pupa")

long_NTs %>%
  filter(instar == "pupa") %>%
  #filter(!is.NA(ignore.reason)) %>%
  group_by(pop, sex, flucT) %>%
  calc.devsumm() %>%
  ggplot(aes(x = flucT, y = avg.mass)) +
  geom_point() +
  geom_line(aes(lty = pop)) +
  y_err_mass(x=0.5) +
  facet_wrap(~sex) +
  theme_bw() +
  labs(caption = "NT pupa")
  # smth is breaking bc this doesnt match the above LOL

long_NTs %>%
  #filter(ignore.reason == "NA") %>%
  filter(instar == "pupa") %>%
  group_by(pop, sex, minT) %>%
  calc.devsumm() %>%
  ggplot(aes(x = minT, y = avg.mass)) +
  geom_point() +
  geom_line(aes(lty = pop)) +
  y_err_mass(x=0.5) +
  facet_wrap(~sex) +
  theme_bw() +
  labs(title = "NT: min temperatures on pupal mass", caption = "NT pupa")
  # this breaks so much LOL

# other y-variables
long_NTs %>%
  #filter(ignore.reason == "NA") %>%
  filter(instar == "pupa") %>%
  group_by(pop, sex, meanT) %>%
  calc.devsumm() %>%
  ggplot(aes(x = meanT, y = avg.tt)) +
  geom_point() +
  geom_line(aes(lty = pop)) +
  y_err_tt(x=0.5) +
  facet_wrap(~sex) +
  theme_bw() +
  labs(title = "NT: mean temperatures on time to pupation", caption = "NT pupa")

long_NTs %>%
  #filter(ignore.reason == "NA") %>%
  #filter(instar == "pupa") %>%
  #filter(sex != "NA") %>%
  group_by(pop, meanT) %>%
  summarise(n_all=n(),
            n_pmd = sum(final.fate == "pmd"),
            pct.surv = 100-round(n_pmd/n_all*100, digits = 1)) %>% 
  ggplot(aes(x = meanT, y = pct.surv)) +
  geom_point() +
  geom_line(aes(lty = pop)) +
  #facet_wrap(~sex) +
  theme_bw() +
  labs(title = "NT: mean temperatures on % survival to pupation", caption = "NT pupa")

```
