---
title: "NTs_survival analyses and modeling"
date: "2024-01-08"
---

# 0. load data & packages
```{r load, message = FALSE}
source("./ntw_helper-functions.R")
rm(list=c(acc_helpers, temps_helpers))

# survival curve stuff
library(survival) # for computing
library(survminer) # for visualising
library(nonnestcox) # for plrtest() to compare CPH models

# need to renumber stuff again at some point but not today
```



# 1.0 KM survival to pupation 

## A. prep data

### - (**) censoring + subsetting
```{r censor data for surv}
# censor data
surv_P <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))
```

```{r filter for KM}
NTl_surv <- surv_P %>% filter(pop == "lab") %>% filter.NTs2()
NTf_surv <- surv_P %>% filter(pop == "field") %>% filter.NTs2()
```

### - create survival objects and fits
```{r survfit}
NTf_sobj <- Surv(NTf_surv$timeto.exit, NTf_surv$censor)
NTf_sfit <- survfit(NTf_sobj ~ treatment, data = NTf_surv)

NTl_sobj <- Surv(NTl_surv$timeto.exit, NTl_surv$censor)
NTl_sfit <- survfit(NTl_sobj ~ treatment, data = NTl_surv)
```

## B. plot survival curves

### - first pass at plotting 

had to separate the L and F bc idk how to do them on the same one
```{r first pass L+F KM plots}
# how to combine these into one?
ggsurvplot(fit = NTl_sfit, data=NTl_surv,
           conf.int=T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NTs: lab survival from hatch-pup", 
           palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)"),
           #strata = pop
           ) +
  labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")

ggsurvplot(fit = NTf_sfit, data=NTf_surv,
           conf.int=T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NTs: field survival from hatch-pup", 
           palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)"),
           #strata = pop
           )

```

### - variations 
(for convenience, just the lab ones unless indicated)
```{r var - fix labels}
## adjust labels to focus on the min T and fluct
# 
# ggsurvplot(fit = NTl_sfit, data=NTl_surv,
#            conf.int=T,
#            xlab = "days from hatching", ylab = "survival probability",
#            title = "NTs: lab survival from hatch-pup", 
#            palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
#            legend.title = "treatment", 
#            legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)"),
#            #strata = pop
#            ) +
#   labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")

# got too lazy to do this
```

```{r var - remove earlier field}
## redo with some of the earlier field ones yeeted too LOL

# remove individuals & refilter
NTf_surv2 <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  filter(pop == "field") %>%
  filter(!(treatment == 337 & ID < 202)) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate."))) %>%
  filter.NTs2()

# set up survival objects
NTf_sobj2 <- Surv(NTf_surv2$timeto.exit, NTf_surv2$censor)
NTf_sfit2 <- survfit(NTf_sobj2 ~ treatment, data = NTf_surv2)

# plot
ggsurvplot(fit = NTf_sfit, data=NTf_surv,
           conf.int=T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NTs: field survival from hatch-pup (adj for 426)", 
           palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           legend.labs = c("26/26 (26±0°C)", "40/26 (33±7°C)", "40/19 (29.5±10.5°C)", "40/33 (36.5±3.5°C)")
           #strata = pop
           )
```


```{r var - plot only 3 groups}
## redo with 26 taken out (3 groups only) → just do if neededbc im lazy
```

# 2.0 some regression models

## - data prep
```{r subset for LM models}
# subset NTs
check_NTstats <- wide_all %>% filter.NTs2()
```

## - models

### - lm(tt.pup ~ x)

```{r model} 
mod_checkNT_ttpup <- lm((jdate.pupa - jdate.hatch) ~ pop + treatment, data=check_NTstats)
  # look into the sigif when expt.group is added...
  # try adding pop*trt?

anova(mod_checkNT_ttpup)
summary(mod_checkNT_ttpup)
```

```{r plot}
#devsumm_L %>% # should redo this with calcing dev only for the group...
long_all %>%
  filter.NTs2() %>% 
  calc.devsumm.trtstg() %>%
  filter(instar == "pupa") %>%
  ggplot(aes(x = trt.stage, y = avg.tt, color = pop)) +
  geom_point() + theme_bw() + y_err_tt(0.5) +
  labs(title = "avg time to pupation")

```

### - lm(mass.pup ~ x)
```{r model}
mod_checkNT_masspup <- lm((mass.pupa) ~ pop + treatment + sex, data=check_NTstats)

anova(mod_checkNT_masspup)
summary(mod_checkNT_masspup)
```

```{r plot}
long_all %>%
  filter.NTs2() %>% 
  calc.devsumm.trtstgsex() %>%
  filter(instar == "pupa") %>%
  ggplot(aes(x = trt.stage, y = avg.mass, color = pop)) +
  geom_point() + theme_bw() + y_err_mass(0.5) +
  facet_wrap(~sex) +
  labs(title = "avg mass at pupation")
```

###  - lm(if.pupa ~ x)

testing if sup x to pup/eclosion ==> this does not work as a model bc i am dumb and u cant do it this way lol

```{r}
#check_suppup <- wide_all %>% filter(if.pupa == "Y")

#mod_suppup <- lm(if.pupa ~ if.sup + treatment + pop, data = check_suppup, na.action=na.omit)
  # does this not work bc Y needs to be a # LOL

# i think this shoudl be part of survival stuff

```

# 4.1 binomial glm(if.pupa ~ x)

formerly titled "binomials (glms): 231130 followups"

checking survival to pup with a binomial glm

## - subset

```{r filter data}
# binomial survival to pup

check_NTstats <- check_NTstats %>% 
  mutate(if.pupa2 = case_when(if.pupa == "Y" ~ 1, TRUE ~ 0),
         tt.pupa = jdate.pupa - jdate.hatch)
```

## - model variations

let's try a bunch of things
```{r model 1}
# going to try a bunch of things
mod_checkNT_survpup1 <-  glm(if.pupa2 ~ trt.stage + pop, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup1)
  # dont rlly know how to interpret these numbers...
```

```{r}
mod_checkNT_survpup2 <-  glm(if.pupa2 ~ trt.stage + mass.pupa + tt.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup2)
  # algo did not converge
```

```{r}
mod_checkNT_survpup3 <- glm(if.pupa2 ~ trt.stage + mass.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup3)
  # algo did not converge
```

```{r}
mod_checkNT_survpup4 <- glm(if.pupa2 ~ trt.stage*mass.pupa, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup4)
  # algo did not converge
```

```{r model 5 that works}
mod_checkNT_survpup5 <- glm(if.pupa2 ~ meanT*flucT + pop, data = check_NTstats, family = "binomial")
summary(mod_checkNT_survpup5)
  # this works!
```

```{r model 5 variations}
# redo this with taking out the 260 group
check_NTstats2 <- check_NTstats %>% 
  filter(meanT != 26) %>%
  mutate(dayT = case_when(TRUE ~ 40),
         nightT = case_when(treatment == 419 ~ 19,
                            treatment == 337 ~ 26,
                            treatment == 433 ~ 33))

mod_checkNT_survpup5a <- glm(if.pupa2 ~ meanT*flucT + pop, data = check_NTstats2, family = "binomial")
summary(mod_checkNT_survpup5a)
```

```{r}
mod_checkNT_survpup5b <- glm(if.pupa2 ~ nightT + pop, data = check_NTstats2, family = "binomial")
summary(mod_checkNT_survpup5b)
```

```{r}
mod_checkNT_survpup6 <- glm(if.pupa2 ~ mass.pupa, data = check_NTstats, family = "binomial")
  # breaks. also breaks when u try to throw sex in here so let's stop here
```

### (?) interpreting model outputs w/ joel

just using (**) to sticky this
```{r anova(model)}
# trying things w/ joel... (see that one slide)

anova(mod_checkNT_survpup1, test = "Chisq") # analysis of deviance (not looking at coef; rather than variance)
  # interp the same as an anova
```

```{r plot.design}
plot.design(if.pupa2 ~ trt.stage + pop, data = check_NTstats)
  # coefs: how diff are they from the intercept
  # anova: are they signif different (compare all tgt simultaneously)
  # this is an easy way to visualise results
```

revisiting glms on 240106: 
```{r visualise}
# redo the glm with 3 groups only to find the predictors

# look at the groups again
plot.design(if.pupa2 ~ trt.stage + pop, data = check_NTstats2)
  # the 426 group is significantly diff from the 419 and 433 groups (higher)
```

```{r model 7 which works}
# look at the trt*pop interaction bt 3 groups
mod_checkNT_survpup7 <- glm(if.pupa2 ~ trt.stage + pop + trt.stage*pop, data = check_NTstats2, family = "binomial")
anova(mod_checkNT_survpup7, test = "Chisq")
  # int and pop is signif but not trt?

# test <- glm(if.pupa2 ~ trt.stage + pop, data = check_NTstats2, family = "binomial") # drop the int, see what happens
# anova(test, test = "Chisq")
  # confirms the above
```

# 5.0 CPH survival modeling 

- things to check as variates: lab/TB diet; if.sup; if.stuck

## - (**) data prep
```{r filter}
NTall_surv <- surv_P %>% filter.NTs2()
```

## - model variations

fit some models and look at them

### - by trt only
```{r everything model}
# try things that i think are relevant
mod_cox_NTsurv01 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ trt.stage + pop + trt.stage*pop + sex, data = NTall_surv)
summary(mod_cox_NTsurv01) # this dies (NAs) prob bc i dont have `sex` for all of them, so...
```

```{r}
# drop sex from the df

NTall_survsex <- NTall_surv %>% drop_na(sex)
mod_cox_NTsurv02 <- coxph(Surv(NTall_survsex$timeto.exit, NTall_survsex$censor) ~ trt.stage + pop + trt.stage*pop + sex, data = NTall_survsex)
summary(mod_cox_NTsurv02) # still gives me NAs LOL
```

```{r a model that works}
# drop sex from the formula 

mod_cox_NTsurv03 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ trt.stage + pop + trt.stage*pop, data = NTall_surv)
summary(mod_cox_NTsurv03) # ok word this kinda works
```

### - break trt into mean/fluct T

extended variations, using mean T/flucT as predictors
```{r}
mod_cox_NTsurv04 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + meanT*flucT + pop + meanT*pop + flucT*pop,
  data = NTall_surv)
# not sure how to handle the interactions tbh... i'm probably doing this wrong!

summary(mod_cox_NTsurv04)
# but at least we get results...
```
just population is signif?

try different nesting

```{r}
# lets drop some stuff
mod_cox_NTsurv05 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + meanT*flucT*pop,
  data = NTall_surv)

summary(mod_cox_NTsurv05)
  # errr this nesting right
```

change incorporation of `pop`

```{r bad model maybe}
mod_cox_NTsurv06 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + meanT*flucT + pop,
  data = NTall_surv)

summary(mod_cox_NTsurv06) 
```
only `pop` is signif here. doesn't feel that right to me, so maybe we should revisit `model 5`

```{r}
mod_cox_NTsurv07 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + meanT*flucT + pop + meanT*flucT*pop,
  data = NTall_surv)

summary(mod_cox_NTsurv07)
```

(make some simpler models to compare to)

```{r}
# additive only
mod_cox_NTsurv08 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + pop,
  data = NTall_surv, x=T)

summary(mod_cox_NTsurv08)
```

in a naive additive way, they all matter

```{r}
# no pop
mod_cox_NTsurv09 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + meanT*flucT,
  data = NTall_surv, x=T)

summary(mod_cox_NTsurv09)

# additive with temp interaction
# mod_cox_NTsurv10 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT + flucT + pop + meanT*flucT,
#   data = NTall_surv)

```

i've probably made enough variations atp so lets compare them all with `plrtest`

(idk if i'm doing this right lol so need to chat with joel)

#### - compare models

lets compare the simple ones first LOL

```{r compare simple coxph}
# double check if these are the same LOL
summary(mod_cox_NTsurv06)
#summary(mod_cox_NTsurv10) # (they are)
```

```{r compare simple models}
plrtest(mod_cox_NTsurv08, mod_cox_NTsurv09, nested = F)
```
### - redoing this and starting even simpler

A. let's try just `trt.stage` and `pop` first

```{r trtstage pop cox models}
coxmod_A1 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ trt.stage, data = NTall_surv)
summary(coxmod_A1)
```
all trts are significant!!! idk what to make of that though. 

lets try testing everything at once... (testing one variate at a time per https://www.sthda.com/english/wiki/cox-proportional-hazards-model)
```{r}
# coxmod_A2 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ trt.stage + pop, data = NTall_surv)
# coxmod_A3 <- coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ trt.stage, data = NTall_surv)

covariates <- c("meanT", "maxT", "flucT", "minT", "pop", "sex", "diet") # some of these vars are basically the same way of subdividing
multicoxfitting <- sapply(covariates,
                          function(x) as.formula(paste('Surv(NTall_surv$timeto.exit, NTall_surv$censor)~', x)))
multicoxmodels <- lapply(multicoxfitting, function(x){coxph(x, data = NTall_surv)})

# extract results
multicoxresults <- lapply(multicoxmodels,
                       function(x){ 
                          x <- summary(x)
                          p.value <- signif(x$wald["pvalue"], digits=2)
                          wald.test <- signif(x$wald["test"], digits=2)
                          beta <-signif(x$coef[1], digits=2);#coeficient beta
                          HR <- signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"], 2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
                                        "p.value")
                          # res<-c(beta, HR, wald.test, p.value, HR.confint.lower, HR.confint.upper) # need to specify other 2 columns or else breaks
                          # names(res[1:6])<-c("beta", "HR (95% CI for HR)", "wald.test", 
                          #               "p.value", "up", "low")
                          return(res)
                          #return(exp(cbind(coef(x),confint(x))))
                         })

# put into df
res <- t(as.data.frame(multicoxresults, check.names = FALSE))
as.data.frame(res)
  # only `test` and `pval` are extracted
  # "trt.stage", "treatment" need to be handled separately
  # is the problem testing cox with a categorical var of >2 types? (yes! see below)


# summary(coxph(Surv(NTall_surv$timeto.exit, NTall_surv$censor) ~ meanT, data = NTall_surv))
```

- google says its bc they're categorical variables: need to dummy var them (https://stackoverflow.com/questions/62876700/how-to-code-categorical-variable-with-three-levels-for-cox-with-random-effects-m)
  - i dont understand setting up contrasts/dummy variables so lets just skip this for now. meanT/flucT + interaction should be good enough anyway
    - nevermind go see this later: https://www.sthda.com/english/articles/40-regression-analysis/163-regression-with-categorical-variables-dummy-coding-essentials-in-r
    
    
- seems like everything but sex is significant but only pop really contributes to the hazard ratio. we can see what happens when we look at interactions tho
  - being lab pop is better
- i feel like diet is prbably more related to being a lab or a field bug tho... should recheck this in the EDA of controls
  
```{r example}
# # working example code
# data(cancer)
# 
# covariates <- c("age", "sex",  "ph.karno", "ph.ecog", "wt.loss")
# univ_formulas <- sapply(covariates,
#                         function(x) as.formula(paste('Surv(time, status)~', x)))
#                         
# univ_models <- lapply( univ_formulas, function(x){coxph(x, data = cancer)})
# 
# # Extract data 
# univ_results <- lapply(univ_models,
#                        function(x){ 
#                           x <- summary(x)
#                           p.value<-signif(x$wald["pvalue"], digits=2)
#                           wald.test<-signif(x$wald["test"], digits=2)
#                           beta<-signif(x$coef[1], digits=2);#coeficient beta
#                           HR <-signif(x$coef[2], digits=2);#exp(beta)
#                           HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
#                           HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
#                           HR <- paste0(HR, " (", 
#                                        HR.confint.lower, "-", HR.confint.upper, ")")
#                           res<-c(beta, HR, wald.test, p.value)
#                           names(res)<-c("beta", "HR (95% CI for HR)", "wald.test", 
#                                         "p.value")
#                           return(res)
#                           #return(exp(cbind(coef(x),confint(x))))
#                          })
# res0 <- t(as.data.frame(univ_results, check.names = FALSE))
# as.data.frame(res0)
```

anyway... let's compare some simple models and start adding a bunch of terms on
```{r}
# skeleton stuff for now... prob should read a bit more on this before setting out tho
mod <- meanT
mod <- meanT + maxT
mod <- mean:max
```


