---
title: "cleaned up NTW"
output: html_notebook
---

because my main stuff is getting tooo messy again!!!


0. load in data, packages
1. clean/reformat data, make subsets
2. look at mass gain/instar time on same plot
  a. calculate stats for larva/adults
  b. group stats by expts (temp, tempxins, NTs)
  c. plot
3. look at mass gain/instar time on separate plot
  a. calculate stats
  b. group stats by expts (temp, tempxins, NTs)
  c. plot
4. kill curves
5. misc: old 2013 data



todo
- work on colors
- plot wand or pupal masses (latter is better) → as a fn of time (focus on the end pt)
- plot individ curves
  - plot sups separately/overlay on them (can be pulled out as a variable)
    - int cuz its the field!!!!
- analyses: lms for pupal dev + time as response vars
  - do a cox ph (trt as predictor)
- figure out to summarise

- focus on pup responses for rn!

# 0. load in packages, data
```{r message=FALSE}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(survival)
library(survminer)
library(googlesheets4)

#data_all <- read.csv("~/Documents/repos/_not-public/1_data/ntw_data/development_v3.csv", header = T)
```

```{r}
# before importing,
  # run cleaning macros (larva sheets)
  # check column/row alignment + proper sorting (adult sheets)

# larva data
data_field <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet = "june field", col_types = "c")
data_labsu <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="june 2023 lab", col_types = "c")
data_labF1 <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="lab F1", col_types = "c")
data_labsp <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="consolidated + tidyed", col_types = "c")

# tents/hatch data
# data_matepairs <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="mating pairs", col_types = "c")
# data_tents <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="tent stats v2", col_types = "c")
# data_hatching <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="hatch stats", col_types = "c")
```



# 1. clean data

## consolidate imported sheets + pre-processing
```{r}
# function for column cleanup, adding some columns (imo separate the column generation to a diff fn. maybe just keep this one tho and just calc a diff one for timeto)
# need to make sure that format is correctly set on gsheet first >_>
fix.columns <- function(data) {
  cleaned_data <- data %>%
    mutate(across(starts_with("date."),  as.Date, format = "%m/%d")) %>%
    #mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
    mutate(mutate(across(starts_with("mass."), as.numeric))
           #,
           # across(starts_with("jdate."), as.numeric),
           # treatment = as.character(treatment),
           # trt.stage = as.character(paste(treatment,instar.enter, sep = "-")), # treatment @ instar
           # grp.trt = paste(expt.group, trt.stage, sep = "-")) # round @ treatment
    )
  
           
  return(cleaned_data)
}

# add IDs
data_field <- mutate(data_field, src = "field")
data_labF1 <- mutate(data_labF1, src = "F1")
data_labsp <- mutate(data_labsp, src = "sp")
data_labsu <- mutate(data_labsu, src = "su")

# consolidate
data_all <- data.table::rbindlist(lapply(list(data_field, data_labF1, data_labsp, data_labsu), fix.columns), fill = T)

### preprocessing + column fixing
# match a column for consistency
data_all$treatment[data_all$treatment=="426"] <- "337"

# filter out individuals that didnt hatch at 26C
data_all <- data_all %>%
    filter(is.na(from.E) | !(from.E == 2 | (from.E == 0 & (treatment == 337 | treatment == 330 | treatment == 267))))

# change field 1sts to hatchlings (uhh maybe okay to keep them as 1sts tbh)
data_all$instar.enter[data_all$trt.stage == "1st"] <- "hatch"
```

## new column creation
```{r}
# survival/development binaries
data_all <- data_all %>% mutate(
  if.stuck = case_when(is.na(date.stuck) ~ "N", TRUE ~ "Y"),
  if.pupa = case_when(is.na(date.LP) & is.na(date.pupa) ~ "N", TRUE ~ "Y"),
  if.sup = case_when(is.na(date.6th) ~ "N", TRUE ~ "Y"),
  sup = case_when(!is.na(date.6th) ~ 6, !is.na(date.7th) ~ 7, TRUE ~ 0)
  )

# mean/fluc Ts
data_all <- data_all %>% mutate(
  meanT = case_when(treatment == 260 | treatment == 267 | treatment == "ctrl" ~ 26,
                    treatment == 337 | treatment == 330 ~ 33,
                    treatment == 433 ~ 36.5,
                    treatment == 419 ~ 29.5),
  flucT = case_when(treatment == 260 | treatment == 330 ~ 0,
                    treatment == 267 | treatment == 337 ~ 7,
                    treatment == 433 ~ 3.5,
                    treatment == 419 ~ 10.5)
  )

# trts and prep for calculating timetos (moved this out from above fn idk if it will work yet)
data_all <- data_all %>%
  mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
  mutate(across(starts_with("jdate."), as.numeric),
         treatment = as.character(treatment),
         trt.stage = as.character(paste(treatment,instar.enter, sep = "-")), # treatment @ instar
         grp.trt = paste(expt.group, trt.stage, sep = "-")) # round @ treatment

# indicate field or lab pops
data_all <- data_all %>%
  mutate(pop = case_when(location == "CC" ~ "field",
                         src == "F1" ~ "F1",
                         TRUE ~ "lab"))

```


## further filtering
```{r}
# filter out MQs
data_MQ <- data_all %>% filter(species == "MQ")
data_all <- data_all %>% filter(is.na(species))

# indicate overly long-lived individuals: things that took > 25 days to die
# these can be omitted later
data_all <- data_all %>% mutate(ignore.reason = case_when(jdate.pmd-jdate.hatch > 25 ~ "slow"))
  # check: table(data_all$ignore.reason)

# extract wordier notes from data and save elsewhere
data_notes <- select(data_all, c("treatment", "ID", "expt.group", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "mass.died", "pupa.deformities", "notes", "ignore.reason", ends_with(".stuck"), ends_with(".culled")))

# remove unneeded columns
data_all <- select(data_all, -c("species", "from.E", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "pupa.deformities", "mass.died", "toss.if", "old.date.pmd", "fate.code", "notes", ends_with(".stuck"), ends_with(".culled"), "eclose-3"))

# write filter functions
filter.temps <- function(data) {
  filtered_data <- data %>% filter(trt.stage == "260-hatch" | trt.stage =="267-hatch" | trt.stage == "330-hatch" | trt.stage == "337-hatch") %>% filter(pop == "lab")
  
  return(filtered_data)
}

filter.accum <- function(data) {
  filtered_data <- data %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch") %>% filter(pop == "lab") %>% mutate(trt.stage = factor(trt.stage, levels = c("260-hatch", "337-hatch", "337-3rd", "337-4th")))
  
  return(filtered_data)
}

filter.NTs <-function(data){
  filtered_data <- data %>% filter(trt.stage == "260-hatch" | trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch") %>% mutate(trt.stage = factor(trt.stage, levels = c("260-hatch", "419-hatch", "337-hatch", "433-hatch")))
  
    return(filtered_data)
}
```


## pivot and final subset
```{r}
# pivot
long_all <- data_all %>% select(-(starts_with("date."))) %>%
                 pivot_longer(cols = starts_with(c("jdate", "mass", "h")),
                        names_to = c(".value", "instar"),
                        #names_sep = ".",
                        values_drop_na = TRUE,
                        names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
                 rename(molt.status = h) %>%
                 drop_na(jdate) # drops NA's if an individual didnt reach a certain stage

# add instar factor levels
long_all$instar <- factor(long_all$instar, c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit"))

# going forward, only use individuals that survived to pupation (includes LPs) 
  # +++ may need to add in individs to ignore
data_pup <- data_all %>% filter(if.pupa == "Y")
long_pup <- long_all %>% filter(if.pupa == "Y")

  # need to check if these work for survival analyses
```

at this point, `data_all` is a consolidated wide version of field, lab, and F1 individuals. "commentary" columns have been removed and the following columns have been added:

- survival/development binaries (to pup, if sup, which sup)
- filtering by src pop, mean/fluct Ts

this dataset still includes: (filter as needed)

- "slow" individuals (that die eventually)
- MQs
- pmds

*use this for doing math bt instars, surv curves* 

`data_long` is pivoted by instar to separate out mass, time to, and headcap status at each instar.

*use this for grouping by/looking at overall instar*



## define labels and aesthetics
```{r}
#all_trts <- c("260-hatch", "267-hatch", "330-hatch", "337-hatch", "337-3rd", "337-4th", "40-19", "40-26")


# A: effect of mean/fluct temp (treatment)
A_trts = c("260-hatch", "267-hatch", "330-hatch", "337-hatch")
A_labels = c("260-hatch"="26°C", "267-hatch"="26±7°C", "330-hatch"="33°C", "337-hatch"="33±7°C")
A_colors = c("260-hatch"="#00C2D1","267-hatch"="#1929B3", "330-hatch"="#F9C639", "337-hatch"="#710A36")



# B: effect of temp x instar (trt.stage)
B_trts = c("260-hatch", "337-hatch", "337-3rd", "337-4th")
B_labels = c("260-hatch"="26°C @ hatch","337-hatch"="33±7°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th")
B_colors = c("260-hatch"="#00C2D1", "337-hatch"="#710A36", "337-3rd"="#C23C1E", "337-4th"="#F3922B")
#CD133F


# C: same DTs, different NTs
C_trts = c("267-hatch", "419-hatch", "337-hatch", "433-hatch")
C_labels = c("260-hatch"="26/26 (26±0°C)", "267-hatch"="33/19 (26±7°C)","419-hatch"="40/19 (29.5±10.5°C)", "337-hatch"="40/26 (33±7°C)", "433-hatch"="40/33 (36.5±3.5°C)")
C_colors = c("267-hatch"="#F4B942", "419-hatch"="#4059AD", "337-hatch"="#6B9AC4", "433-hatch"="#97D8C4")
  # tbh not sure what i want to compare with here... 260 or 267?

# for survival but lowkey need to edit it below
A_hex = c("#00C2D1","#1929B3", "#F9C639", "#710A36")
B_hex = c("#00C2D1", "#710A36", "#C23C1E", "#F3922B")
C_hex = c("#F4B942", "#4059AD", "#6B9AC4", "#97D8C4")
```


# 2.1 validations to run lol

- 260 diurnal vs 26 C
- 25 col vs 260 ctrls generally
- batch effects (maybe clear this in models...)
- effects of lab diet vs TB diet in lab bugs


# 2.2 all individuals vs average dev
```{r}
# PROOF OF CONCEPT.. basically need to yoink this code when plotting all masses vs avg

# long_devtime <- data_all %>% 
#   filter(!(final.fate == "ignore" | final.fate == "pmd")) %>%
#   #data_pup %>% 
#   select(-starts_with(c("h", "date", "mass", "reason."))) %>%
#   mutate("2nd" = jdate.2nd - jdate.hatch,
#          "3rd" = jdate.3rd - jdate.hatch,
#          "4th" = jdate.4th - jdate.hatch,
#          "5th" = jdate.5th - jdate.hatch,
#          "6th" = jdate.6th - jdate.hatch,
#          "7th" = jdate.7th - jdate.hatch, 
#          wander = jdate.wander - jdate.hatch,
#          pupa = jdate.pupa - jdate.wander,
#          #eclose = jdate.eclose - jdate.15
#          eclose = jdate.eclose - jdate.pupa) %>%
#   select(-starts_with("jdate.")) %>%
#   pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupa", "eclose"), # pivot differences by stage
#                 names_to = "instar",
#                 values_to = "days.to") %>%
#   drop_na(days.to)
# 
# long_mass <- long_all %>%
#   filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "6th" | instar == "7th" | instar == "wander") 
# # %>%
# #   group_by(pop, instar, trt.stage) %>%
# #   mutate(logmass = log(mass)) %>%
# #   summarise(avg.mass = mean(na.omit(mass)),
# #             se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
# #             avg.logmass = mean(na.omit(logmass)),
# #             se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
# #             n=n())
# 
# long_alldev <- merge(long_devtime, long_mass, all = T)
# long_alldev <- long_alldev %>% filter(pop!="F1") %>% filter(pop!="ctrl")
# 
# # temps_l <- l_stats %>% filter(trt.stage == "260-hatch" | trt.stage =="267-hatch" | trt.stage == "330-hatch" | trt.stage == "337-hatch")
# # instar_l <- l_stats %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
# # NT_l <- l_stats %>% filter(trt.stage == "260-hatch" | trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch")
# 
# ## sth about the grouping is wroooong :~)
# 
# long_alldev %>% 
#   filter.temps()%>%
#   ggplot(aes(x=days.to, y=mass)) + # cant just use jdate LOL need to match stuff up. do math (time to)
#   geom_line(aes(color=treatment, group=ID)) +
#   scale_color_brewer(palette="BrBG") + theme_bw() +
#   facet_wrap(~pop)
# 
# long_alldev %>% 
#   filter.accum() %>% filter(pop == "lab") %>%
#   ggplot(aes(x=days.to, y=mass)) +
#   geom_line(aes(color=trt.stage, group=ID)) +
#   scale_color_brewer(palette="BrBG") + theme_bw()
#   
# long_alldev %>% 
#   filter.NTs() %>%
#   ggplot(aes(x=days.to, y=mass)) + 
#   geom_line(aes(color=treatment, group=ID)) +
#   scale_color_brewer(palette="BrBG") + theme_bw() +
#   facet_wrap(~pop)

```



# 2.3 look at mass gain/instar time on same plot

## a. calculate stats for adult/larva subsets

adults need to be separated out to be able to look at sex differences

```{r}
#long_alldev

# separate into larv (hatch-5th) and adults
l_devavg <- long_alldev %>%
  filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
  group_by(pop, instar, trt.stage) %>%
    summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))),
            n=n())

a_devavg <- long_alldev %>% 
  group_by(pop, trt.stage, sex, instar) %>%
  filter(instar == "pupa" | instar == "eclose") %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))),
            n=n())

# create subsets + plot
## temps
tempsL_devavg <- l_devavg %>% filter.temps()
tempsA_devavg <- a_devavg %>% filter.temps()
tempsL_alldev <- long_alldev %>% filter.temps()
tempsA_alldev <- long_alldev %>% filter.temps() %>% filter(instar == "pupa" | instar == "eclose")

ggplot(tempsL_devavg, aes(color=trt.stage, x=avg.daysto, y=avg.mass)) +
  geom_line(lty="dashed", lwd=1.5) + geom_point() +
  geom_line(data=tempsL_alldev, aes(x=days.to, y=mass, group=ID), alpha=0.1) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  geom_errorbarh(aes(xmin = avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  #scale_color_brewer(palette="RdBu") + 
  theme_bw() + xlim(0, 25) +
  scale_color_manual(values=c("#0571B0", "#92C5DE", "#F4A582", "#CA0020"), labels=A_labels) +
  labs(title="(temps) average growth (mg/day)", color="treatment")

ggplot(tempsA_devavg, aes(color=trt.stage, x=avg.daysto, y=avg.mass)) +
  geom_line(lty="dashed", lwd=1.5) + geom_point() +
  geom_line(data=tempsA_alldev, aes(x=days.to, y=mass, group=ID), alpha=0.1) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  geom_errorbarh(aes(xmin = avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  #scale_color_brewer(palette="RdBu") + 
  theme_bw() + xlim(0, 25) +
  scale_color_manual(values=c("#0571B0", "#92C5DE", "#F4A582", "#CA0020"), labels=A_labels) +
  labs(title="(temps) average growth (mg/day)", color="treatment")

## accum
accL_devavg <- l_devavg %>% filter.accum()
accA_devavg <- a_devavg %>% filter.accum()
accL_alldev <- long_alldev %>% filter.accum()

ggplot(accL_devavg, aes(color=trt.stage, x=avg.daysto, y=avg.mass)) +
  geom_line(lty="dashed", lwd=1.5) + geom_point() +
  geom_line(data=accL_alldev, aes(x=days.to, y=mass, group=ID), alpha=0.1) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  geom_errorbarh(aes(xmin = avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  scale_color_brewer(palette="PuOr") + 
  theme_bw() + xlim(0, 25) +
  scale_color_manual(values=c("#998EC3", "#FDB863", "#E66101", "#B35806"), labels=B_labels) +
  labs(title="(accumulation) average growth (mg/day)", color="treatment")

## NTs
NTsL_devavg <- l_devavg %>% filter.NTs() %>% filter(trt.stage != "267-hatch")
NTsA_devavg <- a_devavg %>% filter.NTs() %>% filter(trt.stage != "267-hatch")
NTsL_alldev <- long_alldev %>% filter.NTs() %>% filter(trt.stage != "267-hatch")

ggplot(NTsL_devavg, aes(color=trt.stage, x=avg.daysto, y=avg.mass)) +
  geom_line(lty="dashed", lwd=1.5) + geom_point() +
  geom_line(data=NTsL_alldev, aes(x=days.to, y=mass, group=ID), alpha=0.1) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  geom_errorbarh(aes(xmin = avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  #scale_color_brewer(palette="BrBG") + 
  theme_bw() + xlim(0, 25) +
  facet_wrap(~pop) +
  scale_color_manual(values=c("#018571", "#80CDC1", "#E66101", "#A6611A"), labels=C_labels) +
  labs(title="(NTs) average growth (mg/day)", color="treatment")

```



(scrap stuff below here)

```{r}
### pre-filtering out messy data points
  # 231010 so the sup data still doesnt look very good...
  # this removes F1s, sups, slow things

long_f <- long_all %>%
  filter(pop != "F1") %>%
  filter(final.fate != "pmd") %>%
  #filter(if.sup == "N") %>%
  filter(is.na(ignore.reason))

data_f <- data_all %>%
  filter(pop != "F1") %>%
  filter(final.fate != "pmd") %>%
  #filter(if.sup == "N") %>%
  filter(is.na(ignore.reason))
  
```


```{r}
### mass stats

l_mass <- long_f %>%
  filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
  group_by(pop, instar, trt.stage) %>%
  mutate(logmass = log(mass)) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

a_mass <- long_f %>% 
  group_by(pop, trt.stage, sex, instar) %>%
  filter(instar == "pupa" | instar == "eclose") %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            n=n())
```


```{r}
### development time stats

# calculate for all groups
all_devtime <- data_f %>% 
  filter(!(final.fate == "ignore" | final.fate == "pmd")) %>%
  #data_pup %>% 
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         # "6th" = jdate.6th - jdate.hatch,
         # "7th" = jdate.7th - jdate.hatch, 
         wander = jdate.wander - jdate.hatch,
         pupa = jdate.pupa - jdate.wander,
         #eclose = jdate.eclose - jdate.15
         eclose = jdate.eclose - jdate.pupa) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupa", "eclose"), # pivot differences by stage
                names_to = "instar",
                values_to = "days.to") %>%
  drop_na(days.to)

# subset larva
l_dev <- all_devtime %>% 
  filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))))


# subset adults
a_dev <- all_devtime %>% 
  filter(instar == "pupa" | instar == "eclose") %>%
  group_by(pop, trt.stage, sex, instar) %>%
  summarise(avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))))
```

## b. resubset data into expts
```{r}
# combine mass/time data
l_stats <- merge(l_mass, l_dev, all = T) %>%
  #filter(pop != "F1" & (instar != "6th" | instar != "7th")) %>% # bc tbh 6/7 does not look very good lol
  drop_na(avg.daysto) # drop incomplete data while data is being collected

a_stats <- merge(a_mass, a_dev, all = T) %>%
  #filter(pop != "F1" & (instar != "6th" | instar != "7th")) %>% # bc tbh this does not look very good lol
  drop_na(sex) %>% # drop incomplete data while data is being collected
  drop_na() # drop lines w/ only 1-2 individs


# subset: effect of mean/fluct temps
  # tbh i just should do these by what i want LOL
temps_l <- l_stats %>% filter(trt.stage == "260-hatch" | trt.stage =="267-hatch" | trt.stage == "330-hatch" | trt.stage == "337-hatch")
temps_a <- a_stats %>% filter(trt.stage == "260-hatch" | trt.stage =="267-hatch" | trt.stage == "330-hatch" | trt.stage == "337-hatch")

# subset: effect of temp x instar (accumulation)
instar_l <- l_stats %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
instar_l$trt.stage <- factor(instar_l$trt.stage, levels = c("260-hatch", "337-hatch", "337-3rd", "337-4th"))

instar_a <- a_stats %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
instar_a$trt.stage <- factor(instar_a$trt.stage, levels = c("260-hatch", "337-hatch", "337-3rd", "337-4th"))

# subset: effect of different NTs
NT_l <- l_stats %>% filter(trt.stage == "260-hatch" | trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch")
NT_a <- a_stats %>% filter(trt.stage == "260-hatch" | trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch")

# define shapes to use for instars in ggplot
l_shapes <- c(17, 15, 16, 18)
#l_shapes <- c(17, 15, 16, 9, 12, 10)
  #(16, 17, 15, 18, 10, 9)
  # possibly use shape 8
```


## c. plots

```{r}

### temp figs

temps_l %>%
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage, lty = pop)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=l_shapes)+
  geom_line(aes(group = interaction(pop, trt.stage))) +
  #facet_wrap(~pop) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels=A_labels, values = A_colors) + theme_bw() +
  scale_linetype_manual(breaks=c("lab", "field"), values=c(1, 2))

temps_a %>%
  ggplot(aes(x = avg.daysto, y = avg.mass, color=trt.stage, shape=instar, lty=pop)) +
  geom_point(size=3, alpha = 0.75) +
  geom_line(aes(group = interaction(pop, trt.stage))) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  facet_wrap(~sex) +
  labs(title = "(temp) avg adult growth", y = "avg mass (mg)", x = "days since last stage", color = "treatment") +
  scale_color_manual(values = A_colors, labels=A_labels) + theme_bw() +
  scale_linetype_manual(breaks=c("lab", "field"), values=c(1, 2))

```

```{r}

### instar figs

instar_l %>% 
  filter(pop == "lab") %>%
  drop_na() %>% #6ths make this break
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha = 0.75) +
  scale_shape_manual(values=l_shapes)+
  geom_line(aes(group = trt.stage)) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(instar) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels = B_labels, values = B_colors) + theme_bw()
    ## looks kinda wrong w rerun analyses? (lines dont seem like they grouped right..)

# adult mass/dev
instar_a %>% 
  filter(pop == "lab") %>%
  ggplot(aes(x = avg.daysto, y = avg.mass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha = 0.75) +
  geom_line(aes(group = interaction(pop, trt.stage))) +
  facet_wrap(~sex) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  scale_color_manual(labels = B_labels, values = B_colors) + theme_bw() +
  labs(title = "(instar) avg adult growth", y = "avg mass (mg)", x = "days since last stage", color="treatment")
```


```{r}

### NT figs
# only 4 field lol for adults rn


# NT figs w/ 26-7
NT_l %>%
  filter(trt.stage != "260-hatch") %>%
  #filter(instar != "6th") %>% # this is wonky
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage, lty = pop)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=l_shapes) + # use shape 8 if including 6ths
  geom_line(aes(group = interaction(pop, trt.stage))) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(N temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels=C_labels, values = C_colors) + theme_bw() +
  scale_linetype_manual(breaks=c("lab", "field"), values=c(1, 2))


# NT figs w/o 26-7
  # need to adjust the scales bc theres only 3 things
NT_l %>%
  filter(trt.stage != "267-hatch" & trt.stage != "260-hatch") %>%
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage, lty = pop)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=l_shapes)+
  geom_line(aes(group = interaction(pop, trt.stage))) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(N temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels=C_labels, values = C_colors) + theme_bw() +
  scale_linetype_manual(breaks=c("lab", "field"), values=c(1, 2))

# adult mass/dev
NT_a %>%
  filter(trt.stage != "267-hatch" & trt.stage != "260-hatch") %>%
  ggplot(aes(x = avg.daysto, y = avg.mass, shape=instar, color=trt.stage, lty = pop)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=c(16, 17, 15, 18)) + # use shape 8 if including 6ths
  geom_line(aes(group = interaction(pop, trt.stage))) +
  facet_wrap(~sex) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto), height = 175) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  labs(title = "(N temp) avg larval growth", x = "days since hatching", y = "avg mass (mg)", color = "treatment") +
  scale_color_manual(labels=C_labels, values = C_colors) + theme_bw() +
  scale_linetype_manual(breaks=c("lab", "field"), values=c(1, 2))
```



# 3. look at mass gain/instar time on separate plots

## a. calculate stats for larva/adults

### larva
```{r}
# TODO: need to also do math for 6ths
  # jk 231011: taking out the sups bc it makes the math weird

# repivot for easier mathing
  # this is basically all_devtime except w/ renamed "timeto" bc i cant figure out how to make the paste work
all_alt <- data_f %>%
  mutate(timeto.3rd = jdate.3rd-jdate.hatch,
         timeto.4th = jdate.4th-jdate.hatch,
         timeto.5th = jdate.5th-jdate.hatch,
         #timeto.6th = jdate.6th-jdate.hatch,
         timeto.wander = jdate.wander-jdate.hatch,
         timeto.pupa = jdate.pupa-jdate.hatch,
         timeto.eclose = jdate.eclose-jdate.hatch) %>%
  select(-(starts_with(c("date.", "jdate.")))) %>%
  pivot_longer(cols = starts_with(c("timeto", "mass", "h")),
               names_to = c(".value", "instar"),
               #names_sep = ".",
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  rename(molt.status = h)
# drop_na(jdate) # drops NA's if an individual didnt reach a certain stage

# summary statistics
alt_sumstats <- all_alt %>%
  mutate(mass = mass/1000) %>% # convert to g
  mutate(log.mass = log(mass)) %>% # u get neg values if using mass_g
  filter(final.fate != "ignore") %>% 
  filter(instar != "stuck" & instar != "2nd" & instar != "died") %>%
  group_by(pop, meanT, flucT, instar.enter, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(log.mass)),
            se.logmass = sd(na.omit(log.mass))/sqrt(length(na.omit(log.mass))),
            avg.time = mean(na.omit(timeto)),
            se.time = sd(na.omit(timeto))/sqrt(length(na.omit(timeto))),
            n=n())

# fix factor levels
alt_sumstats$instar <- factor(alt_sumstats$instar, levels=c("3rd", "4th", "5th", "wander", "pupa", "eclose"))
alt_sumstats$instar.enter <- factor(alt_sumstats$instar.enter, levels=c("hatch", "3rd", "4th"))
```

### adult

```{r}
# 231011 removing 6ths for now

# calc stats
alt_sexstats <- all_alt %>%
  mutate(mass = mass/1000) %>% # convert to g
  filter(final.fate != "ignore" & sex != "unk") %>% 
  filter(instar == "pupa" | instar == "eclose") %>%
  group_by(pop, meanT, flucT, instar.enter, instar, sex) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.time = mean(na.omit(timeto)),
            se.time = sd(na.omit(timeto))/sqrt(length(na.omit(timeto))),
            n=n())

# fix factor levels
alt_sexstats$instar <- factor(alt_sexstats$instar, levels=c("3rd", "4th", "5th", "wander", "pupa", "eclose"))
alt_sexstats$instar.enter <- factor(alt_sexstats$instar.enter, levels=c("hatch", "3rd", "4th"))
```



## b. group stats by expts (temp, tempxins, NTs)
```{r}
## ugh.. wrong comparison (instar should be to 26)

# larva
temps_la <- filter(alt_sumstats, ((flucT == 0 | flucT == 7) & instar.enter == "hatch"))
instar_la <- filter(alt_sumstats, (meanT == 33 | (meanT == 26 & flucT == 0)))
NT_la <- filter(alt_sumstats, (((meanT == 29.5 | meanT == 36.5) | flucT == 7) & instar.enter == "hatch"))

# adults
temps_aa <- filter(alt_sexstats, ((flucT == 0 | flucT == 7) & instar.enter == "hatch"))
instar_aa <- filter(alt_sexstats, (meanT == 33 | (meanT == 26 & flucT == 0)))
NT_aa <- filter(alt_sexstats, (((meanT == 29.5 | meanT == 36.5) | flucT == 7) & instar.enter == "hatch"))
```


## c. plots

### larval instars
```{r}
### start saving new figs from here

### temp mass
temps_la %>%
  filter(pop == "lab") %>%
  ggplot(aes(x = meanT, y = avg.mass, color = factor(flucT))) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15) +
  facet_wrap(pop~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg) - lab", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))

temps_la %>%
  filter(pop == "field") %>%
  ggplot(aes(x = meanT, y = avg.mass, color = factor(flucT))) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15) +
  facet_wrap(pop~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg) - field", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))

### temp time
temps_la %>%
  filter(pop == "lab") %>%
  ggplot(aes(x = meanT, y = avg.time, color = factor(flucT))) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "temp: time to instar (days) - lab", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))

temps_la %>%
  filter(pop == "field") %>%
  ggplot(aes(x = meanT, y = avg.time, color = factor(flucT))) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "temp: time to instar (days) - field", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))
```

```{r}
jitter <- position_dodge(width=1)

### instar mass
instar_la %>%
  filter(pop == "lab") %>%
  filter(!(meanT == 33 & flucT == 0 & instar.enter == "hatch")) %>%
  ggplot(aes(x = meanT, y = avg.mass, color = instar.enter, shape = factor(flucT))) +
  geom_line() +
  geom_point(position = jitter) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15, position = jitter) +
  facet_wrap(~instar, scales = "free") +
  theme_bw() +
  labs(y = "instar: mass at instar (mg)", color = "instar entered", shape = "fluctuation" )

### instar time
instar_la %>%
  filter(pop == "lab") %>%
  filter(!(meanT == 33 & flucT == 0 & instar.enter == "hatch")) %>%
  ggplot(aes(x = meanT, y = avg.time, color = instar.enter, shape = factor(flucT))) +
  geom_line() +
  geom_point(position = jitter) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15, position = jitter) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "instar: time to instar (days)", color = "instar entered", shape = "fluctuation")
```

```{r}
#if facet_wrap(scales = "free"), set error bar width = 0.5

### NT mass
NT_la %>%
  ggplot(aes(x = meanT, y = avg.mass, color=factor(flucT), shape = pop)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
  facet_wrap(~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))

# NT_la %>%
#   filter(pop == "lab") %>%
#   ggplot(aes(x = meanT, y = avg.mass, color=factor(flucT))) +
#   geom_point(size=1.5) +
#   geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
#   facet_wrap(~instar) +
#   theme_bw() +
#   labs(y = "temp: mass at instar (mg) - lab", color = "fluctuation") +
#   scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
# 
# NT_la %>%
#   filter(pop == "field") %>%
#   ggplot(aes(x = meanT, y = avg.mass, color=factor(flucT))) +
#   geom_point(size=1.5) +
#   geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
#   facet_wrap(~instar) +
#   theme_bw() +
#   labs(y = "temp: mass at instar (mg) - field", color = "fluctuation") +
#   scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))


### NT time
NT_la %>%
  filter(pop == "lab") %>%
  ggplot(aes(x = meanT, y = avg.time, color=factor(flucT), shape = pop)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "temp: time to instar (days)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))

# NT_la %>%
#   filter(pop == "lab") %>%
#   ggplot(aes(x = meanT, y = avg.time, color=factor(flucT))) +
#   geom_point(size=1.5) +
#   geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
#   theme_bw() +
#   facet_wrap(~instar) +
#   labs(y = "temp: time to instar (days) - lab", color = "fluctuation") +
#   scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
# 
# NT_la %>%
#   filter(pop == "field") %>%
#   ggplot(aes(x = meanT, y = avg.time, color=factor(flucT))) +
#   geom_point(size=1.5) +
#   geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
#   theme_bw() +
#   facet_wrap(~instar) +
#   labs(y = "temp: time to instar (days) - field", color = "fluctuation") +
#   scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
```


### adult instars
```{r}
### temp mass
temps_aa %>%
  ggplot(aes(x = meanT, y = avg.mass, color = sex)) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15) +
  facet_wrap(pop~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", lty = "fluctuation", shape = "fluctuation")

### temp time
temps_aa %>%
  ggplot(aes(x = meanT, y = avg.time, color=sex)) +
  geom_line(aes(lty=factor(flucT))) +
  geom_point(aes(shape = factor(flucT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15) +
  theme_bw() +
  facet_wrap(pop~instar, scales = "free") +
  labs(y = "temp: time to instar (days)", lty = "fluctuation", shape = "fluctuation")
```

```{r}
# resume fixing/adding field stuff in here

jitter <- position_dodge(width=0.75)

### instar mass
instar_aa %>%
  filter(pop == "lab") %>%
  ggplot(aes(x = flucT, y = avg.mass, color = instar.enter)) +
  #geom_line() +
  geom_point(aes(shape = sex), size = 2.5, position = jitter) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15, position = jitter) +
  facet_wrap(instar~sex, scales = "free") +
  theme_bw() +
  labs(y = "instar: mass at instar (mg)", color = "instar entered")

### instar time
instar_aa %>%
  filter(pop == "lab") %>%
  ggplot(aes(x = flucT, y = avg.time, color = instar.enter)) +
  #geom_line() +
  geom_point(aes(shape = sex), size = 2, position = jitter) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15, position = jitter) +
  theme_bw() +
  facet_wrap(instar~sex, scales = "free") +
  labs(y = "instar: time to instar (days)")


# post fixing: not sure what that 2nd dot at the bottom of everything is lol 
 # maybe the mean T?
```

```{r}
### NT mass
NT_aa %>%
  ggplot(aes(x = meanT, y = avg.mass, color=factor(flucT), shape = pop)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
  facet_wrap(instar~sex, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))


### NT time
NT_aa %>%
  ggplot(aes(x = meanT, y = avg.time, color=factor(flucT), shape = pop)) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
  theme_bw() +
  facet_wrap(instar~sex, scales = "free") +
  labs(y = "temp: time to instar (days)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
```


# 4. survival curves

## a. prep data

checking survival to pupation from hatching

```{r}
# censor and create subsets

# censor data
surv_pup <-  data_all %>% 
  filter(final.fate != "ignore") %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

# subsets & specify levels
surv_temp <- surv_pup %>%
  filter(flucT == 0 | (flucT == 7 & instar.enter == "hatch")) %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "267-hatch", "330-hatch", "337-hatch")))
  

surv_instar <- surv_pup %>%
  filter((meanT == 33 & flucT == 7) | (meanT == 26 & flucT == 0)) %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "337-hatch", "337-3rd", "337-4th")))

# surv_NT <- surv_pup %>%
#   filter((meanT == 36.5 & flucT == 3.5) | (meanT == 29.5 & flucT == 10.5) | (meanT == 33 & flucT == 7) | (meanT == 26 & flucT == 7) & instar.enter == "hatch") %>%
#   filter(trt.stage != "260-hatch") %>%
#   mutate(trt.stage = factor(trt.stage, c("267-hatch", "419-hatch", "337-hatch", "433-hatch")))

# idk how to get field + lab onto one plot yet soooo

surv_pupNT <- surv_pup

surv_pupNT$trt.stage[surv_pupNT$trt.stage == "337-hatch"] <- "426-hatch"

survf_NT <- surv_pupNT %>%
  filter((meanT == 36.5 & flucT == 3.5) | (meanT == 29.5 & flucT == 10.5) | (meanT == 33 & flucT == 7) | (meanT == 26 & flucT == 0) & instar.enter == "hatch") %>%
  filter(pop == "field") %>%
  #filter(trt.stage != "260-hatch") %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "419-hatch", "426-hatch", "433-hatch")))

survl_NT <- surv_pupNT %>%
  filter((meanT == 36.5 & flucT == 3.5) | (meanT == 29.5 & flucT == 10.5) | (meanT == 33 & flucT == 7) | (meanT == 26 & flucT == 7) | (meanT == 26 & flucT == 0) & instar.enter == "hatch") %>%
  filter(pop == "lab") %>%
  #filter(trt.stage != "260-hatch") %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "267-hatch", "419-hatch", "426-hatch", "433-hatch")))

```

## b. create survival objects and fits

```{r}
temp_sobj <- Surv(surv_temp$timeto.exit, surv_temp$censor)
temp_sfit <- survfit(temp_sobj ~ trt.stage, data = surv_temp)

ins_sobj <- Surv(surv_instar$timeto.exit, surv_instar$censor)
ins_sfit <- survfit(ins_sobj ~ trt.stage, data = surv_instar)

# NT_sobj <- Surv(surv_NT$timeto.exit, surv_NT$censor)
# NT_sfit <- survfit(NT_sobj ~ trt.stage, data = surv_NT)

NTf_sobj <- Surv(survf_NT$timeto.exit, survf_NT$censor)
NTf_sfit <- survfit(NTf_sobj ~ trt.stage, data = survf_NT)

NTl_sobj <- Surv(survl_NT$timeto.exit, survl_NT$censor)
NTl_sfit <- survfit(NTl_sobj ~ trt.stage, data = survl_NT)
```

## c. plot survival curves

```{r}
#### these include sups + slow pupations (no slow deaths?)
  # add conf levels

# define palette (idk how to make the levels work)
# A_hex = c("#00C2D1","#1929B3", "#F9C639", "#710A36")
# B_hex = c("#00C2D1", "#710A36", "#C23C1E", "#F3922B")
# C_hex = c("#F4B942", "#4059AD", "#6B9AC4", "#97D8C4")

# A_labels <- c("260-hatch"="26°C", "267-hatch"="26±7°C", "330-hatch"="33°C", "337-hatch"="33±7°C")
# B_labels = c("260-hatch"="26°C @ hatch","337-hatch"="33±7°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th")
# C_labels = c("267-hatch"="26±7°C (33/19)","419-hatch"="29.5±10.5°C (40/19)", "337-hatch"="33±7°C (40/26)", "433-hatch"="36.5±3.5°C (40/33)")


# temps
  # 260, 267, 330, 337
ggsurvplot(fit = temp_sfit, data=surv_temp,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp x survival from hatching",
           #palette = c("#00C2D1","#1929B3", "#F9C639", "#710A36"),
           legend.title = "treatment", 
           #legend.labs = A_labels
           )

# instar
ggsurvplot(fit = ins_sfit, data=surv_instar,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp:instar x survival from hatching", 
           #palette = c("#00C2D1", "#F3922B", "#C23C1E", "#710A36"),
           legend.title = "treatment-stage entered", 
           #legend.labs = c("26°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th", "337-hatch"="33±7°C @ hatch")
           )
  # 260, 3rd, 4th, hatch


# NTs
ggsurvplot(fit = NT_sfit, data=surv_NT,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NT x survival from hatching", 
           #palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           #legend.labs = c("26±7°C (33/19)", "33±7°C (40/26)", "29.5±10.5°C (40/19)", "36.5±3.5°C (40/33)"),
           #strata = pop
           )


# NTs
ggsurvplot(fit = NTf_sfit, data=survf_NT,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NT x survival from hatching - field", 
           #palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           #legend.labs = c("26±7°C (33/19)", "33±7°C (40/26)", "29.5±10.5°C (40/19)", "36.5±3.5°C (40/33)"),
           #strata = pop
           )

ggsurvplot(fit = NTl_sfit, data=survl_NT,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NT x survival from hatching - lab", 
           #palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           #legend.labs = c("26±7°C (33/19)", "33±7°C (40/26)", "29.5±10.5°C (40/19)", "36.5±3.5°C (40/33)"),
           #strata = pop
           )
  # 267, 337, 419, 433

```




# 5. misc: old 2013 data

## a. prep data

## b. plot data

