---
title: "NTW prelim data analysis"
output: html_notebook
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, echo = TRUE)

library(dplyr)
library(ggplot2)
library(tidyr)
#library(survival) # for kaplan-meier
#library(survminer) # maybe, for visualisation
```

# information & todo

here is code for some analysis of preliminary NTW experiments. as usual i keep making 50 different new files for things bc im indecisive about how to do things

TODO list: (search `// TODO`)

* [done] create `daysto.` columns (from hatching) for all instars/development timepoints
* consolidate todo items up here lol (then redistribute to below... or vice versa)


TODO / statistics to measure:

* [done] log average weight at instars
  * h/m considerations... (esp for grp B)
  * from.E considerations... (only pick 0 and 1?)
* [done for A, B] average length of instars
* [doneish] survival
  * kaplan meier curves -- see:
    * https://rviews.rstudio.com/2017/09/25/survival-analysis-with-r/
    * https://www.emilyzabor.com/tutorials/survival_analysis_in_r_tutorial.html#Packages 
* [done] A vs B comparisons (26 constant)
  * pool relevant A+B individuals --> then stats again?
* who got stuck
* pupa development (looking @ % none? idr what i originally meant here)
* dv coloring (with B only)
* graph a bunch of growth lines + avg growth
* plot time vs mass (dot plot w/ error bars on both sides)


experiment groups:

* A: 2x2 of mean temp = 26 or 33, with 0 degC fluct or 7 degC fluct
* B: 
  * redoing 337 from hatching (A)
  * 2x2(ish) of entering 260 or 337 treatment at different times (hatch, 3rd, 4th)

# load packages

```{r load libraries}
# library(dplyr)
# library(ggplot2)
# library(tidyr)

# for some reason this broke on 230321 so now use the setup chunk

#library(data.table) #for setDT (forget why i was thinking about this one)
#library(survival) # for kaplan-meier
```

# load in data and clean, labels

outputs

* data_all, data_notes, data (surv to pup)
  * maybe rename data to data_ec (surv to eclosion)
* data_A, data_B

```{r data preprocess + clean}
data_all <- read.csv("~/Documents/projects/data/ntw_data/development.csv", header = T)

# save wordier notes elsewhere and remove from analysis dataset
data_notes <- select(data_all, c("from.E", "treatment", "ID", "expt.group", "pupa.deformities", "notes"))
data <- select(data_all, -c("notes", "mass.toast", "pupa.deformities", "mass.died")) # not sure if 'notes' column will mess things up

# remove individuals that didn't survive to pupation
data <- data %>% filter(reason.pmd == "" | reason.pmd == "pupa" | reason.pmd == "toast")

# create julian date columns, additional column fixing
data <- data %>% mutate(across(starts_with("date."),  as.Date, format = "%m/%d/%y")) %>%
  mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
  #relocate(starts_with("jdate."), .after = starts_with("date.")) %>% # this is not working lol
  mutate(across(starts_with("jdate."), as.numeric),
         across(starts_with("mass."), as.numeric), # fix adult.mass
         treatment = as.factor(treatment),
         trt.stage = paste(treatment,instar.enter, sep = "-"))

# pivot and make subsets
long <- data %>% select(-(starts_with("date."))) %>%
                 pivot_longer(cols = starts_with(c("jdate", "mass", "h")),
                        names_to = c(".value", "instar"),
                        #names_sep = ".",
                        values_drop_na = TRUE,
                        names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
                 rename(molt.status = h)

long <- long[long$instar != "in",] # from 'instar.in.trt'?

# create subsets
data_A <- long[long$expt.group == "A",]
data_B <- long[long$expt.group == "B",]

# labels
A_colors <- c("#abd9e9", "#2c7bb6", "#fdae61", "#d7191c")
A_labels <- c("26°C", "26±7°C", "33°C", "33±7°C")

B_trts = c("260-hatch", "337-4th", "337-3rd", "337-hatch")
B_colors = c("#abd9e9", "#fdbe85", "#fd8d3c", "#d94701")
B_labels = c("26°C @ hatch", "33±7°C @ 4th", "33±7°C @ 3rd", "33±7°C @ hatch")

```

# group A

## pt A analysis

```{r A stats}
# for boxplot
growth_A2 <- data_A %>% group_by(treatment, instar) %>%
    filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
    mutate(log.mass = log(mass))

# for boxplot (lots of extra data but its fine)
adultmass_A2 <- data_A %>% group_by(treatment, sex) %>%
  filter(instar == "pupa" | instar == "adult")

# for boxplot (pivot wider later for some reason?)
devtime_A2 <- data %>% filter(expt.group == "A") %>%
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         wander = jdate.wander - jdate.hatch,
         pupate = jdate.pupate - jdate.wander,
         #eclose = jdate.eclose - jdate.15,
         eclose = jdate.eclose - jdate.pupate) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupate", "eclose"), # pivot differences by stage
                names_to = "stage",
                values_to = "days.to",
                values_drop_na = T) %>%
  group_by(treatment, stage)


stats_A2 <- merge(growth_A2, devtime_A2, all = T) %>%
  rename(instar.mass = instar, # mass
         instar.duration = stage) # time
```

## pt A plotting

todo:

* [done, not B] larval growth/dev (maybe combine w/ part B at some point)
* [done] adult masses

```{r A plots}
## larval instar masses

massA_plot <- ggplot(data = stats_A2, aes(x = instar.mass, y = log.mass))

# this is okay
# massA_plot + geom_boxplot(aes(fill = treatment)) +
#   labs(title = "mass at start of instar", x = "instar", y = "log(mass (mg))") +
#   scale_fill_manual(values = A_colors) + theme_bw()
#   # // TODO: rename treatments/key

ggplot(data = stats_A2, aes(y = log.mass)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "mass at start of instar", x = "instar", y = "log(mass (mg))") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw() +
  facet_grid(~instar.mass, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))


## adult masses

adultA_plot <- ggplot(data = adultmass_A2, aes(x = factor(instar, level = c("pupa", "adult")), y = mass))

adultA_plot + geom_boxplot(aes(fill = treatment)) + facet_wrap(~sex) +
  labs(title = "adult masses", y = "mass (mg)", x = "stage") +
  scale_fill_manual(values = A_colors) + theme_bw()
  # // TODO: rename treatments/key


## dev time (larval)
# done
devA.l_plot <- stats_A2 %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(x = instar.duration, y = days.to))

# devA.l_plot + geom_boxplot(aes(fill = treatment)) +
#   labs(title = "larval instar development times", x = "stage", y = "days after hatching") +
#   scale_fill_manual(values = A_colors) + theme_bw()

stats_A2 %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(y = days.to)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "larval instar development times", x = "stage", y = "days after hatching") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw() +
  facet_grid(~instar.duration, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))



## dev time (post-pupation)
# done
devA.a_plot <- stats_A2 %>% filter(instar.duration == "eclose" | instar.duration == "pupate") %>%
  ggplot(aes(x = factor(instar.duration, levels = c("pupate", "eclose")), y = days.to))

devA.a_plot + geom_boxplot(aes(fill = treatment)) +
  labs(title = "post-pupation development times", x = "stage", y = "days after previous stage") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw()

```

# group B

* todo: look at this as a 2x2 (@ hatch, @ 3rd)

## pt B analysis

* [see below] check for batch effects (260 and 337)
  * [done] merge 260A and 337A with B if its ok
* [done?] similar analyses for these groups as in A

```{r B stats}
# just copied over code from pt A and replaced w Bs
growth_B <- data_B %>% group_by(trt.stage) %>%
    filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
    mutate(log.mass = log(mass))

adultmass_B <- data_B %>% group_by(trt.stage, sex) %>%
  filter(instar == "pupa" | instar == "adult") %>%
  mutate(log.mass = log(mass))

devtime_B <- data %>% filter(expt.group == "B") %>%
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         wander = jdate.wander - jdate.hatch,
         pupate = jdate.pupate - jdate.wander,
         #eclose = jdate.eclose - jdate.15
         eclose = jdate.eclose - jdate.pupate) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupate", "eclose"), # pivot differences by stage
                names_to = "stage",
                values_to = "days.to",
                values_drop_na = T) %>%
  group_by(trt.stage, stage)

stats_B <- merge(growth_B, devtime_B, all = T) %>%
  rename(instar.mass = instar, # mass
         instar.duration = stage) # time
```

## B plotting

(this will be the same as A)

```{r B plots}
## larval masses
massB_plot <- ggplot(data = stats_B, aes(x = instar.mass, y = log.mass))

# this is okay
# massB_plot + geom_boxplot(aes(fill = factor(trt.stage, level= B_trts))) +
#   labs(title = "mass at start of instar", x = "instar", y = "log(mass (mg))", fill = "treatment") +
#   scale_fill_manual(values = B_colors) + theme_bw()

ggplot(data = stats_B, aes(y = log.mass)) + geom_boxplot(aes(fill = factor(trt.stage, level= B_trts))) +
  labs(title = "mass at start of instar", x = "instar", y = "log(mass (mg))", fill = "treatment") +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw() +
  facet_grid(~instar.mass, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))


## adult masses (data tbd)
# adultB_plot <- ggplot(data = adultmass_B, aes(x = factor(instar, level = c("pupa", "adult")), y = mass))
# 
# adultB_plot + geom_boxplot(aes(fill = factor(trt.stage, level= B_trts))) + facet_wrap(~sex) +
#   labs(title = "adult masses", y = "mass (mg)", x = "stage", fill = "treatment")

## dev time (larval)
# done
devB.l_plot <- stats_B %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(x = instar.duration, y = days.to))

# devB.l_plot + geom_boxplot(aes(fill = factor(trt.stage, level = B_trts))) +
#   labs(title = "larval instar development times", x = "stage", y = "days after hatching", fill = "treatment")

stats_B %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(y = days.to)) + geom_boxplot(aes(fill = factor(trt.stage, level = B_trts))) +
  labs(title = "larval instar development times", x = "stage", y = "days after hatching", fill = "treatment") +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw() +
  facet_grid(~instar.duration, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))
  # // TODO: take out the stuck ones...?


## dev time (post-pupation)
# data tbd!
# devB.a_plot <- stats_B %>% filter(instar.duration == "eclose" | instar.duration == "pupate") %>%
#   ggplot(aes(x = factor(instar.duration, levels = c("pupate", "eclose")), y = days.to))
# 
# devB.a_plot + geom_boxplot(aes(fill = treatment)) +
#   labs(title = "post-pupation development times", x = "stage", y = "days after previous stage")
```


# comparing A vs B

* to compare batch effects
* maybe do combining of A+B here [this is done in the AB pool analyses]

## AvB analysis

```{r analysis A vs B}
# copying over more code...

compare <- long %>% filter((trt.stage == "337-hatch" | trt.stage == "260-hatch") & (expt.group == "A" | expt.group == "B")) %>%
  rename(group = expt.group) %>%
  mutate(grp.trt = paste(group, trt.stage, sep = "-")) %>%
  relocate(grp.trt, .before = 1)

growth_AvB <- compare %>% group_by(grp.trt) %>%
    filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
    mutate(log.mass = log(mass))

adultmass_AvB <- compare %>% group_by(grp.trt, sex) %>%
  filter(instar == "pupa" | instar == "adult") %>%
  mutate(log.mass = log(mass))

devtime_AvB <- data %>% filter((trt.stage == "337-hatch" | trt.stage == "260-hatch") & (expt.group == "A" | expt.group == "B")) %>%
  rename(group = expt.group) %>% mutate(grp.trt = paste(group, trt.stage, sep = "-")) %>%
  relocate(grp.trt, .before = 1) %>%
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         wander = jdate.wander - jdate.hatch,
         pupate = jdate.pupate - jdate.wander,
         #eclose = jdate.eclose - jdate.15
         eclose = jdate.eclose - jdate.pupate) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupate", "eclose"), # pivot differences by stage
                names_to = "stage",
                values_to = "days.to",
                values_drop_na = T) %>%
  group_by(grp.trt, stage)

stats_AvB <- merge(growth_AvB, devtime_AvB, all = T) %>%
  rename(instar.mass = instar, # mass
         instar.duration = stage) # time

```

## AvB plotting

(do a proper test? compare the distss/means?)

```{r plotting A vs B}
# using pt A plot code

stats_AvB$grp.trt <- factor(stats_AvB$grp.trt, c("A-260-hatch", "B-260-hatch", "A-337-hatch", "B-337-hatch"))
AB_colors <- c("#abd9e9", "#2c7bb6", "#fdae61", "#d7191c")
AB_labels <- c("A 26°C", "B 26°C","A 33±7°C", "B 33±7°C")


## larval instar masses

massAvB_plot <- ggplot(data = stats_AvB, aes(x = instar.mass, y = log.mass))

ggplot(data = stats_AvB, aes(y = log.mass)) + geom_boxplot(aes(fill = grp.trt)) +
  labs(title = "mass at start of instar", x = "instar", y = "log(mass (mg))", fill = "group-treatment") +
  scale_fill_manual(labels = AB_labels, values = AB_colors) + theme_bw() +
  facet_grid(~instar.mass, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))


## adult masses (not plottable rn)

# adultAB_plot <- ggplot(data = adultmass_AB, aes(x = factor(instar, level = c("pupa", "adult")), y = mass))
# 
# adultAB_plot + geom_boxplot(aes(fill = grp.trt)) + facet_wrap(~sex) +
#   labs(title = "adult masses", y = "mass (mg)", x = "stage") +
#   scale_fill_manual(labels = AB_labels, values = AB_colors) + theme_bw()


## dev time (larval)
# done
devAvB.l_plot <- stats_AvB %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(y = days.to))

devAvB.l_plot + geom_boxplot(aes(fill = grp.trt)) + 
  labs(title = "larval instar development times", x = "stage", y = "days after hatching", fill = "group-treatment") +
  scale_fill_manual(labels = AB_labels, values = AB_colors) + theme_bw() +
  facet_grid(~instar.duration, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))



## dev time (post-pupation) (not doable rn)
# devAvB.a_plot <- stats_AvB %>% filter(instar.duration == "eclose" | instar.duration == "pupate") %>%
#   ggplot(aes(x = factor(instar.duration, levels = c("pupate", "eclose")), y = days.to))
# 
# devAvB.a_plot + geom_boxplot(aes(fill = treatment)) +
#   labs(title = "post-pupation development times", x = "stage", y = "days after previous stage") +
#   scale_fill_manual(labels = AB_labels, values = AB_colors) + theme_bw()

```

# AB pooled

## AB pooled analysis

```{r AB pool analysis}
# doing this some groups seem fine to combine
  # // TO DO: futzing around but i think actually keep the grouping separate LOL (i.e. dont list 100 treatments tgt)
    # i totally cannot remember what this means lol
# stealing code from AvB

# larva/adult mass @ instars
  # pupa/adults need to be separated from the larva bc of sex differences
lmass_AB <- long %>% group_by(trt.stage) %>%
    filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
    mutate(log.mass = log(mass))

amass_AB <- long %>% group_by(trt.stage, sex) %>%
  filter(instar == "pupa" | instar == "adult") %>%
  mutate(log.mass = log(mass))

# calculate development time (instar duration), split into larva/adults
devtime_AB <- data %>% 
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         wander = jdate.wander - jdate.hatch,
         pupate = jdate.pupate - jdate.wander,
         #eclose = jdate.eclose - jdate.15
         eclose = jdate.eclose - jdate.pupate) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupate", "eclose"), # pivot differences by stage
                names_to = "stage",
                values_to = "days.to",
                values_drop_na = T) %>%
  group_by(trt.stage, stage)

ldev_AB <- devtime_AB %>% filter(stage == "3rd" | stage == "4th" | stage == "5th" | stage == "wander")
adev_AB <- devtime_AB %>% filter(stage == "pupate" | stage == "eclose")


# merge into larval/adult stats
lstats_AB <- merge(lmass_AB, ldev_AB, all = T) %>%
  rename(instar.mass = instar, # mass @ start of instar
         instar.duration = stage) # duration of instar

astats_AB <- merge(amass_AB, adev_AB, all = T) %>%
  rename(instar.mass = instar,
         instar.duration = stage)


# create subsets for plotting effect of temp and temp @ instar
# temp effect subsets
AB_ltemps <- lstats_AB %>% filter(instar.enter == "hatch")
AB_atemps <- astats_AB %>% filter(instar.enter == "hatch")

# temp+instar effects subsets
AB_linstar <- lstats_AB %>% filter(trt.stage == "260-hatch" | trt.stage == "337-hatch" | trt.stage == "337-3rd" | trt.stage == "337-4th")
AB_ainstar <- astats_AB %>% filter(trt.stage == "260-hatch" | trt.stage == "337-hatch" | trt.stage == "337-3rd" | trt.stage == "337-4th")

```


## AB pooled plots

* todo: need to fix the adult ones (check subsetting + variables)

```{r AB pool plotting}
# using code from AvB
# might have fixed the adult plots. should come back to check
  # post-pupation ones are broken for temps, temp x instar


### plots comparing effect of temps ###
# larval masses
AB_ltemps %>% ggplot(aes(y = log.mass)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "(temp) mass at start of instar", x = "instar", y = "log(mass (mg))") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw() +
  facet_grid(~instar.mass, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))

# adult masses
AB_atemps %>% filter(sex != "") %>% # remove the pt B wanderers for now
  ggplot(aes(x = factor(instar.mass, level = c("pupa", "adult")), y = mass)) + geom_boxplot(aes(fill = treatment)) + facet_wrap(~sex) +
  labs(title = "(temp) adult masses", y = "mass (mg)", x = "stage") +
  scale_fill_manual(values = A_colors, labels=A_labels) + theme_bw()

# larval dev time
AB_ltemps %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(y = days.to)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "(temp) larval instar development times", x = "stage", y = "days after hatching") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw() +
  facet_grid(~instar.duration, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))

# adult dev time
AB_atemps %>% filter(instar.duration == "pupate" | instar.duration == "eclose") %>%
  ggplot(aes(x = factor(instar.mass, levels = c("pupa", "adult")), y = days.to)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "(temp) post-pupation development times", x = "stage", y = "days after previous stage") +
  scale_fill_manual(labels = A_labels, values = A_colors) + theme_bw()


### plots comparing effect of instars ###
# larval masses
AB_linstar %>% ggplot(aes(y = log.mass)) + geom_boxplot(aes(fill = factor(trt.stage, level= B_trts))) +
  labs(title = "(temp x instar) mass at start of instar", x = "instar", y = "log(mass (mg))", fill = "treatment") +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw() +
  facet_grid(~instar.mass, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))

# adult masses
AB_ainstar %>% ggplot(aes(x = factor(instar.mass, level = c("pupa", "adult")), y = mass)) +
  geom_boxplot(aes(fill = factor(trt.stage, level= B_trts))) + facet_wrap(~sex) +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw() +
  labs(title = "(temp x instar) adult masses", y = "mass (mg)", x = "stage", fill = "treatment")

# larval dev
AB_linstar %>% filter(instar.duration == "2nd" | instar.duration == "3rd" | instar.duration == "4th" | instar.duration == "5th" | instar.duration == "wander") %>%
  ggplot(aes(y = days.to)) + geom_boxplot(aes(fill = factor(trt.stage, level = B_trts))) +
  labs(title = "(temp x instar) larval instar development times", x = "stage", y = "days after hatching", fill = "treatment") +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw() +
  facet_grid(~instar.duration, switch="x") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), panel.spacing = unit(0, "lines"))


# dev time (post-pupation) # this looks wrong... maybe bc theres no data in the other 2 cats yet? (337-hatch is mapping to 337-4th)
AB_ainstar %>% filter(instar.duration == "pupa" | instar.duration == "pupate") %>%
  ggplot(aes(x = factor(instar.duration, levels = c("pupate", "eclose")), y = days.to)) + geom_boxplot(aes(fill = treatment)) +
  labs(title = "(temp x instar) post-pupation development times", x = "stage", y = "days after previous stage") +
  scale_fill_manual(labels = B_labels, values = B_colors) + theme_bw()

```



# overall survivorship/kill curves

* who died? (pmd) pre or post 3rd? did all slow growers die?
  * want to add columns for this!
  * how to grp trts? (by mean/fluct/instar? or mean-fluct/instar?)
* visualise: curve + do stacked bars (for an overview)?

## pmd analysis (defunct, do KM instead)

leaving this here for reference, but use KM to look at the data instead

```{r defunct! all pmd stats & plots}
# probably finished

data_pmd <- data_all %>% select(-c("notes", "mass.toast", "pupa.deformities", "mass.died")) %>%
  filter(reason.pmd != "toast", reason.pmd != "squished", reason.pmd != "empty") %>%
  mutate(trt.stage = paste(treatment,instar.enter, sep = "-")) %>%
  group_by(trt.stage, reason.pmd) %>% # grouping by expt.group is wonky (summing is weird)
  summarise(n=n())

data_pmd[data_pmd==""] <- "survived"

# summary table
pmd_stats <- data_pmd %>% pivot_wider(names_from = reason.pmd, values_from = n) %>%
  group_by(trt.stage) %>%
  replace(is.na(.), 0) %>%
  mutate(n = sum(c(survived, pupa, pmd, LP))) %>% select(-survived) %>%
  mutate(pct.pupa = round(pupa/n*100, digits = 1),
         pct.larva = round(pmd/n*100, digits = 1),
         pct.LPI = round(LP/n*100, digits = 1),
         "total pct" = sum(c(pct.pupa, pct.larva, pct.LPI))) %>%
  rename(larva = pmd, LPI = LP)

# individuals (need to manually reorg table to see better)
individial_pmd <- data_all %>% select(-c("notes", "mass.toast", "pupa.deformities", "mass.died")) %>%
  filter(reason.pmd == "pmd" | reason.pmd == "LP" | reason.pmd == "pupa") %>%
  mutate(trt.stage = paste(treatment,instar.enter, sep = "-"), .before = 1) %>%
  mutate(across(starts_with("date."),  as.Date, format = "%m/%d/%y")) %>%
  mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
  #relocate(starts_with("jdate."), .after = starts_with("date.")) %>% # this is not working lol
  mutate(across(starts_with("jdate."), as.numeric),
         "2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         #stuck = jdate.stuck - jdate.hatch,
         #was.stuck = ifelse(date.stuck != "", "1", "NA", #not working like i want it to lol
         wander = jdate.wander - jdate.hatch,
         pupate = jdate.pupate - jdate.wander) %>%
  rename("days.to.2nd" = "2nd", "days.to.3rd" = "3rd","days.to.4th" = "4th","days.to.5th" = "5th", "days.to.wander" = wander,"days.to.pupate" = pupate) %>%
  select(-starts_with(c("jdate.", "h.", "date.", "from.", "instar.", "time.")), -c("treatment", "dv", "mass.adult"))

```

## kaplan-meier curves

* b/g reading
  * general ideas https://datatab.net/tutorial/survival-analysis
      * prob use cox to account for instar entered, trt, stuck in 5th
        * http://www.sthda.com/english/wiki/cox-proportional-hazards-model 
        * https://rstudio-pubs-static.s3.amazonaws.com/5896_8f0fed2ccbbd42489276e554a05af87e.html 
* probably use [this](https://dk81.github.io/dkmathstats_site/rvisual-kaplan-meier.html)
  * https://argoshare.is.ed.ac.uk/healthyr_book/recode-the-data-1.html 
    * look into [cox.zph](https://argoshare.is.ed.ac.uk/healthyr_book/cox-proportional-hazards-regression.html) for stuck?


variables needed

* id, trt, etc (group curves by trt?)
* time to pmd/event (death as LP, P, L)
* censorship for other events (stuck)

```{r}

#km <- data %>% transmute()

```


# code dump

things im not using anymore but can reference. recently added stuff is at the top

```{r}
## old stats calculations for grp A (averaged everything -- got rid of this for boxplots to work)
# growth_A <- data_A %>% group_by(treatment, instar) %>%
#   filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "wander") %>%
#   summarise(n = n(),
#             avg.log.mass = mean(log(mass), na.rm = T),
#             sd.log.mass = sd(log(mass), na.rm = T))

# adultmass_A <- data_A %>% group_by(treatment, sex) %>%
#   filter(instar == "pupa" | instar == "adult") %>%
#   summarise(n=n(),
#             avg.log.mass = mean(mass, na.rm = T),
#             sd.log.mass = sd(mass, na.rm = T))

# devtime_A <- data %>% filter(expt.group == "A") %>%
#   select(-starts_with(c("h", "date", "mass", "reason."))) %>%
#   mutate("2nd" = jdate.2nd - jdate.hatch,
#          "3rd" = jdate.3rd - jdate.hatch,
#          "4th" = jdate.4th - jdate.hatch,
#          "5th" = jdate.5th - jdate.hatch,
#          wander = jdate.wander - jdate.hatch,
#          pupate = jdate.pupate - jdate.wander,
#          eclose = jdate.eclose - jdate.15) %>%
#   select(-starts_with("jdate.")) %>%
#   pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "wander", "pupate", "eclose"), # pivot differences by stage
#                 names_to = "stage",
#                 values_to = "days.to",
#                 values_drop_na = T) %>%
#   group_by(treatment, stage) %>%
#   summarise(n = n(),
#             avg.diff = mean(days.to, na.rm = T),
#             sd.diff = sd(days.to, na.rm = T))

# columns need better names...
# stats_A <- merge(growth_A, devtime_A, by=c("treatment", "n")) %>%
#   rename(instar.mass = instar, # mass
#          instar.duration = stage) # time

## old mass @ start of instar plot for grp A
#vmassA_plot <- ggplot(data = stats_A, aes(x = instar.mass, y = avg.log.mass))
  # doesn't work bc there's no spread lol (no individual data points)

## old growth stats calc for grp A
# growth_A <- data_A %>% group_by(treatment) %>%
#   summarise(n = n(),
#             log.avg.mass.3rd = mean(log(mass.3rd), na.rm=T), sd.mass.3rd = sd(log(mass.3rd), na.rm=T),
#             log.avg.mass.4th = mean(log(mass.4th), na.rm=T), sd.mass.4rd = sd(log(mass.4th), na.rm=T),
#             log.avg.mass.5th = mean(log(mass.5th), na.rm=T), sd.mass.5th = sd(log(mass.5th), na.rm=T),
#             log.avg.mass.wander = mean(log(mass.wander), na.rm=T), sd.mass.wander = sd(log(mass.wander), na.rm=T))


```

