---
title: "NTW prelim data analysis"
output: html_notebook
---

# information

here is code for some analysis of preliminary NTW experiments. as usual i keep making 50 different new files for things bc im indecisive about how to do things

TODO list: (search `// TODO`)

* create function to convert `date.` to julian
* create `daysto.` columns (from hatching) for all instars/development timepoints


statistics to measure:

* log average weight at instars
  * h/m considerations...
* average length of instars
* survival


experiment groups:

* A: 2x2 of mean temp = 26 or 33, with 0 degC fluct or 7 degC fluct
* B: 
  * redoing 337 from hatching (A)
  * 2x2(ish) of entering 260 or 337 treatment at different times (hatch, 3rd, 4th)

# load packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
```

# load in data and clean
```{r}
data_all <- read.csv("~/Documents/projects/data/ntw_data/development.csv", header = T)

# save wordier notes elsewhere and remove from analysis dataset
data_notes <- select(data_all, c("from.E", "treatment", "ID", "expt.group", "pupa.deformities", "notes"))
data <- select(data_all, -c("notes")) #not sure if analsis works with this column

# remove squished/non-existent individuals
data <- data[data$ID != 1002,]
data <- data[data$ID != 1007,]
data <- data[data$ID != 1017,]
```

hi
