---
title: "cleaned up NTW"
output: html_notebook
---

because my main stuff is getting tooo messy again!!!


0. load in data, packages
1. clean/reformat data, make subsets
2. look at mass gain/instar time on same plot
  a. calculate stats for larva/adults
  b. group stats by expts (temp, tempxins, NTs)
  c. plot
3. look at mass gain/instar time on separate plot
  a. calculate stats
  b. group stats by expts (temp, tempxins, NTs)
  c. plot
4. kill curves
5. misc: old 2013 data


# 0. load in packages, data
```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(ggplot2)
library(survival)
library(survminer)
library(googlesheets4)

#data_all <- read.csv("~/Documents/repos/_not-public/1_data/ntw_data/development_v3.csv", header = T)
```

```{r}
# larva data
data_field <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet = "june field", col_types = "c")
data_labsu <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="june 2023 lab", col_types = "c")
data_labF1 <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="lab F1", col_types = "c")
data_labsp <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="consolidated + tidyed", col_types = "c")

# tents/hatch data
data_matepairs <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="mating pairs", col_types = "c")
data_tents <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="tent stats v2", col_types = "c")
data_hatching <- read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="hatch stats", col_types = "c")
```



# 1. clean data

## consolidate imported sheets + some pre-processing
```{r}
# function for column cleanup, adding some columns (imo separate the column generation to a diff fn. maybe just keep this one tho and just calc a diff one for timeto)
# need to make sure that format is correctly set on gsheet first
fix.columns <- function(data) {
  cleaned_data <- data %>%
    mutate(across(starts_with("date."),  as.Date, format = "%m/%d")) %>%
    mutate(across(starts_with("date."), format, "%j", .names = "j{.col}")) %>%
    mutate(across(starts_with("jdate."), as.numeric),
           across(starts_with("mass."), as.numeric),
           treatment = as.character(treatment),
           trt.stage = as.character(paste(treatment,instar.enter, sep = "-")), # treatment @ instar
           grp.trt = paste(expt.group, trt.stage, sep = "-")) # round @ treatment
  
  return(cleaned_data)
}

# add IDs
data_field <- mutate(data_field, src = "field")
data_labF1 <- mutate(data_labF1, src = "F1")
data_labsp <- mutate(data_labsp, src = "sp")
data_labsu <- mutate(data_labsu, src = "su")

# consolidate
data_all <- data.table::rbindlist(lapply(list(data_field, data_labF1, data_labsp, data_labsu), fix.columns), fill = T)

### preprocessing
# match a column for consistency
data_all$treatment[data_all$treatment=="426"] <- "337"

# filter out individuals that didnt hatch at 26C
data_all <- data_all %>%
    filter(is.na(from.E) | !(from.E == 2 | (from.E == 0 & (treatment == 337 | treatment == 330 | treatment == 267))))

# add survival/development binaries
data_all <- data_all %>% mutate(
  if.stuck = case_when(is.na(date.stuck) ~ "N", TRUE ~ "Y")
)

# add meanT/flucT

```


## further filtering
```{r}
# filter out MQs
data_MQ <- data_all %>% filter(species == "MQ")
data_all <- data_all %>% filter(is.na(species))

# extract wordier notes from data and save elsewhere
data_notes <- select(data_all, c("treatment", "ID", "expt.group", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "mass.died", "pupa.deformities", "notes", "ignore.reason", ends_with(".stuck"), ends_with(".culled")))

# remove unneeded columns
data_all <- select(data_all, -c("species", "from.E", "location", "date.collected", "date.stuck", "time.in.approx", "dv", "pupa.deformities", "mass.died", "old.date.pmd", "fate.code", "notes", ends_with(".stuck"), ends_with(".culled"), "eclose-3"))
```


## create and reformat columns
```{r}


# pivot
long_all <- data_all %>% select(-(starts_with("date."))) %>%
                 pivot_longer(cols = starts_with(c("jdate", "mass", "h")),
                        names_to = c(".value", "instar"),
                        #names_sep = ".",
                        values_drop_na = TRUE,
                        names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
                 rename(molt.status = h) %>%
                 drop_na(jdate) # drops NA's if an individual didnt reach a certain stage

# remove extra columns and add instar factor levels
long_all$instar <- factor(long_all$instar, c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit"))
```


## subset data
```{r}
# individuals that survived to pupation (includes LPs)
data_pup <- data_all %>% filter(attempt.pupa == "Y")
long_pup <- long_all %>% filter(attempt.pupa == "Y")
```


## define labels and aesthetics
```{r}
#all_trts <- c("260-hatch", "267-hatch", "330-hatch", "337-hatch", "337-3rd", "337-4th", "40-19", "40-26")


# A: effect of mean/fluct temp (treatment)
A_trts = c("260-hatch", "267-hatch", "330-hatch", "337-hatch")
A_labels = c("260-hatch"="26°C", "267-hatch"="26±7°C", "330-hatch"="33°C", "337-hatch"="33±7°C")
A_colors = c("260-hatch"="#00C2D1","267-hatch"="#1929B3", "330-hatch"="#F9C639", "337-hatch"="#710A36")



# B: effect of temp x instar (trt.stage)
B_trts = c("260-hatch", "337-hatch", "337-3rd", "337-4th")
B_labels = c("260-hatch"="26°C @ hatch","337-hatch"="33±7°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th")
B_colors = c("260-hatch"="#00C2D1", "337-hatch"="#710A36", "337-3rd"="#C23C1E", "337-4th"="#F3922B")
#CD133F


# C: same DTs, different NTs
C_trts = c("267-hatch", "419-hatch", "337-hatch", "433-hatch")
C_labels = c("267-hatch"="26±7°C (33/19)","419-hatch"="29.5±10.5°C (40/19)", "337-hatch"="33±7°C (40/26)", "433-hatch"="36.5±3.5°C (40/33)")
C_colors = c("267-hatch"="#F4B942", "419-hatch"="#4059AD", "337-hatch"="#6B9AC4", "433-hatch"="#97D8C4")
  # tbh not sure what i want to compare with here... 260 or 267?

# for survival but lowkey need to edit it below
A_hex = c("#00C2D1","#1929B3", "#F9C639", "#710A36")
B_hex = c("#00C2D1", "#710A36", "#C23C1E", "#F3922B")
C_hex = c("#F4B942", "#4059AD", "#6B9AC4", "#97D8C4")
```



# 2. look at mass gain/instar time on same plot

## a. calculate stats for adult/larva subsets

adults need to be separated out to be able to look at sex differences

TODO: add in calcs for 6th instar

### mass stats
```{r}
l_mass <- long_all %>%
  filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "6th" | instar == "wander") %>%
  group_by(instar, trt.stage) %>%
  mutate(logmass = log(mass)) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

a_mass <- long_all %>% 
  group_by(trt.stage, sex, instar) %>%
  filter(instar == "pupa" | instar == "eclose") %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            n=n())
```

### development time stats
```{r}
# calculate for all groups
all_devtime <- data_all %>% filter(!(final.fate == "ignore" | final.fate == "pmd")) %>%
  #data_pup %>% 
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("2nd" = jdate.2nd - jdate.hatch,
         "3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         "6th" = jdate.6th - jdate.hatch,
         wander = jdate.wander - jdate.hatch,
         pupa = jdate.pupa - jdate.wander,
         #eclose = jdate.eclose - jdate.15
         eclose = jdate.eclose - jdate.pupa) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("2nd", "3rd", "4th", "5th", "6th", "wander", "pupa", "eclose"), # pivot differences by stage
                names_to = "instar",
                values_to = "days.to") %>%
  drop_na(days.to)

# subset larva
l_dev <- all_devtime %>% 
  filter(instar == "3rd" | instar == "4th" | instar == "5th" | instar == "6th" | instar == "wander") %>%
  group_by(trt.stage, instar) %>%
  summarise(avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))))


# subset adults
a_dev <- all_devtime %>% 
  filter(instar == "pupa" | instar == "eclose") %>%
  group_by(trt.stage, sex, instar) %>%
  summarise(avg.daysto = mean(na.omit(days.to)),
            se.daysto = sd(na.omit(days.to))/sqrt(length(na.omit(days.to))))
```

## b. resubset data into expts
```{r}
# combine mass/time data
l_stats <- merge(l_mass, l_dev, all = T) %>%
  drop_na(avg.daysto) # drop incomplete data

a_stats <- merge(a_mass, a_dev, all = T) %>%
  drop_na(sex) # drop incomplete data


# subset: effect of mean/fluct temps
temps_l <- l_stats %>% filter(trt.stage != "337-3rd" & trt.stage != "337-4th" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
temps_a <- a_stats %>% filter(trt.stage != "337-3rd" & trt.stage != "337-4th" & trt.stage != "419-hatch" & trt.stage != "433-hatch")

# subset: effect of temp x instar (accumulation)
instar_l <- l_stats %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
instar_l$trt.stage <- factor(instar_l$trt.stage, levels = c("260-hatch", "337-hatch", "337-3rd", "337-4th"))

instar_a <- a_stats %>% filter(trt.stage != "267-hatch" & trt.stage != "330-hatch" & trt.stage != "419-hatch" & trt.stage != "433-hatch")
instar_a$trt.stage <- factor(instar_a$trt.stage, levels = c("260-hatch", "337-hatch", "337-3rd", "337-4th"))

# subset: effect of different NTs
NT_l <- l_stats %>% filter(trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch")
NT_a <- a_stats %>% filter(trt.stage == "267-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch")
```


## c. plot

### temp figs
```{r}
temps_l %>%
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=c(16, 17, 15, 18))+
  geom_line(aes(group = trt.stage)) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels=A_labels, values = A_colors) + theme_bw()

temps_a %>%
  ggplot(aes(x = avg.daysto, y = avg.mass, color=trt.stage, shape=instar)) +
  geom_point(size=3, alpha = 0.75) +
  geom_line(aes(group = trt.stage)) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  facet_wrap(~sex) +
  labs(title = "(temp) avg adult growth", y = "avg mass (mg)", x = "days since last stage", color = "treatment") +
  scale_color_manual(values = A_colors, labels=A_labels) + theme_bw()
```



### instar figs
```{r}
instar_l %>% 
  drop_na() %>% #6ths make this break
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha = 0.75) +
  scale_shape_manual(values=c(16, 17, 15, 18))+
  geom_line(aes(group = trt.stage)) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(instar) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels = B_labels, values = B_colors) + theme_bw()
    ## looks kinda wrong w rerun analyses? (lines dont seem like they grouped right..)

# adult mass/dev
instar_a %>% 
  ggplot(aes(x = avg.daysto, y = avg.mass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha = 0.75) +
  geom_line(aes(group = trt.stage)) +
  facet_wrap(~sex) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  scale_color_manual(labels = B_labels, values = B_colors) + theme_bw() +
  labs(title = "(instar) avg adult growth", y = "avg mass (mg)", x = "days since last stage", color="treatment")
```



### NT figs
```{r}
# NT figs w/ 26-7
NT_l %>%
  filter(instar != "6th") %>% # this is wonky
  ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha=0.75) +
  scale_shape_manual(values=c(16, 17, 15, 18)) + # use shape 8 if including 6ths
  geom_line(aes(group = trt.stage)) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
  geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
  labs(title = "(temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
  scale_color_manual(labels=C_labels, values = C_colors) + theme_bw()


# NT figs w/o 26-7
  # need to adjust the scales bc theres only 3 things
# NT_l %>%
#   filter(trt.stage != "267-hatch") %>%
#   ggplot(aes(x = avg.daysto, y = avg.logmass, shape=instar, color=trt.stage)) +
#   geom_point(size=3, alpha=0.75) +
#   scale_shape_manual(values=c(16, 17, 15, 18))+
#   geom_line(aes(group = trt.stage)) +
#   geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto)) +
#   geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass)) +
#   labs(title = "(temp) avg larval growth", x = "days since hatching", y = "avg log(mass (mg))", color = "treatment") +
#   scale_color_manual(labels=C_labels, values = C_colors) + theme_bw()

# adult mass/dev
NT_a %>%
  ggplot(aes(x = avg.daysto, y = avg.mass, shape=instar, color=trt.stage)) +
  geom_point(size=3, alpha=0.75) +
  #scale_shape_manual(values=c(16, 17, 15, 18)) + # use shape 8 if including 6ths
  geom_line(aes(group = trt.stage)) +
  facet_wrap(~sex) +
  geom_errorbarh(aes(xmin= avg.daysto - se.daysto, xmax = avg.daysto + se.daysto), height = 175) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass)) +
  labs(title = "(temp) avg larval growth", x = "days since hatching", y = "avg mass (mg)", color = "treatment") +
  scale_color_manual(labels=C_labels, values = C_colors) + theme_bw()

```


# 3. look at mass gain/instar time on separate plots

## a. calculate stats for larva/adults

### larva
```{r}
# TODO: need to also do math for 6ths

# repivot for easier mathing
  # this is basically all_devtime except w/ renamed "timeto" bc i cant figure out how to make the paste work
all_alt <- data_all %>%
  mutate(timeto.3rd = jdate.3rd-jdate.hatch,
         timeto.4th = jdate.4th-jdate.hatch,
         timeto.5th = jdate.5th-jdate.hatch,
         timeto.6th = jdate.6th-jdate.hatch,
         timeto.wander = jdate.wander-jdate.hatch,
         timeto.pupa = jdate.pupa-jdate.hatch,
         timeto.eclose = jdate.eclose-jdate.hatch) %>%
  select(-(starts_with(c("date.", "jdate.")))) %>%
  pivot_longer(cols = starts_with(c("timeto", "mass", "h")),
               names_to = c(".value", "instar"),
               #names_sep = ".",
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  rename(molt.status = h)
# drop_na(jdate) # drops NA's if an individual didnt reach a certain stage

# summary statistics
alt_sumstats <- all_alt %>%
  mutate(mass = mass/1000) %>% # convert to g
  mutate(log.mass = log(mass)) %>% # u get neg values if using mass_g
  filter(final.fate != "ignore") %>% 
  filter(instar != "stuck" & instar != "2nd" & instar != "died") %>%
  group_by(meanT, fluctT, instar.enter, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.logmass = mean(na.omit(log.mass)),
            se.logmass = sd(na.omit(log.mass))/sqrt(length(na.omit(log.mass))),
            avg.time = mean(na.omit(timeto)),
            se.time = sd(na.omit(timeto))/sqrt(length(na.omit(timeto))),
            n=n())

# fix factor levels
alt_sumstats$instar <- factor(alt_sumstats$instar, levels=c("3rd", "4th", "5th", "6th", "wander", "pupa", "eclose"))
alt_sumstats$instar.enter <- factor(alt_sumstats$instar.enter, levels=c("hatch", "3rd", "4th"))
```

### adult

```{r}
# calc stats
alt_sexstats <- all_alt %>%
  mutate(mass = mass/1000) %>% # convert to g
  filter(final.fate != "ignore" & sex != "unk") %>% 
  filter(instar == "pupa" | instar == "eclose") %>%
  group_by(meanT, fluctT, instar.enter, instar, sex) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.time = mean(na.omit(timeto)),
            se.time = sd(na.omit(timeto))/sqrt(length(na.omit(timeto))),
            n=n())

# fix factor levels
alt_sexstats$instar <- factor(alt_sexstats$instar, levels=c("3rd", "4th", "5th", "6th", "wander", "pupa", "eclose"))
alt_sexstats$instar.enter <- factor(alt_sexstats$instar.enter, levels=c("hatch", "3rd", "4th"))
```



## b. group stats by expts (temp, tempxins, NTs)
```{r}
## ugh.. wrong comparison (instar should be to 26)

# larva
temps_la <- filter(alt_sumstats, ((fluctT == 0 | fluctT == 7) & instar.enter == "hatch"))
instar_la <- filter(alt_sumstats, (meanT == 33 | (meanT == 26 & fluctT == 0)))
NT_la <- filter(alt_sumstats, (((meanT == 29.5 | meanT == 36.5) | fluctT == 7) & instar.enter == "hatch"))

# adults
temps_aa <- filter(alt_sexstats, ((fluctT == 0 | fluctT == 7) & instar.enter == "hatch"))
instar_aa <- filter(alt_sexstats, (meanT == 33 | (meanT == 26 & fluctT == 0)))
NT_aa <- filter(alt_sexstats, (((meanT == 29.5 | meanT == 36.5) | fluctT == 7) & instar.enter == "hatch"))
```


## c. plots

### larval instars
```{r}
### temp mass
temps_la %>%
  ggplot(aes(x = meanT, y = avg.mass, color = factor(fluctT))) +
  geom_line(aes(lty=factor(fluctT))) +
  geom_point(aes(shape = factor(fluctT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15) +
  facet_wrap(~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))

### temp time
temps_la %>%
  ggplot(aes(x = meanT, y = avg.time, color = factor(fluctT))) +
  geom_line(aes(lty=factor(fluctT))) +
  geom_point(aes(shape = factor(fluctT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "temp: time to instar (days)", lty = "fluctuation", shape = "fluctuation", color = "fluctuation") +
  scale_color_manual(values = c("#00BFC4", "#F8766D"))
```

```{r}
jitter <- position_dodge(width=1)

### instar mass
instar_la %>%
  filter(!(meanT == 33 & fluctT == 0 & instar.enter == "hatch")) %>%
  ggplot(aes(x = meanT, y = avg.mass, color = instar.enter, shape = factor(fluctT))) +
  geom_line() +
  geom_point(position = jitter) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15, position = jitter) +
  facet_wrap(~instar, scales = "free") +
  theme_bw() +
  labs(y = "instar: mass at instar (mg)", color = "instar entered", shape = "fluctuation" )

### instar time
instar_la %>%
  filter(!(meanT == 33 & fluctT == 0 & instar.enter == "hatch")) %>%
  ggplot(aes(x = meanT, y = avg.time, color = instar.enter, shape = factor(fluctT))) +
  geom_line() +
  geom_point(position = jitter) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15, position = jitter) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "instar: time to instar (days)", color = "instar entered", shape = "fluctuation")
```

```{r}
#if facet_wrap(scales = "free"), set error bar width = 0.5

### NT mass
NT_la %>%
  ggplot(aes(x = meanT, y = avg.mass, color=factor(fluctT))) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
  facet_wrap(~instar) +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))


### NT time
NT_la %>%
  ggplot(aes(x = meanT, y = avg.time, color=factor(fluctT))) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
  theme_bw() +
  facet_wrap(~instar) +
  labs(y = "temp: time to instar (days)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
```


### adult instars
```{r}
### temp mass
temps_aa %>%
  ggplot(aes(x = meanT, y = avg.mass, color = sex)) +
  geom_line(aes(lty=factor(fluctT))) +
  geom_point(aes(shape = factor(fluctT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15) +
  facet_wrap(~instar, scales = "free") +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", lty = "fluctuation", shape = "fluctuation")

### temp time
temps_aa %>%
  ggplot(aes(x = meanT, y = avg.time, color=sex)) +
  geom_line(aes(lty=factor(fluctT))) +
  geom_point(aes(shape = factor(fluctT)), size=2.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15) +
  theme_bw() +
  facet_wrap(~instar, scales = "free") +
  labs(y = "temp: time to instar (days)", lty = "fluctuation", shape = "fluctuation")
```

```{r}
jitter <- position_dodge(width=0.75)

### instar mass
instar_aa %>%
  ggplot(aes(x = fluctT, y = avg.mass, color = instar.enter)) +
  #geom_line() +
  geom_point(aes(shape = sex), size = 2.5, position = jitter) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.15, position = jitter) +
  facet_wrap(instar~sex, scales = "free") +
  theme_bw() +
  labs(y = "instar: mass at instar (mg)", color = "instar entered")

### instar time
instar_aa %>%
  ggplot(aes(x = fluctT, y = avg.time, color = instar.enter)) +
  #geom_line() +
  geom_point(aes(shape = sex), size = 2, position = jitter) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 0.15, position = jitter) +
  theme_bw() +
  facet_wrap(instar~sex, scales = "free") +
  labs(y = "instar: time to instar (days)")
```

```{r}
### zzzz waiting

### NT mass
NT_aa %>%
  ggplot(aes(x = meanT, y = avg.mass, color=factor(fluctT))) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 1) +
  facet_wrap(instar~sex) +
  theme_bw() +
  labs(y = "temp: mass at instar (mg)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))


### NT time
NT_aa %>%
  ggplot(aes(x = meanT, y = avg.time, color=factor(fluctT))) +
  geom_point(size=1.5) +
  geom_errorbar(aes(ymin = avg.time - se.time, ymax = avg.time + se.time), width = 1) +
  theme_bw() +
  facet_wrap(instar~sex) +
  labs(y = "temp: time to instar (days)", color = "fluctuation") +
  scale_color_manual(values = c("#619CFF", "#00BA38", "#F8766D"))
```


# 4. survival curves

## a. prep data

checking survival to pupation from hatching

```{r}
# censor and create subsets

# censor data
surv_pup <-  data_all %>% 
  filter(final.fate != "ignore") %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

# subsets & specify levels
surv_temp <- surv_pup %>%
  filter(fluctT == 0 | (fluctT == 7 & instar.enter == "hatch")) %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "267-hatch", "330-hatch", "337-hatch")))
  

surv_instar <- surv_pup %>%
  filter((meanT == 33 & fluctT == 7) | (meanT == 26 & fluctT == 0)) %>%
  mutate(trt.stage = factor(trt.stage, c("260-hatch", "337-hatch", "337-3rd", "337-4th")))

surv_NT <- surv_pup %>%
  filter((expt.group == "C" | (meanT == 33 & fluctT == 7) | (meanT == 26 & fluctT == 7)) & instar.enter == "hatch") %>%
  filter(trt.stage != "260-hatch") %>%
  mutate(trt.stage = factor(trt.stage, c("267-hatch", "419-hatch", "337-hatch", "433-hatch")))
```

## b. create survival objects and fits

```{r}
temp_sobj <- Surv(surv_temp$timeto.exit, surv_temp$censor)
temp_sfit <- survfit(temp_sobj ~ trt.stage, data = surv_temp)

ins_sobj <- Surv(surv_instar$timeto.exit, surv_instar$censor)
ins_sfit <- survfit(ins_sobj ~ trt.stage, data = surv_instar)

NT_sobj <- Surv(surv_NT$timeto.exit, surv_NT$censor)
NT_sfit <- survfit(NT_sobj ~ trt.stage, data = surv_NT)
```

## c. plot survival curves

```{r}
# define palette (idk how to make the levels work)
# A_hex = c("#00C2D1","#1929B3", "#F9C639", "#710A36")
# B_hex = c("#00C2D1", "#710A36", "#C23C1E", "#F3922B")
# C_hex = c("#F4B942", "#4059AD", "#6B9AC4", "#97D8C4")

# A_labels <- c("260-hatch"="26°C", "267-hatch"="26±7°C", "330-hatch"="33°C", "337-hatch"="33±7°C")
# B_labels = c("260-hatch"="26°C @ hatch","337-hatch"="33±7°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th")
# C_labels = c("267-hatch"="26±7°C (33/19)","419-hatch"="29.5±10.5°C (40/19)", "337-hatch"="33±7°C (40/26)", "433-hatch"="36.5±3.5°C (40/33)")


# temps
  # 260, 267, 330, 337
ggsurvplot(fit = temp_sfit, data=surv_temp,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp x survival from hatching",
           #palette = c("#00C2D1","#1929B3", "#F9C639", "#710A36"),
           legend.title = "treatment", 
           #legend.labs = A_labels
           )

# instar
ggsurvplot(fit = ins_sfit, data=surv_instar,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp:instar x survival from hatching", 
           #palette = c("#00C2D1", "#F3922B", "#C23C1E", "#710A36"),
           legend.title = "treatment-stage entered", 
           #legend.labs = c("26°C @ hatch", "337-3rd"="33±7°C @ 3rd", "337-4th"="33±7°C @ 4th", "337-hatch"="33±7°C @ hatch")
           )
  # 260, 3rd, 4th, hatch


# NTs
ggsurvplot(fit = NT_sfit, data=surv_NT,
           xlab = "days from hatching", ylab = "survival probability",
           title = "NT x survival from hatching", 
           #palette = c("#F4B942", "#6B9AC4", "#4059AD", "#97D8C4"),
           legend.title = "treatment", 
           #legend.labs = c("26±7°C (33/19)", "33±7°C (40/26)", "29.5±10.5°C (40/19)", "36.5±3.5°C (40/33)")
           )
  # 267, 337, 419, 433

```




# 5. misc: old 2013 data

## a. prep data

## b. plot data

