---
title: "NTW prelim data analysis"
output: html_notebook
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, include = FALSE)
```

# information

here is code for some analysis of preliminary NTW experiments. as usual i keep making 50 different new files for things bc im indecisive about how to do things

TODO list: (search `// TODO`)

* create `daysto.` columns (from hatching) for all instars/development timepoints


statistics to measure:

* log average weight at instars
  * h/m considerations...
* average length of instars
* survival


experiment groups:

* A: 2x2 of mean temp = 26 or 33, with 0 degC fluct or 7 degC fluct
* B: 
  * redoing 337 from hatching (A)
  * 2x2(ish) of entering 260 or 337 treatment at different times (hatch, 3rd, 4th)

# load packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
#library(data.table) #for setDT
```

# load in data and clean
```{r}
data_all <- read.csv("~/Documents/projects/data/ntw_data/development.csv", header = T)

# save wordier notes elsewhere and remove from analysis dataset
data_notes <- select(data_all, c("from.E", "treatment", "ID", "expt.group", "pupa.deformities", "notes"))
data <- select(data_all, -c("notes")) #not sure if analsis works with this column

# remove squished/non-existent individuals
data <- data[data$ID != 1002,]
data <- data[data$ID != 1007,]
data <- data[data$ID != 1017,]

# create julian date columns
data <- data %>% mutate(across(starts_with("date."),  as.Date, format = "%m/%d/%y")) %>%
  mutate(across(starts_with("date."), format, "%j", .names = "j{.col}"), .after = 1) %>% # after isnt working idky
  mutate(across(starts_with("jdate."), as.numeric))

# create subsets
data_A <- data[data$expt.group == "A",]
data_B <- data[data$expt.group == "B",]

```

# stats... pt A

```{r}
avgmass_A <- data_A %>% group_by(treatment) %>%
  summarise(n = n(),
            log.avg.mass.3rd = mean(log(mass.3rd), na.rm=T), sd.mass.3rd = sd(log(mass.3rd), na.rm=T),
            log.avg.mass.4th = mean(log(mass.4th), na.rm=T), sd.mass.4rd = sd(log(mass.4th), na.rm=T),
            log.avg.mass.5th = mean(log(mass.5th), na.rm=T), sd.mass.5th = sd(log(mass.5th), na.rm=T),
            log.avg.mass.wander = mean(log(mass.wander), na.rm=T), sd.mass.wander = sd(log(mass.wander), na.rm=T))

adultmass_A <- data_A %>% group_by(treatment, sex) %>%
  summarise(n=n(),
            avg.mass.pupa = mean(mass.pupa, na.rm=T), sd.mass.pupa = sd(mass.pupa, na.rm=T),
            avg.mass.adult = mean(mass.adult, na.rm=T), sd.mass.pupa = sd(mass.pupa, na.rm=T))

time_A <- mutate(data_A, time.to.2nd = jdate.2nd - jdate.hatch,
                    time.to.3rd = jdate.3rd - jdate.hatch,
                    time.to.4th = jdate.4th - jdate.hatch,
                    time.to.5th = jdate.5th - jdate.hatch,
                    time.to.wander = jdate.wander - jdate.hatch,
                    time.to.pupate = jdate.pupate - jdate.hatch,
                    time.to.eclose = jdate.eclose - jdate.hatch)%>%
       group_by(treatment) %>%
       summarise(n = n(),
                 avg.time.to.2nd = mean(time.to.2nd, na.rm=T), sd.time.to.2nd = sd(time.to.2nd, na.rm=T),
                 avg.time.to.3rd = mean(time.to.3rd, na.rm=T), sd.time.to.3rd = sd(time.to.3rd, na.rm=T),
                 avg.time.to.4th = mean(time.to.4th, na.rm=T), sd.time.to.4th = sd(time.to.4th, na.rm=T),
                 avg.time.to.5th = mean(time.to.5th, na.rm=T), sd.time.to.5th = sd(time.to.5th, na.rm=T),
                 avg.time.to.wander = mean(time.to.wander, na.rm=T), sd.time.to.wander = sd(time.to.wander, na.rm=T),
                 avg.time.to.pupate = mean(time.to.pupate, na.rm=T), sd.time.to.pupate = sd(time.to.pupate, na.rm=T),
                 avg.time.to.eclose = mean(time.to.eclose, na.rm=T), sd.time.to.eclose = sd(time.to.eclose, na.rm=T))

allstats <- merge(avgmass_A, time_A, by=c("treatment", "n"))
```

