---
title: "other_analyses"
date: "2023-12-07"
---

# 0.1 load data & packages

```{r message = FALSE}
library(conflicted)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(tidyr)
library(purrr)
library(ggplot2)

# survival curve stuff
#library(survival) # for computing
#library(survminer) # for visualising

# make lms
#library(lmerTest)

wide_all <- read.csv("~/Documents/repos/_not-public/1_data/ntw_data/clean-gsheets.csv", header = TRUE)

```

## - pivot to long
```{r}
long_all <- wide_all %>% select(-(starts_with("date."))) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch, 
         tt.wander = jdate.wander - jdate.hatch,
         #tt.pupa = jdate.pupa - jdate.wander,
         tt.pupa = jdate.pupa-jdate.hatch,
         tt.15 = jdate.15-jdate.hatch,
         tt.eclose = jdate.eclose-jdate.pupa,
         tt.exit = jdate.exit - jdate.enter,
         tt.surv = jdate.surv - jdate.hatch) %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "h", "tt")),
               names_to = c(".value", "instar"),
               #names_sep = ".",
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  rename(molt.status = h) %>%
  drop_na(jdate) %>% drop_na(tt) %>% # drops NA's if an individual didnt reach a certain stage
  filter(instar != "15")

# add instar factor levels
long_all <- long_all %>% mutate(instar = factor(instar, levels=c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit")))
#long_all$instar <- factor(long_all$instar, c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit"))

#rm(data_all)

```

# 0.2 define helpful functions & aesthetics

## - filtering groups
```{r}
# separate instars
filter.ins.topup <- function(data) {
  filtered_data <- data %>%
      filter(instar == "4th" | instar == "5th" | instar == "6th" | instar == "7th" | instar == "wander" | instar == "pupa")
  
  return(filtered_data)
}
```

# 1.0 lms to check ctrls

(imo some of this needs to be done via the modeling lol)


- 25 col vs 260 ctrls generally
  - also look at survival?
- batch effects (maybe clear this in models...)
- effects of lab diet vs TB diet in lab bugs

## - batch x 260 controls
```{r}

check_lab260 <- wide_all %>%
  filter(treatment == 260 & pop == "lab")

# working on this rn lol

mod_check260_1 <- lm(mass.pupa ~ expt.group + sex, data = check_lab260)

anova(mod_check260_1)

summary(mod_check260_1)

# conclusion: ....
# (also need to toss out some of the other data from the first batch of field stuff....)


mod_check260_2 <- lm(mass.pupa ~ expt.group, data = check_lab260)

anova(mod_check260_2)

#summary(mod_checkctrls2)



check_field260 <- wide_all %>%
  filter(treatment ==  260) %>%
  filter(pop == "field") 

mod_check260_3 <- lm(mass.pupa ~ expt.group + sex, data = check_field260)

anova(mod_check260_3)
  # there should be more experiment.groups...

summary(mod_check260_3)
```

lab: diurnal vs constant 26 doesnt seem to have a big effect on lab bugs (grp C maybe a little bit lol) → batches seem pretty consistent

field: from what survived, seems like batch didn't have an effect. that said, only 1 group really had good survivors lol (grp E)

## - batch x 337 ctrls

```{r}
check_lab337 <- wide_all %>%
  filter(treatment == 337 & pop == "lab")

mod_check337_1 <- lm(mass.pupa ~ expt.group + sex, data = check_lab337)

anova(mod_check337_1)

summary(mod_check337_1)

# with sex included, expt.group isnt sig


mod_check337_2 <- lm(mass.pupa ~ expt.group, data = check_lab337)

anova(mod_check337_2)

#summary(mod_check337_2)



check_field337 <- wide_all %>%
  filter(treatment ==  337) %>%
  filter(pop == "field") 

mod_check337_3 <- lm(mass.pupa ~ expt.group + sex, data = check_field337)

anova(mod_check337_3)
```

expt.group not a significant contribution to pupal mass in lab or field

## - time to pup
```{r}
mod_checktt_pup <- lm((jdate.pupa-jdate.hatch) ~ expt.group + sex + pop + treatment, data = wide_all)
  # oops dont just use 'jdate.pupa' lol

anova(mod_checktt_pup)

summary(mod_checktt_pup)


# just check lab/field
check_pup_lab <- wide_all %>% filter(pop == "lab")

mod_checkpup_lab <- lm(jdate.pupa ~ expt.group + sex + treatment, data = check_pup_lab)

anova(mod_checkpup_lab)

summary(mod_checkpup_lab)



check_pup_field <- wide_all %>% filter(pop == "field")

mod_checkpup_field <- lm(jdate.pupa ~ expt.group + sex + treatment, data = check_pup_field)

anova(mod_checkpup_field)

summary(mod_checkpup_field)

```

expt group an and trt have significant effects on tt pup in lab and field...

## - lab diet x dev

```{r}

# check effects of growth in lab animals (overall)

#check_pup_lab <- wide_all %>% filter(pop == "lab")

mod_diet <- lm(data = check_pup_lab, mass.pupa ~ treatment + sex + diet)

anova(mod_diet)

summary(mod_diet)
# seems like no major effect of lab diet on pupal mass


## visualise

# filtering NTs
NT_trtlabels = c("260"="26/26 (26±0°C)", "419"="40/19 (29.5±10.5°C)", "337"="40/26 (33±7°C)", "433"="40/33 (36.5±3.5°C)")
NT_tstglabels = c("260-hatch"="26/26 (26±0°C)", "419-hatch"="40/19 (29.5±10.5°C)", "337-hatch"="40/26 (33±7°C)", "433-hatch"="40/33 (36.5±3.5°C)")

wide_all %>% filter(pop == "lab") %>%
  filter(sex != "NA") %>%
  filter(trt.stage == "260-hatch" | trt.stage == "337-hatch" | trt.stage == "419-hatch" | trt.stage == "433-hatch") %>%
  mutate(trt.stage = factor(trt.stage, levels = c("260-hatch", "419-hatch", "337-hatch", "433-hatch")),
         treatment=factor(treatment, levels = c(260, 419, 337, 433))) %>%
  filter(expt.group == "D" | expt.group == "E" | expt.group == "F" | expt.group == "H") %>%
  ggplot(aes(y = mass.pupa, x = diet, color = trt.stage)) +
  geom_boxplot(show.legend = FALSE) + 
  facet_wrap(sex~treatment, nrow = 2, ncol = 4, labeller = labeller(treatment = NT_trtlabels)) +
  labs(title = "effect of diet on pupal mass", y = "pupal mass (mg)", x = "diet type") +
  NT_aes()



# compare effect of diet in lab vs field

check_diet <- wide_all %>% filter(expt.group == "E" | expt.group == "F" | expt.group == "H")

mod_diet2 <- lm(mass.pupa ~ treatment + sex + pop + expt.group, data = check_diet)

anova(mod_diet2)

summary(mod_diet2)


```

seems like diet doesnt have a big effect on growth overall in lab animals compared to field animals



 - batch effects
(frankly idr what this was for lol)
```{r}
check_batch <- wide_all %>%
  filter(src != "F1")
```


## - todo: effects of room/tent/etc on fertility

# 2.0 todo?: collective data: x trt - y response
```{r}


```

