---
title: "overall_controls"
date: "2024-01-15"
---

purpose: checking controls over all groups. this is adding visualisations/taking some code from `other_analyses.Rmd`

# 0. load data, helper fns
```{r}
source("./ntw_helper-functions.R")

library(survival)
library(survminer)
```

# 1. checking if lab 260s across all trts are the same to each other
```{r}
# filter lab 260 trts
data_lab260 <- wide_all %>%
  filter(treatment == 260 & pop == "lab") %>%
  filter(expt.group !="C") %>% # N is too small bc most of them culled
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled")

# calculate pup summary stats
summary_lab260 <- data_lab260 %>%
  group_by(expt.group) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))
```

## a. visualise
```{r}
# visualise time to pup, pupal mass across expt groups
summary_lab260 %>% ggplot(aes(y = avg.mass, x = avg.tt, color = expt.group)) +
  geom_point() +
  y_err_mass(y_err= 0.9) + x_err_tt(x_err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)")) +
  labs(title = "avg time to and mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# let's isolate more... just masspup
summary_lab260 %>% ggplot(aes(y = avg.mass, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_mass(y_err= 0.9) + theme_bw() +
  scale_fill_hue(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)")) +
  labs(title = "avg mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "expt group", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# just tt pup
summary_lab260 %>% ggplot(aes(y = avg.tt, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_tt(y_err= 0.9) + theme_bw() +
  scale_fill_hue(labels = c("A (n = 21)",
                                "B (n = 16)",
                                "D (n = 18)",
                                "F (n = 19)")) +
  labs(title = "avg time to pupation for lab controls", 
       y = "average time to pupation from hatching", 
       x = "expt group", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")

# visualise % survival
summary_lab260 %>% ggplot(aes(y = pct.survpup, fill = expt.group, x = expt.group)) +
  geom_col() + theme_bw() +
  #y_err(avg.mass, se.mass) + #SIGH
  # scale_fill_hue(labels = c("A (n = 21)",
  #                            "B (n = 16)",
  #                            "D (n = 18)",
  #                            "F (n = 19)")) +
  geom_text(aes(label = paste0("n = ", n_all), y = 10), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "C omitted bc of culling; D = diurnal, F = diurnal + TB diet")
```
```{r}
# visualise survival dynamics w/ KM
  # add survival binary & time
data_lab260 <- data_lab260 %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# fit model
kmfit.lab260 <- survfit(Surv(data_lab260$tt.exit, data_lab260$status) ~ expt.group, data = data_lab260)

# plot
ggsurvplot(fit = kmfit.lab260, data = data_lab260,
           conf.int=T)
  # well ... this generally fits as expected thus far lol.
  # seems like death time is a bit more wonky under diet
```


LOL okay rip (re: pt plot) let's model it then... (km plot for surv lol)

## b. modeling
```{r}
# mass model
mod.mass <- lm(mass.pupa ~ expt.group, data = data_lab260)
anova(mod.mass)

mod.mass2 <- lm(mass.pupa ~ expt.group*sex, data = data_lab260)
anova(mod.mass2)
  # adding sex doesnt help

mod.mass3 <- lm(mass.pupa ~ sex, data = data_lab260)
anova(mod.mass3)
  # but sex alone is signif, as expected


# quick visualisation
plot.design(mass.pupa ~ expt.group, data = data2_lab260)
  # B is really big? mostly bc it seems like there were 9 f and 7 m though
plot.design(mass.pupa ~ as.factor(sex), data = data2_lab260)
  # as expected
```
word no major effect. sex somewhat confirms it

```{r}
# ttpup model
# a note that only grp F ate TB diet lol

mod.tt <- lm((jdate.pupa - jdate.hatch) ~ expt.group, data = data_lab260)
anova(mod.tt)
  # oops (EG is signif by ***)
summary(mod.tt)

mod.tt2 <- lm((jdate.pupa - jdate.hatch) ~ expt.group*sex, data = data_lab260)
anova(mod.tt2)
  # OOPS.. (only EG is signif by ***)

mod.tt3 <- lm((jdate.pupa - jdate.hatch) ~ sex, data = data_lab260)
anova(mod.tt3)
  # no sig

# lets add the diurnal temps back in lol
data2_lab260 <- data_lab260 %>%
  mutate(flucT = case_when((expt.group == "D" | expt.group == "F") ~ 2.5,
                           TRUE ~ 0))

mod.tt4 <- lm((jdate.pupa - jdate.hatch) ~ flucT, data = data2_lab260)
anova(mod.tt4)
  # diff SS as the expt.group but same level of sig

mod.tt5 <- lm((jdate.pupa - jdate.hatch) ~ flucT*diet, data = data2_lab260)
anova(mod.tt5)
  # flucT and diet both signif
summary(mod.tt5)

mod.tt6 <- lm((jdate.pupa - jdate.hatch) ~ flucT*diet*sex, data = data2_lab260)
anova(mod.tt6)
  # flucT and diet both signif; diet:sex a little sig. sex is not

mod.tt7 <- lm((jdate.pupa - jdate.hatch) ~ diet, data = data2_lab260)
anova(mod.tt7)
summary(mod.tt7)
  # diet and intercept model are both sig lol?

# quick visualisation
plot.design((jdate.pupa - jdate.hatch) ~ expt.group, data = data2_lab260)
  # ... lol
```
errrm so group does have a big effect on ttpup, probably due to the diet + diurnal (sweats;;). see mod 5

let's look at survival...

```{r}
# survival glms

# add some more binaries
data2_lab260 <- data2_lab260 %>%
  mutate(if.pupa2 = case_when(if.pupa == "Y" ~ 0, TRUE ~ 1),
         tt.exit = jdate.exit - jdate.hatch)

# does survival depend on time? mass? diet? temp trt?

```

```{r}
# modeling

# not sure if i should just use KM for anything with time in it...
mod.surv1a <- glm(if.pupa2 ~ mass.pupa*tt.exit, data = data2_lab260, family = binomial(link="logit"))
mod.surv1b <- glm(if.pupa2 ~ mass.pupa*tt.exit, data = data2_lab260, family = binomial)
  # default link for binomial is the logit lol

anova(mod.surv1a)
anova(mod.surv1b)
  # seems like mass + tt dont matter... (but they depend on temp/diet tho?)

mod.surv2 <- glm(if.pupa2 ~ flucT*diet, data = data2_lab260, family = binomial)
anova(mod.surv2)
summary(mod.surv2)
  # diet has a barely signif contribution to overall survival?
  # well that's good, actually

mod.surv3 <- glm(if.pupa2 ~ sex, data = data2_lab260, family = binomial)
anova(mod.surv3)
  # also seems like sex doesnt matter which is good!

mod.surv4 <- glm(if.pupa2 ~ expt.group, data = data2_lab260, family = binomial)
anova(mod.surv4)
  # this is also a good sign i think

# one more quick visualisation
data2_lab260$expt.group <- as.factor(data2_lab260$expt.group)
plot.design(if.pupa2 ~ expt.group, data = data2_lab260)
  # you can see F is the most different and D is also a little different.
  # idk what the numbers mean tho (ranges 0.0 to 0.4. avging 1's and 0's?)

# revisit some of the old models
anova(mod.surv1a, test = "Chisq")
  # u get nothin. thats good i think

# the only model with any signif
anova(mod.surv2, test = "Chisq") # see `?anova.glm`
anova(glm(if.pupa2 ~ diet*flucT, data = data2_lab260, family = binomial), test = "Chisq")
  # so a simpler model is better?

# deciding to skip the cox rn bc it's pretty obvious that survival is different w/in each lab bug cohort. debateable if grp D (diurnal + LD) is diff from A/B though (const + LD)

```

## c. summary...

how did cohort affect lab animals? mostly bc of the temp fluct/diet w/in each cohort ("cohort-related").

- mass wasnt really rlly affected by cohort-related stuff. was significantly affected by sex, as assumed

- tt.pup was significantly affected by fluct + diet in combination? maybe more fluct tho,,, see `mod.tt7`?

- surv seems to be very minorly affected by diet,,, see `mod.surv2`. doesnt super seem to be bc of mass/time (that's good?)

- OVERALL i think this is like. an okay sign of the controls? (i.e. the mean temp wasn't messing things up.) mostly ttpup seems to be affected by all the random stuff i did -- so this'll be important to keep in mind.

# 2. checking field 260s

```{r}
# filter field 260 trts
data_field260 <- wide_all %>%
  filter(treatment == 260 & pop == "field") %>%
  filter(final.fate != "misc" & final.fate != "accidental" & final.fate != "culled")

# calculate pup summary stats
summary_field260 <- data_field260 %>%
  group_by(expt.group) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n_all = n(),
            n_pmd = sum(final.fate == "pmd"), 
            pct.survpup = round(100-(n_pmd/n_all*100), digits=1))
```

## a. visualise

stealing the code from above `lab` stuff lol + adjusting variable names as needed

```{r}
# visualise time to pup, pupal mass across expt groups
summary_field260 %>% ggplot(aes(y = avg.mass, x = avg.tt, color = expt.group)) +
  geom_point() +
  y_err_mass(y_err= 0.9) + x_err_tt(x_err = 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg time to and mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "average time to pupa from hatching", 
       caption = "H = 426 topoff")

# let's isolate more... just masspup
summary_field260 %>% ggplot(aes(y = avg.mass, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_mass(y_err= 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg mass at pupation for lab controls", 
       y = "average pupal mass", 
       x = "expt group", 
       caption = "H = 426 topoff")

# just tt pup
summary_field260 %>% ggplot(aes(y = avg.tt, x = expt.group, fill = expt.group)) +
  geom_col() +
  y_err_tt(y_err= 0.9) + theme_bw() +
  scale_color_hue(labels = c("E (n = 38)",
                                "F (n = 24)",
                                "H (n = 18)")) +
  labs(title = "avg time to pupation for lab controls", 
       y = "average time to pupation from hatching", 
       x = "expt group", 
       caption = "H = 426 topoff")

# visualise % survival
summary_field260 %>% ggplot(aes(y = pct.survpup, fill = expt.group, x = expt.group)) +
  geom_col() + theme_bw() +
  #y_err(avg.mass, se.mass) + #SIGH
  # scale_fill_hue(labels = c("A (n = 21)",
  #                            "B (n = 16)",
  #                            "D (n = 18)",
  #                            "F (n = 19)")) +
  geom_text(aes(label = paste0("n = ", n_all), y = 5), check_overlap = TRUE) +
  labs(title = "% survived to pupation (incl LPI)", 
       y = "% surviving", 
       x = "expt group/cohort", 
       caption = "H = 426 topoff")
```

```{r}
# visualise survival dynamics w/ KM
  # add survival binary & time
data_field260 <- data_field260 %>% 
  mutate(status = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         tt.exit = jdate.exit - jdate.hatch)

# fit model
kmfit.field260 <- survfit(Surv(data_field260$tt.exit, data_field260$status) ~ expt.group, data = data_field260)

# plot
ggsurvplot(fit = kmfit.field260, data = data_field260,
           conf.int=T)
  # they all have huuuuge CIs...
```


# ... checking lab 337s

(population effects will be addressed in NTs only!)




