---
title: "temps_analyses"
date: "2023-12-07"
output: html_notebook
---

# 0.1 load data & packages
```{r message = FALSE}
library(conflicted)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(tidyr)
library(purrr)
library(ggplot2)

# survival curve stuff
library(survival) # for computing
library(survminer) # for visualising

wide_all <- read.csv("~/Documents/repos/_not-public/1_data/ntw_data/clean-gsheets.csv", header = TRUE)

```

## - pivot to long
```{r}
long_all <- wide_all %>% select(-(starts_with("date."))) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch, 
         tt.wander = jdate.wander - jdate.hatch,
         #tt.pupa = jdate.pupa - jdate.wander,
         tt.pupa = jdate.pupa-jdate.hatch,
         tt.15 = jdate.15-jdate.hatch,
         tt.eclose = jdate.eclose-jdate.pupa,
         tt.exit = jdate.exit - jdate.enter,
         tt.surv = jdate.surv - jdate.hatch) %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "h", "tt")),
               names_to = c(".value", "instar"),
               #names_sep = ".",
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  rename(molt.status = h) %>%
  drop_na(jdate) %>% drop_na(tt) %>% # drops NA's if an individual didnt reach a certain stage
  filter(instar != "15")

# add instar factor levels
long_all <- long_all %>% mutate(instar = factor(instar, levels=c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit")))
#long_all$instar <- factor(long_all$instar, c("hatch", "2nd", "3rd", "4th", "5th", "6th", "7th", "stuck", "wander", "15", "pupa", "eclose", "exit"))

#rm(data_all)

```

# 0.2 define helpful functions & aesthetics

## - filtering groups
```{r}
# separate experiment groups

filter.temps2 <- function(data) {
  filtered_data <- data %>% 
    filter(trt.stage == "260-hatch" | trt.stage =="267-hatch" | trt.stage == "330-hatch" | trt.stage == "337-hatch") %>%
    filter(expt.group == "A" | expt.group == "B")
  
  return(filtered_data)
}

# separate instars
filter.ins.topup <- function(data) {
  filtered_data <- data %>%
      filter(instar == "4th" | instar == "5th" | instar == "6th" | instar == "7th" | instar == "wander" | instar == "pupa")
  
  return(filtered_data)
}
```

## - doing math
```{r}
# calc dev summ stats

calc.devsumm <- function(data) {
  summary <- data %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  return(summary)
}

calc.devsumm.trtstg <- function(data) {
  summary <- data %>%
    group_by(pop, trt.stage, instar) %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              logmass = log(mass),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  
  return(summary)
}

calc.devsumm.trtstgsex <- function(data) {
  summary <- data %>%
    group_by(pop, sex, trt.stage, instar) %>%
    summarise(avg.mass = mean(na.omit(mass)),
              se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
              avg.tt = mean(na.omit(tt)),
              se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
              logmass = log(mass),
              avg.logmass = mean(na.omit(logmass)),
              se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
              n=n())
  
  return(summary)
}

```


## - adding geoms
```{r}
# error bars
y_err_logmass <- function(x) {
  list(geom_errorbar(aes(ymin = avg.logmass - se.logmass, ymax = avg.logmass + se.logmass), width = x))
}

y_err_mass <- function(x) {
  list(geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = x))
}

y_err_tt <- function(x){
  list(geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = x))
}

x_err_tt <- function(x){
  list(geom_errorbarh(aes(xmin = avg.tt - se.tt, xmax = avg.tt + se.tt), width = x))
}

# recoloring theme + legend
temp_aes <- function(x)(
  list(theme_bw(), scale_color_manual(values=temp_colors, labels=temp_labels))
)
```


## - define aesthetics
```{r}
# temp: effect of mean/fluct temp (treatment)
temp_trts = c("260-hatch", "267-hatch", "330-hatch", "337-hatch")
temp_labels = c("260-hatch"="26°C", "267-hatch"="26±7°C", "330-hatch"="33°C", "337-hatch"="33±7°C")
temp_colors = c("260-hatch"="#00C2D1","267-hatch"="#1929B3", "330-hatch"="#F9C639", "337-hatch"="#710A36")
```

 - calc timeto (3rd-pup)
```{r}
pivot.ttpup <- function(wide_data) {
  pivoted_data <- wide_data %>%
  select(-starts_with(c("h", "date", "mass", "reason."))) %>%
  mutate("3rd" = jdate.3rd - jdate.hatch,
         "4th" = jdate.4th - jdate.hatch,
         "5th" = jdate.5th - jdate.hatch,
         "6th" = jdate.6th - jdate.hatch,
         "7th" = jdate.7th - jdate.hatch, 
         wander = jdate.wander - jdate.hatch,
         pupa = jdate.pupa - jdate.wander) %>%
  select(-starts_with("jdate.")) %>%
  pivot_longer (cols = c("3rd", "4th", "5th", "wander", "pupa"), # pivot differences by stage
                names_to = "instar",
                values_to = "days.to") %>%
  drop_na(days.to) # drop whoever didnt hit a certain stage
  
  return(pivoted_data)
}
```

# 2.0 super simple stats
```{r}
wide_all %>%
  filter.temps() %>% filter(pop == "lab") %>%
  group_by(treatment) %>%
  summarise(n=n(),
            pct.pmd = round(sum(final.fate == "pmd")/n*100, digits = 1),
            #pct.slow = round(sum(ignore.reason == "slow")/n*100, digits = 1),
            pct.sup = round(sum(if.sup == "Y")/n*100, digits = 1),
            pct.6th = round(sum(sup == 6)/sum(if.sup == "Y")*100, digits = 1),
            pct.7th = round(sum(sup == 7)/sum(if.sup == "Y")*100, digits = 1))
            #n.pmd = sum(final.fate == "pmd"),)

```

## - sup barplots
```{r}
# make these into barplots instead

wide_all %>%
  filter.temps2() %>% filter(pop == "lab") %>%
  mutate(sup = factor(sup)) %>%
  ggplot(aes(x = trt.stage, fill = sup)) +
  geom_bar(stat = "count", position = position_fill(reverse = TRUE)) +
  labs(title = "temps: supernumerary fates", fill = "supernumerary stage", x = "treatment", y = "percent") +
  scale_x_discrete(labels = temp_labels) +
  theme_bw() + scale_fill_brewer(palette = "Greens")
```

# 3.0 overall growth trends; end points

compare overall development of all individuals compared to average growth

## - set up the df (**)
```{r}
# pull out the lab bugs
long_lf <- long_all %>% filter(pop == "lab" | pop == "field") 

# individual dev stats
dev_L <- long_lf %>%
  filter(!(final.fate == "pmd" | final.fate == "misc")) %>%
  filter.ins.topup() %>%
  mutate(logmass = log(mass))

devsumm_L <- dev_L %>%
  group_by(pop, trt.stage, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
            avg.logmass = mean(na.omit(logmass)),
            se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
            n=n())

# for pulling out supernumeraries.. need to tweak
# devsumm_Lsups <- dev_L %>%
#   group_by(pop, sup, trt.stage, instar) %>%
#   summarise(avg.mass = mean(na.omit(mass)),
#             se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
#             avg.tt = mean(na.omit(tt)),
#             se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt))),
#             avg.logmass = mean(na.omit(logmass)),
#             se.logmass = sd(na.omit(logmass))/sqrt(length(na.omit(logmass))),
#             n=n())

# see 'plot NTs' under 'overall growth'
  
```

## - plot temps
```{r}
temps_dev <- dev_L %>% filter(pop == "lab") %>% filter.temps2()
#temps_devsumm <- devsumm_L %>% filter(pop == "lab") %>% filter.temps2()
temps_devsumm <- dev_L %>% filter.temps2() %>% calc.devsumm.trtstg()

temps_devsumm %>%
  filter(instar != "pupa") %>%
  ggplot(aes(x=avg.tt, y=avg.logmass, color=trt.stage)) +
  geom_line() +
  geom_point(data=(temps_dev %>% filter(instar != "pupa")), aes(x=tt, y=logmass, color=trt.stage), alpha = 0.15) + 
  geom_point(aes(shape = instar), size = 3) +
  #y_err_logmass(0.5) + x_err_tt(1) +
  temp_aes() +
  labs(title = "temps: average development from 3rd - wander", color = "treatment", y = "log(mass (mg))", x = "days to instar",
       caption = "lab, no pmd, enter at hatch")

temps_devP <- temps_dev %>% filter(instar == "pupa")
  # tbh can drop lol

temps_devsummP <- dev_L %>%
  filter.temps2() %>%
  group_by(trt.stage, sex, instar) %>%
  summarise(avg.mass = mean(na.omit(mass)),
            se.mass = sd(na.omit(mass))/sqrt(length(na.omit(mass))),
            avg.tt = mean(na.omit(tt)),
            se.tt = sd(na.omit(tt))/sqrt(length(na.omit(tt)))) %>% 
  filter(instar == "pupa")

temps_devsummP %>%
  ggplot(aes(x=avg.tt, y=avg.mass, color=trt.stage)) +
  geom_point(size = 2) +
  geom_point(data=temps_devP, aes(x=tt, y=mass, color=trt.stage), alpha = 0.275) +
  y_err_mass(0.5) + x_err_tt(1) +
  temp_aes() + facet_wrap(~sex) +
  labs(title = "temps: average development to pupa", color = "treatment", y = "mass (mg)", x = "days to instar")
```

# 4.0 survival to pupation

## - censoring + subsetting
```{r}
# censor data
surv_P <-  wide_all %>%
  filter(!(final.fate == "ignore" | final.fate == "culled")) %>%
  mutate(censor = case_when(final.fate == "pmd" ~ 1,
                            TRUE ~ 0),
         timeto.exit = jdate.exit - jdate.hatch) %>%
  select(-starts_with(c("h.", "date.", "mass.", "jdate.", "fate.")))

temps_surv <- surv_P %>% filter.temps2()
```

## - create survival objects and fits
```{r}
temps_sobj <- Surv(temps_surv$timeto.exit, temps_surv$censor)
temps_sfit <- survfit(temps_sobj ~ treatment, data = temps_surv)
```

## - plot survival curves
```{r}
ggsurvplot(fit = temps_sfit, data=temps_surv,
           conf.int = T,
           xlab = "days from hatching", ylab = "survival probability",
           title = "temp: survival from hatch-pup",
           #palette = temp_colors,
           palette = c("#00C2D1","#1929B3", "#F9C639", "#710A36"),
           legend.title = "treatment", 
           legend.labs = temp_labels
           ) +
  labs(caption = "(uses actual cohort; ignores 'ignores' + 'culled')")
```

# 5.0 cox-ph survival stuff?

- check as variables: lab/TB diet; if.sup; if.stuck

##  - if sup x to pup/eclosion

```{r}
#check_suppup <- wide_all %>% filter(if.pupa == "Y")

#mod_suppup <- lm(if.pupa ~ if.sup + treatment + pop, data = check_suppup, na.action=na.omit)
  # does this not work bc Y needs to be a # LOL

# i think this shoudl be part of survival stuff

```