---
title: "tents_data_cleaning"
date: "2023-02-15"
---

# roadmap

1. load data & packages
2. cleaning: prefiltering, fix columns, add columns, value fixing
3. save output

# 1. load data & packages
```{r message = FALSE}
library(tidyverse)
library(googlesheets4)
#conflicts_prefer(dplyr::filter)

here::i_am("_ntw/scripts/import/2023-tents.Rmd")
library(here)
```

## - data import from gsheets

```{r}
# before importing,
  # run cleaning macros,
  # check column/row alignment + proper sorting

dfs_raw <- list(
  pairs = read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="mating pairs", col_types = "c", na = c("#N/A", "NA", "na", "--", "", "#REF!")),
  
  tents = read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="tent stats v2", col_types = "c", range = "A:Q", na = c("#N/A", "NA", "--", "", "#REF!", "unk")),
  
  hatch = read_sheet("https://docs.google.com/spreadsheets/d/1dRgmeeCW1Ou5ayrtXnEL6NBKLB3Leo9ZHec7Y9AFhm0/edit#gid=188618561", sheet="hatch stats", col_types = "c", range = "A:F", na = c("#N/A", "NA", "--", "", "#REF!"))
)
```

# 2. cleaning

matepairs:
- *drop extra na rows*
- *drop hidden columns(?)*
- *replace #N/A with actual NAs*
- *replace "col" and "lab" with actual # values, change column format*
- fill in other tents (i.e cages) w/ sth
- *standardise column names*

## - standardise and consolidate columns
```{r}
# drop unneeded columns/rows and standardise column names
dfs_filtered <- list(
  pairs = dfs_raw[["pairs"]] %>%
    select(-c("sex_A", "sex_B", "id_F2", "id_M2", "id_A-orig", "id_B-orig", "combo2")) %>%
      rename(trt.pair = combo,
         id.tent = tent_id,
         id.pair = pair_id,
         room = location,
         date.paired = date_paired,
         id.f = id_F,
         date.f.died = died_F,
         id.m = id_M,
         date.m.died = died_M,
         trt.f = trt_F,
         date.f.ec = eclose_F,
         track.f = purpose_F,
         trt.m = trt_M,
         date.m.ec = eclose_M,
         track.m = purpose_B) %>%
    drop_na(id.tent),
  
  tents = dfs_raw[["tents"]] %>% 
    select(-starts_with(c("total", "lifetime"))) %>%
    rename(date.coll = date,
         id.tent = tent.ID, # dropping the splits
         n.coll = "eggs collected",
         n.new.f = "females added",
         n.curr.f = "females alive current",
         n.active.f = "females laying", # doesnt actually get used in analysis scripts..
         n.died.f = "females dead",
         n.new.m = "males added",
         n.curr.m = "males alive",
         n.died.m = "males dead",
         notes.tents = notes),
    
  hatch = dfs_raw[["hatch"]] %>%
      rename(date.hatch = date.hatchlings,
         id.tent = "tent id",
         id.coll = "coll-tent", # collection id (date-tent)
         n.hatch = n.hatchlings,
         notes.hatch = notes) %>%
    drop_na(id.tent)
)
    

```


## - consolidation, fixing column format + values

bc i majorly messed up with how i did my hatch data collection so now i am paying the price 👍

the summed # of hatchlings per day is being appended to the daily tent happenings

```{r reshape hatch data}
#data_tents <- merge(data_tentstats, data_hatch, by = c("date.coll", "id.tent"), all = TRUE) 

# hatchinfo <- select(data_tents, c(id.coll, id.tent, date.coll, date.hatch, n.hatch, notes.hatch)) %>%
#   drop_na(date.hatch)

# data_tents <- select(data_tentstats, -c(date.hatch, n.hatch, id.coll, notes.hatch)) %>%
#   distinct()

dfs_cleaned <- list(
  hatch = dfs_filtered[["hatch"]],
  
  tents = 
    full_join(dfs_filtered[["tents"]], 
              dfs_filtered[["hatch"]] %>%
                mutate(n.hatch = as.numeric(n.hatch)) %>%
                group_by(id.tent, date.hatch) %>%
                summarise(n.tothatch = sum(n.hatch)),
              by = c("id.tent", "date.coll" = "date.hatch"), all = TRUE) %>%
    mutate(trt.f = case_when(trt.f == "ctrl" ~ 900, # bc 107-I gets yeeted
                             TRUE ~ as.numeric(trt.f))) %>%
    group_by(id.tent) %>% # fill in missing parent trt info
    fill(c(trt.f, trt.m), .direction = "updown") %>%
    ungroup() %>%
    mutate(trt.f = case_when(id.tent == "301-Z" ~ 260,
                             TRUE ~ as.numeric(trt.f)),
           trt.m = case_when(id.tent == "107-H" ~ 260,
                             #id.tent == "107-I" ~ ,
                             #id.tent == "301-T" ~ ,
                             id.tent == "301-U" ~ 260,
                             TRUE ~ as.numeric(trt.m)))
)

```

(values to fix: nas and stuff.. replacements... unks... etc... better strings to filter on... see above)
```{r formatting}
# consolidate
#data_hatch <- select(data_hatch, -date.coll)

# data_tents <- merge(data_tentstats, data_hatch, by = c("date.coll", "id.tent"), all = TRUE) 

# data_tents <- left_join(data_tentstats, data_hatch, by = c("date.coll" = "date.hatched", "id.tent"), all = TRUE) 

# per-df cleaning
dfs_cleaned <- list_modify(
  dfs_cleaned,
  tents = dfs_cleaned[["tents"]] %>%
    filter(is.na(if.ignore)) %>%
    rename(date = date.coll),
  
  pairs = dfs_filtered[["pairs"]] %>%
    mutate(id.pair = as.numeric(id.pair),
         room = as.numeric(room),
         tent = case_when(is.na(tent) ~ "cage",
                          TRUE ~ as.character(tent)),
         id.tent = paste(room, tent, sep = "-"), # diff from paste(x1, "-", x2) lol idky
         tent.loc = case_when(room == 301 & (tent %in% c("D", "J", "Q", "Y")) ~ "L1",
                              room == 301 & (tent %in% c("C", "L", "O", "W")) ~ "L2",
                              room == 301 & (tent %in% c("B", "N", "T")) ~ "L3",
                              room == 301 & (tent %in% c("A", "V")) ~ "L4",
                              room == 301 & (tent %in% c("Z")) ~ "L5",
                              room == 301 & (tent %in% c("E", "K", "R")) ~ "R1",
                              room == 301 & (tent %in% c("F", "M", "P", "S")) ~ "R2",
                              room == 301 & (tent %in% c("G", "U")) ~ "R3",
                              room == 301 & (tent %in% c("H", "X", "Z")) ~ "R4",
                              TRUE ~ "107")) %>%
    naniar::replace_with_na_all(condition = ~.x %in% c("unk", "na"))
)

# column/value standardising
## column formatting
dfs_cleaned <- 
  dfs_cleaned %>% lapply(\(x){
    x %>%
      mutate(
        across(starts_with("date"),  as.Date, format = "%m/%d/%y"),
        across(starts_with("date"), as.Date, "%j"),
        across(starts_with(c("date", "n.")), as.numeric)) %>%
      rename_with(~ paste0("j", .x), starts_with("date"))
  }
)

## fix trt values
dfs_cleaned[c("tents", "pairs")] <- lapply(
  dfs_cleaned[c("tents", "pairs")], \(x){
    x %>%
      mutate(
           trt.m = case_when(trt.m == "col" ~ 999,
                             trt.m == "ctrl" ~ 900,
                             TRUE ~ as.numeric(trt.m)),
           trt.f = case_when(trt.f == "col" ~ 999,
                             trt.f == "ctrl" ~ 900,
                             TRUE ~ as.numeric(trt.f))
           )
  })

```


```{r add addtl room, pop info}
dfs_cleaned[["tents"]] <- left_join(
  dfs_cleaned[["tents"]], 
  dfs_cleaned[["pairs"]] %>%
    select(id.tent, #room, tent, tent.loc, 
           pop) %>%
    mutate(pop = case_when(pop == "lab + col ctrl" | pop == "lab + 260 ctrl" ~ "lab",
                           TRUE ~ as.character(pop))) %>%
    distinct(), 
  by = "id.tent")
#dfs_cleaned[["tents"]] <- dfs_cleaned[["tents"]][, c(1, 15:16, 2, 17, 3, 14, 4:13, 18:19)]
  # dont need anymore after removing the extra cols...
```

# 3. extracting notes & saving outputs

```{r}
today <- format(Sys.time(), "%y%m%d-%H%M")

# raw
write_csv(dfs_raw[["pairs"]], file = paste0(here("_ntw/data/raw/2023/"), "tent-matings_", today, ".csv")) # "tentpairs"
write_csv(dfs_raw[["tents"]], file = paste0(here("_ntw/data/raw/2023/"), "tent-events_", today, ".csv")) # "tentstats"
write_csv(dfs_raw[["hatch"]], file = paste0(here("_ntw/data/raw/2023/"), "eggs_", today, ".csv")) # "hatching"

# save notes elsewhere
# notes_pairs <- select(data_pairs, c(trt.pair, id.tent, id.pair, date.paired, id.f, id.m, notes))
# notes_tents <- select(data_tents, date, id.tent, room, tent, tent.loc, notes.tents)
# 
# write.csv(notes_pairs, "~/Documents/repos/ntw/2023/data/notes_tentpairs.csv", row.names = FALSE)
# write.csv(notes_pairs, paste0("~/Documents/repos/ntw/2023/data/archive/", today, "_notes_tentpairs.csv"), row.names = FALSE)
# 
# write.csv(notes_tents, "~/Documents/repos/ntw/2023/data/notes_tentstats.csv", row.names = FALSE)
# write.csv(notes_tents, paste0("~/Documents/repos/ntw/2023/data/archive/", today, "_notes_tentstats.csv"), row.names = FALSE)

# save cleaned dfs
dfs_cleaned <- dfs_cleaned %>%
  lapply(\(x){
    x %>%
    select(-starts_with("notes"))
  })

export <- dfs_cleaned[["tents"]] %>% select(-if.ignore)
write.csv(export, file = paste0(here("_ntw/data/clean/2023/"), "tent-info.csv"), row.names = FALSE)
#write.csv(export, file = paste0(here("_ntw/data/archive/2023/"), "tents_", today, ".csv"), row.names = FALSE)


export <- dfs_cleaned[["pairs"]]
write.csv(export, file = paste0(here("_ntw/data/clean/2023/"), "tent-pairs.csv"), row.names = FALSE)
#write.csv(export, file = paste0(here("_ntw/data/archive/2023/"), "pairs_", today, ".csv"), row.names = FALSE)

export <- dfs_cleaned[["hatch"]]
write.csv(export, file = paste0(here("_ntw/data/clean/2023/"), "tent-hatch.csv"), row.names = FALSE)
#write.csv(export, file = paste0(here("_ntw/data/archive/2023/"), "hatch_", today, ".csv"), row.names = FALSE)
```

```{r}
# cleanup
rm(dfs_raw, dfs_filtered, 
   dfs_cleaned, today, export)
```


```{r}
# ref code on running fns across a list lol

# formatear <- function(eq){
#   eq$Volume <- NULL
#   eq$Date <- as.Date(eq$Date)
#   
#   nt <- eq$Adj.Close[1:nrow(eq)-1]
#   nt1 <- eq$Adj.Close[2:nrow(eq)]
#   eq$return <- percent(c(NA, nt1/nt-1), accuracy = 0.0001)
#   return(eq) 
# }

#You can use mget to get list of dataframes and apply the function with lapply.

# clean_list_data <- lapply(mget(dfs), formatear)
# clean_list_data[[1]]
```