---
title: "03-viz/compare-tents"
date: "2025-01-13"
---

viz for comparing 2023 and 2024 fertility data

for 2023, should look at:

- hatch prop vs avg eggs/f per trt (btwn and win)

for 2024, should look at

- avg eggs/f per trt

# setup
```{r}
here::i_am("2024/03-viz/compare-tents.Rmd")
library(here)

library(tidyverse)
library(patchwork)

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "compare-tents_p.R"))
source(here::here(bin_paths24$doviz, "aesthetics_util.R"))
```

## df setup

```{r}
# sums are for across whole existence of cage/tent (`t.duration`)
  # pre-emptively filtering out the field pop bc they never wanna lay eggs
ss_fert <- dall %>%
  group_by(year, pop, trt.temp,
           trt.type, trt.sex) %>%
  summarise(n.reps = n_distinct(id),
            n.f = sum(n.new.f, na.rm=TRUE),
            n.colled = sum(n.coll, na.rm=TRUE),
            #avg.colled = mean(n.coll, na.rm=TRUE),
            se.colled = se(n.coll),
            avg.colledperf = n.colled/n.f,
            n.hatched = sum(n.hatch, na.rm=TRUE),
            #avg.hatched = mean(n.hatch, na.rm=TRUE),
            se.hatched = se(n.hatch),
            p.hatched = n.hatched/n.colled,
            se.phatched = sqrt(p.hatched*(1-p.hatched)/n.colled),
            t.duration = max(jdate) - min(jdate)) %>% 
  mutate(across(ends_with("hatched"), ~ case_when(is.nan(.) ~ NA_real_,
                                                  year == 2023 ~ as.numeric(.))),
         across(starts_with(c("n", "avg", "se", "p")), ~ round(., digits = 3)),
         minT = case_when(trt.temp == 419 ~ 19,
                          trt.temp %in% c(260, 426) ~ 26,
                          TRUE ~ 33),
         minT = factor(minT, levels = c(19, 26, 33)),
         for.comparef = case_when(year == 2024 |
                                    (year == 2023 & (trt.sex == "f" | trt.temp == 260)) ~ "Y")) %>%
  filter(pop == "lab")

# remove the 2023-win-f? (must be the solo f?) -- jk they fall out naturally when u keep the lab pop only

```
summ_longevity <- data_longevity %>%
  mutate(surv = jdate.surv - jdate.eclose) %>%
  group_by(expt.type, pop, #sex, 
           minT) %>%
  summarise(n = n(),
            avg.surv = mean(na.omit(surv)),
            se.surv = sd(na.omit(surv))/sqrt(length(na.omit(surv))))
```

```{r plot}
summ_longevity %>%
  filter(expt.type == "expt") %>%
  #filter(!(pop == "lab" & diet.type == "TB")) %>% 
  ggplot(aes(y = avg.surv, x = as.factor(minT), 
             #group = interaction(pop, sex),
             group = pop)) +
  geom_point(size = 2) +
  geom_line(aes(lty = pop)) +
  geom_errorbar(aes(ymin = avg.surv - se.surv, ymax = avg.surv + se.surv), width = 0.25) +
  geom_text(aes(label = n), hjust = -1, vjust = 0.75) +
  theme_bw() 
  #facet_wrap(~sex)
```


```{r}
# tents math from 2023 for ref... 

summ_hatch <- data_tstats %>%
  group_by(#id.tent,
           trt.f, trt.m, pop,
           #trt.win, trt.btwn, 
           trt.type,
           trt.pair
           ) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.collected = sum(n.coll, na.rm = TRUE),
            se.coll =  sd(na.omit(n.coll))/sqrt(length(na.omit(n.coll))),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            se.hatched = sd(na.omit(n.tothatch))/sqrt(length(na.omit(n.tothatch))),
            n.females = sum(n.new.f, na.rm = TRUE),
            #start = first(jdate), end = last(jdate),
            t.duration = max(jdate) - min(jdate),
            rate.hatch = n.hatched/n.collected,
            se.hatch = sqrt(rate.hatch*(1-rate.hatch)/n.collected),  # check ???
            rate.hatchperf = rate.hatch/n.females,
            se.hatch = sqrt(rate.hatchperf*(1-rate.hatchperf/n.collected)),
            rate.f = round(n.collected/n.females, digits = 1)
            ) %>%
  drop_na(trt.m) %>% # drop tents w/o males
  mutate(trt.win = case_when(trt.f == trt.m & trt.f == 260 ~ "26-26",
                             trt.f == trt.m & trt.f == 419 ~ "40-19",
                             trt.f == trt.m & trt.f == 426 ~ "40-26",
                             trt.f == trt.m & trt.f == 433 ~ "40-33"),
         trt.btwn = case_when(trt.type == "ctrl" ~ "26-26", 
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 419 | trt.m == 419) ~ "40-19",
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 426 | trt.m == 426) ~ "40-26",
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 433 | trt.m == 433) ~ "40-33"),
         # min.f = case_when(trt.f == 260 | trt.f == 426 ~ 26,
         #                   trt.f == 419 ~ 19,
         #                   trt.f == 433 ~ 33), 
         # min.m = case_when(trt.m == 260 | trt.m == 426 ~ 26,
         #                   trt.m == 419 ~ 19,
         #                   trt.m == 433 ~ 33),
         trt.parent = case_when(trt.f == 260 & trt.m == 260 ~ 260,
                                trt.f == 419 | trt.m == 419 ~ 419,
                                trt.f == 426 | trt.m == 426 ~ 426,
                                trt.f == 433 | trt.m == 433 ~ 433),
         # trt.combo = paste0(trt.f, "-", trt.m),
         # min.combo = paste0(min.f, "-", min.m)
         )
```

