---
title: "03-viz/compare-tents"
date: "2025-01-13"
---

viz for comparing 2023 and 2024 fertility data

for 2023, should look at:

- hatch prop vs avg eggs/f per trt (btwn and win)

for 2024, should look at

- avg eggs/f per trt

# setup
```{r}
here::i_am("2024/03-viz/compare-tents.Rmd")
library(here)

library(tidyverse)
library(patchwork)

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "compare-tents_p.R"))
source(here::here(bin_paths24$doviz, "aesthetics_util.R"))
```

## helper functions
```{r helper functions}
# do math
se <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
```

```{r more helper fn testing}
# aes
## lazy error bars: call column names with quotes

get_err <- function(vector){
  # make an if? (if vector = a #, use it (so can just set width = to it), but if not then do math below)
  return(max(vector, na.rm=TRUE)*1.02*0.02) # set err bar width to 2% of max axis + 2%
}

# mean = col w/ "avg", 
geom_yerr <- function(#.data, 
                      mean, se, x
                      #, wd = 0.9
                      ){
  # if(missing(wid)){
  #   wid = max(cur_data()[[mean]])*0.02
  # }
  
  # if(missing(wid)) x <- cur_data_all()[[mean]]
  #   wid <- max(x)*0.02
    
  list(geom_errorbar(aes(ymin = .data[[mean]] - .data[[se]], 
                         ymax = .data[[mean]] + .data[[se]],
                         #width = wd # can i set width to "get_err(x)" here? using whats alrdy declared?
                         #width = get_err(.data[[mean]])
                         #width = get_err(.data)
                         #width = data[[mapping]][[x]][[2]] # maybe sth to do with envir = parent.frame() lol
                         width = get_err(.data[[x]])
                         
                         # if(is.character(x))
                         #   {
                         #   width = get_err(.data[[x]])
                         # } else width = x
                         )))
}

geom_xerr <- function(mean, se, y
                      #, ht = 0.9
                      ){
  list(geom_errorbarh(aes(xmin = .data[[mean]] - .data[[se]], 
                          xmax = .data[[mean]] + .data[[se]],
                          #height = ht,
                          height = get_err(.data[[y]]))))
}
```

holy fuk been trying to figure out how to do custom error bars FOREVER see here:

- https://stackoverflow.com/questions/22309285/how-to-use-a-variable-to-specify-column-name-in-ggplot
- https://stackoverflow.com/questions/62961284/having-trouble-with-a-custom-geom-that-transforms-single-row-aesthetics-to-multi (sorta outside the scope but interesting)

but auto-width is still an issue:

- https://stackoverflow.com/questions/19420903/width-of-error-bars-in-line-plot-using-ggplot2/29059601#29059601 (use 0.02 of axis range)
- https://stackoverflow.com/questions/7705345/how-can-i-extract-plot-axes-ranges-for-a-ggplot2-object
- https://stackoverflow.com/questions/28370249/correct-way-to-specifiy-optional-arguments-in-r-functions
- https://stackoverflow.com/questions/69109106/default-arguments-not-being-recognized-in-custom-function-using-dplyr
- https://stackoverflow.com/questions/53767647/write-a-function-with-default-column-name-inputs-in-dplyrmutate
- https://dplyr.tidyverse.org/articles/programming.html#one-or-more-user-supplied-expressions
- https://stackoverflow.com/questions/79195871/passing-a-global-object-value-to-a-custom-ggplot-function-argument
- https://stackoverflow.com/questions/2641653/pass-a-data-frame-column-name-to-a-function
- https://forum.posit.co/t/extending-ggplot-to-take-lists-and-make-the-internal-elements-of-that-list-available-to-custom-geoms/29870/3

- https://forum.posit.co/t/accessing-ggplots-inherited-data-object-inside-custom-layer/114718 *ughh this is exactly what im trying to do*

## df setup

```{r}
# sums are for across whole existence of cage/tent (`t.duration`)
  # pre-emptively filtering out the field pop bc they never wanna lay eggs
ss_fert <- dall %>%
  group_by(year, pop, trt.temp,
           trt.type, trt.sex) %>%
  summarise(n.reps = n_distinct(id),
            n.f = sum(n.new.f, na.rm=TRUE),
            n.colled = sum(n.coll, na.rm=TRUE),
            #avg.colled = mean(n.coll, na.rm=TRUE),
            se.colled = se(n.coll),
            avg.colledperf = n.colled/n.f,
            #test = n.coll/n.ovi, 
            #se.colledperf = sd(na.omit(n.coll)/na.omit(n.ovi))/sqrt(length(na.omit(n.coll)/na.omit(n.ovi))),
            n.hatched = sum(n.hatch, na.rm=TRUE),
            #avg.hatched = mean(n.hatch, na.rm=TRUE),
            se.hatched = se(n.hatch),
            p.hatched = n.hatched/n.colled,
            se.phatched = sqrt(p.hatched*(1-p.hatched)/n.colled),
            t.duration = max(jdate) - min(jdate)) %>% 
  mutate(across(ends_with("hatched"), ~ case_when(is.nan(.) ~ NA_real_,
                                                  year == 2023 ~ as.numeric(.))),
         across(starts_with(c("n", "avg", "se", "p")), ~ round(., digits = 3)),
         minT = case_when(trt.temp == 419 ~ 19,
                          trt.temp %in% c(260, 426) ~ 26,
                          TRUE ~ 33),
         minT = factor(minT, levels = c(19, 26, 33)),
         for.comparef = case_when(year == 2024 |
                                    (year == 2023 & (trt.sex == "f" | trt.temp == 260)) ~ "Y")) %>%
  filter(pop == "lab")

# remove the 2023-win-f? (must be the solo f?) -- jk they fall out naturally when u keep the lab pop only

```


# plots

```{r}
ss_fert %>%
  filter(year == 2023 #& trt.temp != 260
         ) %>%
  mutate(trt.temp = as.factor(trt.temp)) %>%
  ggplot(aes(y = p.hatched, x = avg.colledperf,
             col = trt.temp, shape = trt.sex)) +
  geom_point(size = 1.5) +
  geom_yerr("p.hatched", "se.phatched", "avg.colledperf") +
  geom_xerr("avg.colledperf", "se.colled", "p.hatched") +
  geom_text(aes(label = n.f), vjust = -0.6, hjust = -0.65,
            show.legend = FALSE) +
  facet_grid(~trt.type, labeller = as_labeller(c(`win` = "both sexes HS'd",
                                                 `btwn` = "1 sex HS'd"))) +
  labs(caption = "2023 hatch proportion/egg laying;\nn = total females, se = se total collected",
       y = "hatch proportion",
       x = "avg eggs collected per day",
       shape = "heat-treated sex",
       color = "larval rearing\ntemperature (°C)") +
  scale_color_manual(values = bin_aes$temp4$col, labels = bin_aes$temp4$labswC) +
  scale_shape_manual(values = bin_aes$sexes$ptfill_all, 
                     #limits = c("f", "m")
                     ) +
  theme_bw()
```

```{r}
ss_fert %>%
  filter(year == 2024) %>%
  mutate(trt.temp = as.factor(trt.temp),
         trt.type = factor(trt.type, levels = c("expt", "ctrl"))) %>%
  ggplot(aes(y = avg.colledperf, x = minT,
             col = trt.temp,)) +
  geom_point(size = 1.5) +
  geom_errorbar(aes(ymin = avg.colledperf - se.colled, 
                    ymax = avg.colledperf + se.colled,
                    width = 0.15)) +
  #geom_yerr("avg.colledperf", "se.colled", "minT") +
  geom_text(aes(label = n.f), vjust = -0.6, hjust = -0.65,
            show.legend = FALSE) +
  facet_grid(~trt.type, labeller = as_labeller(c(`expt` = "40-X°C",
                                                 `ctrl` = "26-26°C"))) +
  labs(caption = "2024 egg laying;\nn = total females; se = se total collected",
       y = "avg eggs collected per day",
       x = "minimum rearing temperature (°C)") +
  scale_color_manual(values = bin_aes$temp4$col) +
  guides(color = "none") +
  theme_bw()
```

```{r}
# 23 and 24
ss_fert %>%
  filter(trt.temp != 260 & trt.sex != "m") %>%
  mutate(trt.temp = as.factor(trt.temp),
         trt.type = "expt") %>% 
  # prob need to redo the math lol and put the trts together in 2023...
  ggplot(aes(y = avg.colledperf, x = minT, 
             col = year, shape = trt.sex))+
  geom_point(size = 1.5) +
  geom_errorbar(aes(ymin = avg.colledperf - se.colled, 
                    ymax = avg.colledperf + se.colled,
                    width = 0.15)) +
  geom_text(aes(label = n.f), vjust = -0.6, hjust = -0.65,
            show.legend = FALSE) +
  geom_line(aes(lty = trt.sex, group = interaction(year, trt.sex))) +
  facet_grid(~trt.type, labeller = as_labeller(c(`expt` = "40-X°C"))) +
  labs(caption = "23v24 egg laying;\nn = total females; se = se total collected",
       y = "avg eggs collected per day",
       x = "minimum rearing temperature (°C)") +
  #scale_color_manual(values = bin_aes$temp4$col) +
  theme_bw()


ss_fert %>%
  filter(trt.temp == 260 | trt.sex == "m") %>%
  mutate(trt.temp = as.factor(trt.temp),
         trt.type = "ctrl") %>%
  ggplot(aes(y = avg.colledperf, x = year, 
             col = trt.temp, shape = trt.sex))+
  geom_point(size = 1.5) +
  geom_errorbar(aes(ymin = avg.colledperf - se.colled, 
                    ymax = avg.colledperf + se.colled,
                    width = 0.1)) +
  geom_line(aes(lty = trt.temp,
                group = interaction(trt.temp))) +
  geom_text(aes(label = n.f), vjust = -0.6, hjust = -0.65,
            show.legend = FALSE) +
  facet_grid(~trt.type, labeller = as_labeller(c(`ctrl` = "26-26°C"))) +
  labs(caption = "23v24 egg laying;\nn = total females; se = se total collected",
       y = "avg eggs collected per day",
       x = "minimum rearing temperature (°C)") +
  scale_color_manual(values = bin_aes$temp4$col) +
  theme_bw()
```


```{r longevity plot}
# not that relevant anymore

summ_longevity <- data_longevity %>%
  mutate(surv = jdate.surv - jdate.eclose) %>%
  group_by(expt.type, pop, #sex, 
           minT) %>%
  summarise(n = n(),
            avg.surv = mean(na.omit(surv)),
            se.surv = sd(na.omit(surv))/sqrt(length(na.omit(surv))))

summ_longevity %>%
  filter(expt.type == "expt") %>%
  #filter(!(pop == "lab" & diet.type == "TB")) %>% 
  ggplot(aes(y = avg.surv, x = as.factor(minT), 
             #group = interaction(pop, sex),
             group = pop)) +
  geom_point(size = 2) +
  geom_line(aes(lty = pop)) +
  geom_errorbar(aes(ymin = avg.surv - se.surv, ymax = avg.surv + se.surv), width = 0.25) +
  geom_text(aes(label = n), hjust = -1, vjust = 0.75) +
  theme_bw() 
  #facet_wrap(~sex)
```


```{r}
# tents math from 2023 for ref... 

summ_hatch <- data_tstats %>%
  group_by(#id.tent,
           trt.f, trt.m, pop,
           #trt.win, trt.btwn, 
           trt.type,
           trt.pair
           ) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.collected = sum(n.coll, na.rm = TRUE),
            se.coll =  sd(na.omit(n.coll))/sqrt(length(na.omit(n.coll))),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            se.hatched = sd(na.omit(n.tothatch))/sqrt(length(na.omit(n.tothatch))),
            n.females = sum(n.new.f, na.rm = TRUE),
            #start = first(jdate), end = last(jdate),
            t.duration = max(jdate) - min(jdate),
            rate.hatch = n.hatched/n.collected,
            se.hatch = sqrt(rate.hatch*(1-rate.hatch)/n.collected),  # check ???
            rate.hatchperf = rate.hatch/n.females,
            se.hatch = sqrt(rate.hatchperf*(1-rate.hatchperf/n.collected)),
            rate.f = round(n.collected/n.females, digits = 1)
            ) %>%
  drop_na(trt.m) %>% # drop tents w/o males
  mutate(trt.win = case_when(trt.f == trt.m & trt.f == 260 ~ "26-26",
                             trt.f == trt.m & trt.f == 419 ~ "40-19",
                             trt.f == trt.m & trt.f == 426 ~ "40-26",
                             trt.f == trt.m & trt.f == 433 ~ "40-33"),
         trt.btwn = case_when(trt.type == "ctrl" ~ "26-26", 
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 419 | trt.m == 419) ~ "40-19",
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 426 | trt.m == 426) ~ "40-26",
                              (trt.type == "hs F" | trt.type == "hs M") & (trt.f == 433 | trt.m == 433) ~ "40-33"),
         # min.f = case_when(trt.f == 260 | trt.f == 426 ~ 26,
         #                   trt.f == 419 ~ 19,
         #                   trt.f == 433 ~ 33), 
         # min.m = case_when(trt.m == 260 | trt.m == 426 ~ 26,
         #                   trt.m == 419 ~ 19,
         #                   trt.m == 433 ~ 33),
         trt.parent = case_when(trt.f == 260 & trt.m == 260 ~ 260,
                                trt.f == 419 | trt.m == 419 ~ 419,
                                trt.f == 426 | trt.m == 426 ~ 426,
                                trt.f == 433 | trt.m == 433 ~ 433),
         # trt.combo = paste0(trt.f, "-", trt.m),
         # min.combo = paste0(min.f, "-", min.m)
         )
```

