---
title: "feas figs and stats"
date: "2025-04-27"
---

goal: to have this all in one place (bc i also forgot to do some lol)

(these are just the aim 1 results)

# setup

## - loading
```{r}
here::i_am("2024/03-viz/feasibility.Rmd")
library(here)

library(tidyverse)
library(patchwork)
library(lme4) # for glms and lms
library(MuMIn) # for model dredging

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "compare-dev_p.R"))
source(here::here(bin_paths24$wrangle, "compare-tents_p.R"))
source(here::here(bin_paths24$doviz, "aesthetics_util.R"))

out_path <- here(bin_paths24$y24, "figs", "feasibility")
```

## - dev stats df prep
```{r general exptal bug filter}
# focusing on exptal bugs only that weren't lab on TB
ntw_expt <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab") & trt.type == "expt") %>%
  mutate(dmass = mass.pupa - mass.eclose,
         rate.pup = mass.pupa/tt.pupa)
```

```{r other subsets}
# omit accidental deaths (so either pmd, pup, LPI)
ntw_expt.surv <- ntw_expt %>%
  filter(is.pup != 2)

# omit unsexed pupa
ntw_expt.sex <- ntw_expt %>%
  filter(!is.na(sex))
```

```{r mumin subsets}
# bc apparently this pkg cant handle NAs bc "all models must be fit to the same data" 
# when comparing fits, so subset:
  # all surv to pups
  # all surv to ecs

# select relevant variables
ntw_mumin.dfs <- list(df = select(ntw_expt.surv, c("year", "pop", "minT", "is.pup", 
                                                   "is.sup", "sup", "is.ec")))

ntw_mumin.dfs <- c(ntw_mumin.dfs, 
                   list(pupd = filter(ntw_mumin.dfs$df, is.pup == 1)),
                   list(ecd = filter(ntw_mumin.dfs$df, is.ec == 1)))
```

## - fert df prep

### - for figs

```{r}
se <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
```


```{r}
# for figs
# sums are for across whole existence of cage/tent (`t.duration`)
  # pre-emptively filtering out the field pop bc they never wanna lay eggs
ss_fert <- dall %>%
  group_by(year, pop, trt.temp,
           trt.type, trt.sex) %>%
  summarise(n.reps = n_distinct(id),
            n.f = sum(n.new.f, na.rm=TRUE),
            n.colled = sum(n.coll, na.rm=TRUE),
            #avg.colled = mean(n.coll, na.rm=TRUE),
            se.colled = se(n.coll),
            avg.colledperf = n.colled/n.f,
            #test = n.coll/n.ovi, 
            #se.colledperf = sd(na.omit(n.coll)/na.omit(n.ovi))/sqrt(length(na.omit(n.coll)/na.omit(n.ovi))),
            n.hatched = sum(n.hatch, na.rm=TRUE),
            #avg.hatched = mean(n.hatch, na.rm=TRUE),
            se.hatched = se(n.hatch),
            p.hatched = n.hatched/n.colled,
            se.phatched = sqrt(p.hatched*(1-p.hatched)/n.colled),
            t.duration = max(jdate) - min(jdate)) %>% 
  mutate(across(ends_with("hatched"), ~ case_when(is.nan(.) ~ NA_real_,
                                                  year == 2023 ~ as.numeric(.))),
         across(starts_with(c("n", "avg", "se", "p")), ~ round(., digits = 3)),
         minT = case_when(trt.temp == 419 ~ 19,
                          trt.temp %in% c(260, 426) ~ 26,
                          TRUE ~ 33),
         minT = factor(minT, levels = c(19, 26, 33)),
         for.comparef = case_when(year == 2024 |
                                    (year == 2023 & (trt.sex == "f" | trt.temp == 260)) ~ "Y")) %>%
  filter(pop == "lab")
```

### - for stats

```{r setting up blank lists for outputs}
dfs <- list() 
mods <- list()
```

## - df prep

```{r general df modifications}
# add binaries for modeling
dall <- dall %>%
  mutate(is.hsF = case_when(trt.sex %in% c("f", "both") ~ 1,
                            TRUE ~ 0),
         is.hsM = case_when(trt.sex %in% c("m", "both") ~ 1,
                            TRUE ~ 0))
```

```{r egg laying}
# ignore the hatching for now
dfs <- list_assign(dfs, 
                   df23 = dall %>%
                     filter(year == 2023) %>%
                     group_by(id) %>%
                     mutate(n.coll.overall = sum(n.coll, na.rm = TRUE),
                            n.f.overall = sum(n.new.f, na.rm=TRUE),
                            n.collperovi.overall = n.coll.overall/n.f.overall,
                            n.hatched.overall = sum(n.hatch, na.rm =TRUE),
                            p.hatched = n.hatched.overall/n.coll.overall,
                            jdate.rel = jdate - first(jdate)
                     ) %>%
                     group_by(jdate.rel, .add = TRUE) %>%
                     mutate(n.coll.daily = sum(n.coll, na.rm = TRUE),
                            #n.hatch.daily = sum(n.hatch, na.rm = TRUE)
                            n.collperovi.daily = n.coll.daily/n.ovi)
)

#View(dfs[["df23"]])
```


# [x] 01. pup survival

## [x] fig

```{r}
p1 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "expt") %>% 
  ggplot(aes(y = prop.survpup, x = minT,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  geom_text(aes(label = n.dev), hjust = -0.25, vjust = -0.5) +
  labs(x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = prop.survpup, x = year,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  geom_text(aes(label = n.dev), hjust = -0.25, vjust = -0.5) +
  facet_wrap(~trt.type) +
  theme_bw()

p2
```

```{r}
# merge
p1 + p2 +
  #labs(caption = "do lower NTs increase survival to pup?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       y = "proportion survived to pupation") &
  guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16)) &
  scale_color_manual(values = "#1B9E77") &
  ylim(c(0, 1))
```


## [x] stats

### - dredge
```{r}
# remove nas
ntw_mumin.dfs <- c(ntw_mumin.dfs, 
                   list(nona = na.omit(ntw_mumin.dfs$df[1:4])))

out.psurv <- list(dredge = dredge(glm(is.pup ~ year*pop*minT, 
                                      data = ntw_mumin.dfs$nona, family = "binomial", na.action = na.fail), 
                                  #extra = c("R^2", F = function(x)
                                  #summary(x)$fstatistic[[1]])
                                  ))
out.psurv$dredge

dredge(glm(is.pup ~ year*pop*minT, data = ntw_mumin.dfs$nona, family = "binomial", na.action = na.fail)) %>%
  subset(., weight > 0.01) %>%
  as.data.frame(.) %>% write.csv(., here(bin_paths24$y24, "figs", "feasibility", "surv_dredge.csv"))

dredge(glm(is.pup ~ year*pop*minT, 
           data = ntw_mumin.dfs$nona, family = "binomial", na.action = na.fail)) %>% subset(., weight > 0.01)

out.psurv <- c(out.psurv,
               dredge.out = list(get.models(out.psurv$dredge, subset = weight > 0.01)))

out.psurv$dredge.out # subset of models based on weight
```

### - manual selection
```{r}
mod.psurv <- list(fullint = glm(is.pup ~ year*pop*minT, 
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  n2way = glm(is.pup ~ (year + pop + minT)^2,
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  fulladd = glm(is.pup ~ year + pop + minT, 
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  null = glm(is.pup ~ 1, 
                                data = ntw_mumin.dfs$df, family = "binomial"))

model.sel(mod.psurv$fullint, mod.psurv$n2way, mod.psurv$fulladd, mod.psurv$null) %>%
  as.data.frame(.) %>%
  write.csv(., here(bin_paths24$y24, "figs", "feasibility", "surv_man-modsel.csv"))

nested(model.sel(mod.psurv$fullint, mod.psurv$n2way, mod.psurv$fulladd, mod.psurv$null))
```



# [ ] 02. supernumerary stuff (later)

## [ ] fig

## [ ] stats


# [x] 03. pup mass/adult mass

## [x] fig

```{r}
# NTs on pupal/adult mass (colored by stage)

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  #filter(!(stage == "ad" & sex == "all")) %>%
  filter(sex != "all") %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = stage, fill = stage, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = stage), size = 2) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(stage
             ~year, scales = "free_y") +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  #filter(!(stage == "ad" & sex == "all")) %>%
  filter(sex != "all") %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.mass, x = year,
             color = stage, fill = stage, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(#aes(alpha = year), 
             size = 2) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_grid(stage
               ~trt.type, scales = "free_y") +
  theme_bw()

p2

# merge
p1 + p2 +
  #labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#D95F02", "#7570B3", "#1B9E77"), labels = c("pupa", "adult", "larva")) &
  scale_fill_manual(values = c("#D95F02", "#7570B3", "#1B9E77"), labels = c("pupa", "adult", "larva")) &
  #scale_alpha_manual(values = c(0.1,1)) &
  #ylim(c(0.75, 7.25)) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())

```

```{r}
# NTs on pupal/adult mass (colored by pop)

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  #filter(!(stage == "ad" & sex == "all")) %>%
  filter(sex != "all") %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = pop, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = pop), size = 2) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  facet_grid(stage
             ~year, scales = "free_y") +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  #filter(!(stage == "ad" & sex == "all")) %>%
  filter(sex != "all") %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.mass, x = year,
             color = pop, 
             shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(#aes(alpha = year), 
             size = 2) +
  labs(x = "year") +
  facet_grid(stage
               ~trt.type, scales = "free_y") +
  theme_bw()

p2

# merge
p1 + p2 +
  #labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       color = "population",
       y = "adult | pupal mass (g)") &
  guides(fill = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(17, 15)) &
  scale_color_manual(values = c("#41b6c4", "#225ea8"), labels = c("field", "lab")) &
  scale_fill_manual(values = c("#41b6c4", "#225ea8")) &
  #ylim(c(0.75, 7.25)) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())

```



## [x] stats

```{r}
# pup mass
mod.pmass_n2way01 <- lm(mass.pupa ~ (year + minT + pop + sex)^2, data = ntw_expt.sex, na.action = na.omit)
anova(mod.pmass_n2way01)

anova(mod.pmass_n2way01) %>%
  write.csv(., here(out_path, "pup-mass_anova.csv"))
```

```{r}
# adult mass
anova(lm(mass.eclose ~ (year*minT*pop*sex), data = ntw_expt.sex, na.action = na.omit))
mod <- anova(lm(mass.eclose ~ (year + minT + pop + sex)^2, data = ntw_expt.sex, na.action = na.omit)) # better
anova(lm(mass.eclose ~ year+minT+pop+sex, data = ntw_expt.sex, na.action = na.omit))

mod
mod %>% 
  write.csv(., here(out_path, "ad-mass_anova.csv"))
```


# [x] 04. pup time/ad longevity

## [x] fig

```{r}
# colored by stg
p1 <- ss_ntw %>% 
  filter(stage %in% c("la", "ad") & trt.type == "expt" & !(sex %in% c("m", "f"))) %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = stage, 
             group = interaction(pop, stage, trt.type)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = stage), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  #facet_wrap(~year) + # orig
  facet_grid(stage~year, scales = "free_y") +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  filter(stage %in% c("la", "ad") & trt.type == "ctrl" & !(sex %in% c("m", "f"))) %>%
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.tt, x = year,
             color = stage, 
             group = interaction(pop, stage, trt.type)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  #geom_point(aes(alpha = year), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  geom_point(#fill = NA, 
             size = 2) +
  labs(x = "year") +
  #facet_wrap(~trt.type) + # orig
  facet_grid(stage~trt.type, scales = "free_y") +
  theme_bw()

p2

# merge
p1 + p2 +
  #labs(caption = "do lower NTs reduce pup dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "time to pupation / adult longevity (days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  #scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#1B9E77",  "#7570B3", "#D95F02"), labels = c("larva", "adult", "pupa")) &
  scale_fill_manual(values = c("#1B9E77",  "#7570B3", "#D95F02"), labels = c("larva", "adult", "pupa")) &
  #ylim(c(15, 37)) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())

```

```{r}
# colored by pop
p1 <- ss_ntw %>% 
  filter(stage %in% c("la", "ad") & trt.type == "expt" & !(sex %in% c("m", "f"))) %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = pop, 
             group = interaction(pop, stage, trt.type)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(#alpha = year, 
                 color = pop), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  #geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (°C)") +
  #facet_wrap(~year) + # orig
  facet_grid(stage~year, scales = "free_y") +
  theme_bw()

p1


p2 <- ss_ntw %>% 
  filter(stage %in% c("la", "ad") & trt.type == "ctrl" & !(sex %in% c("m", "f"))) %>%
  mutate(trt.type = "control (26/26 °C)") %>%
  ggplot(aes(y = avg.tt, x = year,
             color = pop, 
             group = interaction(pop, stage, trt.type)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  #geom_point(aes(alpha = year), size = 2) +
  geom_text(aes(label = n), hjust = -0.25, vjust = -0.5, show.legend = FALSE) +
  geom_point(#fill = NA, 
             size = 2) +
  labs(x = "year") +
  #facet_wrap(~trt.type) + # orig
  facet_grid(stage~trt.type, scales = "free_y") +
  theme_bw()

p2

# merge
p1 + p2 +
  #labs(caption = "do lower NTs reduce pup dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population", color = "population",
       y = "adult longevity | time to pupation (days)") &
  guides(fill = "none") &
  scale_linetype_manual(values = c("dashed", "solid"), labels = c("field", "lab")) &
  #scale_shape_manual(values = c(19)) &
  scale_color_manual(values = c("#41b6c4", "#225ea8"), labels = c("field", "lab")) &
  scale_fill_manual(values = c("#41b6c4", "#225ea8")) &
  #ylim(c(15, 37)) &
  theme(strip.background.y = element_blank(), strip.text.y = element_blank())

```



## [x] stats

```{r}
# time to pup
anova(lm(tt.pupa ~ (year + minT + pop + sex)^2, data = ntw_expt, na.action = na.omit)) # sex is ns, so focus on 2way

mod <- anova(lm(tt.pupa ~ (year + minT + pop)^2, data = ntw_expt, na.action = na.omit))
mod

mod %>%
  write.csv(., here(out_path, "pup-time_anova.csv"))
```

```{r}
# ad long
anova(lm(tt.surv ~ (year + minT + pop + sex)^2, data = ntw_expt.sex, na.action = na.omit)) # sex is ns so drop

mod <- anova(lm(tt.surv ~ (year + minT + pop)^2, data = ntw_expt.sex, na.action = na.omit))

mod

mod %>%
  write.csv(., here(out_path, "ad-longevity_anova.csv"))
```



# [ ] 05. fertility/fecundity


## [x] fig

```{r}
ss_fert %>%
  filter(year == 2023 #& trt.temp != 260
         ) %>%
  mutate(trt.temp = as.factor(trt.temp),
         trt.type = factor(trt.type, levels = c("win", "btwn"))) %>%
  ggplot(aes(y = p.hatched, x = avg.colledperf,
             col = trt.temp, shape = trt.sex)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = p.hatched + se.phatched, ymin = p.hatched - se.phatched, width = 300*0.04)) +
  geom_errorbarh(aes(xmax = avg.colledperf + se.colled, xmin = avg.colledperf - se.colled, height = 0.2*0.025)) +
  geom_text(aes(label = n.f), vjust = -0.7, hjust = -0.65,
            show.legend = FALSE) +
  facet_grid(~trt.type, labeller = as_labeller(c(`win` = "both sexes HS'd",
                                                 `btwn` = "1 sex HS'd"))) +
  labs(#caption = "2023 hatch proportion/egg laying;\nn = total females, se = se total collected",
       y = "hatch proportion (overall)",
       x = "mean eggs collected per female (overall)",
       shape = "heat-treated sex",
       color = "larval rearing\ntemperature (°C)") +
  scale_color_manual(#values = bin_aes$temp4$col, 
                     values = c("#4393ce", "darkgoldenrod1", "#fd8d3c", "#d94701"),
                     labels = bin_aes$temp4$labswC) +
  scale_shape_manual(#values = bin_aes$sexes$ptfill_all, 
                     values = c(4, 16, 1),
                     #limits = c("f", "m")
                     ) +
  theme_bw()
```


## [x] fertility stats

```{r warning=FALSE}
mods <- list_assign(mods, 
                    y23.overall = glm(n.collperovi.overall ~ trt.temp*is.hsF*is.hsM, 
                                       data = dfs$df23, family = "poisson")) 

mod <- anova(mods$y23.overall, test = "Chisq")
mod

mod %>%
  write.csv(here(out_path, "fecundity_anova.csv"))
```


## [ ] fecundity stats

```{r}
# logit transform response
dfs <- list_assign(dfs, 
                   hatchstats = dfs$df23 %>%
                     mutate(z.logit = log(p.hatched/(1-p.hatched)), # dont use lol
                            z.mod = p.hatched + 0.005,
                            z.logit2 = log((z.mod/(1-z.mod))))
)
```

```{r}
mod <- anova(lm(z.logit2 ~ as.factor(trt.temp)*is.hsF*is.hsM, 
   data = dfs$hatchstats, na.action = na.omit))
mod

mod %>%
  write.csv(here(out_path, "fitness.csv"))

```



