---
title: "2024 ntw dev analyses"
date: "2024-09-23"
---

# roadmap & loading

= (pulling a lot of of stuff from the 2023 analyses lol)

## - questions

- [ ] how do 2024 ctrls compare to 2023 ctrls?


```{r message = FALSE}
source("./helpers_ntw.R") # need run cleaning.Rmd beforehand if gsheet has been updated

library(ggrepel)

# INCASE: bring this back if u need it and are breaking things
  # library(tidyverse)
  # conflicted::conflicts_prefer(dplyr::filter)
  # all_wide <- read.csv("~/Documents/repos/ntw/2024/data/ntw.csv", header = TRUE)
```



# 1. generic dev analyses (dev trajectories, overall outcomes)

- drop pmds in the #s but incl injured bugs i think.... though omit injured from barplot math

## - dev outcomes barplot (sup, pmd)

table (for sanity check)

```{r}
outcomes_table <- wide_all %>%
  filter(!is.na(surv.outcome)) %>% # excludes things that are still developing
  mutate(diet = case_when(diet == "LD" ~ "lab diet",
                          diet == "TB" ~ "tobacco diet"),
         diet = factor(diet, levels = c("tobacco diet", "lab diet"))
         ) %>%
  group_by(pop, diet, labs.trt) %>%
  summarise(n = n(),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.misc = sum(na.omit(surv.outcome == 2)),
            n.surv = n - n.pmd - n.misc,
            pct.pmd = round(n.pmd/(n-n.misc)*100, digits = 1),
            n.sup = sum(na.omit(sup > 0)),
            pct.sup = round(n.sup/n*100, digits = 1),
            pct.6th = round(sum(na.omit(sup == 6))/n.sup*100, digits = 1),
            pct.7th = round(sum(na.omit(sup == 7))/n.sup*100, digits = 1),
            pct.8th = round(sum(na.omit(sup == 8))/n.sup*100, digits = 1)
            )

outcomes_table
```

barplot

```{r}
# data formatting and filtering
outcomes <- wide_all %>%
  mutate(sup = case_when(surv.outcome == 1 ~ "pmd",
                         (surv.outcome == 0 & (is.na(sup) | sup == 0)) ~ "5",
                         surv.outcome == 0 & sup > 5  ~ as.character(sup)),
         sup = factor(sup, levels = c("pmd", "5", "6", "7", "8")),
         diet = case_when(diet == "LD" ~ "lab diet",
                          diet == "TB" ~ "tobacco diet"),
         diet = factor(diet, levels = c("tobacco diet", "lab diet"))
         #trt = factor(sup)
         ) %>%
  filter(surv.outcome == 0 | surv.outcome == 1) # excludes things w/ alt deaths/bugs still developing
  #filter(!is.na(sup) | surv.outcome == 2 | !is.na(surv.outcome)) 
```

```{r}  
# plots
  # needs outcomes_table from prev prev section


## all bugs
outcomes %>%
  ggplot(aes(x = labs.trt, fill = sup)) +
  geom_bar(stat = "count", 
           #position = "stack", 
           position = position_fill(reverse = TRUE)
           ) +
  facet_wrap(pop ~ diet) +
  labs(title = "developmental fates",
       x = "treatment", y = "proportion",
       fill = "# larval instars/\ndeath") +
  geom_text(data = outcomes_table, aes(x = labs.trt, y = 0.05, label = n), inherit.aes = FALSE) +
  scale_fill_brewer(palette = "Blues") +
  theme_bw() + guides(x = guide_axis(angle = 45))

## only looking at alive ones
outcomes %>%
  filter(sup != "pmd") %>% #View()
  #filter(!(pop == "lab" & diet == "tobacco diet")) %>%
  ggplot(aes(x = labs.trt, fill = sup)) +
  geom_bar(stat = "count", 
           #position = "stack", 
           position = position_fill(reverse = TRUE)
           ) +
  facet_wrap(pop~diet) +
  labs(title = "2024 developmental fates",
       x = "treatment", y = "proportion",
       fill = "# larval instars") +
  geom_text(data = outcomes_table, aes(x = labs.trt, y = 0.05, label = n.surv), inherit.aes = FALSE) + 
  #scale_fill_brewer(palette = "Blues") +
  scale_fill_manual(values = c("#BDD7E7", "#6BAED6", "#2171B5")) + # keep consistent w prev
  theme_bw() + guides(x = guide_axis(angle = 45))
```

- todo  
  - 2024-10-02: 
    - maybe flip the order so dev stuff is on the bottom? (or else just look at only counts for suriving things)
    
    
## - (wip) daily dev

# 2. main development stats (big 3)

## - df prep

= look at avg pupal mass and avg time to pup (see if need to filter out diet stuff later)

df prep

```{r}
# remove pmd
dev_nopmd <- wide_all %>%
  filter(surv.outcome != 1 | is.na(surv.outcome)) # incl in progress stuff
```

calculations

```{r}
dev_summary <- dev_nopmd %>%
  group_by(pop, diet, labs.trt, 
           trt.type, minT, #sex #revisit this when theres more LOL
           ) %>%
  summarise(n = n(), # todo: 2024-10-08 this doesnt look right tbh LOL... maybe bc of the in-progress ones?
            avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))))

surv_summary <- wide_all %>%
  filter(surv.outcome != 2) %>% # omit accidental deaths
  group_by(pop, diet, labs.trt, 
           trt.type, minT) %>%
  summarise(n = n(),
            n.pmd = sum(surv.outcome == 1), 
            prop.survpup = round(1-(n.pmd/n), digits=2))
```

## - plots (all data)

- todo
  - 2024-10-03: `geom_text_repel` is temp until numbers are finalised

```{r}
# survival
surv_summary %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  #geom_text(aes(label = n), vjust = 1, hjust = -0.5) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "survival to pupation",
       y = "proportion survived to pupation", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type") +
  guides(shape = "none")

# pupal mass
dev_summary %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass), width = 0.25) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "pupal mass",
       y = "avg mass at pupation (mg)", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type")+
  guides(shape = "none")

# time to pup
dev_summary %>%
  ggplot(aes(y = avg.tt, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt), width = 0.25) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "time to pupation", 
       y = "avg time to pupation (days)", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type") +
  guides(shape = "none")

```

- notes 2024-10-03:
  - looks super wonky bc not enough things have gotten to pup still LMFAO
  - def need to exclude the lab bug diet type variation.... (see below)
  

## - plots (focused)

= per mtg recommendations, focus on lab bugs on LD vs field bugs on TB; omit the 26 constant trt for clarity


filtering

```{r}
# needs outputs of 'df prep'

surv_summf <- surv_summary %>%
  filter(!(pop == "lab" & diet == "TB") & trt.type == "expt")

dev_summf <- dev_summary %>%
  filter(!(pop == "lab" & diet == "TB") & trt.type == "expt")
```



plotting

```{r}
# survival
surv_summf %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_line(aes(lty = pop, group = pop)) +
  #geom_text(aes(label = n), vjust = 1, hjust = -0.5) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "survival to pupation",
       y = "proportion survived to pupation", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population") +
  ylim(0, 0.38)

# pupal mass
dev_summf %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass), width = 0.15) +
  geom_line(aes(lty = pop, group = pop)) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "pupal mass",
       y = "avg mass at pupation (mg)", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population")

# time to pup
dev_summf %>%
  ggplot(aes(y = avg.tt, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt), width = 0.15) +
  geom_line(aes(lty = pop, group = pop)) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "time to pupation", 
       y = "avg time to pupation (days)", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population")

```




# 3. survival

- this will all be for fun lol

