---
title: "2024 ntw dev analyses"
date: "2024-09-23"
---

# roadmap & loading

= (pulling a lot of of stuff from the 2023 analyses lol)

2025-02-20: mostly superceded by the 23v24 analyses now!!!

## - questions

- [ ] how do 2024 ctrls compare to 2023 ctrls?


```{r message = FALSE}
here::i_am("2024/03-viz/ntw.Rmd")
library(here)

library(tidyverse)
library(ggrepel)

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "ntw.R"))

# INCASE: bring this back if u need it and are breaking things
  # library(tidyverse)
  # conflicted::conflicts_prefer(dplyr::filter)
  # all_wide <- read.csv("~/Documents/repos/ntw/2024/data/ntw.csv", header = TRUE)
```



# 1. generic dev analyses (dev trajectories, overall outcomes)

- drop pmds in the #s but incl injured bugs i think.... though omit injured from barplot math

## - dev outcomes barplot (sup, pmd)



barplot


```{r}  
# plots
  # needs outcomes_table 


## all bugs
outcomes %>%
  ggplot(aes(x = labs.trt, fill = sup)) +
  geom_bar(stat = "count", 
           #position = "stack", 
           position = position_fill(reverse = TRUE)
           ) +
  facet_wrap(pop ~ diet) +
  labs(title = "developmental fates",
       x = "treatment", y = "proportion",
       fill = "# larval instars/\ndeath") +
  geom_text(data = outcomes_table, aes(x = labs.trt, y = 0.05, label = n), inherit.aes = FALSE) +
  scale_fill_brewer(palette = "Blues") +
  theme_bw() + guides(x = guide_axis(angle = 45))

## only looking at alive ones
outcomes %>%
  filter(sup != "pmd") %>% #View()
  #filter(!(pop == "lab" & diet == "tobacco diet")) %>%
  ggplot(aes(x = labs.trt, fill = sup)) +
  geom_bar(stat = "count", 
           #position = "stack", 
           position = position_fill(reverse = TRUE)
           ) +
  facet_wrap(pop~diet) +
  labs(title = "2024 developmental fates",
       x = "treatment", y = "proportion",
       fill = "# larval instars") +
  geom_text(data = outcomes_table, aes(x = labs.trt, y = 0.05, label = n.surv), inherit.aes = FALSE) + 
  #scale_fill_brewer(palette = "Blues") +
  scale_fill_manual(values = c("#BDD7E7", "#6BAED6", "#2171B5")) + # keep consistent w prev
  theme_bw() + guides(x = guide_axis(angle = 45))
```

- todo  
  - 2024-10-02: 
    - maybe flip the order so dev stuff is on the bottom? (or else just look at only counts for suriving things)
    
    
## - (wip) daily dev

# 2. main development stats (big 3)

= look at avg pupal mass and avg time to pup (see if need to filter out diet stuff later)

## - plots (all data)

- todo
  - 2024-10-03: `geom_text_repel` is temp until numbers are finalised

```{r}
# survival
surv_summary %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  #geom_text(aes(label = n), vjust = 1, hjust = -0.5) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "survival to pupation",
       y = "proportion survived to pupation", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type") +
  guides(shape = "none")

# pupal mass
dev_summary %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass), width = 0.25) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "pupal mass",
       y = "avg mass at pupation (mg)", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type")+
  guides(shape = "none")

# time to pup
dev_summary %>%
  ggplot(aes(y = avg.tt, x = as.factor(minT), color = trt.type, shape = diet)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt), width = 0.25) +
  geom_line(aes(lty = diet, group = interaction(diet, trt.type))) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  facet_wrap(~pop) +
  theme_bw() +
  labs(title = "time to pupation", 
       y = "avg time to pupation (days)", x = "minimum temperature (°C)",
       shape = "treatment group", color = "diet type") +
  guides(shape = "none")

```

- notes 2024-10-03:
  - looks super wonky bc not enough things have gotten to pup still LMFAO
  - def need to exclude the lab bug diet type variation.... (see below)
  

## - plots (focused)

= per mtg recommendations, focus on lab bugs on LD vs field bugs on TB; omit the 26 constant trt for clarity


```{r}
# survival
surv_summf %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_line(aes(lty = pop, group = pop)) +
  #geom_text(aes(label = n), vjust = 1, hjust = -0.5) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "survival to pupation",
       y = "proportion survived to pupation", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population") +
  ylim(0, 0.38)

# pupal mass
dev_summf %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass), width = 0.15) +
  geom_line(aes(lty = pop, group = pop)) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "pupal mass",
       y = "avg mass at pupation (mg)", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population")

# time to pup
dev_summf %>%
  ggplot(aes(y = avg.tt, x = as.factor(minT), color = pop, shape = pop)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt), width = 0.15) +
  geom_line(aes(lty = pop, group = pop)) +
  geom_text_repel(aes(label = n), show.legend = FALSE) +
  theme_bw() +
  labs(title = "time to pupation", 
       y = "avg time to pupation (days)", x = "minimum temperature (°C)",
       shape = "population", color = "population", lty = "population")

```




# 3. survival

- this will all be for fun lol

