---
title: "2025 ntw tidying"
date: "2025-11-26"
---

# # preamble

prepping ntw data for 2025 wrangling

final code is purled to a `.R` script, rerun `purl` step as needed to update the wrangling for analyses


```{r set paths, eval=FALSE, include=FALSE, purl=FALSE}
# run this prior to other pathing stuff if things are breaking
here::i_am("2025/ntw/data-tidying.Rmd")
```

```{r for purling, eval=FALSE, include=FALSE, purl=FALSE}
# rerun this chunk if purled .R script needs to be updated.
# save and apply updates to current .Rmd before purling!

knitr::purl("data-tidying.Rmd", "tidy-data.R")

# creates an archive
knitr::purl("data-tidying.Rmd", 
            paste0(here(bin_paths$y25$ntw), "/archive/cleaning/", 
                   format(Sys.time(), "%y%m%d-%H%M"), "_tidy-data.R"))
```

# load libs & pkgs

```{r about}
# cleaning/ntw-2025.R

# knitted tidying code for 2025 ntw dev figs/analyses.
# source() this in corresponding analysis scripts.

# this script compiles ntw data from 2023-25 into cohesive units for easier wrangling.
```

```{r load libs}
library(here)
library(tidyverse)
library(data.table) # for rbindlist

source(here::here("set-paths.R"))

source(here::here(bin_paths$y25$ntw, "utils.R"))
```


# # larval data (all yrs)

## loading
```{r load larval data}
dfs_imported <- list(
  y23 = read.csv(here::here(bin_paths$y23$data, "clean-ntw.csv"), header = TRUE) %>%
    mutate(year = 2023) %>%
    mutate(ID = case_when(ID == 1379 & !is.na(jdate.pmd) ~ 1379.1,
                          TRUE ~ as.numeric(ID))),
  y24 = read.csv(here::here(bin_paths$y24$data, "ntw.csv"), header = TRUE) %>%
    mutate(year = 2024),
  y25 = read.csv(here::here(bin_paths$y25$ntwdata, "clean-larva.csv"), header = TRUE) %>%
    mutate(year = 2025)
)
```

## standardising

```{r filtering and column cleanup}
dfs_filtered <- list(
  
  y23 = dfs_imported[["y23"]] %>%
    filter(instar.enter == "hatch",
           pop %in% c("field", "lab"),
           !(treatment %in% c("ctrl", "267", "330")),
           reason.ignore != "lost" | is.na(reason.ignore),
           ) %>%
    rename(id = ID,
           jdate.LPI = jdate.LP,
           trt = treatment,
           cohort = expt.group) %>%
    select(-c(starts_with(c("date", "h.", "if.")), ends_with("2nd"), 
              "instar.enter",  "jdate.collected", "jdate.enter", 
              "jdate.stuck", "jdate.15", "jdate.died",
              "parent.tent", "src", "trt.stage", "grp.trt")),

  y24 = dfs_imported[["y24"]] %>%
    filter(instar.enter == "hatch") %>%
    select(-c("trt.type", "jdate.enter", "instar.enter", ends_with("2nd"))),
  
  y25 = dfs_imported[["y25"]] %>%
    filter(is.na(species),
           !(notes.ignore == "empty") | is.na(notes.ignore)) %>%
    rename(pop = population,
           trt = treatment) %>%
    select(-c("species", "jdate.enter", "instar.enter", ends_with("2nd")))
)

# stdise column types
dfs_stdised <- lapply(dfs_filtered, \(x){
  x %>%
  mutate(trt = as.numeric(trt),
         id = as.numeric(id),
         across(starts_with("jdate"), as.numeric))
})
```

```{r standardising values by yr}
# individual fixes: recoding + adding missing columns
# stdising date meanings:
  # cull = non-pmd larval death
  # surv = death after eclosion

# 2023: redo trt values, survival coding/notes, sup coding; add trt group
dfs_stdised[["y23"]] <- dfs_stdised[["y23"]] %>% 
  mutate(mass.8th = NA, jdate.8th = NA,
         trt = case_when(trt == 337 ~ 426,
                         TRUE ~ trt),
         trt.group = case_when(pop == "lab" & diet == "TB" ~ 1000,
                               pop == "lab" & diet == "LD" ~ 2000,
                               TRUE ~ 100),
         fate.dev = case_when(!is.na(jdate.eclose) ~ "ec",
                              !is.na(jdate.pupa) | !is.na(jdate.LPI) ~ "pup",
                              reason.ignore %in% c("wet diet", "accidental", "hot larva", "cut") | 
                                final.fate %in% c("accidental", "culled") ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         jdate.culled = case_when(final.fate %in% c("accidental", "culled") & fate.dev != "pup" ~ as.numeric(jdate.exit))
         ) %>%
  select(-c("final.fate", "reason.ignore"))

# 2024: redo survival coding, trt group
dfs_stdised[["y24"]] <- dfs_stdised[["y24"]] %>% 
  mutate(jdate.LPI = NA, jdate.exit = NA, 
         trt.group = case_when(id < 1000 ~ 100,
                               id > 2000 ~ 2000,
                               TRUE ~ 1000),
         fate.dev = case_when(!is.na(jdate.eclose) ~ "ec",
                              !is.na(jdate.pupa) ~ "pup",
                              !is.na(jdate.culled) | surv.outcome == 2  ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         jdate.culled = NA,
         jdate.surv = case_when(is.na(jdate.eclose) ~ NA_real_,
                                TRUE ~ as.numeric(jdate.surv))
         ) %>%
  select(-"surv.outcome")

# 2025: redo survival coding/notes; add tempts
dfs_stdised[["y25"]] <- dfs_stdised[["y25"]] %>% 
  mutate(fate.dev = case_when(!is.na(jdate.eclose) & !is.na(jdate.surv) ~ "ec",
                              !is.na(jdate.pupa) | !is.na(jdate.LPI) ~ "pup",
                              !is.na(jdate.culled) | !is.na(notes.ignore)  ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         jdate.culled = case_when(fate.dev == "other" ~ as.numeric(jdate.culled),
                                  FALSE ~ NA_real_),
         meanT = case_when(trt == 260 ~ 26,
                           trt == 419 ~ 29.5,
                           trt == 426 ~ 33,
                           trt == 433 ~ 36.5),
         flucT = case_when(trt == 260 ~ 0,
                           trt == 419 ~ 10.5,
                           trt == 426 ~ 7,
                           trt == 433 ~ 3.5),
         maxT = case_when(trt == 260 ~ 26,
                          TRUE ~ 40),
         minT = case_when(trt == 260 | trt == 426 ~ 26,
                          trt == 419 ~ 19,
                          trt == 433 ~ 33)) %>%
  filter(!is.na(status)) %>%
  select(-c(starts_with("notes"), "status"))
```

```{r standardisng values for all}
# redo survival, sup coding (counts after reaching 5th);
# fix exit dates
dfs_stdised <- lapply(dfs_stdised, \(x){
  x %>%
    mutate(fate.code = case_when(fate.dev %in% c("ec", "pup") ~ 1,
                                 fate.dev == "pmd" ~ 0,
                                 fate.dev == "other" ~ 2),
           sup = case_when(!is.na(jdate.8th) ~ 8,
                           !is.na(jdate.7th) ~ 7,
                           !is.na(jdate.6th) ~ 6,
                           #!is.na(jdate.5th) ~ 5,
                           #FALSE ~ as.numeric(sup)
                            ),
           jdate.exit = case_when(!is.na(jdate.surv) ~ as.numeric(jdate.surv),
                                  !is.na(jdate.eclose) ~ as.numeric(jdate.eclose),
                                  !is.na(jdate.LPI) ~ as.numeric(jdate.LPI),
                                  !is.na(jdate.pupa) & is.na(jdate.eclose) ~
                                    as.numeric(jdate.pupa + 36),
                                  !is.na(jdate.culled) ~ as.numeric(jdate.culled),
                                  !is.na(jdate.pmd) ~ as.numeric(jdate.pmd))
    )
}) #%>% data.table::rbindlist(., fill = TRUE) %>% View()

```

## prep for final export

```{r prep dev data for export}
# combine dfs and remove short-lived larvae
dfs <- list(
  wide = data.table::rbindlist(dfs_stdised, fill = TRUE) %>%
              filter(jdate.exit - jdate.hatch > 1 #| is.na(jdate.exit)
                     ) %>%
              select(-c("id.cage")) # cage not relevant for larval stuff
)


dfs <- list_modify(dfs,
                   long = dfs$wide %>% 
                     mutate(tt.3rd = jdate.3rd - jdate.hatch,
                            tt.4th = jdate.4th - jdate.hatch,
                            tt.5th = jdate.5th - jdate.hatch,
                            tt.6th = jdate.6th - jdate.hatch,
                            tt.7th = jdate.7th - jdate.hatch,
                            tt.8th = jdate.8th - jdate.hatch,
                            tt.wander = jdate.wander - jdate.hatch,
                            tt.pupa = jdate.pupa - jdate.hatch,
                            tt.eclose = jdate.eclose - jdate.pupa,
                            tt.surv = jdate.surv - jdate.eclose,
                            #tt.exit = jdate.exit - jdate.hatch
                            ) %>%
                     pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
                                  names_to = c(".value", "instar"),
                                  values_drop_na = TRUE,
                                  names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
                     
                     # drop empty dev times but keep masses
                     mutate(tt = case_when(is.na(tt) & !is.na(mass) ~ 999,
                                           TRUE ~ as.numeric(tt))) %>%
                     drop_na(tt) %>%
                     mutate(tt = na_if(tt, 999)) %>%
                     # remove short-lived eclosed bugs
                     filter(!(instar == "surv" & tt == 0))
) 
```

# # adult data

preamble:
- for each yr, have adult counts (when added/removed), egg collection counts, egg hatch counts.
- bc my method of tracking the daily egg/moth #s is very convoluted, need to split + wrangle the dfs in a certain order before finally collating into the final df that can be used in `summarise()` for data analysis
  - note that 2023 data is also formatted a little differently than 2025 data (i.e. pairing was more deliberate), so it needs some additional pre-wrangling
- see also some random TODOs left thruout this section,,,

roadmap:
1. create stdised `info` (about mating pairs) and `events` dfs
2. `events` gets created in the following manner to avoid incorrect duplication of jdates:
  a. `events` is first generated from `info` (to get counts of moths removed/added)
  b. later, `events` is overwritten by merged collection + hatching data, with
     moth counts extracted from `events` appended on
3. `events` then gets used to generate the daily running count of eggs/moths/etc
  a. counts are determined separately per cage and then recollated again
     (avoids wonkiness with `cumsum()` at the first/last values of the df)
4. 2023/2025 are handled separately, then combined for final export.


## # fns for info generation
```{r tent data common functions}
# to assign cage mating info
# note: data needs to be "wide" (...f, ...m) to determine mate types
.GenerateCageMateInfo <- function(x){
  x %>%
    mutate(mate.type = case_when(trt.f == trt.m ~ "within",
                                 trt.f != trt.m ~ "between",
                                 is.na(trt.m) ~ "virgin f"),
           # if both m+f are from col, then mate.pop = "col", otherwise "lab"
           mate.pop = case_when(mate.type == "within" | is.na(pop.m) ~ pop.f,
                                mate.type == "between" & !(pop.f == "field" | pop.m == "field") ~ "lab"), 
           mate.trt = case_when(mate.type == "within" | is.na(trt.m) ~ trt.f,
                                mate.type == "between" & trt.f > trt.m ~ trt.f,
                                mate.type == "between" & trt.m > trt.f ~ trt.m),
           mate.col = case_when(pop.f == "col" & pop.m == "col" ~ "both",
                                pop.f == "col" & pop.m == "lab" ~ "f",
                                pop.m == "col" & pop.f == "lab" ~ "m" ),
           mate.hs = case_when(trt.f > trt.m | is.na(trt.m) ~ "f",
                               trt.m > trt.f | is.na(trt.f) ~ "m",
                               TRUE ~ "both"))
} 


# to determine daily m/f counts for cage events
.GenerateCageMothCounts <- function(x){
  x %>%
    group_by(cage, jdate) %>%
    summarise(f.added = sum(sex == "f" & event == "added"),
              m.added = sum(sex == "m" & event == "added"),
              f.removed = sum(sex == "f" & event == "removed"),
              m.removed = sum(sex == "m" & event == "removed")) %>%
    drop_na(cage)
}

# appends cage mate info + daily moth events to daily egg events
# needs dfs_stdised$info and $events to exist alrdy
.AppendCageMothEvents <- function(x){
  x %>%
  merge(., dfs_stdised$events, all = TRUE) %>%
  merge(., dfs_stdised$info %>%
          select(c("cage", starts_with("mate"))) #%>% View()
        ,
        all = TRUE, by = "cage") %>%
  mutate(across(starts_with(c("eggs", "f.", "m.")), replace_na, 0))
}


# gets daily running counts of new/dead moths + eggs collected
# splits $events by cage to avoid cumsum wonkiness, then recombines

.GenerateCageDailyEvents <- function(x){
  x %>%
    split(f = dfs_stdised$events$cage) %>% 
    lapply(., \(x){
      x %>%
        arrange(jdate) %>%
        mutate(f.diff = lag(f.added, default = 0) - lag(f.removed, default = 0),
               f.curr = 0,
               f.curr = cumsum(f.diff + f.curr),
               m.diff = lag(m.added, default = 0) - lag(m.removed, default = 0),
               m.curr = 0,
               m.curr = cumsum(m.diff + m.curr),
               
               # fix egg counts: excludes eggs from hatch proportion if no males present
               ## TODO: just kidding not doing this anymore bc females can sequester sperm!
               eggs.exclude = case_when(#f.curr > 0 & m.curr == 0 | 
                                          mate.type == "virgin f" ~ 
                                          as.numeric(eggs.collected + eggs.exclude),
                                        TRUE ~ as.numeric(eggs.exclude)),
               eggs.collected = case_when(#f.curr > 0 & m.curr == 0 |
                                            mate.type == "virgin f" ~ 0,
                                          TRUE ~ as.numeric(eggs.collected)),
               eggs.coll = eggs.collected + eggs.exclude) %>%
        select(-ends_with(c(".removed", ".diff"))) %>%
        rename(eggs.fert = eggs.collected,
               eggs.unfert = eggs.exclude)
    }) %>%
    rbindlist()
}
```

## # 2025 data
### loading

```{r load 2025 tents data}
dfs_imported <- list(
  ad25 = read.csv(here::here(bin_paths$y25$ntwdata, "clean-adults.csv"), header = TRUE),
  coll25 = read.csv(here::here(bin_paths$y25$ntwdata, "clean-coll.csv"), header = TRUE),
  hatch25 = read.csv(here::here(bin_paths$y25$ntwdata, "clean-hatch.csv"), header = TRUE)
)
```

### stdise cage/egg events
```{r prep 2025 tent data for wrangling}
### need to rerun the WHOLE chunk at once if updating $info to avoid dups later on in $events

dfs_stdised <- list(
  
  # cage info (mating pairs)
  # TODO maybe could tidyr::complete this but i dont get the nesting lol
  info =
  merge(
    dfs_imported[["ad25"]] %>% filter(sex == "f") %>%
      select(cage, id, pop, trt, sex) %>%
      pivot_wider(names_from = "sex", values_from = c("pop", "trt"), 
                  names_sep = ".") %>% 
       select(-id) %>% distinct(),
    dfs_imported[["ad25"]] %>% filter(sex == "m") %>%
      select(cage, id, pop, trt, sex) %>%
      pivot_wider(names_from = "sex", values_from = c("pop", "trt"), 
                  names_sep = ".") %>% 
       select(-id) %>% distinct(),
    all = TRUE) %>% 
    drop_na(cage) %>% #View() #for events23 futzing
    .GenerateCageMateInfo() %>% #View()
    # some fixes
    filter(!(cage == "A8" & pop.m == "lab")) %>%
    mutate(mate.pop = case_when(cage %in% c("A6", "B16", "C7") ~ "lab",
                                TRUE ~ as.character(mate.pop))),
  
  # cage timelines (moth additions/removals)
  events = dfs_imported[["ad25"]] %>%
    pivot_longer(starts_with("jdate"),
                 names_to = c(".value", "event"),
                 names_sep = "\\.") %>% 
    drop_na("jdate") %>% #View()
    group_by(cage, jdate) %>%
    .GenerateCageMothCounts() #%>% View()
)


# cage timelines (append egg hatching + collection info)
dfs_stdised$events <- 
  dfs_imported[["coll25"]] %>%
  rename(jdate = jdate.collected) %>%
  merge(., dfs_imported[["hatch25"]] %>% 
          select(c("cage", ends_with(".hatched"))) %>%
          rename(jdate = jdate.hatched) %>%
          group_by(jdate, cage) %>%
          summarise(eggs.hatched = sum(eggs.hatched)) %>%
          drop_na(),
        all = TRUE) %>% #View()
  .AppendCageMothEvents()
```

### collate

TODO
- somehow account for def m/f's that prob = unviable eggs?

```{r}
# troubleshooting tents: A2, B2, C14, C4
dfs_stdised <- list_modify(
  dfs_stdised, 
  tents = 
    dfs_stdised$events[order(dfs_stdised$events$cage, dfs_stdised$events$jdate),] %>%
    .GenerateCageDailyEvents()
)
```

### export
```{r}
dfs_export <- list(
  )
```


## # 2023 data

### loading
```{r load 2023 tent data}
dfs_imported <- list(
  ad23 = 
    read.csv(here::here(bin_paths$y23$data, "clean-tentpairs.csv"), header = TRUE) %>%
    mutate(across(starts_with("date"), as.Date, "%Y-%m-%d")) %>%
    # minor fixes
    mutate(id.tent = case_when(id.tent == "107-cage" & is.na(trt.f) ~ NA_character_,
                               TRUE ~ as.character(id.tent))) %>%
    filter(!(id.f %in% c(282, 284))),
  
  coll23 = 
    read.csv(here::here(bin_paths$y23$data, "clean-tentstats.csv"), header = TRUE) %>%
    mutate(across(starts_with("date"), as.Date, "%Y-%m-%d")) %>% 
    filter(id.tent != "107-I"),
  
  hatch23 = 
    read.csv(here::here(bin_paths$y23$data, "clean-hatchstats.csv"), header = TRUE) %>% 
    mutate(across(starts_with("date"), as.Date, "%m/%d/%y"))) %>%
  
  lapply(., \(x){
    x %>%
      rename(cage = id.tent) %>%
      mutate(across(starts_with("date"), as.Date, "%j", .names = "j{.col}"),
             across(starts_with("jdate"), as.numeric)) %>%
      select(-starts_with("date"))
  })
```

```{r}
# filter out useful columns
dfs_filtered <- list(
  ad23 = dfs_imported[["ad23"]] %>%
    select(c("cage", "id.pair", starts_with("jdate"), ends_with(c(".m", ".f"))),
           -starts_with("track")),
  coll23 = dfs_imported[["coll23"]] %>%
    select(c("cage", "n.coll", "jdate")),
  hatch23 = dfs_imported[["hatch23"]] %>%
    select(c("cage", "n.hatch", "jdate.hatch")))
```

### stdise cage/egg events

#### reformat 2023 data

(match 2023 data to 2025 formatting)

```{r 2023 adult info fixes}
## run entire chunk to properly update dfs_stdised$ad23 (to build final tent events dfs)

# rename jdate cols, add uids to ctrls/cols
dfs_filtered$ad23 <- dfs_filtered$ad23 %>%
  rename_with(., 
              ~paste("jdate", str_extract(.,"[a-z]*$"), 
                     str_sub(., 7, 7), sep = "."), 
              starts_with(c("jdate.m", "jdate.f"))) %>%
  mutate(id.m = case_when(id.m %in% c("col", "ctrl") ~ 
                            paste0(id.m, "-m", sprintf("%03d", id.pair)),
                          #id.f == "146" & id.m == "253" ~ NA_character_,
                          TRUE ~ as.character(id.m)),
         id.f = case_when(id.f %in% c("col", "ctrl") ~ 
                            paste0(id.f, "-f", sprintf("%03s", id.pair)),
                          #id.f == "244" & id.m == "224" ~ NA_character_,
                          TRUE ~ as.character(id.f)))

# fixing dup'd individuals
dfs_stdised <- list(
  ad23 = merge(
    dfs_filtered$ad23 %>%
      select(c("cage", contains("pair"), ends_with(".m"))) %>%
      pivot_longer(matches("^(?!jdate).*m", perl = T),
                   names_to = c(".value", "sex"), names_sep = "\\.") %>% 
      rename_with(., ~ gsub(".m", "", .)),
    dfs_filtered$ad23 %>%
      select(c("cage", contains("pair"), ends_with(".f"))) %>%
      pivot_longer(matches("^(?!jdate).*f", perl = T),
                   names_to = c(".value", "sex"), names_sep = "\\.") %>% 
      rename_with(., ~ gsub(".f", "", .)),
    all = TRUE) %>% 
    drop_na(id))

# adding in other trt info
dfs_stdised$ad23 <- dfs_stdised$ad23 %>%
  mutate(ctrl = case_when(grepl("col", id) ~ "col",
                          grepl("ctrl", id) ~ "diurnal"), 
         pop = case_when(grepl("col", id) ~ "col",
                         !is.na(ctrl) | as.numeric(id) > 1000 ~ "lab",
                         as.numeric(id) < 1000 ~ "field"),
         if.fridge = case_when(jdate.paired != jdate.ec ~ "x"))

```

#### generate dfs
```{r prep 2023 tent data for wrangling}
## rerun the whole chunk to regenerate dfs_tents$events, $info
dfs_stdised <- list_modify(
  dfs_stdised,
  
  info = 
    dfs_stdised[["ad23"]] %>%
    pivot_wider(id_cols = c("cage", "id.pair", #"ctrl", "if.fridge"
                            ), 
                names_from = "sex", values_from = c("trt", "pop"),
                names_sep = ".") %>% #View()
    select(c("cage", starts_with(c("trt", "pop")))) %>%
    distinct() %>%
    .GenerateCageMateInfo() %>% #View() %>%
    # fixes (NOT solo f tents: 301P, 301U, 107H...)
    filter(!(mate.type == "virgin f" & !(cage %in% c("107-cage", "301-T"))),
           !(cage == "301-M" & pop.f == "lab")) ,
  
  events = 
    dfs_filtered[["ad23"]] %>%
    pivot_longer(ends_with(c(".f", ".m")),
                 names_to = c(".value", "sex"),
                 names_pattern = "([a-z]*.[a-z]*).(m|f)") %>% 
    drop_na(id) %>% 
    pivot_longer(starts_with("jdate"),
                 names_to = c(".value", "event"),
                 names_sep = "\\.") %>%
    mutate(event = case_when(event == "paired" ~ "added",
                             event == "died" ~ "removed")
    ) %>% 
    drop_na(jdate, event) %>% 
    .GenerateCageMothCounts()
)

# overwrite moth events (append egg info)
dfs_stdised$events <- 
  dfs_filtered[["coll23"]] %>%
  rename(eggs.collected = n.coll) %>%
  merge(., dfs_filtered[["hatch23"]] %>% 
          rename(jdate = "jdate.hatch") %>%
          group_by(cage, jdate) %>%
          summarise(eggs.hatched = sum(n.hatch)) %>%
          drop_na(),
        all = TRUE) %>% 
  .AppendCageMothEvents() %>%
  mutate(eggs.exclude = 0) # a column in 2025 but not 2023
```

### collate
```{r}
dfs_stdised <- list_modify(
  dfs_stdised,
  tents = 
    dfs_stdised$events[order(dfs_stdised$events$cage, dfs_stdised$events$jdate),] %>%
    .GenerateCageDailyEvents()
)
```

### export
```{r}
dfs_export <- list_modify(
  dfs_export,
  tents23 = dfs_stdised$tents %>% 
    merge(.,
          dfs_stdised$info %>% select(c("cage", starts_with("trt"))),
          all = TRUE) %>%
    mutate(year = 2023)
)
```

## prep for final export

```{r export tent data}
dfs <- list_modify(
  dfs,
  tents = rbindlist(dfs_export, fill = TRUE)
)
```




# # cleanup
```{r remove intermediate objects}
rm(dfs_imported, dfs_filtered, dfs_stdised,
   dfs_export)
```

