---
title: "entsoc 2024 figs and analyses"
date: "2024-10-22"
---

# setup

goal: making figs/running stats for entsoc poster, focusing on lab bugs

```{r}
here::i_am("2024/03-stats/entsoc.Rmd")
library(here)

library(tidyverse)
library(patchwork) # for plot_layout

source(here::here("set-paths.R"))
source(here::here(bin_paths$y24$wrangle, "entsoc.R"))
```

# 2. adult, larval dev

## - survival

```{r surv fig}
plot.surv <- #ss_dev.expt %>%
  ss_dev %>%
  filter(trt.type == "expt") %>%
  ggplot(aes(y = prop.survpup, x = as.factor(minT), group = stage,
             #color = trt.type
             )) +
  geom_point(size = 2, shape = 16) +
  geom_line(lty = 2) +
  #geom_text(aes(label = n.surv), vjust = 2.25) +
  theme_bw() +
  labs(y = "proportion survived to pupation",
       x = "larval nighttime temperature (°C)") +
  ylim(c(0, 1))

plot.surv
```


## - mass

```{r mass fig}
# after running stats: sex matters for adults lol

plot.devmass <- #ss_dev.expt %>%
  ss_dev %>%
  filter(trt.type == "expt") %>%
  #filter(!(stage %in% c("ad.sex", "pu"))) %>% # m/f aren't that different
  #filter(stage %in% c("ad.sex", "pu")) %>%
  filter(stage == "ad" & !is.na(sex) | stage == "pu") %>%
  # mutate(stage = case_when(stage == "la" ~ "pu",
  #                          TRUE ~ as.character(stage)),
  #        stage = factor(stage, levels = c("pu", "ad.sex"))) %>%
  # mutate(sex = case_when(is.na(sex) ~ "etc",
  #                        TRUE ~ sex)) %>%
  ggplot(aes(y = avg.mass, x = as.factor(minT), group = interaction(stage, sex),
             #group = stage,
             color = sex, shape = sex,
             #color = trt.type
             )) +
  geom_point(size = 2) +
  geom_line(aes(lty = stage)) +
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.1) +
  theme_bw() +
  labs(y = "mass (g)",
       x = "larval nighttime temperature (°C)") +
  scale_linetype_manual(labels = c("pupa", "adult"), values = c(2, 1)) +
  scale_shape_manual(#breaks = ~ .x[!is.na(.x)],
                       values = c(17, 15), limits = c("f", "m"),
                       na.translate = TRUE, na.value = 16
                       ) +
  scale_color_discrete(#breaks = ~ .x[!is.na(.x)],
                       limits = c("f", "m"),
                       na.translate = TRUE, na.value = "black"
                       ) +
  ylim(c(1.15, 4.6)) # 1.25-4.6 if omitting sex
  
plot.devmass
```



## - time

```{r dev time fig} 
plot.devtime <- #ss_dev.expt %>% 
  ss_dev %>%
  filter(trt.type == "expt") %>%
  #filter(!(stage %in% c("ad.sex", "pu"))) %>% # m/f not that different
  filter(stage %in% c("ad", "la") & is.na(sex)) %>% 
  ggplot(aes(y = avg.tt, x = as.factor(minT), group = stage,
             #color = trt.type
             )) +
  geom_point(size = 2, shape = 16) +
  geom_line(aes(lty = stage)) +
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt),
                width = 0.08) +
  theme_bw() +
  scale_linetype_manual(labels = c(#"adult\n(eclosion → death)", "larva\n(hatching → pupation)",
                                   "larva", "adult"
                                   #"pupa (pupation → eclosion)"
                                   ),
                        values = c(4, 1#, 3
                                   )
                        ) +
  labs(y = "development time / longevity (days)",
       x = "larval nighttime temperature (°C)")

plot.devtime
```


## - combined plot

```{r}
# library(patchwork)
# check that lty/shape values match across plots for merge to work


(plot.surv | plot.devmass | plot.devtime) +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(1:1)) &
  theme(legend.position = "top") &
  scale_linetype_manual(labels = c("larva", "pupa", "adult", "adult"), values = c(4,2,1,1), 
                        drop = FALSE) # good enough man idk

#ggsave("./figs/entsoc/big 3.png", dpi = "retina")
```




# 3. tent stats


```{r}
# a note that colors were formerly
# c("#f15a29", "#b2325c", "#4d276a")
```


### - win hatch

```{r win hatch rates plot}
plot.whatch <- ss_hatchrates %>%
  filter(trt.ref %in% c("win", "ctrl")) %>%
  ggplot(aes(y = prop.hatched, x = n.laidf, color = as.factor(trt.hs))) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin= prop.hatched - se.hatchprop, ymax = prop.hatched + se.hatchprop), 
                width = 8
                ) +
  geom_errorbarh(aes(xmin = n.laidf - se.laidf, xmax = n.laidf + se.laidf), 
                 height = 0.005
                 ) +
  #geom_errorbarh(aes(xmin = n.laid - se.laid, xmax = n.laid + se.laid), height = 0.01) +
  geom_text(aes(label = n.fs), vjust = -0.7, hjust = -0.7, show.legend = FALSE) +
  theme_bw() +
  labs(color = "larval temperature treatment",
       y = "proportion eggs hatched",
       x = "# eggs laid per female")+
  guides(color = guide_legend(nrow = 2, byrow = TRUE)) +
  scale_color_manual(labels = c("26-26°C", "40-19°C", "40-26°C", "40-33°C"),
                     #values = c("#0b809f", "#4d266b", "#b12256", "#f88e28")
                     #values = c("#4393c3", "#f4a582", "#d6604d", "#b2182b")
                     values = c("#4393ce", "#fdae61", "#d73027", "#a50026")
                     )

plot.whatch # y: 0-0.18?, x: 0-200
```


### - btwn hatch

```{r btwn hatch rates plot}
plot.bhatch <- ss_hatchrates %>%
  filter(trt.ref == "btwn") %>%
  ggplot(aes(y = prop.hatched, x = n.laidf, color = as.factor(trt.hs), shape = trt.sex)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin= prop.hatched - se.hatchprop, ymax = prop.hatched + se.hatchprop), 
                width = 8
                ) +
  geom_errorbarh(aes(xmin = n.laidf - se.laidf, xmax = n.laidf + se.laidf), 
                 height = 0.005
                 ) +
  geom_text(aes(label = n.fs), vjust = -0.9, hjust = -0.8, show.legend = FALSE) +
  theme_bw() +
  labs(color = "larval temperature\ntreatment",
       shape = "treated sex",
       y = "proportion eggs hatched",
       x = "# eggs laid per female") +
  scale_color_manual(labels = c("40-19°C", "40-26°C", "40-33°C"),
                     #values = c("#4d266b", "#b12256", "#f88e28"),
                     #values = c("#f4a582", "#d6604d", "#b2182b"),
                     #values = c("#fdae61", "#f46d43", "#a50026"),
                     values = c("#fdae61", "#d73027", "#a50026")) +
  guides(color = "none", # omit for plot_layout to work
         shape = guide_legend(nrow = 2)
         ) + 
  scale_shape_manual(values = c(17, 15))

plot.bhatch # y: 0-0.12, x = 0-350
```
### - combined plot

```{r}
(plot.whatch | plot.bhatch) +
  plot_layout(guides = "collect", 
              axes = "collect", 
              widths = c(1:1)
              ) &
  theme(legend.position = "bottom",
        legend.justification = "left"
        ) &
  # scale_color_manual(labels = c("26-26°C", "40-19°C", "40-26°C", "40-33°C"),
  #                    values = c("#4393ce", "#fdae61", "#d73027", "#a50026")
  #                    )
  ylim(c(-0.005, 0.18)) &
  xlim(c(0, 340))

#ggsave("./figs/entsoc/tents-2.png",dpi = "retina")
```


# 4. some archived stuff

- archived 2025-01-20 from `2024/analyses_ntw-compare.Rmd` when i first tried to combine 23v24 for entsoc lol (before I decided to just focus on 2023 data)
  - moved code into this doc 2025-02-13
- needs the dfs (all ntw bugs and 2023 ntw bugs) to run properly lolol, so see thw wrangling from there.


## - summary stats

// todo 
- 2024-10-17: keeping bugs separate by year for now, will see how they look separate before trying to combine
- 2025-01-20: filter out bugs on their correct diet types only for now

```{r}
# sep by sups (bc the sups have their own dev trajectory...)
ss_wide.sup <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT, sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) # `minT` is ok here

# all dev trajectories together
ss_wide <- all
  filter(!(pop == "lab" & diet == "TB")) %>%
  group_by(year, 
           pop, #diet, 
           trt.type, minT,
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)),
            n.surv = n.all - n.pmd,
            prop.survpup = round(1-(n.pmd/n.all), digits=2),
            n.inprog = sum(is.na(surv.outcome))) #%>% 
  #mutate_all(~ifelse(is.nan(.), NA, .)) # breaks `minT`

```


prereqs: a cleaned up `d23`

## - stage-specific stats

```{r}
# slapping sth on for entsocing
ss_larva <- d23 %>%
  filter(pop == "lab") %>%
  group_by(year, pop, diet, trt.type, minT, 
           #sup
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            n.all = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n.all - n.pmd, 
            prop.survpup = round(1-(n.pmd/n.all), digits=2)
            ) %>%
  mutate_all(~ifelse(is.nan(.), NA, .)) %>%
  mutate(minT = factor(minT, levels = c("19", "26", "33"))) %>%
  filter(!(pop == "lab" & diet == "TB") &
           trt.type == "expt")
```




## - figs


*notes:* 

- use lab bugs on LD + field on TB only!!!! (this is why u cant just reuse ur LB figs lol)
- 2025-01-20: below figs assume all combos of diet types and pops are included in the `all` df.

```{r}
# (messing around w 2023 data only for entsoc,,)
# using LB code lol

# survival
p1 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = prop.survpup, x = minT, shape = pop, color = pop,
             )) +
  geom_point(size=3) + 
  geom_line(aes(lty = pop, group = pop)) +
  geom_text(aes(label = n.all), hjust = -0.5, show.legend = FALSE) +
  theme_bw() +
  labs(title = "23v24 ntw pupal survival",
       y = "proportion survived to pupation", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year) +
  ylim(c(0, 1))

p1
```

```{r}
# avg tt pup
pd <- position_dodge(0.3)

p2 <-
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.tt, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3, #position = pd
             ) +  
  geom_errorbar(aes(ymax = avg.tt + se.tt, ymin = avg.tt - se.tt),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop), #position = pd
            ) +
  theme_bw() +
  ylim(c(18, 34)) +
  labs(title = "23v24 ntw larval development",
       y = "time to pupation (days)", 
       x = "minimum temperature (°C)") +
  facet_wrap(~year)

p2
```

```{r}
# avg mass
p3 <- 
  ss_wide %>% 
  filter(trt.type == "expt" & ((pop == "field" & diet == "TB") | (pop == "lab" & diet == "LD"))) %>%
  ggplot(aes(y = avg.mass, x = minT, shape = pop, color = pop
             )) +
  geom_point(size = 3) +  
  geom_errorbar(aes(ymax = avg.mass + se.mass, ymin = avg.mass - se.mass),
                width = 0.25, #position = pd
                ) +
  geom_line(aes(lty = pop, group = pop)) +
  theme_bw() +
  #ylim(c(1750, 4750)) +
  labs(title = "23v24 ntw pupal mass",
       y = "mass at pupation (mg)", 
       x = "minimum temperature (°C)",
       #caption = "2023 data only; lab+LD and field+TB"
       ) +
  facet_wrap(~year)

p3
```

```{r}
# merged plot
p1 + p2 + p3 +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(1:1)) &
  theme(legend.position = "top") &
  labs(lty = "population", shape = "population",
       title = NULL) &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16))
  #guides(shape = "none", lty = "none")
```




