---
title: "mean/flucts 2023 v 2024 data comparisons"
date: "2025-02-13"
---

# note to self

- 2025-02-13: this is a draft!!!!!!!!!!!!
  - this is meant to compare the meanT = 26/33 vs  flucT = 0/7 groups. i realised they should nawt be in the same doc as the `ntw-compare` stuff bc i alrdy got enough going on in that .Rmd lol
  - draft state bc need to import all the relevant data wrangling things before this sheet can work at all LOL




# 0. roadmap

goal: compare 2023 and 2024 ntw development data. needs some wrangling lol (esp 2023)

wip: combine fertility data...

# 1. load data & packages

```{r}
source("./helpers_ntw-compare.R")

test <- d23
#d23 <- test
```


# 6. const temp figs

## - survival

```{r}
# NTs on pupal survival 

p1 <- ss_cvf %>% 
  filter(stage == "la") %>% #View()
  ggplot(aes(y = prop.survpup, x = trt.type,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type)
             )) +
  geom_line() +
  geom_point(size = 2) +
  labs(x = "minimum temperature (째C)") +
  facet_wrap(~year) +
  theme_bw()

p1
```

```{r}
p2 <- ss_ntw %>% 
  filter(stage == "la" & trt.type == "ctrl") %>% 
  mutate(trt.type = "control (26/26 째C)") %>%
  ggplot(aes(y = prop.survpup, x = year,
             color = stage, 
             lty = pop, shape = year,
             group = interaction(pop, trt.type))) +
  geom_line() +
  geom_point(size = 2) +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs increase survival to pup?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(lty = "population",
       y = "proportion survived to pupation") &
  guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(1, 16)) &
  scale_color_manual(values = "#1B9E77") &
  ylim(c(0, 1))

```
## - dev time/longevity

```{r}
# NTs on development time/longevity
# see this
  # https://stackoverflow.com/questions/78193007/how-do-i-control-the-alpha-of-a-geom-point-fill

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.tt, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (째C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "la" | stage == "ad") & trt.type == "ctrl") %>% 
  ggplot(aes(y = avg.tt, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.tt - se.tt, ymax = avg.tt + se.tt), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs reduce dev time?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "development time / longevity (days)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(21, 24, 22)
                     #c(16, 17, 15)
                     #c(1, 2, 0)
                     ) &
  scale_color_manual(values = c("#1B9E77", "#7570B3")) &
  scale_fill_manual(values = c("#1B9E77", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(5, 35))

```

## - mass

```{r}
# NTs on pupal/adult mass

p1 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "expt") %>% #View()
  ggplot(aes(y = avg.mass, x = minT,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year, color = stage), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "minimum temperature (째C)") +
  facet_wrap(~year) +
  theme_bw()

p1

p2 <- ss_ntw %>% 
  mutate(sex = case_when(is.na(sex) ~ "both",
                         TRUE ~ as.character(sex))) %>%
  filter(!(stage == "ad" & sex == "all")) %>%
  filter((stage == "pu" | stage == "ad") & trt.type == "ctrl") %>% 
  ggplot(aes(y = avg.mass, x = year,
             color = stage, fill = stage, shape = sex,
             group = interaction(pop, stage, sex)
             )) +
  geom_errorbar(aes(ymin = avg.mass - se.mass, ymax = avg.mass + se.mass), width = 0.25) +
  geom_line(aes(lty = pop), alpha = 1) +
  geom_point(aes(alpha = year), size = 2) +
  geom_point(fill = NA, size = 2) +
  labs(x = "year") +
  facet_wrap(~trt.type) +
  theme_bw()

p2

# merge
p1 + p2 +
  labs(caption = "do lower NTs lead to larger mass?") +
  plot_layout(guides = "collect",
              axes = "collect",
              widths = c(2:1)) &
  theme(legend.position = "top") &
  labs(#lty = "population",
       y = "mass (g)") &
  #guides(color = "none") &
  scale_linetype_manual(values = c("dashed", "solid")) &
  scale_shape_manual(values = c(24, 22)) &
  scale_color_manual(values = c("#D95F02", "#7570B3")) &
  scale_fill_manual(values = c("#D95F02", "#7570B3")) &
  scale_alpha_manual(values = c(0.1,1)) &
  ylim(c(1.1, 7.25))

```