---
title: "02-wrangle/compare-tents"
date: "2025-02-27"
---

# 0. preamble

wrangling data for 2023 v 2024 tent data comparisons. 

final code is purled to a `.R` script, manually rerun `purl()` as needed to update the wrangling script for analyses

```{r eval=FALSE, include=FALSE, purl=FALSE}
knitr::purl("compare-tents.Rmd", "compare-tents_p.R")
```


# 1. loading & defining

```{r about}
# wrangle/compare-tents.R

# knitted wrangling code for 23v24 tent comparison figs/analyses.
# source() this in corresponding analysis scripts.
```

```{r load utils}
library(tidyverse)

here::i_am("2024/02-wrangle/compare-tents.Rmd")
library(here)

source(here::here("set-paths.R"))
#source(here::here(bin_paths24$wrangle, "compare-tents_util.R")) # idk if need this yet
```

```{r load 2324 tent data}
d23 <- read.csv(here(bin_paths23$data, "clean-tentstats.csv"), header = TRUE) %>% mutate(year = 2023)
d24 <- read.csv(here(bin_paths24$data, "clean-fertility.csv"), header = TRUE) %>% mutate(year = 2024)

# drop a tent from 2023 that was reared up differently
d23 <- filter(d23, id.tent != "107-I")
```



# 2. standardising cols, format; col fixes

## - renaming

```{r rename cols}
# renaming columns
d23 <- d23 %>%
  rename(n.hatch = n.tothatch,
         id.room = room,
         id = id.tent,
         id.tent = tent,
         n.ovi = females.laying,
         notes = notes.tents) %>%
  select(-c("tent.loc", "if.ignore", "notes"))

d24 <- d24 %>%
  rename(id = uid2,
         date = date.coll) %>%
  select(-c("combo", "notes"))

jic <- list(d23 = d23,
            d24 = d24)
```


## - fixing d23 egg counts

2025-02-28: lowkey seems like i dont actually need this anymore lol (apparently i fixed stuff at some point) but keeping for ref
2025-03-20: no u *do* need this for doing stats, manual calcs throw in some extra negs that need to be fixed here

ripped this code from `entsoc` btw

```{r d23 egg fixes}
# determine actual # of ovipositing females
  # 2025-02-28 dont actually need this anymore bc "females.laying" is manually calculated in the gsheet!
# dtents <- dtents %>%
#   mutate_at(c(6:14), ~replace_na(., 0)) %>% 
#   group_by(id.tent) %>%
#   mutate(jdate.rel = jdate - first(jdate)) %>%
#   ungroup() %>%
#   group_by(id.tent, trt.f, trt.m, trt.sex, jdate.rel) %>%
#   arrange(jdate.rel, .by_group = TRUE) %>%
#   mutate(#n.laying = n.curr.f + n.died.f, .after = "n.curr.f",
#          n.ovi = dplyr::lag(n.curr.f, default = first(n.curr.f)) + n.died.f, .after = "n.curr.f") %>%
#   ungroup()


# where n.ovipositing females <= 0 but n.collected > 0, add the # eggs collected to the day before
# (or last date of ovipositing females)

# first, figure out the rows
#cond <- (d23$n.coll > 0 & d23$n.ovi == 0)
cond <- (d23$n.coll > 0 & d23$n.ovi <= 0)

ovifix <- d23 %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

negovi <- ovifix[ovifix$n.ovi <= 0, ]

lastovi <- ovifix[ovifix$n.curr.f > 0 & ovifix$n.ovi >= 1, ]

#test <- c(negovi$index, lastovi$index)

# then, edit values
# summ_rate$n.coll[negovi$index] <- 0
d23 <- d23 %>% 
  mutate(n.coll = case_when(rownames(.) %in% negovi$index ~ 0,
                            rownames(.) %in% lastovi$index ~ n.coll + 1,
                            !(rownames(.) %in% c(negovi$index, lastovi$index)) ~ as.numeric(n.coll)),
         n.ovi = case_when(n.ovi < 0 ~ 0, # needs to be ≥ 0 for modeling math
                           TRUE ~ as.numeric(n.ovi))
  )


# check: ovifix2 should be blank
cond2 <- (d23$n.coll > 0 & d23$n.ovi <= 0)

ovifix2 <- d23 %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

# removing extra 0's
# dtents <- dtents %>%
#   filter(!(n.coll == 0 & n.ovi == 0))
  # tbh not sure why i did this in the old version. this drops a few #s here and there
```

## - merging, standardising dfs
```{r merge 2324 dfs}
# have a merged version of 23-24
dall <- d24 %>%
  rename(n.new.f = n.new,
         n.curr.f = n.curr,
         n.rem.f = n.rem,
         n.died.f = n.died,
         trt.sex = sex) %>%
  mutate(trt.f = trt,
         trt.temp = trt) %>%
  bind_rows(., d23) %>%
  select(names(d23))
```

```{r standardise column format}
# column standardising
fix.cols <- function(data){
  data <- data %>%
    mutate(date = as.Date(date, format = "%Y-%m-%d"),
           jdate = as.numeric(as.Date(date, "%j")),
           year = as.factor(year))
}

dall <- fix.cols(dall)
d24 <- fix.cols(d24)

# apply column fixing to dfs,
# keep separate to stitch them together in certain ways later
# temp <- list(d23 = d23, d24 = d24, d242 = d242)
# temp <- lapply(temp, fix.cols)
# list2env(temp, envir = .GlobalEnv)
#all <- merge(d23, d242, all = TRUE)
```

testing map/lappy to apply fn to more 3 dfs at once lol. want the dfs separated eventually bc only some of them are getting merged

- https://stackoverflow.com/questions/47140756/map-with-purrr-multiple-dataframes-and-have-those-modified-dataframes-as-the-out

```{r map testing, eval=FALSE, include=FALSE, purl=FALSE}
# make d23, d24, d242
test <- list(d23 = d23, d24 = d24, d242 = d242)
  # naming here skips needing to do: names(df) <- c(...)

#lapply(test, fix.cols) %>% list2env(., .GlobalEnv) # idt the . pronoun works like this lol

test <- lapply(test, fix.cols)
list2env(test, envir = .GlobalEnv)
```


# 3. adding extra columns

```{r add more cols}
d24 <- d24 %>%
  mutate(minT = as.factor(case_when((trt == 260 | trt == 426) ~ 26,
                                    trt == 419 ~ 19,
                                    trt == 433 ~ 33)),
         trt.type = case_when(trt == 260 ~ "ctrl",
                              TRUE ~ "expt"))

#d23 <- d23 %>%
dall <- dall %>%
  mutate(id.room = as.factor(case_when(year == 2024 ~ 107,
                                       TRUE ~ as.numeric(id.room))),
         trt.type = case_when(year == 2024 & trt.f == 260 ~ "ctrl",
                              year == 2024 & trt.f > 260 ~ "expt",
                              trt.m != trt.f ~ "btwn",
                              TRUE ~ "win"
                              #trt.m == trt.f ~ "win",
                              #TRUE ~ "btwn"
         ),
         #trt.type = factor(trt.type, levels = c("win", "btwn", "expt", "ctrl")),
         trt.sex = case_when(trt.m > trt.f ~ "m",
                             is.na(trt.m) | year == 2024 | trt.m < trt.f ~ "f",
                             TRUE ~ "both"),
         trt.temp = case_when(#trt.m > trt.f ~ trt.m, 
                               is.na(trt.m) | year == 2024 | trt.m < trt.f ~ trt.f,
                               TRUE ~ trt.m)) #%>%
  #mutate(across(c(trt.f, trt.m, trt.temp), as.factor))
```


# cleanup

```{r cleanup}
rm(#moreinfo, temp,
   fix.cols, jic)

rm(cond, ovifix, negovi, lastovi, cond2, ovifix2)
```

