---
title: "02-wrangle/compare-tents"
date: "2025-02-27"
---

# 0. preamble

wrangling data for 2023 v 2024 tent data comparisons. 

final code is purled to a `.R` script, manually rerun `purl()` as needed to update the wrangling script for analyses

```{r eval=FALSE, include=FALSE, purl=FALSE}
#knitr::purl("compare-tents.Rmd", "compare-tents_p.R")
```


# 1. loading & defining

```{r about}
# wrangle/compare-tents.R

# knitted wrangling code for 23v24 tent comparison figs/analyses.
# source() this in corresponding analysis scripts.
```

```{r load utils}
library(tidyverse)

here::i_am("2024/02-wrangle/compare-tents.Rmd")
library(here)

source(here::here("set-paths.R"))
#source(here::here(bin_paths24$wrangle, "compare-tents_util.R")) # idk if need this yet
```

```{r load 2324 tent data}
d23 <- read.csv(here(bin_paths23$data, "clean-tentstats.csv"), header = TRUE) %>% mutate(year = 2023)
d24 <- read.csv(here(bin_paths24$data, "clean-fertility.csv"), header = TRUE) %>% mutate(year = 2024)

jic <- list(d23 = d23,
            d24 = d24)
```

# 2. standardising cols, format

```{r rename cols}
# renaming columns
d23 <- d23 %>%
  rename(n.hatch = n.tothatch,
         id.room = room,
         id = id.tent,
         id.tent = tent,
         n.ovi = females.laying,
         notes = notes.tents) %>%
  select(-c("tent.loc", "if.ignore", "notes"))

d24 <- d24 %>%
  rename(id = uid2,
         date = date.coll,
         # n.new.f = n.new,
         # n.curr.f = n.curr,
         ) %>%
  select(-c("combo", "notes"))
```

```{r standardise column format}
# column standardising
fix.cols <- function(data){
  data <- data %>%
    mutate(date = as.Date(date, format = "%Y-%m-%d"),
           jdate = as.numeric(as.Date(date, "%j")),
           year = as.factor(year))
}

d23 <- fix.cols(d23)
d24 <- fix.cols(d24)

rm(fix.cols)
```


# 3. adding/fixing stuff

## - adding extra columns

```{r add more cols}
d24 <- d24 %>%
  mutate(minT = as.factor(case_when((trt == 260 | trt == 426) ~ 26,
                                    trt == 419 ~ 19,
                                    trt == 433 ~ 33)),
         trt.type = case_when(trt == 260 ~ "ctrl",
                              TRUE ~ "expt"))

d23 <- d23 %>%
  mutate(trt.type = case_when(trt.m == trt.f ~ "win",
                              TRUE ~ "btwn"),
         trt.sex = case_when(trt.m > trt.f ~ "m",
                             trt.m < trt.f ~ "f",
                             TRUE ~ "both"),
         trt.temp = as.factor(case_when(trt.m > trt.f ~ trt.m,
                              trt.m < trt.f ~ trt.f,
                              TRUE ~ trt.m)))
```


## - correcting d23 egg counts

2025-02-28: lowkey seems like i dont actually need this anymore lol (apparently i fixed stuff at some point) but keeping for ref

```{r d23 egg count fix, eval=FALSE, include = FALSE, purl=FALSE}
# determine actual # of ovipositing females
  # 2025-02-28 dont actually need this anymore bc "females.laying" is manually calculated in the gsheet!
# dtents <- dtents %>%
#   mutate_at(c(6:14), ~replace_na(., 0)) %>% 
#   group_by(id.tent) %>%
#   mutate(jdate.rel = jdate - first(jdate)) %>%
#   ungroup() %>%
#   group_by(id.tent, trt.f, trt.m, trt.sex, jdate.rel) %>%
#   arrange(jdate.rel, .by_group = TRUE) %>%
#   mutate(#n.laying = n.curr.f + n.died.f, .after = "n.curr.f",
#          n.ovi = dplyr::lag(n.curr.f, default = first(n.curr.f)) + n.died.f, .after = "n.curr.f") %>%
#   ungroup()


# where n.ovipositing females = 0 but n.collected > 0, add the # eggs collected to the day before
# (or last date of ovipositing females)

# first, figure out the rows
cond <- (d23$n.coll > 0 & d23$n.ovi == 0)

ovifix <- d23 %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

negovi <- ovifix[ovifix$n.ovi < 0, ]

lastovi <- ovifix[ovifix$n.curr.f > 0 & ovifix$n.ovi >= 1, ]

#test <- c(negovi$index, lastovi$index)

# then, edit values
# summ_rate$n.coll[negovi$index] <- 0
d23 <- d23 %>% 
  mutate(n.coll = case_when(rownames(.) %in% negovi$index ~ 0,
                            rownames(.) %in% lastovi$index ~ n.coll + 1,
                            !(rownames(.) %in% c(negovi$index, lastovi$index)) ~ as.numeric(n.coll)
                            ))

rm(cond, ovifix, negovi, lastovi)

# removing extra 0's
# dtents <- dtents %>%
#   filter(!(n.coll == 0 & n.ovi == 0))
  # tbh not sure why i did this in the old version. this drops a few #s here and there
```

# cleanup

```{r cleanup}
rm(moreinfo, 
   jic)
```

