---
title: "02-wrangle/compare-dev"
date: "2025-02-19"
---

# 0. preamble

wrangling data for 2023 v 2024 development comparisons. 

final code is purled to a `.R` script, manually rerun `purl()` as needed to update the wrangling script for analyses

```{r eval=FALSE, include=FALSE, purl=FALSE}
knitr::purl("compare-dev.Rmd", "compare-dev_p.R")
```

other notes:

- 2025-02-13: commenting out/moved out some of the 26/33 vs ±0/7 stuff.


# 1. loading & defining

```{r about}
# wrangle/compare-dev.R

# knitted wrangling code for ntw development comparison figs/analyses.
# source() this in corresponding analysis scripts.

print(paste0("last updated: ", format(Sys.time(), "%Y-%m-%d")))
```

```{r load utils}
library(tidyverse)
library(conflicted)
conflicts_prefer(dplyr::filter)

here::i_am("2024/02-wrangle/compare-dev.Rmd")
library(here)

source(here::here("set-paths.R"))
source(here::here(bin_paths$y24$wrangle, "compare-dev_util.R"))
```

```{r load 2324 dev data}
d23 <- read.csv(here(bin_paths$y23$data, "clean-ntw.csv"), header = TRUE) %>% mutate(year = 2023)
d24 <- read.csv(here(bin_paths$y24$data, "ntw.csv"), header = TRUE) %>% mutate(year = 2024)

test <- d23
#d23 <- test
```

- adult longevity data was very convolutedly calculated and i am now suffering the consequences

```{r load 23 longevity data}
source(here::here(bin_paths23$y23, "helpers_tents.R"))

d23_longevity <- data_longevity %>%
  filter(!is.na(jdate.died)) %>%
  select(id, sex, trt, jdate.ec, jdate.died) %>%
  rename(jdate.lec = jdate.ec,
         jdate.lsurv = jdate.died)

rm(data_tstats, data_hatch, data_longevity,
   labels.alltrts, labels.exptrts, RYB,
   x_err_ncoll, x_err_ncollf, y_err_hrate)
```


# 2. formatting

```{r filtering 2023 data}
# pick relevant data and rename columns/values to match 2024

# troubleshooting: compare column names
# setdiff(names(d24), names(d23)) # differences
# setdiff(names(d23), names(d24))
# intersect(names(d24), names(d23)) # same

# filter rows and rename columns
d23 <- d23 %>%
  filter(reason.ignore != "lost" | is.na(reason.ignore) #| !is.na(final.fate)
         ) %>% 
  filter(src != "F1") %>% 
  filter(trt.stage %in% c("260-hatch", "267-hatch", "330-hatch", "337-hatch", "419-hatch","433-hatch")) %>%
  filter(expt.group %in% c("A", "B", "C", "D", "E", "F", "H")) %>%
  rename(trt = treatment,
         id = ID,
         cohort = expt.group,
         notes.ignore = reason.ignore)

# standardise values
d23$trt[d23$trt == 337] <- 426

# d23 <- d23 %>%
#   mutate(final.fate = case_when(notes.ignore %in% c("hot larva", "cut", "culled larva", "wet diet") | final.fate == "culled" ~ "other",
#                                 TRUE ~ as.character(final.fate))
#          )

# more filtering (remaining "lost" things and early field pmds; <1 day deaths)
d23 <- d23 %>%
  filter(!(notes.ignore == "early field" & final.fate == "pmd") | is.na(notes.ignore)) %>%
  # filter(jdate.exit - jdate.enter > 1 |
  #          (is.na(jdate.exit) & (is.na(jdate.pmd) | (jdate.pmd - jdate.enter > 1)))) %>% View()
  mutate(notes.ignore = case_when(notes.ignore == "early field" & id == 217 ~ NA_character_,
                                  TRUE ~ notes.ignore))

#testa <- d23
```

```{r 2023 LPI and exit date handling}
# address LPIs/post-pupal deaths to avoid wonkiness
#write.csv(d23, "~/Desktop/d23temp.csv")

# append longevity data to d23
d23_longevity[d23_longevity$id == 1415, "sex"] <- "f" # fix data

d23 <- merge(d23, d23_longevity, all = TRUE)
#testb <- d23

# redo exit dates
d23 <- d23 %>%
  mutate(is.LPI = case_when(!is.na(date.LP) ~ 1,
                            TRUE ~ 0),
         jdate.pupa = case_when(!is.na(jdate.LP) ~ jdate.LP, # treat LPs like pups
                                TRUE ~ jdate.pupa),
         jdate.15 = case_when(!is.na(jdate.eclose) & is.na(jdate.pupa) ~ NA_integer_,
                              TRUE ~ jdate.15),
         jdate.surv = case_when(!is.na(jdate.lsurv) ~ jdate.lsurv,
                                TRUE ~ as.numeric(jdate.surv))) %>%
  select(-c("jdate.lec", "jdate.lsurv")) %>%
         # correcting exit dates...
  mutate(across(starts_with("jdate"), as.numeric)) %>% 
  mutate(jdate.exit2 = case_when(!is.na(jdate.surv) ~ jdate.surv,
                                 !is.na(jdate.eclose) ~ jdate.eclose,
                                 !is.na(jdate.pmd) ~ jdate.pmd,
                                 !is.na(jdate.exit) ~ jdate.exit)) %>% 
  mutate(jdate.exit2 = case_when(is.na(jdate.exit2) & !is.na(jdate.pupa) ~ jdate.pupa + 70,  # assume dead, non-diapausing
                                 TRUE ~ jdate.exit2)) %>% # anything else got lost
         # correcting jdate.pmd: any pups should not have an empty jdate 
  mutate(jdate.pmd = case_when(!is.na(jdate.pupa) ~ NA_real_, # only pmd if no pup
                               TRUE ~ jdate.pmd))

```

```{r matching 2023 binaries to 2024}
d23 <- d23 %>%
  mutate(final.fate = case_when(notes.ignore %in% c("culled larva", "cut", "hot larva", "wet diet") ~ "other"))

# match values to d24
d23 <- d23 %>%
  mutate(sup = case_when(!is.na(jdate.7th) ~ 7,
                         !is.na(jdate.6th) ~ 6,
                         # (is.na(jdate.pmd) & !is.na(jdate.pupa) |
                         #    is.na(jdate.6th) & !is.na(jdate.pupa)) ~ 0), # for things that died as pups
                          is.na(jdate.pmd) & (!is.na(jdate.pupa) | !is.na(jdate.LP)) ~ 0),
                         #is.na(jdate.pmd) & !is.na(jdate.pupa) ~ 0),
         # case_when(sup = 0 & !is.na(jdate.5th) ~ NA_real_,
         #                 TRUE ~ as.numeric(sup)),
         surv.outcome = case_when(final.fate == "other" | 
                                    #(is.na(jdate.pmd) & is.na(jdate.pupa) & is.na(jdate.LP)) ~ 2,
                                    (is.na(jdate.pmd) & is.na(jdate.pupa)) ~ 2,
                                  !is.na(date.pupa) | !is.na(date.LP) ~ 1,
                                  #!is.na(date.pupa) & !is.na(jdate.exit2) ~ 1,
                                  !is.na(date.pmd) ~ 0,
                                  #TRUE ~ 2
                                  ),
         # surv.outcome = case_when(!is.na(date.pupa) | !is.na(date.LP) ~ 0,
         #                          final.fate == "pmd" ~ 1,
         #                          final.fate == "other" ~ 2), # treat injuries the same as this
         flucT = case_when(flucT == 2.5 ~ 0,
                           TRUE ~ as.numeric(flucT)),
         trt.type = case_when(meanT == 26 & flucT == 0 ~ "ctrl",
                              TRUE ~ "expt"),
         trt = as.numeric(trt),
         #id = paste(id, year, sep = "-")
         )

# troubleshooting
# filter(d23, is.na(surv.outcome)) %>% View()
#test %>% group_by(final.fate) %>% summarise(n = n())
#test %>% group_by(reason.ignore) %>% summarise(n = n())
#filter(test, is.na(final.fate)) %>% View()

test2 <- d23
#d23<-test2

# how to handle LPIs? (in 23 and 24?)

d23 <- d23 %>%
  select(-c("jdate.collected", "jdate.15", "jdate.stuck",
            "jdate.exit", "jdate.exit2", "jdate.LP", "jdate.died")) %>% # d23 exclusives
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "instar.enter", starts_with(c("jdate", "mass")), 
           "sex",
           ends_with("T", ignore.case = FALSE), "trt.type",
           "sup", "surv.outcome", "year"
          ))

```


```{r fine-tune 2024}
d24 <- d24 %>%
  #mutate(id = paste(id, year, sep = "-")) %>%
  select(-"jdate.culled") %>%
  filter(instar.enter == "hatch")
```


# 3. merging

```{r final df}
all <- merge(d24, d23, all = TRUE) %>%
  drop_na(id) %>%
  select(-c("instar.enter")) %>%
  mutate(is.pup = surv.outcome)


# drop things that died in 1 day
  # 2025-01-23: can i do this??????
all <- filter(all, jdate.pmd - jdate.enter > 1 | is.na(jdate.pmd - jdate.enter))

# add in exit dates to more easily parse out things still developing (for 2024 data)
# and other convenience things (i.e. more dev binaries)
  # for dev binaries, 0 = no, 1 = yes
all <- all %>% 
  mutate(
        # jdate.exit = case_when(!is.na(jdate.pmd) ~ jdate.pmd,
        #                         !is.na(jdate.pupa) ~ jdate.pupa,
        #                         TRUE ~ NA_integer_),
         minT = factor(minT, levels = c(19, 26, 33)),
         year = factor(year, levels = c(2023, 2024)),
         is.sup = case_when((!is.na(jdate.pupa) & is.na(sup)) | sup == 0 ~ 0,
                            sup > 1 ~ 1),
         is.ec = case_when(!is.na(jdate.eclose) ~ 1,
                               is.pup == 1 & is.na(jdate.eclose) ~ 0),
         for.ntw = case_when(#year == 2023 & cohort %in% c("A", "B") ~ "N",
                             trt %in% c(260, 419, 426, 433) ~ "Y",
                             TRUE ~ "N"),
         for.const = case_when(#flucT %in% c(0, 7) & meanT %in% c(26, 33) ~ "Y",
                               trt %in% c(260, 267, 330, 426) ~ "Y",
                               TRUE ~ "N"),
                               # trt.type == "ctrl" | (meanT == 33 & flucT %in% c(0, 7)) ~ "Y",
                               # TRUE ~ "N")
         )

```


# 4. figs df prep (math)

## - pivot

2025-01-20: filtering out diet wonkiness here lol. we can revisit it later if needed

```{r}
all_wide <- all %>%
  filter(!(pop == "lab" & diet == "TB")) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
         tt.4th = jdate.4th - jdate.hatch,
         tt.5th = jdate.5th - jdate.hatch,
         tt.6th = jdate.6th - jdate.hatch,
         tt.7th = jdate.7th - jdate.hatch,
         tt.8th = jdate.8th - jdate.hatch, # omit if only d23
         tt.wander = jdate.wander - jdate.hatch,
         tt.pupa = jdate.pupa - jdate.hatch,
         tt.eclose = jdate.eclose - jdate.pupa,
         tt.surv = jdate.surv - jdate.eclose,
         #tt.lsurv = jdate.lsurv - jdate.lec, # for d23 longevity
         tt.pmd = jdate.pmd - jdate.hatch
         )  
# %>%
#   # drop extra d23 longevity columns
#   mutate(tt.surv = case_when(is.na(tt.surv) ~ as.numeric(tt.lsurv),
#                              TRUE ~ as.numeric(tt.surv))) %>%
#   #filter(year == 2023 & sex == "f" & !is.na(jdate.lsurv)) %>% View()
#   select(-c("jdate.lsurv", "jdate.lec", "tt.lsurv"))

all_long <- all_wide %>%
  pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
               names_to = c(".value", "instar"),
               values_drop_na = TRUE,
               names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
  drop_na(jdate) %>%
  drop_na(tt) # drops NA's if an individual didnt reach a certain stage


# filtering by expt type
ntw_wide <- filter(all_wide, for.ntw == "Y")
#cvf_wide <- filter(all_wide, for.const == "Y")
```

## - stage-specific summary stats

2025-01-20 troubleshooting:
- https://stackoverflow.com/questions/73533852/create-multiple-dataframes-based-on-the-dates-of-another-dataframe-in-r
- https://stackoverflow.com/a/24376207/3358227

```{r}
# see/load utility fns in separate .R

# apply calcs to data
#ss_all <- lapply(list(ntw_wide, cvf_wide), calc.devstats)

# ss_ntw <- Reduce(full_join, ss_all[[1]]) %>% 
#   calc.ssadj()

# ss_cvf <- Reduce(full_join, ss_all[[2]]) %>%
#   calc.ssadj()

ss_ntw <- ntw_wide %>%
  calc.devstats() %>%
  Reduce(full_join, .) %>%
  calc.ssadj()
```

## - dev ss variations

```{r}
ss_moredev <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab")) %>% 
  calc.ssmoredev()
```

## - dev outcomes

//todo
- 2025-02-04: try to group by surv outcome maybe... (instead of splitting into 2 dfs)

```{r}
# pmds + sup of those that pup'ed (i.e. pmd + pup5 + pup6...)
ss_devall <- ntw_wide %>%
  mutate(trt = as.factor(trt)) %>%
  filter(!(diet == "TB" & pop == "lab")) %>%
  group_by(year, diet, pop, trt) %>%
  summarise(n.tot = n(),
            n.pmd = sum(na.omit(is.pup == 0)),
            n.misc = sum(na.omit(is.pup == 2)), # accident deaths
            n.pup = sum(na.omit(is.pup == 1)),
            n.sup = sum(na.omit(is.pup == 1 & sup > 0)), # surviving sups
            n.dev = n.tot - n.misc, # = pmd + pups of any sup
            # pups of 5+ instars
            n.5th = sum(na.omit(sup == 0 & is.pup == 1)),
            n.6th = sum(na.omit(sup == 6 & is.pup == 1)),
            n.7th = sum(na.omit(sup == 7 & is.pup == 1)),
            n.8th = sum(na.omit(sup == 8 & is.pup == 1)),
            p.pmd = n.pmd/n.dev, 
            p.sup = n.sup/n.dev, # any pup with 6+ instars
            p.5th = n.5th/n.dev, # pup in 5... etc
            p.6th = n.6th/n.dev,
            p.7th = n.7th/n.dev,
            p.8th = n.8th/n.dev,
            #tot = n.dev - (n.5th + n.6th + n.7th + n.8th), # = pmds
            #p.tot = p.pmd + p.5th + p.6th + p.7th + p.8th
            )

## troubleshooting...
# ntw_wide %>%
#   filter(is.pup == 1) %>% View()
 # check: pup but not ec; pmd = d before pup


# sup of those that pmd'd (i.e pmd6 + pmd7 + ...)
  # 2025-05-04: lowkey math doesnt rly make sense so commenting out
# ss_devpmd <- ntw_wide %>%
#   mutate(trt = as.factor(trt)) %>%
#   filter(!(diet == "TB" & pop == "lab")) %>%
#   filter(is.pup == 0) %>%
#   group_by(year, diet, pop, trt) %>%
#   summarise(n.tot = n(),
#             #n.5th = sum(na.omit(is.na(sup))),
#             n.6th = sum(na.omit(sup == 6)),
#             n.7th = sum(na.omit(sup == 7)),
#             n.8th = sum(na.omit(sup == 8)),
#             n.5th = n.tot - n.6th - n.7th - n.8th, # (anyone that died before 6th...)
#             p.5th = round(n.5th/n.tot, digits = 3),
#             p.6th = round(n.6th/n.tot, digits = 3),
#             p.7th = round(n.7th/n.tot, digits = 3),
#             p.8th = round(n.8th/n.tot, digits = 3)
#             )

# pivots
ss_devall2 <- ss_devall %>%
  pivot_longer(cols = starts_with(c("n.", "p.")),
               names_to = c(".value", "stage"),
               names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
  filter(stage %in% c("pmd", "5th", "6th", "7th"))

# ss_devpmd2 <- ss_devpmd %>%
#   pivot_longer(cols = starts_with(c("n.", "p.")),
#                names_to = c(".value", "stage"),
#                names_pattern = ("([a-z]*)\\.(\\d*[a-z]*)")) %>%
#     filter(stage %in% c("5th", "6th", "7th"))
```



# 5. stats df prep (filtering)

2025-02-26: moved these into the stats script for easier manipulation, just changing the chunk header for now

- filter bugs for modeling: 
  - exptal bugs only (D = 40, N = 19/26/40)
  - field on TB and lab on LD

```{r general ntw filter, purl=FALSE}
ntw_expt <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab") & trt.type == "expt") %>%
  mutate(dmass = mass.pupa - mass.eclose,
         rate.pup = mass.pupa/tt.pupa)
```

```{r specific ntw filters, purl=FALSE}
# omit accidental deaths (so either pmd, pup, LPI)
ntw_expt.surv <- ntw_expt %>%
  filter(is.pup != 2)

# omit unsexed pupa
ntw_expt.sex <- ntw_expt %>%
  filter(!is.na(sex))
```

# 6. cleanup

```{r cleanup}
rm(test, test2)

rm(calc.devstats, calc.ssadj, calc.ssmoredev)
 
rm(all, d23, d23_longevity, d24)
```

