---
title: "wrangle/entsoc.Rmd"
date: "2025-02-20"
---

# 0. preamble

wrangling data for 2024 entsoc poster.

final code is purled to a `.R` script, rerun `purl()` step as needed to update the wrangling for analyses

```{r eval=FALSE, include=FALSE, purl=FALSE}
knitr::purl("entsoc.Rmd")
```

other notes:

- copied/pasted sections of analyses_ntw-compare.Rmd focusing on the 2023 data for this poster

# 1. loading & defining

```{r about}
# wrangle/entsoc.R

# knitted wrangling code for entsoc figs/analyses.
# source() this in corresponding analysis scripts.
```

```{r load utils}
library(tidyverse)

here::i_am("2024/02-wrangle/entsoc.Rmd")
library(here)

source(here::here("set-paths.R"))
```

```{r load data}
# 2023 dev data
d23 <- read.csv(here::here(bin_paths23$data,"clean-ntw.csv"), header = TRUE) %>% mutate(year = 2023)

# 2023 tent data
source(here::here(bin_paths23$y23, "helpers_tents.R"))

dtents <- data_tstats 
#%>% select(-c("trt.type", "trt.pair"))

dad <- data_longevity

rm(data_tstats, data_hatch, data_longevity,
   labels.alltrts, labels.exptrts, RYB,
   x_err_ncoll, x_err_ncollf, y_err_hrate)
```

```{r convenience functions}
se <- function(x){
  sd(na.omit(x))/sqrt(length(na.omit(x)))
}
```

# 2. df formatting

```{r dev data clean up}
# filter NTW data and column renaming
d23 <- d23 %>%
  filter(reason.ignore != "lost" | src != "F1" #| !is.na(final.fate)
  ) %>% 
  filter(trt.stage %in% c("260-hatch", "337-hatch", "419-hatch","433-hatch")) %>%
  filter(expt.group %in% c("C", "D", "E", "F", "H")) %>%
  rename(trt = treatment,
         id = ID,
         cohort = expt.group,
         notes.ignore = reason.ignore)

# standardise values
d23$trt[d23$trt == 337] <- 426

d23 <- d23 %>%
  mutate(final.fate = case_when(notes.ignore %in% c("hot larva", "cut", "culled larva", "wet diet") ~ "other",
                                TRUE ~ as.character(final.fate)))

# simplify values
d23 <- d23 %>%
  mutate(sup = case_when(sup = 0 & !is.na(jdate.5th) ~ NA_real_,
                         TRUE ~ as.numeric(sup)),
         surv.outcome = case_when(!is.na(date.pupa) | !is.na(date.LP) ~ 0,
                                  final.fate == "pmd" ~ 1,
                                  final.fate == "other" ~ 2), # group injuries together
         flucT = case_when(flucT == 2.5 ~ 0,
                           TRUE ~ as.numeric(flucT)),
         trt.type = case_when(meanT == 26 & flucT == 0 ~ "ctrl",
                              TRUE ~ "expt"),
         trt = as.numeric(trt)
  )

### TROUBLESHOOTING
backup_d23.init <- d23

# pick relevant columns
d23 <- d23 %>%
  select(c("cohort", "pop", "diet", "trt", "id", # identifying info
           "instar.enter", starts_with(c("jdate", "mass")), 
           "sex",
           ends_with("T", ignore.case = FALSE), "trt.type",
           "sup", "surv.outcome"
  )) %>%
  select(-c("jdate.collected", "jdate.stuck", "jdate.LP", "jdate.died",
            ends_with("7th"), ends_with("15")))


# drop things that died in 1 day
d23 <- filter(d23, jdate.pmd - jdate.enter > 1 | is.na(jdate.pmd - jdate.enter))


# add exit dates and other convenience things for later
d23 <- d23 %>% 
mutate(jdate.exit = case_when(!is.na(jdate.pmd) ~ jdate.pmd,
                              !is.na(jdate.pupa) ~ jdate.pupa,
                              TRUE ~ NA_integer_),
       #minT = factor(minT, levels = c(19, 26, 33))
       )


### TROUBLESHOOTING
backup_d23.all <- d23

# more heavy-handed filtering:
  # focus on lab bugs for entsoc

d23 <- filter(d23, pop == "lab" & diet == "LD")
```

```{r dev data formatting}
# change mass to g
d23 <- d23 %>%
  mutate(mass.pupa = mass.pupa/1000,
         mass.eclose = mass.eclose/1000)

# adult data
dad <- dad %>%
  rename(jdate.eclose = jdate.ec,
         jdate.surv = jdate.died) %>%
  mutate(mass.eclose = mass.eclose/1000,
         for.surv = case_when(track.reason %in% c("surv-fert", "surv") ~ 1),
         for.fert = case_when(track.reason %in% c("surv-fert", "fert") ~ 1),
         trt.type = case_when(maxT == minT ~ "ctrl",
                              TRUE ~ "expt")) %>%
  select(-c(track.reason, starts_with("date")))
```


```{r tent data formatting}
dtents <- dtents %>%
  filter(pop == "lab") %>%
  #rename(trt.typeold = trt.type) %>%
  mutate(trt.mateto = case_when(trt.m != trt.f | trt.m + trt.f == 520 ~ "ctrl",
                                TRUE ~ "hs"),
         trt.sex = case_when(trt.m > trt.f ~ "m",
                             trt.f > trt.m ~ "f",
                             trt.m == trt.f ~ "both"),
         trt.hs = case_when(trt.m > trt.f ~ as.numeric(trt.m),
                            trt.f > trt.m ~ as.numeric(trt.f),
                            trt.m + trt.f > 600 ~ as.numeric(trt.m),
                            TRUE ~ 260),
         trt.isctrl = case_when(trt.m + trt.f == 520 ~ 1,
                              TRUE ~ 0),
         trt.ref = case_when(trt.m + trt.f == 520 ~ "ctrl",
                             trt.m != trt.f ~ "btwn",
                             trt.m == trt.f ~ "win")
         ) %>%
  select(-pop)
```

# 3. development
## a. viz prep

```{r calc dev summ stats}
# larval stats
ss_la <- d23 %>%
  group_by(trt.type, minT, 
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.pupa - jdate.hatch)),
            se.tt = sd(na.omit(jdate.pupa - jdate.hatch))/sqrt(length(na.omit(jdate.pupa - jdate.hatch))),
            # avg.mass = mean(na.omit(mass.pupa)),
            # se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),
            #n = n(),
            n = n() - sum(na.omit(surv.outcome == 2)),
            n.pmd = sum(na.omit(surv.outcome == 1)), 
            n.surv = n - n.pmd, 
            prop.survpup = round(1-(n.pmd/n), digits=2)
            ) %>% 
  #mutate_all(~ifelse(is.nan(.), NA, .))
  mutate(stage = "la")

# pupal dev (eclosion time)
ss_pu <- d23 %>%
  group_by(trt.type, minT) %>%
  summarise(avg.tt = mean(na.omit(jdate.eclose - jdate.pupa)),
            se.tt = sd(na.omit(jdate.eclose - jdate.pupa))/sqrt(length(na.omit(jdate.eclose - jdate.pupa))),
            avg.mass = mean(na.omit(mass.pupa)),
            se.mass = sd(na.omit(mass.pupa)/sqrt(length(na.omit(mass.pupa)))),) %>%
  mutate(stage = "pu")

# adult stats
ss_ad.sex <- dad %>%
  filter(!is.na(for.surv), pop == "lab") %>%
  group_by(#diet, # gets too parsed out w/ diet lol
           trt.type, minT, sex
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.surv - jdate.eclose)),
            se.tt = sd(na.omit(jdate.surv - jdate.eclose))/sqrt(length(na.omit(jdate.surv - jdate.eclose))),
            avg.mass = mean(na.omit(mass.eclose)),
            se.mass = sd(na.omit(mass.eclose)/sqrt(length(na.omit(mass.eclose)))),
            n = sum(na.omit(!is.na(jdate.eclose)))
            ) %>%
  mutate(stage = "ad")

ss_ad <- dad %>%
  filter(!is.na(for.surv), pop == "lab") %>%
  group_by(
           trt.type, minT
           ) %>%
  summarise(avg.tt = mean(na.omit(jdate.surv - jdate.eclose)),
            se.tt = sd(na.omit(jdate.surv - jdate.eclose))/sqrt(length(na.omit(jdate.surv - jdate.eclose))),
            avg.mass = mean(na.omit(mass.eclose)),
            se.mass = sd(na.omit(mass.eclose)/sqrt(length(na.omit(mass.eclose)))),
            n = sum(na.omit(!is.na(jdate.eclose)))
            ) %>%
  mutate(stage = "ad")
```

```{r create final dev df}
# combine dfs
ss_dev <- full_join(ss_la, ss_pu)
ss_dev <- full_join(ss_dev, ss_ad)
ss_dev <- full_join(ss_dev, ss_ad.sex)

ss_dev$stage <- factor(ss_dev$stage, levels = c("la", "pu", "ad"))

ss_dev.expt <- filter(ss_dev, trt.type == "expt")
ss_dev.ctrl <- filter(ss_dev, trt.type == "ctrl")
```

## b. stats prep

```{r final stats df}
# df for exptal bugs only (for modeling)
d23.expt <- filter(d23, trt.type == "expt")
```

```{r subsetting}
# new df for survival data: drops accidental deaths/na outcomes (0 = eclose, 1 = death)
d_surv <- d23.expt %>%
  rename(status = surv.outcome) %>%
  filter(status < 2)
```


# 4. fertility

calcs are copy/pasted/tweaked from `/2023/analyses_tents.Rmd`

## a. viz prep

```{r fixing calcs}
# determine actual # of ovipositing females
dtents <- dtents %>%
  mutate_at(c(6:14), ~replace_na(., 0)) %>% 
  group_by(id.tent) %>%
  mutate(jdate.rel = jdate - first(jdate)) %>%
  ungroup() %>%
  group_by(id.tent, trt.f, trt.m, trt.sex, jdate.rel) %>%
  arrange(jdate.rel, .by_group = TRUE) %>%
  mutate(#n.laying = n.curr.f + n.died.f, .after = "n.curr.f",
         n.ovi = dplyr::lag(n.curr.f, default = first(n.curr.f)) + n.died.f, .after = "n.curr.f") %>%
  ungroup()


# where n.ovipositing females = 0 but n.collected > 0, add the # eggs collected to the day before
# (or last date of ovipositing females)
# first, figure out the rows
cond <- (dtents$n.coll > 0 & dtents$n.ovi == 0)

ovifix <- dtents %>%
  mutate(index = as.numeric(rownames(.)), .before = 1) %>%
  subset(cond | lead(cond) | lead(cond, n=2)) #%>% View() 

negovi <- ovifix[ovifix$females.laying < 0, ]

lastovi <- ovifix[ovifix$n.curr.f > 0 & ovifix$n.ovi >= 1, ]

test2 <- c(negovi$index, lastovi$index)

# then, edit values
#summ_rate$n.coll[negovi$index] <- 0
dtents <- dtents %>% 
  mutate(n.coll = case_when(rownames(.) %in% negovi$index ~ 0,
                            rownames(.) %in% lastovi$index ~ n.coll + 1,
                            !(rownames(.) %in% c(negovi$index, lastovi$index)) ~ as.numeric(n.coll)
                            )) %>%
  select(-females.laying)

rm(cond, ovifix, negovi, lastovi)

# removing extra 0's
# dtents <- dtents %>%
#   filter(!(n.coll == 0 & n.ovi == 0))
  # tbh not sure why i did this in the old version. this drops a few #s here and there

```


ref: `summ_rates_all` calcs in `analyses_tents.Rmd`

```{r calc fecundity rates (# laid)}
# i dont need this lol

# # this calcs how # of tents, # eggs laid, # ovi f changes daily per trt
# ss_ovirates <- dtents %>%
#   group_by(trt.ref, trt.sex, trt.hs, jdate.rel) %>%
#   summarise(n.tents = n(), #idk why i use mutate in the orig lol
#          eggs.perovi = sum(n.coll)/sum(n.ovi),
#          eggsovi.pertent = eggs.perovi / n.tents) %>%
#   mutate_all(~ifelse(is.nan(.), NA, .))

```


ref: `summ_hatch` calcs in `analyses_tents.Rmd`

```{r calc fertility rates (# hatched)}
# want: # eggs laid per female (overall), prop hatched in tent
ss_hatchrates <- dtents %>%
  group_by(trt.ref, trt.sex, trt.hs) %>%
  summarise(n.tents = n_distinct(id.tent),
            n.laid = sum(n.coll, na.rm = TRUE),
            #se.laid = se(n.coll),
            n.fs = sum(n.new.f, na.rm = TRUE), # = total # fs in a tent
            n.laidf = round(n.laid/n.fs, digits = 2),
            #se.laidf = se(sum(n.coll, na.rm = TRUE)/sum(n.new.f)), # doesnt work
            se.laidf = sqrt(n.laidf*(1-n.laidf/n.laid)), # hmmmm.
            n.hatched = sum(n.tothatch, na.rm = TRUE), # needed downstream
            #se.hatched = se(n.tothatch),
            prop.hatched = n.hatched/n.laid,
            se.hatchprop = sqrt(prop.hatched*(1-prop.hatched)/n.laid),
            # prop.hatchf = prop.hatched/n.fs,
            # se.hatchfprop = sqrt(prop.hatchf*(1-prop.hatchf)/n.laid)
            )
```

## b. stats prep

```{r}
# calc avg eggs/F (fecundity) and overall prop hatch PER TENT
d_repro <- dtents %>%
  filter(trt.ref != "ctrl") %>%
  group_by(id.tent, trt.ref, trt.sex, trt.hs) %>%
  summarise(n.laid = sum(n.coll, na.rm = TRUE),
            n.fs = sum(n.new.f, na.rm = TRUE), # = total # fs in a tent
            avg.laidf = round(n.laid/n.fs, digits = 2), # fecundity stat
            log.fec = log(avg.laidf + 1),
            n.hatched = sum(n.tothatch, na.rm = TRUE),
            prop.hatched = n.hatched/n.laid, # fert stat
            log.fert = log(prop.hatched + 1) 
            )

# imo need log.fert and avg.laidf is ok... just looking at 0's. but should check qqplots later

d_win <- d_repro %>%
  filter(trt.ref == "win")
d_btwn <- d_repro %>%
  filter(trt.ref == "btwn")
```

# 5. remove extra dfs

```{r cleanup}
rm(backup_d23.all, backup_d23.init)

rm(dad, dtents)
```




