---
title: "260227 LB figs n stats"
date: "2026-02-18"
---

# preamble and loading

figs for LB/committee update
  

```{r load libs and utils, warning=FALSE}
library(tidyverse)
library(patchwork)
library(survival)
library(survminer)

here::i_am("outputs/2026-02_LB/results.Rmd")
library(here)

source(here("_ntw/scripts/utils/wrangle.R"))
source(here("_ntw/scripts/utils/analysis.R"))
source(here("_ntw/scripts/utils/analysis-stats.R"))

source(here("_tdt/scripts/utils/wrangle.R"))
```

```{r load data, warning=FALSE}
source(here("_ntw/scripts/R/tidy-dev.R"))
dfs_ntw <- list(
  wide = dfs_tidy$wide,
  long = dfs_tidy$long
) %>%
  lapply(., \(x){
    x %>%
    mutate(across(c(starts_with("trt"), "year"),
                  as.factor))
  })

source(here("_tdt/scripts/R/tidy.R"))
dfs_tdt <- list(
  wide = dfs_tidy$wide,
  long = dfs_tidy$long
) %>%
  lapply(., \(x){
    x %>%
      filter(status %in% c(0, 1)) %>% # drop pmc bugs
      mutate(across(starts_with("trt"), as.factor))
  })

rm(dfs_tidy)
```

# # 2x2 figs
- pupation big 3

## wrangle

```{r}
FilterNTW1 <- function(x){
  x %>%
    filter(year == 2023,
           pop == "lab" & diet == "LD",
           is.ntw %in% c(0, 2))
}

# pivot
dfs_ntw1 <- list(
  ss_dev = dfs_ntw$long %>%
    FilterNTW1() %>%
    filter(instar == "pupa") %>% 
    mutate(mass = mass/1000,
           devrate = 1/tt,
           growthrate = mass/tt) %>%
    group_by(trt, trt.meanT, trt.flucT) %>% 
    summarise(n = n(),
              avg.mass = mean(mass, na.rm = TRUE),
              se.mass = se(mass),
              avg.devrate = mean(devrate, na.rm = TRUE),
              se.devrate = se(devrate),
              avg.growthrate = mean(growthrate, na.rm = TRUE),
              se.growthrate = se(growthrate)),
  
  ss_surv = dfs_ntw$wide %>%
    FilterNTW1() %>%
    group_by(trt, trt.meanT, trt.flucT) %>%
    summarise(n.surv = sum(is.pup == 1),
              p.surv = n.surv/sum(is.pup %in% c(0, 1)),
              se.surv = se.prop(p.surv, n.surv))
)
```

## big 3

```{r}
.plotfn <- function(x){
  x %>%
    #filter(trt.flucT == 0) %>%
    ggplot(aes(x = trt.meanT,
               lty = trt.flucT, shape = trt.flucT,
               group = trt.flucT)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(width = 0.2, lty = "solid",
                  show.legend = FALSE)
}

p1 <- dfs_ntw1$ss_surv %>%
  .plotfn() +
  aes(y = p.surv,
      ymax = p.surv + se.surv,
      ymin = p.surv - se.surv) +
  ylim(c(0, 1)) +
  labs(y = "proportion survived to pupation") 

p2 <- dfs_ntw1$ss_dev %>%
  .plotfn() +
  aes(y = avg.mass,
      ymax = avg.mass + se.mass,
      ymin = avg.mass - se.mass) +
  labs(y = "pupal mass (g)") +
  ylim(c(3.9, 6.7))

p3 <- dfs_ntw1$ss_dev %>%
  .plotfn() +
  aes(y = avg.devrate,
      ymax = avg.devrate + se.devrate,
      ymin = avg.devrate - se.devrate) +
  labs(y = "development rate (1/days)")

(p1 + p2 + p3) +
  plot_layout(axes = "collect", 
              guides = "collect",
              widths = c(1, 1, 1)
              ) &
  labs(x = "mean temperature (°C)",
       lty = "fluctuation (°C)", shape = "fluctuation (°C)") &
  scale_shape_manual(labels = c("±0", "±7"), values = c(19, 1)) &
  scale_linetype_discrete(labels = c("±0", "±7")) &
  theme_bw() &
  theme(legend.position = "bottom")

```

## tpc-ish

- aka rly just the growth rate
- (to show effects of fluct)

```{r}
dfs_ntw1$ss_dev %>%
  ggplot(aes(y = avg.growthrate,
             x = trt.meanT,
             shape = trt.flucT, lty = trt.flucT,
             group = trt.flucT)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.growthrate + se.growthrate,
                    ymin = avg.growthrate - se.growthrate),
                width = 0.075, lty = "solid") +
  scale_shape_manual(labels = c("±0", "±7"), values = c(19, 1)) +
  scale_linetype_discrete(labels = c("±0", "±7")) +
  labs(y = "growth rate (g/day)",
       x = "mean temperature (°C)",
       shape = "fluctuation (°C)", lty = "fluctuation (°C)") +
  theme_bw()
```


# # ntw figs

## wrangle

```{r}
FilterNTW2 <- function(x){
  x %>%
    FilterForNTWTrts() %>%
    filter(!(pop == "lab" & diet == "TB"),
           pop != "col",
           is.ntw %in% c(1,2))
}

dfs_ntw2 <- list(
  ss_dev = dfs_ntw$long %>%
    FilterNTW2() %>%
    filter(instar == "pupa") %>%
    mutate(mass = mass/1000,
           devrate = 1/tt) %>%
    group_by(trt, trt.type, pop) %>%
    summarise(n = n(),
              avg.mass = mean(mass, na.rm = TRUE),
              se.mass = se(mass),
              avg.devrate = mean(devrate, na.rm = TRUE),
              se.devrate = se(devrate)),
  
  ss_surv = dfs_ntw$wide %>%
    FilterNTW2() %>%
    group_by(trt, trt.type, pop) %>%
    summarise(n = n(),
              n2 = sum(is.pup %in% c(0, 1)),
              n.surv = sum(is.pup == 1, na.rm = TRUE),
              p.surv = n.surv/sum(is.pup %in% c(0, 1)),
              se.surv = se.prop(p.surv, n.surv))
)
```

## big 3

```{r}
.plotfn <- function(x){
  x %>%
    ggplot(aes(x = trt, lty = pop, shape = pop,
             group = interaction(pop, trt.type))) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(width = 0.1, lty= "solid", show.legend = FALSE) +
    facet_grid(~ trt.type, scales = "free_x", space = "free_x",
               labeller = as_labeller(c(`ctrl` = "26±0°C (control)",
                                        `expt` = "40-X°C (nighttime warming)"))
               ) +
    theme_bw() +
    scale_x_discrete(labels = c("260" = "26", "419" = "19", "426" = "26", "433" = "33")) +
    scale_linetype_manual(values = c("field" = "dashed", "lab" = "solid")) +
    scale_shape_manual(values = c("field" = 1, "lab" = 19)) +
    labs(x = "minimum temperature (°C)",
         lty = "population", shape = "population")
}

dfs_ntw2$ss_surv %>%
  .plotfn() +
  aes(y = p.surv,
      ymax = p.surv + se.surv,
      ymin = p.surv - se.surv) +
  ylim(c(0, 1)) +
  labs(y = "proportion survived to pupation")

dfs_ntw2$ss_dev %>%
  .plotfn() +
  aes(y = avg.mass,
      ymax = avg.mass + se.mass,
      ymin = avg.mass - se.mass) +
  ylim(c(0, 6.25)) +
  labs(y = "pupal mass (g)") 

dfs_ntw2$ss_dev %>%
  .plotfn() +
  aes(y = avg.devrate,  
      ymax = avg.devrate + se.devrate,
      ymin = avg.devrate - se.devrate) +
  labs(y = "development rate (1/days)")
```




# # tdt figs

- const ctrls survival to 3rd (probably need to use KM curve or sth.. bc 4ths dont even survive to 3rd lol)
- const ctrls dev to 4th

## kms

### ctrls + temp
- how control bugs fared

```{r}
km_df <- dfs_tdt$long %>%
  filter(subset == "fromenter" & trt.duration == 0)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df) + 
  labs(caption = "from enter")
```


### 40s + recovery duration
- how 0, 24, 48h hot bugs fared

```{r}
# from enter
km_df <- dfs_tdt$long %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df) + 
  labs(caption = "from enter")
  
# from return
km_df <- dfs_tdt$long %>%
  filter(subset == "fromreturn" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df) +
  labs(caption = "from return")

# another flavor (being ctrl vs expt)
km_df <- dfs_tdt$long %>%
  filter(subset == "fromreturn" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.type, data = km_df)

ggsurvplot(fit = km_fit, data = km_df) +
  labs(caption = "from return")

```

### recovery + temp
```{r}
# from enter
km_df <- dfs_tdt$long %>%
  filter(subset == "fromenter" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.recover") + 
  labs(caption = "from enter")
  
# from return
km_df <- dfs_tdt$long %>%
  filter(subset == "fromreturn" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.recover") +
  labs(caption = "from return")
```

### recovery + duration
```{r}
# from enter
km_df <- dfs_tdt$long %>%
  filter(subset == "fromenter" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") + 
  labs(caption = "from enter") +
    scale_color_hue(direction = -1)
  
# from return
km_df <- dfs_tdt$long %>%
  filter(subset == "fromreturn" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") +
  labs(caption = "from return") +
    scale_color_hue(direction = -1)
```

## time to exit

```{r}
# wrangle
dfs_tdt <- list_modify(
  dfs_tdt,
  ss_surv = dfs_tdt$wide %>%
  #filter(trt.duration %in% c(24, 48)) %>%
    filter(fate == "died" & trt.enter == 40) %>%
    mutate(exitfromreturn = dh.exit - dh.return) %>%
  group_by(across(starts_with("trt"))) %>%
  summarise(avg.ttexit.ret = mean(exitfromreturn, na.rm = TRUE),
            se.ttexit.ret = se(exitfromreturn),
            avg.ttexit.ent = mean(dt.exit, na.rm = TRUE),
            se.ttexit.ent = se(dt.exit)
            )
)
```

### from enter

```{r eval=FALSE}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.ttexit.ent,
               x = trt.recover)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.ttexit.ent + se.ttexit.ent,
                      ymin = avg.ttexit.ent - se.ttexit.ent),
                  width = 0.25, lty = "solid") +
    labs(y = "time to death (h)",
         x = "recovery temperature (°C)",
         #caption = "from enter"
         ) +
    theme_bw()
}

## facet by duration
dfs_tdt$ss_surv %>%
  .plotfn() +
    facet_grid(~ trt.duration, scales = "free_x", space = "free_x")


## not faceted
dfs_tdt$ss_surv %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)")
```
### from return

```{r}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.ttexit.ret,
               x = trt.recover)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.ttexit.ret + se.ttexit.ret,
                      ymin = avg.ttexit.ret - se.ttexit.ret),
                  width = 0.25, lty = "solid") +
    labs(y = "time to death (h)",
         x = "recovery temperature (°C)",
         caption = "from return"
         ) +
    theme_bw()
}

## facet by duration
dfs_tdt$ss_surv %>%
  .plotfn() +
    facet_grid(~ trt.duration, scales = "free_x", space = "free_x")

dfs_tdt$ss_surv %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_grid(~ trt.type, scales = "free_x", space = "free_x",
             labeller = as_labeller(c(`ctrl` = "40°C constant",
                                      `expt` = "40°C + recovery"))) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)")


## not faceted
dfs_tdt$ss_surv %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)")
```


# # stats
(do separately here bc the stats df is differnt from the fig df)

## 2x2
```{r}
# df prep, exclude culled
df <- dfs_ntw$wide %>% 
  filter(is.pup != 2) %>%
  FilterNTW1() %>%
  mutate(tt.pupa = jdate.pupa - jdate.hatch)

mods_ntw1 <- list(
  surv = glm(is.pup ~ trt.meanT*trt.flucT, 
             family = "binomial", data = df),
  dev = lm(tt.pupa ~ trt.meanT*trt.flucT,
           data = df),
  mass = lm(mass.pupa ~ trt.meanT*trt.flucT*sex,
           data = df %>% filter(sex %in% c("m", "f")))
)

mods_ntw1 %>% GetModelResults()
```


## ntw dev
```{r ntw df prep}
# hot bugs only; exclude culled
dfs_ntw2 <- list_modify(
  dfs_ntw,
  df = dfs_ntw$wide %>%
    FilterNTW2() %>%
    filter(trt != 260,
           is.pup != 2) %>%
    mutate(tt.pupa = jdate.pupa - jdate.hatch)
)
```


```{r}
# survival
mods_ntw2 <- list(
  surv_all = glm(is.pup ~ year*pop*trt.minT,
             family = "binomial", data = dfs_ntw2$df),
  surv_noyr = glm(is.pup ~ pop*trt.minT,
                 family = "binomial", data = dfs_ntw2$df)
)

mods_ntw2 %>%
  GetModelResults()
```

```{r}
# pupal mass
mods_ntw2 <- list(
  mass_all = lm(mass.pupa ~ year*pop*trt.minT*sex,
                data = dfs_ntw2$df),
  mass_noyr = lm(mass.pupa ~ pop*trt.minT*sex,
                 data = dfs_ntw2$df)
)

mods_ntw2 %>%
  GetModelResults()
```

```{r}
# dev rate
mods_ntw2 <- list(
  ttpup_all = lm(tt.pupa ~ year*pop*trt.minT,
                data = dfs_ntw2$df),
  ttpup_noyr = lm(tt.pupa ~ pop*trt.minT,
                 data = dfs_ntw2$df)
)

mods_ntw2 %>%
  GetModelResults()
```


## recovery

- mean survival time anovas
- comparing hot vs not (different Z's):
  - if hot, surv? `time*temp` (from enter)
  - if hot and surv to 3rd, uhhh. something with return mass?


```{r}
# df prep
dfs_tdt <- list_modify(
  dfs_tdt,
  df = dfs_tdt$wide %>%
    mutate(tt.exit = dh.exit - dh.recover)
)
```

```{r}
# hot bugs only
mods_rec <- list(
  ifctrl = lm(tt.exit ~ trt.type, data = dfs_tdt$df),
  rectrt = lm(tt.exit ~ trt.duration*trt.recover,
              data = dfs_tdt$df %>% filter(trt.duration != 0))
)

mods_rec %>% GetModelResults()
```

