---
title: "03-stats/ntw-compare"
date: "2025-02-19"
---

# preamble

dates in this doc are for when i first did this stuff bc a lot of the modeling is based off inputs from joel mtgs on/before corresponding dates LOL

## loading

```{r}
here::i_am("2024/03-stats/compare-dev.Rmd")
library(here)

library(tidyverse)
#library(lme4) # for lmer.. for tent stuff
library(MuMIn) # for survival nested model comparisons

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "compare-dev_p.R"))
rm(filter.2pops, se)
```

## df prep

```{r general exptal bug filter}
# focusing on exptal bugs only that weren't lab on TB
ntw_expt <- ntw_wide %>%
  filter(!(diet == "TB" & pop == "lab") & trt.type == "expt") %>%
  mutate(dmass = mass.pupa - mass.eclose,
         rate.pup = mass.pupa/tt.pupa)
```

```{r other subsets}
# omit accidental deaths (so either pmd, pup, LPI)
ntw_expt.surv <- ntw_expt %>%
  filter(is.pup != 2)

# omit unsexed pupa
ntw_expt.sex <- ntw_expt %>%
  filter(!is.na(sex))
```

```{r mumin subsets}
# bc apparently this pkg cant handle NAs bc "all models must be fit to the same data" 
# when comparing fits, so subset:
  # all surv to pups
  # all surv to ecs

# select relevant variables
ntw_mumin.dfs <- list(df = select(ntw_expt.surv, c("year", "pop", "minT", "is.pup", 
                                                   "is.sup", "sup", "is.ec")))

ntw_mumin.dfs <- c(ntw_mumin.dfs, 
                   list(pupd = filter(ntw_mumin.dfs$df, is.pup == 1)),
                   list(ecd = filter(ntw_mumin.dfs$df, is.ec == 1)))
# colSums(is.na(ntw_mumin.dfs$pupd)) > 0 # check
```

# pupal survival

-   what to do w/ modeling from 250213 joel mtg:
    -   look into nested model comparisons (try MuMin first, then AICcmodaug). compare the interaction (temp, pop, year), then additive, then simplest
-   qns:
    -   check surv to pup
    -   of the surv, who's sup?
    -   pup to eclosion?
-   mumin pdf reading notes
    -   arm.glm?

## 250213

scrappy attempts pre 250213 joel mtg, ignore these for the most part

```{r}
mod.surv_minT <- glm(is.pup ~ minT*pop, data = ntw_expt.surv, family = "binomial") # need to add `year`

mod.surv_meanT <- glm(is.pup ~ meanT*pop, data = ntw_expt.surv, family = "binomial")
  # ignore meanT bc it's confounded w minT

anova(mod.surv_minT, mod.surv_meanT, test = "Chisq") 
```

## 250226 mumining

actual modeling attempts w/ MuMIn and things. see the `mumin` df prep code chunk

roadmap: 

1. generate set of candidate models (full interaction, full additive, null) 
2. ~~use mumin::nested?~~

refs: 

- <https://sites.google.com/site/rforfishandwildlifegrads/home/mumin_usage_examples> 
- <https://rdrr.io/cran/MuMIn/man/nested.html>

so when i google "model.sel vs dredge mumin" apparently i should `dredge` first, pick some nice candidates base off dredge results, then throw them into `model.sel`, see also <https://uoftcoders.github.io/rcourse/lec09-model-selection.html#a_more_realistic_example>

```{r dredge no sup}
# need to remove nas & have model fitting break when fitting to NA data to compare fits correctly,
# so either the global option options(na.action = "na.fail") before dredge() or 
# set that option in the global model function call

# remove nas
ntw_mumin.dfs <- c(ntw_mumin.dfs, 
                   list(nona = na.omit(ntw_mumin.dfs$df[1:4])))

out.psurv <- list(dredge = dredge(glm(is.pup ~ year*pop*minT, 
                                      data = ntw_mumin.dfs$nona, family = "binomial", na.action = na.fail), 
                                  #extra = c("R^2", F = function(x)
                                  #summary(x)$fstatistic[[1]])
                                  ))
out.psurv$dredge

out.psurv <- c(out.psurv,
               dredge.out = list(get.models(out.psurv$dredge, subset = weight > 0.01)))

out.psurv$dredge.out[["64"]]
out.psurv$dredge.out[["128"]]
```

```{r manual selection no sup}
# fit candidate models for pup surv (looking generally at sup, ignoring is.sup stuff for now)
mod.psurv <- list(fullint = glm(is.pup ~ year*pop*minT, 
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  n2way = glm(is.pup ~ (year + pop + minT)^2,
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  fulladd = glm(is.pup ~ year + pop + minT, 
                                data = ntw_mumin.dfs$df, family = "binomial"),
                  null = glm(is.pup ~ 1, 
                                data = ntw_mumin.dfs$df, family = "binomial"))

# picking models..
out.psurv <- c(out.psurv,
               modsel = list(model.sel(mod.psurv$fullint, mod.psurv$n2way, 
                                       mod.psurv$fulladd, mod.psurv$null)))
out.psurv$modsel

nested(out.psurv$mod.sel)
anova(mod.psurv$fullint, test = "Chisq")
anova(mod.psurv$n2way, test = "Chisq")
anova(mod.psurv$fullint, mod.psurv$n2way, test = "Chisq") # no sig diff!
```

idrk any of this means so stopping here for today lol

post-mtg: data gives support that the 2-way interactions matter while the 3-way is kinda meh but not a bad model either!

## 250306 sups & carryover

```{r}
# if you survived, were you sup?

ntw_mumin.dfs <- c(ntw_mumin.dfs, 
                   list(nona.sup = na.omit(ntw_mumin.dfs$pupd[1:5])))

out.psurv <- c(out.psurv, 
               dredge.sup = list(dredge(glm(is.sup ~ year*pop*minT, 
                                            data = ntw_mumin.dfs$nona.sup, family = "binomial", na.action = na.fail)))
               )

out.psurv$dredge.sup

out.psurv <- c(out.psurv,
               dredge.outsup = list(get.models(out.psurv$dredge.sup, subset = weight > 0.05)))

out.psurv$dredge.outsup

out.psurv <- c(out.psurv,
               modsel.sup = list(model.sel(out.psurv$dredge.outsup[["56"]], out.psurv$dredge.outsup[["64"]],
                                           out.psurv$dredge.outsup[["48"]], out.psurv$dredge.outsup[["128"]])))

out.psurv$modsel.sup


```

## 250512 random yr feas followup

try to model `year` as a random effect per feasibility suggestions

```{r}
mod <- glmer(is.pup ~ pop*minT + (1|year), data = ntw_mumin.dfs$df, family = "binomial", na.action = "na.fail")

summary(mod)
dredge(mod)


# Fixed effects:
#               Estimate Std. Error z value Pr(>|z|)    
# (Intercept)    -1.5555     0.3407  -4.566 4.98e-06 ***
# poplab          1.4851     0.3246   4.575 4.76e-06 ***
# minT26          0.7202     0.2848   2.529   0.0114 *  
# minT33          0.5656     0.3005   1.882   0.0598 .  
# poplab:minT26   0.6559     0.4468   1.468   0.1421    
# poplab:minT33  -0.7972     0.4279  -1.863   0.0625 .

# Model selection table 
#     (Int) mnT pop mnT:pop df   logLik  AICc delta weight
# 8 -1.5550   +   +       +  7 -425.006 864.2  0.00  0.982
# 4 -1.5390   +   +          5 -431.024 872.1  7.97  0.018

# try reordering minTs relative to 26...
mod2 <- ntw_mumin.dfs$df %>%
  mutate(minT = factor(minT, levels = c(33, 26, 19)),
         pop = factor(pop, levels = c("lab", "field"))) %>%
  glmer(is.pup ~ pop*minT + (1|year), data = ., family = "binomial", na.action = "na.fail")

summary(mod2)
# when lab is ref,
  # when 26 is ref, 19 = * and 33 = ns
  # when 33 is ref, 26 = ns, 19 = *
# (lots of stuff moves around depending on whos the ref lol)
```

so pop definitely matters overall... and temp matters sorta depending on whos the ref pop/temp

```{r}
# compare rand yr to fixed yr (in the 2way)

df <- ntw_mumin.dfs$df #%>%
  # mutate(minT = factor(minT, levels = c(33, 26, 19)),
  #        pop = factor(pop, levels = c("lab", "field")))

mod <- lme4::glmer(is.pup ~ pop*minT + (1|year), data = df, family = "binomial", na.action = "na.fail")
mod2 <- glm(is.pup ~ (year + pop + minT)^2, data = df, family = "binomial", na.action = "na.fail")

anova(mod, mod2, test = "Chisq")
# they are def different models lol

model.sel(mod, mod2)
# glmer has lower negative LL tho glm has lower AIC
```

soooo tldr:
- not exactly clear which model is better (yr as rand vs yr as fixed?)... and that the glm/glmer are different
- predictor significance varies in the glmer depending on the reference level (of `pop` and `minT`)



# pupal mass

## 250213

what affects pupal mass? Tmin, sex, year, pop?

(numbering is just based on when i ran them lol)

```{r}
# 01: test the full complex int mod vs the full additive model
# mods.pmass <- list(int_01 = lm(mass.pupa ~ year*minT*pop*sex, data = ntw_expt.sex),
#                    add_01 = lm(mass.pupa ~ year + minT + pop + sex, data = ntw_expt.sex))
# names(mods.pmass)

mod.pmass_int01 <- lm(mass.pupa ~ year*minT*pop*sex, data = ntw_expt.sex)
mod.pmass_add01 <- lm(mass.pupa ~ year + minT + pop + sex, data = ntw_expt.sex)

anova(mod.pmass_int01, mod.pmass_add01, test = "Chisq") # no sig diff btwn add/int: add is better

# just look at the variables
anova(mod.pmass_int01)
anova(mod.pmass_add01)
```

```{r}
# 02: look at int mod again
# based on plotting the pupal masses by temp/sex/yr/pop. which ints matter?

# drop year
mod.pmass_int02 <- lm(mass.pupa ~ minT*pop*sex, data = ntw_expt.sex) 

anova(mod.pmass_int01, mod.pmass_int02, test = "Chisq") # temp*pop*sex is better

anova(mod.pmass_int02)

# drop sex
mod.pmass_int03 <- lm(mass.pupa ~ minT*pop, data = ntw_expt.sex)

anova(mod.pmass_int02, mod.pmass_int03, test = "Chisq") # no sig diff, so keep sex?

# which terms matter?
anova(mod.pmass_int02)
```

-   `mass.pupa ~ minT*pop*sex` seems good, with the following sig ints:

    -   minT; pop = `***`

    -   sex, minT:pop = `*`

```{r}
# 03: look into the additive mods some more
# drop year
mod.pmass_add02 <- lm(mass.pupa ~ minT + pop + sex, data = ntw_expt.sex)

anova(mod.pmass_int02, mod.pmass_add02, test = "Chisq") # no sig diff
```

idt this one is right lol bc year def makes a difference

## 250220

per 250220 joel mtg: try the 2 way interactions only....

```{r}
# mods.pmass <- list(mods.pmass,
#                    n2way_01 =lm(mass.pupa ~ (year + minT + pop + sex)^2, data = ntw_expt.sex))
  # ONE DAY... we'll figure it out shoving models into lists lol

mod.pmass_n2way01 <- lm(mass.pupa ~ (year + minT + pop + sex)^2, data = ntw_expt.sex)

anova(mod.pmass_n2way01)
# *** yr, minT, pop
# * sex, minT:pop
```

# time to pup

## 250213

```{r}
# 01 full models
mod.pdev_int01 <- lm(tt.pupa ~ year*minT*pop*sex, data = ntw_expt)
mod.pdev_add01 <- lm(tt.pupa ~ year+minT+pop+sex, data = ntw_expt)

anova(mod.pdev_int01, mod.pdev_add01, test = "Chisq")

anova(mod.pdev_int01)
anova(mod.pdev_add01)
```

intmod is better?

## 250220

per 250220 joel mtg, try 2 way interactions

```{r}
mod.pdev_n2way01 <- lm(tt.pupa ~ (year + minT + pop + sex)^2, data = ntw_expt)

anova(mod.pdev_n2way01)
# *** = yr, minT, pop, yr:minT
# * = yr:pop

# sex isn't rly relevant here yet so omit
mod.pdev_n2way02 <- lm(tt.pupa ~ (year + minT + pop)^2, data = ntw_expt)
anova(mod.pdev_n2way02)
# *** = yr, minT, pop
# ** = year:minT
# * = yr:pop


# yr matters, is it bc of the field bug timing?
mod.pdev_n2way03a <- lm(tt.pupa ~ (year*minT), data = filter(ntw_expt, pop == "lab"))
mod.pdev_n2way03b <- lm(tt.pupa ~ (year*minT), data = filter(ntw_expt, pop == "field"))
anova(mod.pdev_n2way03a)
anova(mod.pdev_n2way03b)
# actually, it's the lab bugs lol. prob bc we keep selecting for faster bugs in colony

```
