---
title: "stats/entsoc.Rmd"
date: "2025-02-20"
---

# setup

```{r}
here::i_am("2024/03-stats/entsoc.Rmd")
library(here)

library(tidyverse)
library(lme4)

source(here::here("set-paths.R"))
source(here::here(bin_paths24$wrangle, "entsoc.R"))
```

# pupal survival

```{r surv stats}
mod_surv <- glm(status ~ as.factor(minT), data = d_surv, family = "binomial")
mod_surv2 <- glm(status ~ as.factor(minT)*sex, data = d_surv, family = "binomial")

summary(mod_surv)
anova(mod_surv, test = "Chisq") # minT ***
```


# pupal, adult mass

```{r pupal mass stats}
# trying other possible influencing factors first
mod_pmass1 <- lm(mass.pupa ~ as.factor(minT)*sex, data = d23)
mod_pmass2 <- lm(mass.pupa ~ as.factor(minT), data = d23) # better per AIC

anova(mod_pmass1) # minT ***, sex ns (0.19), int ns (0.42)
anova(mod_pmass2) # minT ***

AIC(mod_pmass1) # 351.1274
AIC(mod_pmass2) # 348.6995
```

```{r adult mass stats}
mod_amass1 <- lm(mass.eclose ~ as.factor(minT)*sex, data = d23) # better per AIC
mod_amass2 <- lm(mass.eclose ~ as.factor(minT), data = d23)

anova(mod_amass1) # minT ***, sex **
anova(mod_amass2) # minT ***

AIC(mod_amass1) # 214.7664
AIC(mod_amass2) # 220.7273
```



# larval dev, adult longevity

```{r larval dtime stats}
mod_ltime1 <- lm((jdate.pupa - jdate.hatch) ~ as.factor(minT)*as.factor(sup), data = d23) # better per AIC
mod_ltime2 <- lm((jdate.pupa - jdate.hatch) ~ as.factor(minT), data = d23)

anova(mod_ltime1) # minT ***, sup ***, int ns (0.75)
anova(mod_ltime2) # minT ***

AIC(mod_ltime1) # 480.7616
AIC(mod_ltime2) # 576.1871
```


```{r adult dtime stats}
mod_atime1 <- lm((jdate.surv - jdate.eclose) ~ as.factor(minT)*sex, data = d23) 
mod_atime2 <- lm((jdate.surv - jdate.eclose) ~ as.factor(minT), data = d23) # better per AIC

anova(mod_atime1) # minT ns (0.55), sex ns (0.81)
anova(mod_atime2) # minT ns

AIC(mod_atime1) # 158.7356
AIC(mod_atime2) # 156.8043
```

# fertility

## - btwn trts
```{r modeling btwn}
# no random effs bc of tent grouping lol

# fecundity 
mod_bfec <- lm(avg.laidf ~ trt.hs*trt.sex, data = d_btwn)
mod_bfec.log <- lm(log.fec ~ trt.hs*trt.sex, data = d_btwn)

# fertility
mod_bfert <- lm(prop.hatched ~ trt.hs*trt.sex, data = d_btwn)
mod_bfert.log <- lm(log.fert ~ trt.hs*trt.sex, data = d_btwn)
```

```{r checking btwn mods}
anova(mod_bfec) # trt (0.95), sex (.17), int (0.82) ns 
anova(mod_bfec.log) # trt (0.79), sex (.97), int (0.64) ns 
AIC(mod_bfec) # 177.0551
AIC(mod_bfec.log) # 61.76643

anova(mod_bfert) 
anova(mod_bfert.log)  
AIC(mod_bfert) # better?? not rly tho - they are pretty close anyway lol
AIC(mod_bfert.log)
```

```{r qq btwn mods}
# fecundity
qqnorm(resid(mod_bfec), main = "QQ btwn fecundity")
qqline(resid(mod_bfec))

qqnorm(resid(mod_bfec.log), main = "QQ log(btwn fecundity)") # this is better (sorta)
qqline(resid(mod_bfec.log))

# fertility 
# they both rly suck bc of a shitton of 0s lol
qqnorm(resid(mod_bfert), main = "QQ btwn fertility")
qqline(resid(mod_bfert))

qqnorm(resid(mod_bfert.log), main = "QQ log(btwn fertility)")
qqline(resid(mod_bfert.log))
```


Analysis of Variance Table

Response: avg.laidf
               Df Sum Sq Mean Sq F value Pr(>F)
trt.hs          1     48    48.0  0.0039 0.9517
trt.sex         1  27487 27487.0  2.2042 0.1685
trt.hs:trt.sex  1    680   680.4  0.0546 0.8200
Residuals      10 124702 12470.2               
Analysis of Variance Table

Response: log.fec
               Df Sum Sq Mean Sq F value Pr(>F)
trt.hs          1  0.242  0.2421  0.0732 0.7922
trt.sex         1  0.005  0.0052  0.0016 0.9693
trt.hs:trt.sex  1  0.784  0.7844  0.2372 0.6368
Residuals      10 33.075  3.3075               
[1] 177.0551
[1] 61.76643


Analysis of Variance Table

Response: prop.hatched
               Df   Sum Sq   Mean Sq F value Pr(>F)
trt.hs          1 0.002134 0.0021337  0.3753 0.5553
trt.sex         1 0.007761 0.0077610  1.3653 0.2726
trt.hs:trt.sex  1 0.000566 0.0005664  0.0996 0.7595
Residuals       9 0.051161 0.0056845               
Analysis of Variance Table

Response: log.fert
               Df   Sum Sq   Mean Sq F value Pr(>F)
trt.hs          1 0.001633 0.0016333  0.3449 0.5715
trt.sex         1 0.006554 0.0065543  1.3841 0.2696
trt.hs:trt.sex  1 0.000402 0.0004023  0.0849 0.7773
Residuals       9 0.042620 0.0047355               
[1] -25.0981
[1] -27.47261



## - win trts
```{r modeling win}
# fecundity 
mod_wfec.log <- lm(log.fec ~ trt.hs, data = d_win)

# fertility
mod_wfert <- lm(prop.hatched ~ trt.hs, data = d_win)
```


```{r checking win mods}
anova(mod_wfec.log) # trt (0.94) ns

anova(mod_wfert) # trt (0.21) ns

```


```{r qqplots win}
# fecundity
qqnorm(resid(mod_wfec.log)) 
qqline(resid(mod_wfec.log))

# fertility 
qqnorm(resid(mod_wfert))
qqline(resid(mod_wfert))

```
