---
title: "tdt viz"
date: "2026-01-06"
---

generate tdt viz

TODO
- before
  - import data
  - import utils
  - make subsets(?)
- analysis
  - all the kms

# preamble

```{r}
# pathfinding
here::i_am("_tdt/scripts/analysis-viz.Rmd")
library(here)

# libs and utils
library(tidyverse)
library(survival) # for Surv 
library(survminer) # for ggsurvplot

source(here("_tdt/scripts/utils/wrangle.R"))
source(here("_tdt/scripts/utils/analysis-viz.R"))
```

```{r data prep}
# import data
source(here("_tdt/scripts/R/tidy.R"))

# format col types for easier plotting...?

# drop pmc bugs
dfs_tidy <- dfs_tidy %>% lapply(., \(x){
  x %>% 
    filter(status %in% c(0, 1)) %>%
    mutate(across(starts_with("trt"), as.factor))
})
```
# ## kms/survival curves

# kms ctrls

## by rec temp
results
- no major diffs in KMs for 40 ctrls across cohorts

```{r}
# by trt
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.duration == 0)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "cohort"
           ) %++%
  AddKMLines(km_df) + 
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) + 
  labs(caption = "rec duration = 0h")
```

## by cohort
```{r}
# by cohort
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.duration == 0,
         trt.recover == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ cohort, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "trt.recover"
           ) %++%
  AddKMLines(km_df) +
  labs(caption = "rec 0h @ 40C")

# looking at the 40C grp only -- no major diffs
```


# kms exptal
## by rec temp
```{r}
# from entering
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40) # ohh this is where im hitting issues w the subsetting...
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "cohort"
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.recover, 
             trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = enter")
```

```{r}
# from return
km_df <- dfs_tidy$long %>%
  filter(subset == "fromreturn" & trt.enter == 40) 
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "cohort"
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.return, 
             trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = return to 40C")
```

```{r}
# for STOR consult: differences in surv curves depending on where you start
# km_df <-
#   list(fromenter = dfs_tidy$long %>% filter(trt.enter == 40 & subset == "fromenter"),
#        fromreturn = dfs_tidy$long %>% filter(trt.enter == 40 & subset == "fromreturn"))
km_df <- dfs_tidy$long %>% filter(trt.enter == 40, subset != "fromrecover")
km_fit <- surv_fit(Surv(tt.exit, status) ~ trt.recover, data = km_df)

ggsurvplot_facet(fit = km_fit, data = km_df,
                 facet.by = c("subset", "trt.duration")) +
  AddKMLines(df = filter(km_df, subset == "fromenter"), dt.enter, trt.duration, subset) +
  AddKMLines(df = filter(km_df, subset == "fromreturn"), dt.return, trt.duration, subset) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "(varying t0)")


```



## by rec dur
```{r}
# effect of trt.duration
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "cohort"
           #facet.by = "trt.recover"
           ) %++%
  AddKMLines(km_df, dt.recover, cohort
             )
```
## by cohort
```{r}
# effect of cohort
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ cohort, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "trt.recover",
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.recover, trt.duration
             )
```

# ## development stuff

# avg time to death

(copied from from 2026-02 LB `results.Rmd`)

## from enter
```{r}
# from enter
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") + 
  labs(caption = "from enter") +
    scale_color_hue(direction = -1)
  
# from return
km_df <- dfs_tidy$long %>%
  filter(subset == "fromreturn" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") +
  labs(caption = "from return") +
    scale_color_hue(direction = -1)
```


## from return


# surv props

## overall
- using simulated 40C data

```{r}
# % bugs dying at each timept
# (N gets reset to the current timept)
dfs_tidy$widesim %>%
  CalcSurvProps(., trt.recover, trt.duration) %>% View()
  ggplot(aes(y = p, x = timept,
             fill = status)) +
  geom_col() +
  facet_grid(trt.duration ~ trt.recover, drop = TRUE) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(caption = "N = total bugs at start of timept")
  
# running count of bugs thru expt
dfs_tidy$widesim %>%
  CalcSurvProps(., trt.recover, trt.duration) %>% 
  #filter(status == "surv") %>%
  ggplot(aes(y = n.initial, x = timept,
             group = interaction(trt.recover, trt.duration),
             )) +
  geom_point() + geom_line() +
  facet_grid(trt.duration ~ trt.recover, drop = TRUE) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(caption = "N = total bugs at enter")
```

# development

(futzing around w this for 2026-02 LB)

## trajectories of cold ctrls

```{r}
dfs_tidy$wide %>%
  CalcDevSS() %>%
  filter(trt.duration == 0,
                instar %in% c("3rd", "4th", "5th")) %>%
  ggplot(aes(y = avg.logmass, x = avg.tt,
             col = trt, shape = trt)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                    ymin = avg.logmass - se.logmass),
                width = 0.25) +
  geom_errorbarh(aes(xmax = avg.tt + se.tt,
                     xmin = avg.tt - se.tt),
                 height = 0.1) +
  labs(caption = "mean±se; development trajectories from\n3rd-5th instars (ctrls)",
       y = "ln mass (mg)",
       x = "time to instar (days)") +
  scale_color_hue(direction = -1)
```

## trajectories of hot return bugs (return - 3rd)
- uhh interesting but not quite it. i think just do the instars separately

```{r eval=FALSE}
dfs_tidy$wide %>%
  CalcDevSS() %>% 
  filter(#trt.type == "expt",
         trt.recover != 40,
         instar %in% c("return", "3rd")) %>%
  mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  ggplot(aes(y = avg.logmass,x = avg.tt, 
             shape = instar,
             color = trt.recover,
             group = trt)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                    ymin = avg.logmass - se.logmass,
                    width = 0.5)) +
  facet_grid(~ trt.duration)
```

## masses of hot return bugs (@ return/3rd)

### util fns

```{r}
# note -- 2ndary versions below also sorta depend on these original ones lol
.calcfn <- function(x){
  x %>%
    filter(!(trt.type == "expt" & jdate.2nd > jdate.return)) %>%
    mutate(trt.to3rd = case_when(trt.type == "expt" & is.na(jdate.3rd) ~ "expt no 3rd",
                                 trt.type == "expt" & !is.na(jdate.3rd) ~ "expt 3rd",
                                 TRUE ~ "ctrl"
                                 
                                 # trt.type == "expt" & is.na(jdate.3rd) ~ "no",
                                 # TRUE ~ "yes"
    )) %>%
    CalcDevSS() #%>% 
  #filter(instar %in% c("return", "3rd")) %>%
  #mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt.to3rd,
               color = trt.duration,
               shape = trt.duration,
               #shape = instar
               )) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass,
                      width = 0.25)) +
    ## omit this when x = trt.duration bc its too messy.
    # geom_text(aes(label = n),
    #           hjust = -0.7, vjust = -0.6, size = 4,
    #           show.legend = FALSE) +
    facet_grid(~ trt.recover, scales = "free_x", drop = TRUE) +
    labs(y = "mass (mg)",
         x = "expt group",
         color = "recovery\nduration (h)",
         shape = "recovery\nduration (h)")
}
```

### by trt-outcome group
```{r}
# parsed out expt bugs that made it to 3rd
dfs_tidy$wide %>%
  .calcfn() %>% 
  filter(instar == "return", 
         trt.type != "ctrl") %>%
  .plotfn() +
  labs(caption = "masses @ return")


dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "3rd") %>%
  .plotfn() +
  labs(caption = "masses @ 3rd")
```

```{r eval=FALSE}
# using ln(mass)
# need .calcfn from utils

# i like the actual mass more (than the ln) bc theyre so tiny -- 
# but dumping the ln one just in case..

.plotfn2 <- function(x){
  x %>%
    ggplot(aes(y = avg.logmass, x = trt.to3rd,
               color = trt.duration,
               shape = trt.duration,
               #shape = instar
               )) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                      ymin = avg.logmass - se.logmass,
                      width = 0.25)) +
    facet_grid(~ trt.recover, scales = "free_x", drop = TRUE) +
    labs(y = "ln mass (mg)",
         x = "expt group",
         color = "recovery\nduration (h)",
         shape = "recovery\nduration (h)")
}

dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "return", 
         trt.type != "ctrl") %>%
  .plotfn2() +
  labs(caption = "masses @ return")

dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "3rd") %>%
  .plotfn2() +
  labs(caption = "masses @ 3rd")
```


### by trt duration

#### util fns
```{r}
# (needs og .plotfn and .calcfn util fns)
## v2 of these fns mostly just overwrite some of the filtering/axes for this view bc im lazy
## also omit geom_text from og .plotfn if not already bc messy

.calcfn2 <- function(x){
  x %>%
    .calcfn() %>% 
    filter(trt != 40,
           instar %in% c("3rd", "return")
    ) %>%
    mutate(instar = factor(instar, levels = c("return", "3rd")))
}

.plotfn2 <- function(x){
  x %>%
    .plotfn() +
    aes(x = trt.duration,
        color = trt.recover,
        shape = trt.to3rd) +
    facet_grid(#instar 
      ~ trt.recover) +
    scale_color_hue(direction = -1) +
    labs(x = "recovery duration",
         color = "recovery temp")
}
```


```{r}
# all bugs grouped tgt
dfs_tidy$wide %>%
  filter(!(trt.type == "expt" & jdate.2nd > jdate.return)) %>%
  CalcDevSS() %>%
  filter(trt != 40,
         instar %in% c("3rd", "return"),
         !(trt.duration == 0 & instar == "return")
         ) %>%
  mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  .plotfn() +
  aes(x = trt.duration,
      color = trt.recover,
      shape = trt.type
      ) +
  facet_grid(~ instar, scales = "free_x", drop = TRUE) +
  labs(caption = "(unparsed expt bugs)")

# parsed out if expt bugs made it to 3rd
dfs_tidy$wide %>%
  .calcfn2() %>%
  filter(instar == "3rd") %>%
  .plotfn2() +
  #idk this used to have 3 of them but also nonsensical if u think about it...
  #scale_shape_manual(values = c(19, 17, 2)) + 
  scale_shape_discrete(labels = c("ctrl", "expt")) +
  labs(caption = "mass at 3rd",
       shape = "treatment type")

dfs_tidy$wide %>%
  .calcfn2() %>%
  filter(trt.duration != 0, instar == "return") %>%
  .plotfn2() +
  scale_shape_manual(values = c(17, 2), labels = c("yes", "no")) +
  labs(caption = "mass at return",
       shape = "survived to 3rd?")
```