---
title: "tdt viz"
date: "2026-01-06"
---

generate tdt viz

# TODO
- before
  - import data
  - import utils
  - make subsets(?)
- analysis
  - all the kms

# preamble

```{r message=FALSE}
# pathfinding
here::i_am("_tdt/scripts/analysis-viz.Rmd")
library(here)

# libs and utils
library(tidyverse)
library(survival) # for Surv 
library(survminer) # for ggsurvplot

source(here("_tdt/scripts/wrangle.R"))
source(here("_tdt/scripts/analysis-viz_utils.R"))
```

```{r data prep, message=FALSE}
# import data
source(here("_tdt/scripts/R/tidy.R"))

# format col types for easier plotting...?

# drop pmc bugs
dfs_tidy <- dfs_tidy %>% lapply(., \(x){
  x %>% 
    filter(status %in% c(0, 1)) %>%
    mutate(across(starts_with("trt"), as.factor))
})
```

# # kms/survival curves

# ctrls

## cohort x rec temp
results
- no major diffs in KMs for 40 ctrls across cohorts

```{r}
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.duration == 0)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "cohort"
           ) %++%
  AddKMLines(km_df) + 
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) + 
  labs(caption = "rec duration = 0h")
```

## cohort
```{r}
# by cohort
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.duration == 0,
         trt.recover == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ cohort, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "trt.recover"
           ) %++%
  AddKMLines(km_df) +
  labs(caption = "rec 0h @ 40C")

# looking at the 40C grp only -- no major diffs
```


# exptal

## rec temp x rec dur
```{r}
# from entering (drop cohort A)
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40) # ohh this is where im hitting issues w the subsetting...
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "cohort",
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.enter, 
             trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = enter")


# from entering (w/ cohort A)
km_df <- dfs_tidy$longall %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = c("cohort", "trt.duration")
           ) %++%
  AddKMLines(km_df, dt.enter, 
             cohort, trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = enter")
```

```{r}
# from return (no cohort A)
km_df <- dfs_tidy$long %>%
  filter(subset == "fromreturn" & trt.enter == 40) 
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "cohort"
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.return, 
             trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = return to 40C")


# from return (with cohort A)
km_df <- dfs_tidy$longall %>%
  filter(subset == "fromreturn" & trt.enter == 40) 
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = c("cohort", "trt.duration")
           ) %++%
  AddKMLines(km_df, dt.return, 
             cohort, trt.duration 
             ) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "t0 = return to 40C")
```

```{r}
# for STOR consult: differences in surv curves depending on where you start
# km_df <-
#   list(fromenter = dfs_tidy$long %>% filter(trt.enter == 40 & subset == "fromenter"),
#        fromreturn = dfs_tidy$long %>% filter(trt.enter == 40 & subset == "fromreturn"))
km_df <- dfs_tidy$long %>% filter(trt.enter == 40, subset != "fromrecover")
km_fit <- surv_fit(Surv(tt.exit, status) ~ trt.recover, data = km_df)

ggsurvplot_facet(fit = km_fit, data = km_df,
                 facet.by = c("subset", "trt.duration")) +
  AddKMLines(df = filter(km_df, subset == "fromenter"), dt.enter, trt.duration, subset) +
  AddKMLines(df = filter(km_df, subset == "fromreturn"), dt.return, trt.duration, subset) +
  scale_color_manual(values = c("#002F70", "#3C6FC4", "#DE738B", "#8E063B")) +
  labs(caption = "(varying t0)")
```



## cohort x rec dur
```{r}
# TODO testing
.test_AddKMLines <- function(){
  df <- dfs_tidy$longall %>%
    filter(subset == "fromenter" & trt.enter == 40) %>%
    group_by(cohort) %>%
    summarise(dt.enter, dt.recover, dt.return) %>%
    drop_na() %>% unique() %>%
    mutate(across(starts_with("dt."), ~ . - dt.recover),
           across(starts_with("dt."), ~ replace(.x, which(.x <0), NA))) %>%
    unique() %>%
    
    pivot_longer(starts_with("dt"),
                 names_to = "from", names_prefix = "dt\\.",
                 values_to = "x") %>% 
    mutate(col = case_when(from == "recover" ~ "skyblue",
                           from == "return" ~ "orange",
                           from == "enter" ~ "red"),
           lty = case_when(from == "recover" ~ "dashed",
                           from == "return" ~ "solid",
                           from == "enter" ~ "solid"))
  
  return(
    list(geom_vline(data = df,
                    aes(xintercept = x,
                        # lty = lty,
                        # col = col,
                        group = cohort),
                    show.legend = FALSE)
    )
  )
}

# TOFIX AddKMLines() shows all lines for all cohorts...

# effect of trt.duration (without cohort A)
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "cohort"
           #facet.by = "trt.recover"
           ) %++%
  AddKMLines(km_df, dt.recover, 
             cohort
             )

# (with cohort A)
km_df <- dfs_tidy$longall %>%
  filter(subset == "fromenter" & trt.enter == 40)
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.duration, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "cohort"
           #facet.by = "trt.recover"
           ) %++%
  AddKMLines(km_df, dt.recover, # NOTE can filter this...
             cohort
             ) #%++%
  #.test_AddKMLines()
```
## cohort x duration

```{r}
# effect of cohort
  # (cohort A's 24/48h recovery isnt the same as B+C, so is omitted)
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.enter == 40) 
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ cohort, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "trt.recover",
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.recover, trt.duration
             )

# with cohort A
km_df <- dfs_tidy$longall %>%
  filter(subset == "fromreturn" & trt.enter == 40) 
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ cohort, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           #facet.by = "trt.recover",
           facet.by = "trt.duration"
           ) %++%
  AddKMLines(km_df, dt.return, trt.duration
             ) +
  labs(caption = "t0 = from return")
```
# hot exptal only

## rec temp x duration

(copied from from 2026-02 LB `results.Rmd`)

```{r}
# from enter
km_df <- dfs_tidy$long %>%
  filter(subset == "fromenter" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") + 
  labs(caption = "from enter") +
    scale_color_hue(direction = -1)
  
# from return
km_df <- dfs_tidy$long %>%
  filter(subset == "fromreturn" & trt.type == "expt")
km_fit <- survfit(Surv(km_df$tt.exit, km_df$status) ~ trt.recover, data = km_df)

ggsurvplot(fit = km_fit, data = km_df,
           facet.by = "trt.duration") +
  labs(caption = "from return") +
    scale_color_hue(direction = -1)
```


# # (WIP) duration/times

(orig from 2026-02 LB `results.Rmd`)

## df prep
```{r}
dfs <- list(
  noA = dfs_tidy$wide,
  all = dfs_tidy$wideall
) %>%
  lapply(., \(x){
    x %>%
      #filter(trt.duration %in% c(24, 48)) %>%
      filter(fate == "died" & trt.enter == 40) %>%
      mutate(exitfromreturn = dt.exit - dt.return,
             timeat40 = case_when(trt == 40 ~ dt.exit,
                                  trt.duration != 0 & cohort == "A" ~ exitfromreturn + 48,
                                  trt.duration != 0 & cohort != "A" ~ exitfromreturn + 24))
  })

dfs <- list_modify(
  dfs,
  noA = dfs$noA %>% group_by(across(starts_with("trt"))),
  all = dfs$all %>% group_by(across(c(grp, starts_with("trt"))))
) %>%
  lapply(., \(x){
    x %>%
      summarise(avg.ttexit.ret = mean(exitfromreturn, na.rm = TRUE),
                se.ttexit.ret = se(exitfromreturn),
                avg.ttexit.ent = mean(dt.exit, na.rm = TRUE),
                se.ttexit.ent = se(dt.exit),
                avg.timeat40 = mean(timeat40, na.rm = TRUE),
                se.timeat40 = se(timeat40)
      ) %>%
      ungroup()
  })
```

## time to death vs rec temp/dur

### from enter
```{r}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.ttexit.ent,
               x = trt.recover)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.ttexit.ent + se.ttexit.ent,
                      ymin = avg.ttexit.ent - se.ttexit.ent),
                  width = 0.25, lty = "solid") +
    labs(y = "time to death (h)",
         x = "recovery temperature (°C)",
         #caption = "from enter"
         ) +
    theme_bw()
}

## facet by duration
dfs$noA %>%
  .plotfn() +
    facet_grid(~ trt.duration, scales = "free_x", space = "free_x")

## not faceted
dfs$noA %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)",
       caption = "cohort B+C; from entering")

## facet by grp
dfs$all %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_wrap(~ grp, labeller = as_labeller(c(`A` = "cohort A (48h)", 
                                             `B+C` = "cohort B + C (24h)"))) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)",
       caption = "from entering;\n(initial hours exposure to 40C before recovering)")
```

### from return
```{r}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.ttexit.ret,
               x = trt.recover)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.ttexit.ret + se.ttexit.ret,
                      ymin = avg.ttexit.ret - se.ttexit.ret),
                  width = 0.25, lty = "solid") +
    labs(y = "time to death (h)",
         x = "recovery temperature (°C)",
         caption = "cohort B+C; from return",
         lty = "recovery\nduration (h)",
         shape = "recovery\nduration (h)",
         color = "recovery\nduration (h)"
         ) +
    theme_bw()
}

## facet by duration
dfs$noA %>%
  .plotfn() +
    facet_grid(~ trt.duration, scales = "free_x", space = "free_x")

dfs$noA %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_grid(~ trt.type, scales = "free_x", space = "free_x",
             labeller = as_labeller(c(`ctrl` = "40°C constant",
                                      `expt` = "40°C + recovery")))

## not faceted
dfs$noA %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration)

## facet by grp
dfs$all %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_wrap(~ grp, labeller = as_labeller(c(`A` = "cohort A (48h)", 
                                             `B+C` = "cohort B + C (24h)"))) +
  labs(caption = "from return;\n(initial hours exposure to 40C before recovering)")
```

## time at 40C vs exposure (TODO/WIP)

- WIP!!! cant ducking do math) 
- maybe only compare to the bugs that survived after the initial 24h?

```{r}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.timeat40,
               x = trt.recover)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.timeat40 + se.timeat40,
                      ymin = avg.timeat40 - se.timeat40),
                  width = 0.25, lty = "solid") +
    labs(y = "total exposure to 40°C (h)",
         x = "recovery temperature (°C)",
         caption = "until death"
         ) +
    theme_bw()
}

## facet by duration
dfs$noA %>%
  .plotfn() +
    facet_grid(~ trt.duration, scales = "free_x", space = "free_x")

dfs$noA %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_grid(~ trt.type, scales = "free_x", space = "free_x",
             labeller = as_labeller(c(`ctrl` = "40°C constant",
                                      `expt` = "40°C + recovery"))) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)")

## facet by grp
dfs$all %>%
  .plotfn() +
  aes(lty = trt.duration, shape = trt.duration, color = trt.duration,
      group = trt.duration) +
  facet_wrap(~ grp, labeller = as_labeller(c(`A` = "A (48h)", `B+C` = "B + C (24h)"))) +
  labs(lty = "recovery\nduration (h)",
       shape = "recovery\nduration (h)",
       color = "recovery\nduration (h)",
       caption = "exposure before death; (initial hours exposure to 40C before recovering)")
```

## exit time diff from ctrl (WIP!!!)


```{r}
# try 1: just trying stuff out... (figuring out how to do the math)
# some funky df prep
  ## (note that avg.diff is the same whether doing:
  ## avg.ttexit.expt - avg.ttexit.ctrl or
  ## avg(tt.exit.expt - avg.ttexit.ctrl))

dfs <- list(
  # get avg.exit times of 40C ctrl groups per cohort
  ctrls = dfs_tidy$wideall %>%
    filter(trt == 40) %>%
    select(c(grp, starts_with("trt"), dt.exit)) %>% 
    group_by(across(c(grp, starts_with("trt")))) %>%
    summarise(avg.ttexit.ctrl = mean(dt.exit, na.rm = TRUE),
              se.ttexit.ctrl = se(dt.exit)) %>%
    ungroup()
  )

dfs <- list_modify(
  dfs,
  # individual survival time data for expt bugs
  expt_counts = dfs_tidy$wideall %>%
    filter(trt != 40,
           trt.enter == 40,
           fate == "died") %>%
    merge(., dfs$ctrls %>% select(c(grp, avg.ttexit.ctrl, #se.ttexit.ctrl
                                     )), 
          by = "grp", 
          all = TRUE) %>%
    mutate(timeat40 = case_when(grp != "A" ~ dh.exit - dh.return + 24,
                                grp == "A" ~ dh.exit - dh.return + 48,
                                #trt == 40 ~ dh.exit - dh.enter
                                ),
           diff.fromctrl = case_when(trt != 40 ~ timeat40 - avg.ttexit.ctrl,
                                     #TRUE ~ timeat40
                                     ))
  )

dfs <- list_modify(
  dfs,
  expt_ss = dfs$expt_counts %>%
    # mutate(grp = case_when(#trt == 40 ~ "ctrl",
    #                        cohort == "A" ~ "A",
    #                        cohort != "A" ~ "B+C"
    #                        )) %>% 
    group_by(across(c("grp", starts_with("trt")))) %>%
    mutate(n = n(),
           avg.diff = case_when(trt != 40 ~ mean(diff.fromctrl, na.rm = TRUE),
                                #trt == 40 ~ avg.ttexit.ctrl
                                ),
           se.diff = case_when(trt != 40 ~ se(diff.fromctrl),
                               #trt == 40 ~ se.ttexit.ctrl
                               ),
           # # dont need to do this LOL
           # avg.ttexit.expt = case_when(trt != 40 ~ mean(timeat40, na.rm = TRUE)), 
           #  # makes this "too" unique 
           #  # (will lead to overplotting avg.ttexit.ctrl)
           # diff.overall = case_when(trt != 40 ~ avg.ttexit.expt - avg.ttexit.ctrl,
           #                          #trt == 40 ~ avg.ttexit.ctrl)
                                    ) %>%
    select(c(#cohort,
             grp,
             starts_with("trt"), n, 
             contains(c("avg", "se"#, "overall"
                        )), 
             #-contains("ctrl")
             )) %>%
    unique()
)

# plot
  # the ctrl points make no sense here LOL bc this should show the diff, not the raw!
dfs$expt_ss %>%
  #filter(grp != "ctrl") %>%
  ggplot(aes(x = trt.recover, col = trt.duration,
             y = avg.diff,
             group = interaction(trt.duration#, cohort
                                 ))) +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = avg.diff + se.diff,
                    ymin = avg.diff - se.diff), width = 0.25) +
  facet_wrap(~ grp, labeller = as_labeller(c(`A` = "A (48h)", `B+C` = "B + C (24h)")))

```

```{r}
# try 2: compare individs vs avgs; trying out other ways to do math

# df prep
dfs <- list(
  counts = dfs_tidy$wide %>%
    # get avg.exit times of ctrl groups per cohort
    filter(trt == 40) %>%
    select(c(cohort, dt.exit)) %>% 
    group_by(cohort) %>%
    summarise(avg.ttexit.ctrl = mean(dt.exit, na.rm = TRUE)) %>%
    ungroup() %>%
    
    # merge w/ recovery bugs and calc diff
    merge(., dfs_tidy$wide %>% filter(fate == "died",
                                      trt != 40,
                                      trt.enter == 40),
          by = "cohort", all = TRUE) %>% 
    mutate(#diff.fromctrl = dt.exit - avg.ttexit.ctrl, # NOTE i think this is wrong..
           
           # NOTE (i think this is also wrong)
           timeat40 = case_when(trt.duration != 0 ~ dh.exit - dh.return + 24),
           diff.fromctrl = timeat40 - avg.ttexit.ctrl, # = difference in total time at 40C 
           ) %>%
    filter(!is.na(diff.fromctrl))
)

dfs <- list_modify(
  dfs,
  ss = dfs$counts %>% 
    group_by(across(c(starts_with("trt")))) %>%
    summarise(n = n(),
              avg.diff = mean(diff.fromctrl, na.rm = TRUE),
              se.diff = se(diff.fromctrl)) %>%
    ungroup()
)


# all data + avgs
dfs$counts %>% 
  ggplot(aes(x = trt.recover)) +
  geom_point(aes(y = diff.fromctrl),
             alpha = 0.5, size = 1) +
  geom_point(data = dfs$ss, 
             aes(y = avg.diff), color = "red") +
  geom_errorbar(data = dfs$ss, 
                aes(ymax = avg.diff + se.diff,
                    ymin = avg.diff - se.diff),
                width = 0.25, color = "red") +
  facet_wrap(~trt.duration)

# focus on the ss
dfs$ss %>%
  ggplot(aes(y = avg.diff, x = trt.recover, col = trt.duration,
             group = trt.duration)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.diff + se.diff,
                    ymin = avg.diff - se.diff),
                width = 0.25) +
  geom_hline(yintercept = 0, lty = "longdash", alpha = 0.5, show.legend = NA) +
  geom_text(aes(label = n),
            vjust = -1, 
            hjust = -0.75, show.legend = FALSE) +
  labs(y = "difference (h)",
       captions = "(diff in total time at 40C vs ctrls)")
  
```


# # development stuff

# surv props

## overall (sim)
- using simulated 40C data

```{r eval=FALSE}
# % bugs dying at each timept
# (N gets reset to the current timept)
dfs_tidy$widesim %>%
  CalcSurvProps(., trt.recover, trt.duration) %>% #View()
  ggplot(aes(y = p, x = timept,
             fill = status)) +
  geom_col() +
  facet_grid(trt.duration ~ trt.recover, drop = TRUE) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(caption = "N = total bugs at start of timept")
  
# running count of bugs thru expt
dfs_tidy$widesim %>%
  CalcSurvProps(., trt.recover, trt.duration) %>% 
  #filter(status == "surv") %>%
  ggplot(aes(y = n.initial, x = timept,
             group = interaction(trt.recover, trt.duration),
             )) +
  geom_point() + geom_line() +
  facet_grid(trt.duration ~ trt.recover, drop = TRUE) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(caption = "N = total bugs at enter")
```

## (WIP) stage of death

```{r eval=FALSE}
# died at which point in the time course? (testing...)
# TOFIX 
  # math depends on the trt type + cohort tho...
  # should be dt.exit for trt != 40?, not dt.return lol?
dfs_tidy$wideall %>%
  filter(trt.enter == 40) %>%
  mutate(stage.died = case_when((trt == 40 & dt.exit < 24) |
                                  (trt != 40 & (dt.return < dt.return + 24)) ~ "24h",
                                (trt == 40 & dt.exit < 48) |
                                  (trt != 40 & (dt.return < dt.return + 48)) ~ "48h",
                                (trt == 40 & dt.exit < 72) |
                                  (trt != 40 & (dt.return < dt.return + 72)) ~ "72h",
                                (trt == 40 & dt.exit < 96) |
                                  (trt != 40 & (dt.return < dt.return + 96)) ~ "96h",
                                (trt == 40 & dt.exit > 96) |
                                  (trt != 40 & (dt.return > dt.return + 96))~ "> 96h",
                                is.na(dt.recover) ~ "before recovery",
                                is.na(dt.return) & (dt.exit < dt.recover + 24) ~ "24h (rec)",
                                TRUE ~ "48h (rec)")) %>%
  mutate(stage.died = factor(stage.died, levels = c("before recovery",
                                                    "24h (rec)", "48h (rec)",
                                                    "24h", "48h", "72h", "96h", "> 96h"))) %>%
  ggplot(aes(x = trt.recover, fill = stage.died)) +
  geom_bar(position = "fill") +
  facet_grid(grp ~ trt.duration, drop = TRUE, scales = "free_x", space = "free")
```

```{r}
# hours of exposure to 40C
list(
  dfs_tidy$wideall %>%
    filter(trt.enter == 40,
           cohort == "A") %>%
    mutate(stage.died = case_when(dt.exit < 24 ~ "24h",
                                  dt.exit < 48 ~ "48h*",
                                  (trt == 40 & dt.exit < 72) |
                                    (trt != 40 & dt.exit < dt.return + 24) ~ "72h",
                                  (trt == 40 & dt.exit < 96) |
                                    (trt != 40 & dt.exit < dt.return + 48) ~ "96h",
                                  (trt == 40 & dt.exit > 96) |
                                    (trt != 40 & dt.exit > dt.return + 48) ~ "> 96h",
                                  is.na(dt.return) & (dt.exit < dt.recover + 24) ~ "rec 24h",
                                  is.na(dt.return) & (dt.exit < dt.recover + 48) ~ "rec 48h"
                                  )),
  dfs_tidy$wideall %>%
    filter(trt.enter == 40,
           cohort != "A") %>%
    mutate(stage.died = case_when(dt.exit < 24 ~ "24h",
                                  (trt == 40 & dt.exit < 48) |
                                    (trt != 40 & dt.exit < dt.return + 24) ~ "48h",
                                  (trt == 40 & dt.exit < 72) |
                                    (trt != 40 & dt.exit < dt.return + 24) ~ "72h",
                                  (trt == 40 & dt.exit < 96) |
                                    (trt != 40 & dt.exit < dt.return + 48) ~ "96h",
                                  (trt == 40 & dt.exit > 96) |
                                    (trt != 40 & dt.exit > dt.return + 48) ~ "> 96h",
                                  is.na(dt.return) & (dt.exit < dt.recover + 24) ~ "rec 24h",
                                  is.na(dt.return) & (dt.exit < dt.recover + 48) ~ "rec 48h"
                                  ))
) %>%
  rbindlist() %>% 
  # levels kinda wonky bc "rec" isnt the same for cohorts A / B+C
  mutate(grp = factor(grp, levels = c("B+C", "A")),
         stage.died = factor(stage.died, levels = c("24h", "48h*",
                                                    "rec 24h", "rec 48h", 
                                                    "48h", "72h", "96h", "> 96h"))) %>%
  ggplot(aes(x = trt.recover, fill = stage.died)) +
  geom_bar(position = "fill") +
  facet_grid(grp ~ trt.duration, # NOTE prob should be trt.exposure ~ trt.duration but its wonky
             #trt.exposure ~ trt.duration,
             drop = TRUE, scales = "free_x", space = "free",
             labeller = labeller(trt.duration = c(`0` = "0h recovery",
                                                  `24` = "initial HS + 24h recovery",
                                                  `48` = "initial HS + 48h recovery"),
                                 grp = c(`A` = "48h initial HS",
                                         `B+C` = "24h initial HS"),
                                 )) +
  scale_fill_manual(breaks = c("24h", "48h", "rec 24h", "rec 48h", "72h", "96h", "> 96h"),
                    values = c("24h" = "#FED976", "48h*" = "#FEB24C", "48h" = "#FEB24C", 
                               "rec 24h" = "#9ECAE1", "rec 48h" = "#6BAED6",
                               "72h" = "#FC4E2A", "96h" = "#E31A1C", "> 96h" = "#B10026"),
                    labels = c("0-24 h", "24-48 h",
                               "rec 0-24 h", "rec 24-48 h", "48-72 h", "72-96 h", "> 96 h")
                    ) +
  labs(fill = "period died\n(hrs at 40C)",
       y = "proportion",
       x = "recovery temp (°C)",
       title = "survival at 40C (hrs)",
       caption = "*24h or **48h initial HS before recovery") +
  theme_bw()
```


# dev trajectories

(futzing around w this for 2026-02 LB)

## cold ctrls

- TODO:

```{r}
dfs_tidy$wide %>%
  CalcDevSS() %>%
  filter(trt.duration == 0,
                instar %in% c("3rd", "4th", "5th")) %>%
  ggplot(aes(y = avg.logmass, x = avg.tt,
             col = trt, shape = trt)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                    ymin = avg.logmass - se.logmass),
                width = 0.25) +
  geom_errorbarh(aes(xmax = avg.tt + se.tt,
                     xmin = avg.tt - se.tt),
                 height = 0.1) +
  labs(caption = "mean±se; development trajectories from\n3rd-5th instars (ctrls)",
       y = "ln mass (mg)",
       x = "time to instar (days)") +
  scale_color_hue(direction = -1)
```

## hot return bugs (return - 3rd)
- uhh interesting but not quite it. i think just do the instars separately

```{r eval=FALSE}
dfs_tidy$wide %>%
  CalcDevSS() %>% 
  filter(#trt.type == "expt",
         trt.recover != 40,
         instar %in% c("return", "3rd")) %>%
  mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  ggplot(aes(y = avg.logmass,x = avg.tt, 
             shape = instar,
             color = trt.recover,
             group = trt)) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                    ymin = avg.logmass - se.logmass,
                    width = 0.5)) +
  facet_grid(~ trt.duration)
```

# mass @ return

(futzing around w this for 2026-02 LB)

## expt bugs (@ return/3rd)

### util fns

```{r}
# note -- 2ndary versions below also sorta depend on these original ones lol
.calcfn <- function(x){
  x %>%
    filter(!(trt.type == "expt" & jdate.2nd > jdate.return)) %>%
    mutate(trt.to3rd = case_when(trt.type == "expt" & is.na(jdate.3rd) ~ "expt no 3rd",
                                 trt.type == "expt" & !is.na(jdate.3rd) ~ "expt 3rd",
                                 TRUE ~ "ctrl"
                                 
                                 # trt.type == "expt" & is.na(jdate.3rd) ~ "no",
                                 # TRUE ~ "yes"
    )) %>%
    CalcDevSS() #%>% 
  #filter(instar %in% c("return", "3rd")) %>%
  #mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  
}

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt.to3rd,
               color = trt.duration,
               shape = trt.duration,
               #shape = instar
               )) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass,
                      width = 0.25)) +
    ## omit this when x = trt.duration bc its too messy.
    # geom_text(aes(label = n),
    #           hjust = -0.7, vjust = -0.6, size = 4,
    #           show.legend = FALSE) +
    facet_grid(~ trt.recover, scales = "free_x", drop = TRUE) +
    labs(y = "mass (mg)",
         x = "expt group",
         color = "recovery\nduration (h)",
         shape = "recovery\nduration (h)")
}
```

### by trt-outcome group
```{r}
# parsed out expt bugs that made it to 3rd
dfs_tidy$wide %>%
  .calcfn() %>% 
  filter(instar == "return", 
         trt.type != "ctrl") %>%
  .plotfn() +
  labs(caption = "masses @ return")


dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "3rd") %>%
  .plotfn() +
  labs(caption = "masses @ 3rd")
```

```{r eval=FALSE}
# using ln(mass)
# need .calcfn from utils

# i like the actual mass more (than the ln) bc theyre so tiny -- 
# but dumping the ln one just in case..

.plotfn2 <- function(x){
  x %>%
    ggplot(aes(y = avg.logmass, x = trt.to3rd,
               color = trt.duration,
               shape = trt.duration,
               #shape = instar
               )) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = avg.logmass + se.logmass,
                      ymin = avg.logmass - se.logmass,
                      width = 0.25)) +
    facet_grid(~ trt.recover, scales = "free_x", drop = TRUE) +
    labs(y = "ln mass (mg)",
         x = "expt group",
         color = "recovery\nduration (h)",
         shape = "recovery\nduration (h)")
}

dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "return", 
         trt.type != "ctrl") %>%
  .plotfn2() +
  labs(caption = "masses @ return")

dfs_tidy$wide %>%
  .calcfn() %>%
  filter(instar == "3rd") %>%
  .plotfn2() +
  labs(caption = "masses @ 3rd")
```


### by trt duration

#### util fns
```{r}
# (needs og .plotfn and .calcfn util fns)
## v2 of these fns mostly just overwrite some of the filtering/axes for this view bc im lazy
## also omit geom_text from og .plotfn if not already bc messy

.calcfn2 <- function(x){
  x %>%
    .calcfn() %>% 
    filter(trt != 40,
           instar %in% c("3rd", "return")
    ) %>%
    mutate(instar = factor(instar, levels = c("return", "3rd")))
}

.plotfn2 <- function(x){
  x %>%
    .plotfn() +
    aes(x = trt.duration,
        color = trt.recover,
        shape = trt.to3rd) +
    facet_grid(#instar 
      ~ trt.recover) +
    scale_color_hue(direction = -1) +
    labs(x = "recovery duration",
         color = "recovery temp")
}
```


```{r}
# all bugs grouped tgt
dfs_tidy$wide %>%
  filter(!(trt.type == "expt" & jdate.2nd > jdate.return)) %>%
  CalcDevSS() %>%
  filter(trt != 40,
         instar %in% c("3rd", "return"),
         !(trt.duration == 0 & instar == "return")
         ) %>%
  mutate(instar = factor(instar, levels = c("return", "3rd"))) %>%
  .plotfn() +
  aes(x = trt.duration,
      color = trt.recover,
      shape = trt.type
      ) +
  facet_grid(~ instar, scales = "free_x", drop = TRUE) +
  labs(caption = "(unparsed expt bugs)")

# parsed out if expt bugs made it to 3rd
dfs_tidy$wide %>%
  .calcfn2() %>%
  filter(instar == "3rd") %>%
  .plotfn2() +
  #idk this used to have 3 of them but also nonsensical if u think about it...
  #scale_shape_manual(values = c(19, 17, 2)) + 
  scale_shape_discrete(labels = c("ctrl", "expt")) +
  labs(caption = "mass at 3rd",
       shape = "treatment type")

dfs_tidy$wide %>%
  .calcfn2() %>%
  filter(trt.duration != 0, instar == "return") %>%
  .plotfn2() +
  scale_shape_manual(values = c(17, 2), labels = c("yes", "no")) +
  labs(caption = "mass at return",
       shape = "survived to 3rd?")
```