---
title: "final ntw results"
date: "2026-05-24"
---

# preamble and setup

```{r pathfinding, message=FALSE}
here::i_am("_ntw/scripts/analysis/mss/final-res.Rmd")
```

for collated viz + stats results for ntw mss

-   notes on the viz:
    -   main viz = focuses on trt type (NTW vs control), minimum temp, population
    -   supp viz = includes year and sex (if applicable)
-   notes on the stats:
    -   main stats = focuses on the NTW bugs and population, minimum temp as predictors
    
main results/figs:

1. survival: hatch to pupation; hatch to eclosion
2. mass: pupal mass; eclosion mass
3. time: time to pupation; pupal duration
4. adult longevity
5. adult fitness



# loading

```{r data, message=FALSE, warning=FALSE}
source(here::here("_ntw/scripts/analysis/mss/load.R"))
```

## utils
```{r utilities}
# for (main) plots
p <- list()
# ps for supp plots...?

# (temp) function to save results
ResToCsv <- function(res, filename){
  res <- as.data.frame(res)
  today <- format(Sys.time(), "%y%m%d")
  path <- paste0("_ntw/out/mss-stats/", filename, "_", today, ".csv")
  
  write.csv(res, here::here(path), row.names = TRUE)
}
```

## legend fns
```{r}
# development plots (pop vs trt.type by minT)
Plot_Dev <- function(df){
  df %>%
    ggplot(data = df %>% arrange(desc(trt)), # plot ctrl over NTW if overplotted
           mapping = aes(x = trt.minT,
                         shape = pop,
                         lty = pop,
                         col = trt.type,
                         group = interaction(pop, trt.type)
           )) +
    labs(lty = "population", shape = "population",
         col = "larval temperature\ntreatment",
         x = p_scales$xlab_minT
    ) +
    scale_shape_manual(values = p_scales$shp_pop) +
    scale_linetype_manual(values = p_scales$lty_pop) +
    scale_color_manual(values = p_scales$cols_trttype,  
                       labels = p_scales$labs_trttype) +
    scale_x_discrete(labels = p_scales$labs_minT) +
    guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid")))) +
    theme_bw()
}

# fitness plots (trted sex vs minT)
Plot_Fit <- function(df){
  df %>%
    ggplot(aes(x = trt.minT,
               lty = trt.mate, 
               shape = trt.mate,
               group = trt.mate,
               col = trt.mate)) +
    labs(col = "heat-treated sex", lty = "heat-treated sex", shape = "heat-treated sex",
         x = p_scales$xlab_minT) +
    scale_shape_manual(values = p_scales$shp_trtsex, labels = p_scales$labs_trtmate) +
    scale_color_manual(values = p_scales$cols_trtsex, labels = p_scales$labs_trtmate) +
    scale_linetype_manual(values = p_scales$lty_trtsex, labels = p_scales$labs_trtmate) +
    scale_x_discrete(labels = c(p_scales$labs_trt)) +
    guides(shape = guide_legend(override.aes = list(linetype = c("blank", "solid",
                                                                 "3223", "4224")))) +
    theme_bw()
}
```

# ## development

# (1) survival

## survival to pup

```{r plot prep pup survival}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = prop.pup,
        ) +
    geom_line(show.legend = TRUE) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    ylim(c(0, 1))
}
```

```{r main fig pup survival}
p <- list_modify(
  p,
  surv_la =
    dfs_viz$wide %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>%
    .plotfn() +
    labs(#y = "% survived to pupation",
      title = "hatch to pupation"
    )
)

p$surv_la
```

### supp
```{r supp fig pup survival}
# supp
dfs_viz$wide %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  labs(y = "% survival to pupation") +
  facet_wrap(~year)
```

### stats
```{r main stats pup survival}
mods <- list(
  m1 = glm(is.pup ~ year*pop*trt.minT,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m1, test = "Chisq") # can drop the 3-way

mods <- list_modify(
  mods,
  m2 = glm(is.pup ~ (year + pop + trt.minT)^2,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m2, test = "Chisq") #%>% ResToCsv("pup-surv")
summary(mods$m2)
```

## survival to eclosion
```{r main fig adult survival}
# TODO include sample size?
p <- list_modify(
  p,
  surv_pup = 
    dfs_viz$wide %>%
    filter(is.na(is.LPI),
           #!is.na(sex) # can keep?
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>%
    Plot_Dev() +
    aes(y = prop.ec) +
    geom_point(size = 2) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = prop.ec + se.ec,
                      ymin = prop.ec - se.ec),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(#y = "% pupa eclosed",
         title = "pupa to eclosion",
         ) +
    ylim(c(0, 1))
)

p$surv_pup
```

## collated main fig

```{r}
p$surv_la + p$surv_pup +
  plot_layout(axes = "collect", guides = "collect") &
  labs(y = "% survival") &
  theme(plot.title = element_text(hjust = 0.5))
```


# (2) mass 

## pupal mass

drop LPIs bc mass isnt accurate

```{r plot prep pup mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "pupa",
         is.na(is.LPI)
         )

.plotfn <- function(x) {
  x %>%
    Plot_Dev() +
    aes(y = avg.mass) +
    geom_point(size = 2) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.5, lty = "solid", show.legend = FALSE,
                  ) +
    labs(#y = "pupal mass (g)",
         #y = "mass (g)",
         title = "pupal mass (g)"
         ) 
}
```

```{r main figs pup mass}
p <- list_modify(
  p,
  mass_pup = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    ylim(c(1, 6.4)),
  
  mass_pupsex = # sexes parsed
    df %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    ylim(c(1, 6.8))
)

p$mass_pup

p$mass_pupsex # use this, per stats
```

### supp

```{r supp figs pup mass}
# sexes combined
df %>%
  PrepAdultSS() %>%
  ungroup(starts_with("trt")) %>% 
  group_by(trt, trt.type, trt.minT, .add = TRUE) %>%
  CalcDevSS() %>% #View()
  .plotfn() +
  facet_wrap(~ year)
  #ylim(c(1.8, 6.7))

# sexes parsed (use this one)
df %>%
  PrepAdultSS() %>%
  ungroup(starts_with("trt")) %>% 
  group_by(trt, trt.type, trt.minT, sex, .add = TRUE) %>%
  CalcDevSS() %>%
  .plotfn() +
  facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
             scales = "free") 
  #ylim(c(1.8, 7.2))
```


### stats
```{r main stats pup mass}
df <- dfs_stats$dev_expt %>%
  filter(!is.na(mass.pupa),
         !is.na(sex),
         is.na(is.LPI)
         )

mods <- list(
  m1 = lm(mass.pupa ~ year*pop*trt.minT*sex,
           data = df)
)

anova(mods$m1) # ok to just focus on the 2 ways of trt/pop/year and main effect of sex

mods <- list_modify(
  mods,
  m2 = lm(mass.pupa ~ (year + pop + trt.minT + sex)^2,
           data = df),
  m3 = lm(mass.pupa ~ (year + pop + trt.minT)^2 + sex,
           data = df)
)

anova(mods$m3, mods$m2, mods$m1)
## adding model complexity doesnt do much
AIC(mods$m3, mods$m2, mods$m1)
## m3 = lowest AIC

anova(mods$m3) #%>% ResToCsv("pup-mass")
summary(mods$m3)
```


## eclosion mass (WIP)

```{r plot prep eclosion mass}
df <- dfs_viz$long %>%
  filter(instar %in% c("pupa", "eclose", "change", "prop")) %>%
  group_by(across(c(pop, diet, starts_with("trt"), instar))) %>%
  CalcDevSS()

.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = avg.mass) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.25, lty = "solid", show.legend = FALSE)
}
```

```{r main fig pup mass change}

p <- list_modify(
  p,
  mass_perat =
    df %>%
    filter(instar == "prop") %>%
    .plotfn() +
    labs(#y = "pupal:eclosion mass",
         #y = "mass ratio",
         title = "pupa:eclosion mass"),
  
  mass_pepct =
    df %>%
    mutate(avg.mass = avg.mass * 100,
           se.mass  = se.mass * 100) %>%
    filter(instar == "change") %>%
    .plotfn() +
    labs(#y = "% mass lost\n(pupa to eclosion)",
         #y = "% mass lost",
         title = "pupa to eclosion\n(% mass lost)")
)

p$mass_perat

p$mass_pepct
```

## collated main fig

```{r}
# TODO finalise pup:ec metric to use
p$mass_pup + p$mass_perat + p$mass_pepct +
    plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank() # TODO remove once decided on P:E
        )
```

# (3) development time

## time to pupation
### figs main + supp
```{r plot prep time to pup}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = avg.devrate) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "development rate (1/days)")
}
```

```{r main fig time to pup}
p <- list_modify(
  p,
  dev_la =
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    filter(instar == "pupa") %>%
    .plotfn() +
    labs(title = "hatch to pupation",
         ) #+
    #ylim(c(0.025, 0.05))
)

p$dev_la
```

```{r supp fig time to pup}
dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa") %>%
  .plotfn() +
  facet_wrap(~ year)
```


### stats
```{r main stats pup dev rate}
mods <- list(
  m1 = lm(1/tt.pupa ~ year*pop*trt.minT,
           data = dfs_stats$dev_expt)
)

anova(mods$m1) # ok to focus on 2ways

mods <- list_modify(
  mods,
  m2 = lm(1/tt.pupa ~ (year + pop + trt.minT)^2,
           dfs_stats$dev_expt)
)

#AIC(mods$m2, mods$m1)
anova(mods$m2) #%>% ResToCsv("pup-devrate")
summary(mods$m2)
```


## pupal duration

```{r main fig pup duration}
p <- list_modify(
  p,
  dev_pup = 
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    filter(instar == "eclose") %>%
    # mutate(avg.tt = avg.tt - 15,
    #        se.tt = se.tt - 15) %>%
    Plot_Dev() +
    aes(y = avg.tt) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(#y = "pupal development rate (1/days)",
         y = "duration (days)",
         title = "pupal duration",
         ) #+
    #ylim(c(0.025, 0.05))
)

p$dev_pup
```

## collated main fig

```{r}
p$dev_la + p$dev_pup + p$dev_ad1 +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5))
```


# (4) longevity (WIP)

## longevity

```{r plot prep adult longevity}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = avg.tt) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    labs(#y = "adult longevity (days)",
         y = "duration (days)",
         )
}
```


```{r main fig adult longevity}
p <- list_modify(
  p,
  # w/o parsing of mate status
  dev_ad1 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long",
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    CalcDevSS() %>%
    .plotfn() +
    labs(title = "adult longevity"), 
  
  dev_ad2 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long") %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    group_by(sex, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    facet_wrap(~sex) +
    labs(title = "adult longevity"),
  
  # w/ mate status parsing
  dev_ad3 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long",
           #year != 2024
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    group_by(if.mated, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    #geom_point(aes(size = sqrt(n))) +
    facet_wrap(~if.mated, labeller = as_labeller(c(`0` = "unmated",
                                                   `1` = "mated"))) +
    labs(title = "adult longevity")
)

p$dev_ad1
p$dev_ad2
p$dev_ad3
```

### supp

```{r supp fig adult longevity}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long") %>%
  CalcDevSS_ad() %>%
  .plotfn() +
  labs(y = "adult longevity (days)") +
  facet_wrap(~ year)
```

### stats (WIP)

```{r main stats adult longevity}
# TODO:
# include sex? parse the mating? drop 2024? (same qns as the viz)
mods <- list(
  m1 = lm(tt.long ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt),
  m2 = lm(tt.long ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt %>% filter(if.mated == 1))
)

anova(mods$m1) # can drop if looking at everyone
anova(mods$m2) # sex matters if looking at mated bugs only...
```

### results table
```{r}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long") %>%
  # TODO include sex?
  CalcDevSS_ad() %>% 
  select(c(year, pop, trt.minT, trt.type, n,
           ends_with("tt"))) #%>% ResToCsv("longevity")
```

# ## (5) fitness 

## utils
```{r}
```

## # eggs laid

```{r main fig eggs laid}
pd = position_dodge(0.35)

.plotfn <- function(x){
  x %>%
    Plot_Fit() +
    aes(#y = avg.coll.perf, 
      y = avg.sqrt.coll.perf, 
    ) +
    geom_point(size = 2, position = pd) +
    geom_line(position = pd, show.legend = TRUE) +
    # very overwhelming w error bars lol
    # geom_errorbar(aes(ymax = avg.coll.perf + se.coll.perf,
    #                   ymin = avg.coll.perf - se.coll.perf,),
    #               lty = "solid", width = 0.25) +
    geom_errorbar(aes(ymax = avg.sqrt.coll.perf + se.sqrt.coll.perf,
                      ymin = avg.sqrt.coll.perf - se.sqrt.coll.perf),
                  lty = "solid", width = 0.5, position = pd,
                  show.legend = FALSE) +
    # geom_text(aes(label = n.f.total),
    #           vjust = -0.25, hjust = -0.25, position = "jitter",
    #           show.legend = FALSE) +
    # geom_text(aes(label = paste0(n.cages.total, "C, ", n.f.total, "F")),
    #           vjust = -0.5, hjust = -0.5, #position = "jitter",
    #           show.legend = FALSE
    #           ) +
    labs(#caption = "mean±se; lab only; N = # F",
      #y = "sqrt [eggs laid per female]", 
      title = "sqrt [eggs laid per female]"
    )
}

p <- list_modify(
  p,
  eggs_laid = 
    dfs_viz$eggs %>%
    .plotfn() +
    facet_grid(~year),
  
  eggs_laidnoyr = dfs_viz$eggs_noyr %>%
    .plotfn()
)

p$eggs_laid
p$eggs_laidnoyr
```

## # eggs hatched

```{r main fig eggs hatched}
.plotfn <- function(x){
  x %>%
    Plot_Fit() +
    aes(y = p.hatch) +
    geom_errorbar(aes(ymax = p.hatch + se.hatch,
                      ymin = p.hatch - se.hatch,),
                  lty = "solid", width = 0.25, show.legend = FALSE) +
    # geom_text(aes(label = n.total.coll),
    #           vjust = -0.5, hjust = -0.5, position = "jitter",
    #           show.legend = FALSE) +
    # geom_text(aes(label = paste0(n.cages.total, "C,", n.f.total, "F")),
    #       vjust = -0.5, hjust = -0.5, size = 3
    #       #position = "jitter"
    #       ) +
    #geom_point(aes(size=sqrt(n.total.coll)) +, # looks dumb
    geom_point(size = 2) +
    geom_line() +
    labs(#caption = "lab only; overall hatch prop. per mating combination",
      #y = "proportion eggs hatched",
      title = "proportion eggs hatched")
}

p <- list_modify(
  p,
  eggs_hatch = 
    dfs_viz$eggs %>%
    .plotfn() +
    facet_wrap(~ year),
  
    eggs_hatchnoyr = 
    dfs_viz$eggs_noyr %>%
    .plotfn()
)

p$eggs_hatch
p$eggs_hatchnoyr # do not combine years... LOL
```

## results table
```{r}
# TODO drop the colony F + expt M tents (from 2025?) but 
# this wasnt tracked in 2023...
dfs_viz$eggs %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg results")

dfs_viz$eggs_noyr %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg-results_noyr")
```

## collated main fig
```{r collated eggs viz}
# yrs faceted
p$eggs_laid + p$eggs_hatch +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        legend.position = "bottom")

# years combined
p$eggs_laidnoyr + p$eggs_hatchnoyr +
  labs(caption = "years combined") +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        legend.position = "bottom")
```


# ## dump

## (ref) supp dev patchwork figs


```{r eval = FALSE}
# add facet var
dfs_viz[c("wide", "long")] <- dfs_viz[c("wide", "long")] %>%
  lapply(., \(x){
    x %>%
      mutate(trt.facet = case_when(trt == 260 ~ "ctrl",
                                   pop == "lab" ~ "lab",
                                   pop == "field" ~ "field"),
             trt.facet = factor(trt.facet))
  })

# legend fn for patchwork plots (trt.type/yr)
AddLeg_Patchwork <- function(){
  list(
    labs(lty = "population", shape = "population"),
    scale_shape_manual(values = p_scales$shp_pop),
    scale_linetype_manual(values = p_scales$lty_pop),
    guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
  )
}
```

```{r supp fig pup survival patchwork, eval = FALSE}
.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "% survived to pupation") +
    AddLeg_Patchwork() +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

p1 <- dfs_viz$wide %>%
  filter(trt == 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = year, 
      group = pop)

p2 <- dfs_viz$wide %>%
  filter(trt != 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = trt,
      group = year,
      col = year) +
  labs(x = p_scales$xlab_minT) +
  scale_x_discrete(labels = p_scales$labs_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r supp fig pup dev rate patchwork, eval = FALSE}
df <- dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa")

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.devrate,
               shape = pop, lty = pop
               )) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "larval development rate (1/days)") +
    scale_color_brewer(palette = "Dark2") +
    AddLeg_Patchwork() +
    theme_bw()
}

p1 <- df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, group = interaction(pop, year), col = year) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0.029, 0.054))
```

```{r supp fig adult longevity patchwork, eval = FALSE}
df <- dfs_viz$long %>%
    filter(if.fridge == 0,
           instar == "long") %>%
    CalcDevSS_ad()

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               shape = pop, lty = pop
               )
           ) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    facet_grid(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "adult longevity (days)",) +
    theme_bw() +
    AddLeg_Patchwork() +
    ylim(c(4, 17.5))
}

p1 <-  df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year,
      group = pop) +
  labs(x = "year")

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt,
      col = year,
      group = interaction(year, pop)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1,2),
              axes = "collect", guides = "collect")
```