---
title: "final ntw results"
date: "2026-05-24"
---

# preamble and setup

```{r pathfinding, message=FALSE}
here::i_am("_ntw/scripts/analysis/mss/final-res.Rmd")
```

for collated viz + stats results for ntw mss

-   notes on the viz:
    -   main viz = focuses on trt type (NTW vs control), minimum temp, population
    -   supp viz = includes year and sex (if applicable)
-   notes on the stats:
    -   main stats = focuses on the NTW bugs and population, minimum temp as predictors
    
main results/figs:

1. survival: hatch to pupation; hatch to eclosion
2. mass: pupal mass; eclosion mass
3. time: time to pupation; pupal duration
4. adult longevity
5. adult fitness



# loading

```{r data, message=FALSE, warning=FALSE}
source(here::here("_ntw/scripts/analysis/mss/load.R"))
```

## utils
```{r utilities}
# for (main) plots
p <- list()
# ps for supp plots...?

# (temp) function to save results
ResToCsv <- function(res, filename){
  res <- as.data.frame(res)
  today <- format(Sys.time(), "%y%m%d")
  path <- paste0("_ntw/out/mss-stats/", filename, "_", today, ".csv")
  
  write.csv(res, here::here(path), row.names = TRUE)
}
```


# ## development

## legend fns
```{r}

# legend for development performance plots (pop vs trt.type by minT)
## main figs
AddLeg_OnePanel <- function(){
  list(
    labs(lty = "population", shape = "population",
         #col = "temperature treatment"
         col = "larval temperature\ntreatment"
         ),
    scale_shape_manual(values = c("field.ctrl" = 17, "field.expt" = 17,
                                  "lab.ctrl" = 16, "lab.expt" = 16), 
                       breaks = c("field.expt", "lab.expt"),
                       labels = c("field", "lab")),
    scale_linetype_manual(values = p_scales$lty_pop),
    scale_color_manual(values = p_scales$cols_trttype, 
                       labels = p_scales$labs_trttype),
    scale_x_discrete(labels = p_scales$labs_minT),
    guides(#color = guide_legend(override.aes = list(shape = c(16, 16))),
           shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
    )
}

## supp figs (add facet by year)
AddLeg_OnePanel2 <- function(){
  list(
    labs(x = p_scales$xlab_minT,
         lty = "population", shape = "population",
         col = "larval temperature\ntreatment"),
      scale_shape_manual(values = p_scales$shp_pop),
      scale_color_manual(values = p_scales$cols_trttype, 
                         labels = p_scales$labs_trttype),
      scale_linetype_manual(values = p_scales$lty_pop),
      guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
  )
}
```

# (1) survival

## survival to pup
```{r main fig pup survival}
df <- dfs_viz$wide %>%
  PrepSurvProps() %>%
  ungroup(year) %>%
  CalcSurvProps()

p <- list_modify(
  p,
  surv_la =
    df %>%
    #filter(trt != 260) %>%
    ggplot(aes(y = prop.pup, x = trt.minT,
               shape = interaction(pop, trt.type), 
               lty = pop,
               col = trt.type,
               group = interaction(pop, trt.type))) +
    geom_point(size = 2, position = position_dodge(0.3)) +
    geom_line(position = position_dodge(0.3), show.legend = TRUE) +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE,
                  position = position_dodge(0.3)) +
    # geom_point(data = filter(df, trt == 260),
    #            mapping = aes(),
    #            size = 2, position = position_dodge(0.3),
    # ) +
    # geom_errorbar(filter(df, trt == 260),
    #               mapping = aes(ymax = prop.pup + se.pup,
    #                             ymin = prop.pup - se.pup),
    #               width = 0.5, position = position_dodge(0.3),
    #               lty = "solid", show.legend = FALSE) +
    labs(#y = "% survived to pupation",
         title = "hatch to pupation",
         x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    theme_bw() +
    ylim(c(0, 1))
)

p$surv_la
```

### stats
```{r main stats pup survival}
mods <- list(
  m1 = glm(is.pup ~ year*pop*trt.minT,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m1, test = "Chisq") # can drop the 3-way

mods <- list_modify(
  mods,
  m2 = glm(is.pup ~ (year + pop + trt.minT)^2,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m2, test = "Chisq") #%>% ResToCsv("pup-surv")
summary(mods$m2)
```

### supp
```{r supp fig pup survival}
pd = position_dodge(0.3)

dfs_viz$wide %>%
  CalcSurvProps_def() %>%
  ggplot(aes(y = prop.pup,
             shape = pop, lty = pop,
             x = trt.minT,
             col = trt.type,
             group = interaction(pop, trt.type))
         ) +
  geom_point(size = 2, position = pd) +
  geom_line(position = pd) +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  position = pd, width = 0.75, 
                  lty = "solid", show.legend = FALSE) +
  labs(y = "% survival to pupation") +
  AddLeg_OnePanel2() +
  facet_wrap(~year) +
  theme_bw()
```

## survival to eclosion
```{r main fig adult survival}
# TODO include sample size?
p <- list_modify(
  p,
  surv_pup = 
    dfs_viz$wide %>%
    filter(is.na(is.LPI),
           #!is.na(sex) # can keep?
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>%
    ggplot(aes(y = prop.ec,
               shape = interaction(pop, trt.type), 
               lty = pop,
               col = trt.type,
               group = interaction(pop, trt.type),
               x = trt.minT)) +
    geom_point(size = 2) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = prop.ec + se.ec,
                      ymin = prop.ec - se.ec),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(#y = "% pupa eclosed",
         title = "pupa to eclosion",
         x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    ylim(c(0, 1)) +
    theme_bw()
)

p$surv_pup
```

## collated fig

```{r}
p$surv_la + p$surv_pup +
  plot_layout(axes = "collect", guides = "collect") &
  labs(y = "% survival") &
  theme(plot.title = element_text(hjust = 0.5))
```


# (2) mass 

## pupal mass

drop LPIs bc mass isnt accurate

```{r main fig pup mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "pupa",
         is.na(is.LPI)
         )

pd <- position_dodge(0.1)

.plotfn <- function(x) {
  x %>%
    ggplot(aes(x = trt.minT, y = avg.mass,
               shape = interaction(pop, trt.type), 
               lty = pop, col = trt.type,
               group = interaction(pop, trt.type))) +
    geom_point(size = 2,
               position = pd) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  position = pd, width = 0.5, 
                  lty = "solid", show.legend = FALSE,
                  ) +
    labs(#y = "pupal mass (g)",
         #y = "mass (g)",
         title = "pupal mass (g)", 
         x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    theme_bw()
}


p <- list_modify(
  p,
  mass_pup = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    ylim(c(1, 6.4)),
  
  mass_pupsex = # sexes parsed
    dfs_viz$long %>%
    filter(!is.na(sex),
           instar == "pupa") %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    ylim(c(1, 6.8))
)

p$mass_pup

p$mass_pupsex # use this, per stats
```

### stats
```{r main stats pup mass}
df <- dfs_stats$dev_expt %>%
  filter(!is.na(mass.pupa),
         !is.na(sex),
         is.na(is.LPI)
         )

mods <- list(
  m1 = lm(mass.pupa ~ year*pop*trt.minT*sex,
           data = df)
)

anova(mods$m1) # ok to just focus on the 2 ways of trt/pop/year and main effect of sex

mods <- list_modify(
  mods,
  m2 = lm(mass.pupa ~ (year + pop + trt.minT + sex)^2,
           data = df),
  m3 = lm(mass.pupa ~ (year + pop + trt.minT)^2 + sex,
           data = df)
)

anova(mods$m3, mods$m2, mods$m1)
## adding model complexity doesnt do much
AIC(mods$m3, mods$m2, mods$m1)
## m3 = lowest AIC

anova(mods$m3) #%>% ResToCsv("pup-mass")
summary(mods$m3)
```

### supp
```{r supp fig pup mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "pupa",
         is.na(is.LPI)
         )

.plotfn <- function(x) {
  x %>%
    ggplot(aes(x = trt.minT, y = avg.mass,
               lty = pop, shape = pop,
               col = trt.type,
               group = interaction(pop, trt.type))) +
    geom_point(size = 2,
               #aes(size = sqrt(n)) # points are way too big
               ) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "pupal mass (g)",) +
    AddLeg_OnePanel2() +
    theme_bw()
}

# sexes combined
df %>%
  PrepAdultSS() %>%
  ungroup(starts_with("trt")) %>% 
  group_by(trt.type, trt.minT, .add = TRUE) %>%
  CalcDevSS() %>% #View()
  .plotfn() +
  facet_wrap(~ year)
  #ylim(c(1.8, 6.7))

# sexes parsed (use this one)
df %>%
  PrepAdultSS() %>%
  ungroup(starts_with("trt")) %>% 
  group_by(trt.type, trt.minT, sex, .add = TRUE) %>%
  CalcDevSS() %>%
  .plotfn() +
  facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
             scales = "free") 
  #ylim(c(1.8, 7.2))
```

## eclosion mass (WIP)

```{r main fig pup mass change}
df <- dfs_viz$long %>%
  filter(instar %in% c("pupa", "eclose", "change", "prop")) %>%
  group_by(across(c(pop, diet, starts_with("trt"), instar))) %>%
  CalcDevSS()

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.mass, x = trt.minT,
               lty = pop, col = trt.type,
               shape = interaction(pop, trt.type), 
               group = interaction(pop, trt.type))) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    theme_bw()  
}

p <- list_modify(
  p,
  mass_perat =
    df %>%
    filter(instar == "prop") %>%
    .plotfn() +
    labs(#y = "pupal:eclosion mass",
         #y = "mass ratio",
         title = "pupa:eclosion mass"),
  
  mass_pepct =
    df %>%
    mutate(avg.mass = avg.mass * 100,
           se.mass  = se.mass * 100) %>%
    filter(instar == "change") %>%
    .plotfn() +
    labs(#y = "% mass lost\n(pupa to eclosion)",
         #y = "% mass lost",
         title = "pupa to eclosion\n(% mass lost)")
)

p$mass_perat

p$mass_pepct
```
## collated fig

```{r}
# TODO finalise pup:ec metric to use
p$mass_pup + p$mass_perat + p$mass_pepct +
    plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank() # TODO remove once decided on P:E
        )
```

# (3) development time

## time to pupation
```{r main fig pup dev rate}
p <- list_modify(
  p,
  dev_la =
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    filter(instar == "pupa") %>%
    ggplot(aes(y = avg.devrate, x = trt.minT,
               lty = pop, col = trt.type,
               shape = interaction(pop, trt.type), 
               group = interaction(pop, trt.type))) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(#y = "larval development rate (1/days)",
         y = "development rate (1/days)",
         title = "hatch to pupation",
         x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    #ylim(c(0.025, 0.05)) +
    theme_bw()
)

p$dev_la
```

### stats
```{r main stats pup dev rate}
mods <- list(
  m1 = lm(1/tt.pupa ~ year*pop*trt.minT,
           data = dfs_stats$dev_expt)
)

anova(mods$m1) # ok to focus on 2ways

mods <- list_modify(
  mods,
  m2 = lm(1/tt.pupa ~ (year + pop + trt.minT)^2,
           dfs_stats$dev_expt)
)

#AIC(mods$m2, mods$m1)
anova(mods$m2) #%>% ResToCsv("pup-devrate")
summary(mods$m2)
```

### supp
```{r supp fig pup dev rate}
dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa") %>%
  ggplot(aes(y = avg.devrate,
             shape = pop, lty = pop,
             x = trt.minT,
             col = trt.type,
             group = interaction(pop, trt.type)
  )) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                    ymin = avg.devrate - se.devrate),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  labs(y = "larval development rate (1/days)") +
  AddLeg_OnePanel2() +
  facet_wrap(~ year) +
  theme_bw()
```


## pupal duration
```{r main fig pup duration}
p <- list_modify(
  p,
  dev_pup = 
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    filter(instar == "eclose") %>%
    # mutate(avg.tt = avg.tt - 15,
    #        se.tt = se.tt - 15) %>%
    ggplot(aes(y = avg.tt, x = trt.minT,
               lty = pop, col = trt.type,
               shape = interaction(pop, trt.type), group = interaction(pop, trt.type))) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(#y = "pupal development rate (1/days)",
         y = "duration (days)",
         title = "pupal duration",
         x = p_scales$xlab_minT) +
    AddLeg_OnePanel() +
    #ylim(c(0.025, 0.05)) +
    theme_bw()
)

p$dev_pup
```

## collated fig

```{r}
p$dev_la + p$dev_pup + p$dev_ad1 +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5))
```


# (4) longevity (WIP)

## longevity

```{r main fig adult longevity}
.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               shape = interaction(pop, trt.type),
               x = trt.minT, 
               lty = pop,
               col = trt.type,
               group = interaction(trt.type, pop)
    )) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    AddLeg_OnePanel() +
    labs(#y = "adult longevity (days)",
         y = "duration (days)",
         title = "adult longevity",
         x = p_scales$xlab_minT) +
    theme_bw() 
}

p <- list_modify(
  p,
  # w/o parsing of mate status
  dev_ad1 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long",
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    CalcDevSS() %>%
    .plotfn(), 
  
  dev_ad2 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long") %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    group_by(sex, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    facet_wrap(~sex),
  
  # w/ mate status parsing
  dev_ad3 = dfs_viz$long%>%
    filter(if.fridge == 0,
           instar == "long",
           #year != 2024
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>% 
    group_by(if.mated, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    #geom_point(aes(size = sqrt(n))) +
    facet_wrap(~if.mated, labeller = as_labeller(c(`0` = "unmated",
                                                   `1` = "mated")))
)

p$dev_ad1
p$dev_ad2
p$dev_ad3
```

### stats (WIP)

```{r main stats adult longevity}
# TODO:
# include sex? parse the mating? drop 2024? (same qns as the viz)
mods <- list(
  m1 = lm(tt.long ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt),
  m2 = lm(tt.long ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt %>% filter(if.mated == 1))
)

anova(mods$m1) # can drop if looking at everyone
anova(mods$m2) # sex matters if looking at mated bugs only...
```

### results table
```{r}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long") %>%
  # TODO include sex?
  CalcDevSS_ad() %>% 
  select(c(year, pop, trt.minT, trt.type, n,
           ends_with("tt"))) #%>% ResToCsv("longevity")
```

### supp
```{r supp fig adult longevity}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long") %>%
  CalcDevSS_ad() %>%
  ggplot(aes(y = avg.tt,
             shape = pop, lty = pop,
             x = trt.minT,
             col = trt.type,
             group = interaction(pop, trt.type)
  )) +
  geom_point(size = 2,
             #aes(size = sqrt(n)) # looks bad..
             ) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.tt + se.tt,
                    ymin = avg.tt - se.tt),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  labs(y = "adult longevity (days)") +
  AddLeg_OnePanel2() +
  facet_wrap(~ year) +
  theme_bw()
```

# ## (5) fitness 

## legend fn
```{r}
# plot fn
.plotfn <- function(x){
  x %>%
    ggplot(aes(x = trt.minT,
               lty = trt.mate, 
               shape = trt.mate,
               group = trt.mate,
               col = trt.mate)) +
    #facet_grid(~year) +
    theme_bw() +
    scale_shape_manual(values = p_scales$shp_trtsex, labels = p_scales$labs_trtmate) +
    scale_color_manual(values = p_scales$cols_trtsex, labels = p_scales$labs_trtmate) +
    scale_linetype_manual(values = p_scales$lty_trtsex, labels = p_scales$labs_trtmate) +
    scale_x_discrete(labels = c(p_scales$labs_trt)) +
    labs(col = "heat-treated sex", lty = "heat-treated sex", shape = "heat-treated sex",
         x = p_scales$xlab_minT) +
    guides(shape = guide_legend(override.aes = list(linetype = c("blank", "solid",
                                                                 "3223", "4224"))))
}
```

## # eggs laid

```{r main fig eggs laid}
pd = position_dodge(0.35)

.plotfn2 <- function(){
  list(
    aes(#y = avg.coll.perf, 
      y = avg.sqrt.coll.perf, 
    ),
    geom_point(size = 2, position = pd),
    geom_line(position = pd, show.legend = TRUE),
    # very overwhelming w error bars lol
    # geom_errorbar(aes(ymax = avg.coll.perf + se.coll.perf,
    #                   ymin = avg.coll.perf - se.coll.perf,),
    #               lty = "solid", width = 0.25),
    geom_errorbar(aes(ymax = avg.sqrt.coll.perf + se.sqrt.coll.perf,
                      ymin = avg.sqrt.coll.perf - se.sqrt.coll.perf),
                  lty = "solid", width = 0.5, position = pd,
                  show.legend = FALSE),
    # geom_text(aes(label = n.f.total),
    #           vjust = -0.25, hjust = -0.25, position = "jitter",
    #           show.legend = FALSE),
    # geom_text(aes(label = paste0(n.cages.total, "C, ", n.f.total, "F")),
    #           vjust = -0.5, hjust = -0.5, #position = "jitter",
    #           show.legend = FALSE
    #           ),
    labs(#caption = "mean±se; lab only; N = # F",
      #y = "sqrt [eggs laid per female]", 
      title = "sqrt [eggs laid per female]"
    )
  )
}

p <- list_modify(
  p,
  eggs_laid = 
    dfs_viz$eggs %>%
    .plotfn() +
    .plotfn2() +
    facet_grid(~year),
  
  eggs_laidnoyr = dfs_viz$eggs_noyr %>%
    .plotfn() +
    .plotfn2()
)

p$eggs_laid
p$eggs_laidnoyr
```

## # eggs hatched

```{r main fig eggs hatched}
.plotfn2 <- function(){
  list(
    aes(y = p.hatch),
    geom_errorbar(aes(ymax = p.hatch + se.hatch,
                      ymin = p.hatch - se.hatch,),
                  lty = "solid", width = 0.25, show.legend = FALSE),
    # geom_text(aes(label = n.total.coll),
    #           vjust = -0.5, hjust = -0.5, position = "jitter",
    #           show.legend = FALSE),
    # geom_text(aes(label = paste0(n.cages.total, "C,", n.f.total, "F")),
    #       vjust = -0.5, hjust = -0.5, size = 3
    #       #position = "jitter"
    #       ),
    #geom_point(aes(size=sqrt(n.total.coll))), # looks dumb
    geom_point(size = 2),
    geom_line(),
    labs(#caption = "lab only; overall hatch prop. per mating combination",
      #y = "proportion eggs hatched",
      title = "proportion eggs hatched")
  )
}

p <- list_modify(
  p,
  eggs_hatch = 
    dfs_viz$eggs %>%
    .plotfn() +
    .plotfn2() +
    facet_wrap(~ year),
  
    eggs_hatchnoyr = 
    dfs_viz$eggs_noyr %>%
    .plotfn() +
    .plotfn2()
)

p$eggs_hatch
p$eggs_hatchnoyr # do not combine years... LOL
```

## results table
```{r}
# TODO drop the colony F + expt M tents (from 2025?) but 
# this wasnt tracked in 2023...
dfs_viz$eggs %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg results")

dfs_viz$eggs_noyr %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg-results_noyr")
```

## collated fig
```{r collated eggs viz}
# yrs faceted
p$eggs_laid + p$eggs_hatch +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        legend.position = "bottom")

# years combined
p$eggs_laidnoyr + p$eggs_hatchnoyr +
  labs(caption = "years combined") +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.y = element_blank(),
        legend.position = "bottom")
```


# ## dump

## (ref) supp dev patchwork figs


```{r eval = FALSE}
# legend fn for patchwork plots (trt.type/yr)
AddLeg_Patchwork <- function(){
  list(
    labs(lty = "population", shape = "population"),
    scale_shape_manual(values = p_scales$shp_pop),
    scale_linetype_manual(values = p_scales$lty_pop),
    guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
  )
}
```

```{r supp fig pup survival patchwork, eval = FALSE}
.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "% survived to pupation") +
    AddLeg_Patchwork() +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

p1 <- dfs_viz$wide %>%
  filter(trt == 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = year, 
      group = pop)

p2 <- dfs_viz$wide %>%
  filter(trt != 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = trt,
      group = year,
      col = year) +
  labs(x = p_scales$xlab_minT) +
  scale_x_discrete(labels = p_scales$labs_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r supp fig pup dev rate patchwork, eval = FALSE}
df <- dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa")

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.devrate,
               shape = pop, lty = pop
               )) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "larval development rate (1/days)") +
    scale_color_brewer(palette = "Dark2") +
    AddLeg_Patchwork() +
    theme_bw()
}

p1 <- df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, group = interaction(pop, year), col = year) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0.029, 0.054))
```

```{r supp fig adult longevity patchwork, eval = FALSE}
df <- dfs_viz$long %>%
    filter(if.fridge == 0,
           instar == "long") %>%
    CalcDevSS_ad()

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               shape = pop, lty = pop
               )
           ) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    facet_grid(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "adult longevity (days)",) +
    theme_bw() +
    AddLeg_Patchwork() +
    ylim(c(4, 17.5))
}

p1 <-  df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year,
      group = pop) +
  labs(x = "year")

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt,
      col = year,
      group = interaction(year, pop)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1,2),
              axes = "collect", guides = "collect")
```