---
title: "2025 ntw tents tidying"
date: "2025-11-26"
---

# # preamble

prepping ntw data for 2025 wrangling

final code is purled to a `.R` script, rerun `purl` step as needed to update the wrangling for analyses

```{r set paths, eval=FALSE, include=FALSE, purl=FALSE}
# run this prior to other pathing stuff if things are breaking
here::i_am("_ntw/scripts/tidy-tents.Rmd")
```

```{r for purling, eval=FALSE, include=FALSE, purl=FALSE}
# rerun this chunk if purled .R script needs to be updated.
# save and apply updates to current .Rmd before purling!

knitr::purl("tidy-tents.Rmd", here::here("_ntw/scripts/R/tidy-tents.R"))

# creates an archive
# knitr::purl("data-tidying.Rmd", 
#             paste0(here::here(bin_paths$y25$ntw), "/archive/cleaning/", 
#                    format(Sys.time(), "%y%m%d-%H%M"), "_tidy-data.R"))
```

## load libs & pkgs

```{r about}
# ntw/tidy-tents.R

# knitted tidying code for 2025 ntw dev figs/analyses.
# source() this in corresponding analysis scripts.

# this script compiles ntw data from 2023-25 into cohesive units for easier wrangling.
```

```{r load libs}
library(data.table) # rbindlist
library(tidyverse)
library(here)

conflicted::conflicts_prefer(
  dplyr::select,
  dplyr::filter) 
```

## roadmap

preamble:
- for each yr, have adult counts (when added/removed), egg collection counts, egg hatch counts.
- bc my method of tracking the daily egg/moth #s is very convoluted, need to split + wrangle the dfs in a certain order before finally collating into the final df that can be used in `summarise()` for data analysis
  - note that 2023 data is also formatted a little differently than 2025 data (i.e. pairing was more deliberate), so it needs some additional pre-wrangling
- see also some random TODOs left thruout this section,,,

roadmap:
1. create stdised `info` (about mating pairs) and `events` dfs (these dfs will get recycled)
2. `events` gets created in the following manner to avoid incorrect duplication of jdates:
  a. `events` is first generated from `info` (to get counts of moths removed/added)
  b. later, `events` is overwritten by merged collection + hatching data, with
     moth counts extracted from `events` appended on
  c. hatching events are aligned with collection date (i.e. hatch date is dropped)
3. `events` then gets used to generate the daily running count of eggs/moths/etc
  a. counts are determined separately per cage and then recollated again
     (avoids wonkiness with `cumsum()` at the first/last values of the df)
4. 2023/2025 are wrangled separately (bc of column name diffs), then combined for final export.
5. 2024 is also handled manually bc of column mismatches that prevent usage of of the info generation functions. (i.e lack of m info)



## fns for info generation

note: revisiting fertile eggs
- currently, all collected eggs treated as fertile as long as males were present in the cage at some point.
- F can sequester sperm, not sure how long this lasts, though.

```{r tent data common functions}
# to assign cage mating info
## note: data needs to be "wide" (...f, ...m) to determine mate types
.GenerateCageMateInfo <- function(x){
  x %>%
    mutate(mate.type = case_when(trt.f == trt.m ~ "within",
                                 trt.f != trt.m ~ "between",
                                 is.na(trt.m) ~ "virgin"),
           # if both m+f are from col, then mate.pop = "col", otherwise "lab"
           mate.pop = case_when(mate.type == "within" | is.na(pop.m) ~ pop.f,
                                mate.type == "between" & !(pop.f == "field" | pop.m == "field") ~ "lab"), 
           mate.trt = case_when(mate.type == "within" | is.na(trt.m) ~ trt.f,
                                mate.type == "between" & trt.f > trt.m ~ trt.f,
                                mate.type == "between" & trt.m > trt.f ~ trt.m),
           mate.col = case_when(pop.f == "col" & pop.m == "col" ~ "both",
                                pop.f == "col" & pop.m == "lab" ~ "f",
                                pop.m == "col" & pop.f == "lab" ~ "m" ),
           mate.ishs = case_when(trt.f > trt.m | is.na(trt.m) ~ "f",
                               trt.m > trt.f | is.na(trt.f) ~ "m",
                               trt.f == trt.m & trt.f == 260 ~ "none",
                               TRUE ~ "both"),
           mate.istrt = case_when(mate.ishs %in% c("none", "both") ~ "both",
                                  TRUE ~ as.character(mate.ishs)),
           trt.type = case_when(mate.trt == 260 ~ "ctrl",
                                TRUE ~ "expt"),
           trt.ishs.f = case_when(trt.f == 260 ~ 0,
                               trt.f > 260 ~ 1),
           trt.ishs.m = case_when(trt.m == 260 ~ 0,
                               trt.m > 260 ~ 1))
} 


# to determine daily m/f counts for cage events
.GenerateCageMothCounts <- function(x){
  x %>%
    group_by(cage, jdate) %>%
    summarise(f.added = sum(sex == "f" & event == "added"),
              m.added = sum(sex == "m" & event == "added"),
              f.removed = sum(sex == "f" & event == "removed"),
              m.removed = sum(sex == "m" & event == "removed")) %>%
    drop_na(cage)
}

# appends cage mate info + daily moth events to daily egg events
## needs dfs_stdised$info and $events to exist alrdy
.AppendCageMothEvents <- function(x){
  x %>%
  merge(., dfs_stdised$events, all = TRUE) %>%
  merge(., dfs_stdised$info %>%
          select(c("cage", starts_with("mate"))) #%>% View()
        ,
        all = TRUE, by = "cage") %>%
  mutate(across(starts_with(c("eggs", "f.", "m.")), replace_na, 0))
}


# gets daily running counts of new/dead moths + eggs collected
## splits $events by cage to avoid cumsum wonkiness at the ends of the range, then recombines
.GenerateCageDailyEvents <- function(x){
  x %>%
    split(f = dfs_stdised$events$cage) %>% 
    lapply(., \(x){
      x %>%
        arrange(jdate) %>%
        mutate(f.diff = dplyr::lag(f.added, default = 0) - dplyr::lag(f.removed, default = 0),
               f.curr = 0,
               f.curr = cumsum(f.diff + f.curr),
               m.diff = dplyr::lag(m.added, default = 0) - dplyr::lag(m.removed, default = 0),
               m.curr = 0,
               m.curr = cumsum(m.diff + m.curr),
               
               # fix egg counts: excludes eggs from hatch proportion if no males present
               ## TODO: just kidding not doing this anymore bc females can sequester sperm!
               eggs.exclude = case_when(#f.curr > 0 & m.curr == 0 | # revisit?
                                          mate.type == "virgin" ~ 
                                          as.numeric(eggs.collected + eggs.exclude),
                                        TRUE ~ as.numeric(eggs.exclude)),
               eggs.collected = case_when(#f.curr > 0 & m.curr == 0 | # revisit?
                                            mate.type == "virgin" ~ 0,
                                          TRUE ~ as.numeric(eggs.collected)),
               eggs.coll = eggs.collected + eggs.exclude) %>%
        select(-ends_with(c(".removed", ".diff"))) %>%
        rename(eggs.fert = eggs.collected,
               eggs.unfert = eggs.exclude)
    }) %>%
    rbindlist()
}

# setup empty lists
  # dfs_stdised gets overwritten/is hardcoded into .AppendCageEvents so dont create here
sapply(c("dfs_imported", "dfs_export"),
       \(x) assign(x, list(), envir = .GlobalEnv))
```

# # 2025 data
## loading

```{r load 2025 tents data}
dfs_imported <- list_modify(
  dfs_imported,
  ad25 = read.csv(here("_ntw/data/clean/2025/tent-matings.csv"), header = TRUE),
  coll25 = read.csv(here("_ntw/data/clean/2025/tent-eggs.csv"), header = TRUE),
  hatch25 = read.csv(here("_ntw/data/clean/2025/tent-events.csv"), header = TRUE)
)
```

## stdise cage/egg events

- to get `info`: convert `ad25` from long to wide (i.e. as pairs; extract m/f separately then combine)

```{r prep 2025 tent data for wrangling}
### need to rerun the WHOLE chunk at once if updating $info to avoid dups later on in $events

dfs_stdised <- list(
  
  # cage info (mating pairs)
  # TODO maybe could tidyr::complete this but i dont get the nesting lol
  info =
  merge(
    dfs_imported[["ad25"]] %>% filter(sex == "f") %>%
      select(cage, id, pop, trt, sex) %>%
      pivot_wider(names_from = "sex", values_from = c("pop", "trt"), 
                  names_sep = ".") %>%
       select(-id) %>% distinct(),
    dfs_imported[["ad25"]] %>% filter(sex == "m") %>%
      select(cage, id, pop, trt, sex) %>%
      pivot_wider(names_from = "sex", values_from = c("pop", "trt"), 
                  names_sep = ".") %>% 
       select(-id) %>% distinct(),
    all = TRUE) %>% 
    drop_na(cage) %>% #View() #for events23 futzing
    .GenerateCageMateInfo() %>% #View()
    # some fixes
    filter(!(cage == "A8" & pop.m == "lab")) %>%
    mutate(mate.pop = case_when(cage %in% c("A6", "B16", "C7") ~ "lab",
                                TRUE ~ as.character(mate.pop))),
  
  # cage timelines (moth additions/removals)
  events = dfs_imported[["ad25"]] %>%
    pivot_longer(starts_with("jdate"),
                 names_to = c(".value", "event"),
                 names_sep = "\\.") %>% 
    drop_na("jdate") %>% #View()
    group_by(cage, jdate) %>%
    .GenerateCageMothCounts() #%>% View()
)


# cage timelines (append egg hatching + collection info)
dfs_stdised$events <- 
  dfs_imported[["coll25"]] %>%
  rename(jdate = jdate.collected) %>%
  merge(., dfs_imported[["hatch25"]] %>% 
          select(c("jdate.collected", "cage", "eggs.hatched")) %>%
          rename(jdate = jdate.collected) %>%
          group_by(jdate, cage) %>%
          summarise(eggs.hatched = sum(eggs.hatched)) %>%
          drop_na(),
        all = TRUE) %>% #View()
  .AppendCageMothEvents()
```

## collate

```{r}
# troubleshooting tents: A2, B2, C14, C4
dfs_stdised <- list_modify(
  dfs_stdised, 
  tents = 
    dfs_stdised$events[order(dfs_stdised$events$cage, dfs_stdised$events$jdate),] %>%
    .GenerateCageDailyEvents()
)
```

## export
```{r}
dfs_export <- list_modify(
  dfs_export,
  tents25 = dfs_stdised$tents %>% 
    merge(.,
          dfs_stdised$info %>% select(c("cage", starts_with("trt"))),
          all = TRUE) %>%
    mutate(year = 2025)
  )
```


# # 2023 data

## loading
```{r load 2023 tent data}
dfs_imported <- list_modify(
  dfs_imported,
  ad23 = 
    read.csv(here("_ntw/data/clean/2023/tent-matings.csv"), header = TRUE) %>%
    # drop solo m/f's
        mutate(id.tent = case_when(id.tent == "107-cage" & sex == "f" ~ NA_character_,
                               TRUE ~ as.character(id.tent))) %>%
    filter(!(id %in% c(282, 284))), 
  
  coll23 = 
    read.csv(here("_ntw/data/clean/2023/tent-events.csv"), header = TRUE) %>%
    filter(id.tent != "107-I"),
  
  hatch23 = 
    read.csv(here("_ntw/data/clean/2023/tent-eggs.csv"), header = TRUE)
  )
  
dfs_imported[4:6] <- dfs_imported[4:6] %>%
  lapply(., \(x){
    x %>%
      rename(cage = id.tent)
  })
```

```{r}
# filter out useful columns
dfs_filtered <- list(
  ad23 = dfs_imported[["ad23"]] %>%
    select(-"track"),
    # select(c("cage", "id.pair", starts_with("jdate"), ends_with(c(".m", ".f"))),
    #        -starts_with("track")),
  coll23 = dfs_imported[["coll23"]] %>%
    select(c("cage", "n.coll", "jdate")),
  hatch23 = dfs_imported[["hatch23"]] %>%
    select(c("cage", "n.hatch", "jdate.coll")))
```

## stdise cage/egg events

- to get `info`: parse out sex-specific mate pair info

### generate dfs
```{r prep 2023 tent data for wrangling}
## rerun the whole chunk to regenerate dfs_tents$events, $info
dfs_stdised <- list(
  info = 
    dfs_filtered[["ad23"]] %>%
    pivot_wider(id_cols = c("cage", "id.pair", #"ctrl", "if.fridge"
                            ), 
                names_from = "sex", values_from = c("trt", "pop"),
                names_sep = ".") %>% #View()
    select(c("cage", starts_with(c("trt", "pop")))) %>%
    distinct() %>%
    .GenerateCageMateInfo() %>% #View() %>%
    # fixes (NOT solo f tents: 301P, 301U, 107H...)
    filter(!(mate.type == "virgin" & !(cage %in% c("107-cage", "301-T"))),
           !(cage == "301-M" & pop.f == "lab")) ,
  
  events = 
    dfs_filtered[["ad23"]] %>%
    pivot_longer(starts_with("jdate"),
                 names_to = c(".value", "event"),
                 names_sep = "\\.") %>%
    mutate(event = case_when(event == "paired" ~ "added",
                             event == "died" ~ "removed")
    ) %>% 
    drop_na(jdate, event) %>% 
    .GenerateCageMothCounts()
)

# overwrite moth events (append egg info)
dfs_stdised$events <- 
  dfs_filtered[["coll23"]] %>%
  rename(eggs.collected = n.coll) %>%
  merge(., dfs_filtered[["hatch23"]] %>% 
          rename(jdate = jdate.coll) %>%
          group_by(cage, jdate) %>%
          summarise(eggs.hatched = sum(n.hatch)) %>%
          drop_na(),
        all = TRUE) %>% 
  .AppendCageMothEvents() %>%
  mutate(eggs.exclude = 0) # a column in 2025 but not 2023
```

## collate
```{r}
dfs_stdised <- list_modify(
  dfs_stdised,
  tents = 
    dfs_stdised$events[order(dfs_stdised$events$cage, dfs_stdised$events$jdate),] %>%
    .GenerateCageDailyEvents()
)
```

## export
```{r}
dfs_export <- list_modify(
  dfs_export,
  tents23 = dfs_stdised$tents %>% 
    merge(.,
          dfs_stdised$info %>% select(c("cage", starts_with("trt"))),
          all = TRUE) %>%
    mutate(year = 2023)
)
```

# # 2024 data

- mainly just matching the columns for 2023 + 2025 bc all F are virgin

## import
```{r}
dfs_imported = list_modify(
  dfs_imported,
  ad24 = read.csv(here("_ntw/data/clean/2024/tent-matings.csv"), header = TRUE),
  coll24 = read.csv(here("_ntw/data/clean/2024/tent-events.csv"), header = TRUE)
)
```

## stdise

standardise manually bc 2024 dfs arent compatible with the info generation fns (`NA` info for males/differences in egg info breaks the fns)

```{r}
dfs_stdised = list(
  info = dfs_imported[["ad24"]] %>% 
    # fix 2099: sub in its new cage for longevity calcs
    filter(!(id == 2099 & cage == "BH1")) %>% 
    mutate(cage = case_when(id == 2099 ~ "BH1",
                            TRUE ~ cage)) %>%
    
    # match outcomes of.GenerateCageMateInfo()
    # mutate(trt.m = NA,
    #        pop.m = NA) %>% 
    # doesnt quite work.. bc cols.m are all NA?
    rename(pop.f = pop,
           trt.f = trt,
           id.f = id) %>%
    mutate(mate.type = "virgin",
           mate.pop = pop.f,
           mate.trt = trt.f,
           mate.col = NA,
           # mate.ishs = case_when(trt.f > 400 ~ "f"), 
           # mate.ishs = case_when(trt.f == 260 ~ NA_character_,
           #                       TRUE ~ "f"),
           mate.ishs = "f",
           mate.istrt = "f",
           trt.type = case_when(trt.f > 400 ~ "expt",
                                TRUE ~ "ctrl"),
           trt.ishs.f = case_when(trt.f > 400 ~ 1,
                               TRUE ~ 0),
    ),
  
  events = dfs_imported[["coll24"]] %>%
    # skip .AppendCageMothEvents
      ## m info added at export
    mutate(across(c("eggs.collected", 
                    "f.added", "f.active", "f.removed", "f.died"), replace_na, 0),
           #f.removed = f.removed + f.died,
           eggs.unfert = eggs.collected) %>%
    select(-c("f.curr", "f.died", "f.removed",
              "pop", "sex", "trt")) %>%
    
    # skip .GenerateCageDailyEvents
    rename(f.curr = f.active, # math alrdy calculated here
           eggs.coll = eggs.collected) 
)
```

## collate and export
```{r}
# add on missing cols for merging
dfs_export <- list_modify(
  dfs_export,
  tents24 = dfs_stdised$events %>%
    merge(., dfs_stdised$info %>% 
            select(c("cage", starts_with(c("trt", "mate")))) %>% distinct(), 
          by = c("cage")) %>%
    mutate(year = 2024,
           # missing .m stuff
           trt.ishs.m = NA, trt.m = NA,
           m.curr = 0, m.added = 0)
)
```

# # prep for final export

```{r export tent data}
# specify levels 
dfs_tidy <- list(
  tents = rbindlist(dfs_export, fill = TRUE)
  )
```

# # cleanup
```{r remove intermediate objects}
rm(dfs_imported, dfs_filtered, dfs_stdised,
   dfs_export,
   .GenerateCageDailyEvents, .GenerateCageMateInfo, .GenerateCageMothCounts,
   .AppendCageMothEvents)
```
