---
title: "tdt viz"
date: "2025-08-14"
---

# # preamble

- 2025-07-16
  - [>] prob need to rerun A now that ive fixed the decimal hours calcs finally LO L L...
- 2025-07-17 re: jdate vs dh
  - cohort B: after updates to cleaning script, `dh` is calculated in days (should be hours, ideally)... jdate is calculated in days (so just be careful w axes)
  - this will probably break some of the cohort A analyses lol
  
  
- WIP TO MOVE
  - R1 df prep (move to its own section in R1)
  - R1: kms 06-22
  - R1: km redos (06-27)
  - R1: surv props
  - R2/3 df prep
  - R2 kms
  - R2: surv props
  - R3: kms
  - R3: surv props

## loading

```{r}
# libraries
library(tidyverse)
library(survival) # for fitting 
library(survminer) # for plotting

# pathfinding
here::i_am("2025/tdt/analysis-viz.Rmd")
library(here)

source(here::here("set-paths.R"))
source(here(bin_paths$y25$tdt, "analysis-utils.R")) # helper functions
```

```{r}
# data import
source(here(bin_paths$y25$tdt, "wrangle_p.R"))

dfs <- bin_wrangled$km # for cohort B+ for R2+
```

## df prep
```{r}
dfs <- list_assign(dfs, 
                   group = dfs$data %>% 
                     filter(trt.duration > 0) %>%
                     group_by(cohort, trt.recover, trt.duration) %>% 
                     summarise(day.enter, day.recover, day.return) %>%
                     drop_na() %>% unique()
                   )
```

# # R3

## = 2025-08-11 pre mtg

### kms 

```{r}
# expt + ctrls
obj <- Surv(dfs$data$exit.fromenter, dfs$data$km.status)
fit <- survfit(obj ~ trt.recover + trt.duration, data = dfs$data)

ggsurvplot(fit, data = dfs$data,
           #facet.by = "trt.recover",
           facet.by = c("cohort", "trt.duration")
           #facet.by = "trt.duration"
           #color = "strata"
           ) +
  geom_vline(data = dfs$group, aes(xintercept = day.enter), color = "red") +
  geom_vline(data = dfs$group, aes(xintercept = day.recover), color = "skyblue") +
  geom_vline(data = dfs$group, aes(xintercept = day.return), color = "orange") +
  #scale_color_manual(values = c("#4B1D91", "#AB1488", "#ECC000", "#E7D39A")) + 
  scale_color_manual(values = c("#4B1D91", "#00A3B6","#F9C25C", "#A71B4B")) +  
  labs(caption = "R2-3 tdts, time = hrs from enter")
```

### survival props

```{r}
dfs <- list_assign(dfs, r1 = list())

dfs$r1 <- list_assign(dfs$r1, 
                      survprop1 = dfs$data %>%
                        calc.surv_ssB2(),
                      survprop1a = dfs$data %>%
                        calc.surv_ssB2a(),
                      survprop1b = dfs$data %>%
                        calc.surv_ssB2b(), ## cohort splits
                      )
```

```{r}
# (just blindly copied over code from surv props B attempt 2)

dfs$r1$survprop1 %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.duration, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "R2+3 proportion survived (cumulative)")

dfs$r1$survprop1 %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.5, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "R2+3 proportion survived (cumulative)") +
  theme_bw()

dfs$r1$survprop1a %>% # comment out cohort facet
# dfs$r1$survprop1b %>% # to use with cohort facet
  filter(status == "died") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.5, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(#cohort
             ~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "R2+3 proportion died (cumulative)") +
  theme_bw()


dfs$r1$survprop1a %>%
  filter(status == "died") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.8, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "R2+3 proportion died (each timept)") +
  theme_bw()

## the above but with cohort facet (comment out the other cohort as needed for faceting)
dfs$r1$survprop1b %>%
  filter(status == "died") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(cohort, 
                                 trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.8, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(cohort
             ~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "R2+3 proportion died (each timept)") +
  theme_bw()

```

=====

# # R2

## = 2025-07-16 cohort B fups

cohort B only

```{r}
dfs <- list_assign(dfs, r1 = list())

dfs$r1 <- list_assign(dfs$r1, 
                      allB = dfs$data %>% filter(cohort == "B"),
                      hotB = dfs$data %>% filter(is.40C == "Y"),
                      recB = dfs$data %>% filter(cohort == "B" & trt.duration > 0))
```

### km R2 only
```{r}
# survfit doesnt like lists lol
obj <- Surv(dfs$r1$allB$exit.fromenter, dfs$r1$allB$km.status)
fit <- survfit(obj ~ trt.recover + trt.duration, data = dfs$r1$allB)

group <- dfs$subsets$recs %>% 
  group_by(cohort, trt.recover, trt.duration) %>% 
  summarise(day.enter, day.recover, day.return) %>%
  filter(cohort == "B") %>%
  drop_na() %>% unique()

ggsurvplot(fit, data = dfs$r1$allB,
           #facet.by = "trt.recover",
           facet.by = "trt.duration",
           #color = "strata"
           ) +
  geom_vline(data = group, aes(xintercept = group$day.enter), color = "red") +
  geom_vline(data = group, aes(xintercept = group$day.recover), color = "skyblue") +
  geom_vline(data = group, aes(xintercept = group$day.return), color = "orange") +
  #scale_color_manual(values = c("#4B1D91", "#AB1488", "#ECC000", "#E7D39A")) + 
  scale_color_manual(values = c("#4B1D91", "#00A3B6","#F9C25C", "#A71B4B")) +  
  labs(caption = "R2 tdt, time = hrs from enter")
```

### km both rounds 

```{r}
obj <- Surv(dfs$data$exit.fromenter, dfs$data$km.status)
fit <- survfit(obj ~ trt.recover + trt.duration, data = dfs$data)
```

```{r}
group <- dfs$subsets$recs %>% 
  group_by(cohort, trt.recover, trt.duration) %>% 
  summarise(day.recover, day.return) %>%
  drop_na() %>% unique()

ggsurvplot(fit, data = dfs$data,
           facet.by = c("cohort", "trt.duration")) +
  geom_vline(data = group, aes(xintercept = day.recover), color = "skyblue") +
  geom_vline(data = group, aes(xintercept = day.return), color = "orange") +
  scale_color_manual(values = c("#4B1D91", "#00A3B6","#F9C25C", "#A71B4B")) 

  
  ggsurvplot(fit, data = dfs$data,
             facet.by = c("cohort", "trt.recover")) +
  geom_vline(data = group, aes(xintercept = day.recover), color = "skyblue") +
  geom_vline(data = group, aes(xintercept = day.return), color = "orange") +
  scale_color_manual(values = c("orange", "skyblue", "#009E73"))


ggsurvplot(fit, data = dfs$data,
           facet.by = "cohort",
           #palette = "jco",
           #palette = c("red","blue", "red", "red", "red", "red", "red", "red", "red", "red"),
           # palette = c("#4B1D91", "#4B1D91", "#4B1D91", "#00A3B6","#00A3B6","#00A3B6", "#F9C25C","#F9C25C","#F9C25C", "#A71B4B",
           #             "#4B1D91", "#4B1D91", "#4B1D91", "#00A3B6","#00A3B6","#00A3B6", "#F9C25C","#F9C25C","#F9C25C", "#A71B4B"),
           # linetype = c(1, 2, 3, 1, 2, 3, 1, 2, 3, 1,
           #              1, 2, 3, 1, 2, 3, 1, 2, 3, 1)
           linetype = "strata",
           font.legend = c(5)
           ) +
  geom_vline(data = group, aes(xintercept = day.recover), color = "skyblue") +
  geom_vline(data = group, aes(xintercept = day.return), color = "orange") +
  scale_color_manual(values = c("#4B1D91", "#4B1D91", "#4B1D91", "#00A3B6","#00A3B6","#00A3B6", "#F9C25C","#F9C25C","#F9C25C", "#A71B4B")) +
  scale_linetype_manual(values = c(2, 1, 3, 2, 1, 3, 2, 1, 3, 2,
                                   2, 1, 3, 2, 1, 3, 2, 1, 3, 2)) # not quite lining up right due to the cohort strata..
```


### survival props

#### attempt 2

```{r}
dfs$r1 <- list_assign(dfs$r1, survprop2 = dfs$r1$allB %>%
                        calc.surv_ssB2(),
                      survprop2a = dfs$r1$allB %>%
                        calc.surv_ssB2a(),
                      )
```

```{r}
dfs$r1$survprop2 %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.duration, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion survived (cumulative)")

dfs$r1$survprop2 %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.5, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion survived (cumulative)") +
  theme_bw()

## not sure why 26 is jumping up at hs0 lol

dfs$r1$survprop2 %>%
  filter(status == "died") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.5, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion died (cumulative)") +
  theme_bw()


dfs$r1$survprop2a %>%
  filter(status == "died") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "hs72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.8, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion died (each timept)") +
  theme_bw()

```

#### attempt 1

```{r}
dfs$r1 <- list_assign(dfs$r1, survprop = dfs$r1$allB %>%
                        calc.surv_ssB()
                      )
```

```{r}
## TODO 2025-07-20 mull over these a little more... not sure if the math is right

dfs$r1$survprop %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.duration, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion survived")

dfs$r1$survprop %>%
  filter(status == "surv") %>%
  filter(trt >= 40) %>% #View()
  #filter(timept != "return72") %>% # kinda interesting... ideally this should be more symmetrical tho (i.e. +48 hr instead of +24)
  drop_na() %>%
  mutate(trt.duration = as.factor(trt.duration), trt.recover = as.factor(trt.recover)) %>%
  ggplot(aes(x = timept, y = prop, 
             color = trt.recover,
             lty = trt.duration,
             group = interaction(trt.recover, trt.duration)
         ))+
  geom_point() +
  geom_line() +
  geom_text(aes(label = n),
            vjust = 1.5, show.legend = FALSE) +
  #geom_point(aes(interaction = c("prop", "timept"))) +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  labs(y = "proportion survived")
# 48h at 19C is better than the other temps; 24h is worse for 19C than the other temps ?
# 48h seems to slow things down.. 
```

=====

# # R1 archive

R1 is handled too differently so its analysis stuff gets moved here

## = df prep

```{r}
# set up surv objects
survObj <- list()

# model fits
fits <- list()

# for dfs
testing <- list()

dfs <- bin_wrangled$kmA
```

## 2025-06-27 post mtg fups

TODO

- [ ] look at survival props for 48h intervals
- [ ] reviz km relative to recovery times...

### look at surv props

#### attempt 3

2025-06-29

```{r}
## code survival
testing <- list_assign(testing, r3 = list())


testing$r3 <- list_assign(testing$r3, df = dfs$data %>%
                         filter(is.40C == "Y") %>%
                         mutate(# ctrls should not have return/recover times
                                # dh.recover = case_when(trt < 100 ~ NA_real_,
                                #                        TRUE ~ dh.recover),
                                # dh.return = case_when(trt < 100 ~ NA_real_,
                                #                       TRUE ~ dh.return),
                                
                                dh.enter48 = dh.enter + 2,
                                dh.return48 = case_when(trt > 100 ~ dh.return + 2,
                                                        trt < 100 ~ dh.enter48 + 2),
                                
                                status.enter = case_when(dh.exit > dh.enter ~ 1,
                                                         TRUE ~ 0),
                                status.enter48 = case_when(dh.exit > dh.enter48 ~ 1,
                                                           TRUE ~ 0),
                                status.return = case_when(trt > 100 & dh.exit > dh.return ~ 1,
                                                          trt < 100 ~ NA_real_,
                                                          TRUE ~ 0),
                                status.return48 = case_when(dh.exit > dh.return48 ~ 1,
                                                            TRUE ~ 0)
                         ))

```

##### attempt 3.6 (pt plot **)

making point plots

```{r}
testing$r3 <- list_assign(testing$r3, 
                          countsw = testing$r3$df %>% 
                            group_by(trt.duration, trt.recover) %>%
                            summarise(n.enter = sum(status.enter == 1),
                                      n.enter48 = sum(status.enter48 == 1),
                                      n.return = sum(status.return == 1),
                                      n.return48 = sum(status.return48 == 1)))

testing$r3 <- list_assign(testing$r3,
                          countsl = testing$r3$countsw %>%
                            pivot_longer(cols = starts_with("n."),
                                         names_to = c(".value", "timept"), 
                                         names_sep = "\\."))

testing$r3 <- list_assign(testing$r3, ss6 = 
                          testing$r3$countsl %>%
  # mutate(grp.trtdur = paste0(trt.recover, "_", trt.duration),
  #        grp.durtime = paste0(trt.duration, "_", timept)) %>%
  group_by(trt.recover) %>% mutate(tot.trt = sum(n, na.rm = TRUE)) %>% 
  group_by(trt.duration) %>% mutate(tot.dur = sum(n, na.rm = TRUE)) %>%
  group_by(timept) %>% mutate(tot.timept = sum(n, na.rm = TRUE)) %>%
  group_by(trt.recover, trt.duration) %>% mutate(tot.trtdur = sum(n, na.rm = TRUE)) %>%
  group_by(trt.recover, timept) %>% mutate(tot.trttime = sum(n, na.rm = TRUE)) %>%
  group_by(trt.duration, timept) %>% mutate(tot.durtime = sum(n, na.rm = TRUE)) %>%
  mutate(across(starts_with("tot."), ~ na_if(., 0))) %>%
  #select(-n) %>%
  unique()
  )


testing$r3 <- list_assign(testing$r3, ss6b = 
                            testing$r3$countsw %>%
                            group_by(trt.duration, trt.recover) %>%
                            mutate(prop.surv.enter48 = n.enter48/n.enter,
                                   prop.surv.return = n.return/n.enter48,
                                   prop.surv.return48 = case_when(trt.duration == 0 ~ n.return48/n.enter48,
                                                                  TRUE ~ n.return48/n.return),
                                   prop.died.enter48 = 1 - prop.surv.enter48,
                                   prop.died.return = 1 - prop.surv.return,
                                   prop.died.return48 = 1 - prop.surv.return48,
                                   ) %>% 
                            pivot_longer(cols = starts_with(c("prop")),
                                         names_to = c(".value", "status", "timept"), names_sep = "\\.") %>% 
                            mutate(n = case_when(timept == "enter48" ~ n.enter,
                                                 timept == "return" ~ n.enter48,
                                                 timept == "return48" ~ 
                                                   case_when(trt.duration == 0 ~ n.enter48,
                                                             TRUE ~ n.return)),
                                   se = sqrt((prop * (1 - prop))/ n)
                                   ## no se for proportions lol
                                   ) %>%
                            select(-starts_with("n.")) #%>% View()
)

```

```{r}
testing$r3$ss6b %>%
  mutate(trt.recover = factor(trt.recover, levels = c(19, 26, 33, 40)),
         trt.duration = factor(trt.duration)) %>%
  filter(status == "died") %>%
  ggplot(aes(y = prop, x = trt.recover,
             color = trt.duration,
             group = trt.duration)) +
  geom_point() +
  geom_line() +
  #geom_errorbar(aes(ymin = prop -  se, ymax = prop + se)) +
  facet_grid(~timept, drop = TRUE)

testing$r3$ss6b %>%
  mutate(trt.recover = factor(trt.recover, levels = c(19, 26, 33, 40)),
         trt.duration = factor(trt.duration)) %>%
  filter(status == "died",
         timept != "enter48"
         #trt.recover != 40
         ) %>%
  drop_na() %>%
  ggplot(aes(y = prop, x = trt.recover,
             color = trt.duration,
             group = trt.duration)) +
  geom_text(aes(label = n), vjust = -0.5, hjust = -0.2,
            show.legend = FALSE) + 
  geom_point() +
  geom_line() +
  #geom_errorbar(aes(ymin = prop -  se, ymax = prop + se)) +
  facet_grid(~timept, drop = TRUE, scales = "free",
             labeller = as_labeller(c(`return` = "0h after recovery",
                                   `return48` = "48h after recovery"))) +
  ylim(c(0, 1)) +
  labs(y = "proportion died",
       caption = "n = initial N")
```


```{r eval = FALSE}
### not quite right but getting there...
  ### would be way easier to just make a fig per timept then combine the timepts later??

testing$r3$countsw %>%
  mutate(prop = n.enter48/n.enter,
         trt.recover = factor(trt.recover),
         trt.duration = factor(trt.duration)) %>%
  ggplot(aes(y = prop, x = trt.recover,
             lty = trt.duration,
             group = trt.duration)) +
  geom_point() +
  geom_line()


testing$r3$ss6 %>%
  filter(timept != "enter") %>%
  ### maybe just summarise on the fly...
  mutate(trt.recover = factor(trt.recover),
         trt.duration = factor(trt.duration),
         prop = n/tot.trttime) %>%
  ggplot(aes(y = prop, x = trt.recover,
             lty = trt.duration,
             group = interaction(trt.duration)
             )) +
  geom_point() +
  geom_line() +
  facet_wrap(~timept)

```


##### attempt 3.5

```{r}
# trying to just get the raw counts

testing$r3 <- list_assign(testing$r3, counts = testing$test3 %>% 
                            select(c(trt.duration, trt.recover, starts_with("status"))) %>%
                            pivot_longer(cols = starts_with("status"),
                                         names_to = c(".value", "timept"),
                                         names_sep = "\\.") %>%
                            drop_na() %>%
                            mutate(status = case_when(status == 1 ~ "surv",
                                                      status == 0 ~ "died"))
                          )

```

```{r}
testing$r3$counts %>%
  ggplot(aes(x = timept, fill = interaction(status, trt.recover))) +
  geom_bar(position = "fill") +
  facet_grid(~trt.duration, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45))

testing$r3$counts %>%
  ggplot(aes(x = timept, fill = interaction(status, trt.duration))) +
  geom_bar(position = "fill") +
  facet_grid(~trt.recover, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45))

testing$r3$counts %>%
  ggplot(aes(x = as.factor(trt.recover), fill = interaction(status, trt.duration))) +
  geom_bar(position = "fill") +
  facet_grid(~timept, scales = "free") +
  scale_x_discrete(guide = guide_axis(angle = 45))

## this is it (more or less. needs some sample size #s in there, also missing the 24/48 LOL...)
  # tmp %>% # counts data
  # select(-c(trt.recover, contains(c("surv", "died")))) %>% 
  #   unique() %>% 
  #   summarise(enter = sum(n.tot.enter), 
  #             enter48 = sum(n.tot.enter48), 
  #             return = sum(n.tot.return, na.rm = TRUE), 
  #             return48 = sum(n.tot.return48))

testing$r3$counts %>%
  filter(!(timept == "enter" & trt.recover == 40 & status == "died")) %>%
  mutate(status = factor(status, levels = c("surv", "died"))) %>%
  ggplot(aes(x = status, 
             fill = interaction(trt.recover))) +
  geom_bar(position = "fill") +
  facet_grid(~timept, scales = "free",
             labeller = as_labeller(c(`enter` = "enter (N = 303)", `enter48` = "enter48 (N = 161)",
                                      `return` = "return (N = 109)", `return48` = "return48 (N = 68)"))) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  scale_fill_manual(values = c("#8DA0CB", "#66C2A5", "#E78AC3", "#FC8D62")) + # RColorBrewer::brewer.pal(4, "Set2")
  theme_bw()


### kinda better...
testing$r3$counts %>%
  #filter(!(timept == "enter" & trt.recover == 40 & status == "died")) %>%
  filter(timept != "enter") %>%
  mutate(status = factor(status, levels = c("surv", "died")),
         trt.duration = factor(trt.duration, levels = c("0", "24", "48")),
         trt.recover = factor(trt.recover)) %>%
  ggplot(aes(#x = interaction(as.factor(trt.duration), status), 
             #fill = as.factor(trt.recover)),
             x = interaction(trt.duration, trt.recover),
             fill = interaction(status, trt.recover)
         ))+
  geom_bar(position = "fill") +
  facet_grid(#~interaction(trt.duration, timept), # interesting but also a Lot
             ~timept,
             scales = "free",
             # labeller = as_labeller(c(`enter` = "enter (N = 303)", `enter48` = "enter48 (N = 161)",
             #                          `return` = "return (N = 109)", `return48` = "return48 (N = 68)"))
             ) +
  scale_x_discrete(guide = guide_axis(angle = 45), # _manual if only changing the guides
                   labels = c("24h", "48h", "24h", "48h", "24h", "48h", "0h")
                   ) +
  scale_fill_manual(values = c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", 
                               "#FDBF6F", "#FF7F00", "#FB9A99", "#E31A1C")) + # RColorBrewer::brewer.pal(8, "Paired")
  labs(y = "survival outcome", x = "recovery duration (h)", fill = "outcome x test temperature") +
  theme_bw()


testing$r3$counts %>%
  #filter(!(timept == "enter" & trt.recover == 40 & status == "died")) %>%
  filter(timept != "enter") %>%
  mutate(status = factor(status, levels = c("surv", "died")),
         trt.duration = factor(trt.duration, levels = c("0", "24", "48")),
         trt.recover = factor(trt.recover)) %>%
  ggplot(aes(#x = interaction(as.factor(trt.duration), status), 
             #fill = as.factor(trt.recover)),
             x = trt.duration,
             fill = trt.recover,
             alpha = status,
             group = interaction(status, trt.recover)
         ))+
  geom_bar(position = "fill") +
  facet_grid(#~interaction(trt.duration, timept), # interesting but also a Lot
             ~timept,
             scales = "free",
             # labeller = as_labeller(c(`enter` = "enter (N = 303)", `enter48` = "enter48 (N = 161)",
             #                          `return` = "return (N = 109)", `return48` = "return48 (N = 68)"))
             ) +
  # scale_x_discrete(guide = guide_axis(angle = 45), # _manual if only changing the guides
  #                  labels = c("24h", "48h", "24h", "48h", "24h", "48h", "0h")
  #                  ) +
  scale_fill_manual(values = c("#8DA0CB", "#66C2A5", "#E78AC3", "#FC8D62")) +
  scale_alpha_manual(values = c(0.7, 1)) +
  labs(y = "survival outcome", x = "recovery duration (h)", fill = "test temperature") +
  theme_bw()

testing$r3$counts %>%
  #filter(!(timept == "enter" & trt.recover == 40 & status == "died")) %>%
  filter(timept != "enter") %>%
  mutate(status = factor(status, levels = c("surv", "died")),
         trt.duration = factor(trt.duration, levels = c("0", "24", "48")),
         trt.recover = factor(trt.recover)) %>%
  ggplot(aes(#x = interaction(as.factor(trt.duration), status), 
             #fill = as.factor(trt.recover)),
             x = trt.recover,
             fill = trt.duration,
             alpha = status,
             #group = interaction(status, trt.recover)
         ))+
  geom_bar(position = "fill") +
  facet_grid(#~interaction(trt.duration, timept), # interesting but also a Lot
             ~timept,
             scales = "free",
             ) +
  scale_fill_manual(values = c( "#FC8D62","#66C2A5", "#8DA0CB")) +
  scale_alpha_manual(values = c(0.68, 1)) +
  labs(y = "survival outcome", x = "test temperature",
       fill = "recovery duration (h)", 
       alpha = "survival outcome"
       ) +
  theme_bw()

### meh not rlly it
testing$r3$counts %>%
  #filter(!(timept == "enter" & trt.recover == 40 & status == "died")) %>%
  filter(timept != "enter") %>%
  mutate(status = factor(status, levels = c("surv", "died")),
         trt.duration = factor(trt.duration, levels = c("0", "24", "48")),
         trt.recover = factor(trt.recover)) %>%
  ggplot(aes(
             #x = interaction(status, trt.duration),
             x = trt.duration,
             fill = trt.recover,
             #group = interaction(status, trt.recover)
         ))+
  geom_bar(#position = "fill",
           stat = "count", position = "dodge") +
  facet_grid(#~interaction(trt.duration, timept), # interesting but also a Lot
             status~timept,
             scales = "free", space = "free" ## for this to work it needs to be with geom_col on the summed data rip
             ) +
  #scale_fill_manual(values = c( "#FC8D62","#66C2A5", "#8DA0CB")) +
  labs(y = "survival outcome", x = "test temperature",
       fill = "recovery duration (h)", 
       alpha = "survival outcome"
       ) +
  scale_fill_manual(values = c("#8DA0CB", "#66C2A5", "#E78AC3", "#FC8D62")) +
  scale_x_discrete(guide = guide_axis(angle = 45)) +
  theme_bw()
```


##### attempt 3.4 (**)

```{r}
testfn.surv_ss4b <- function(longdata) {
  
  counts <- longdata %>%
    #tmp <- testing$test3 %>%
    group_by(trt.duration) %>%
    mutate(n.tot.enter = sum(status.enter == 1),
           n.tot.enter48 = sum(status.enter48 == 1),
           n.tot.return = sum(status.return == 1),
           n.tot.return48 = sum(status.return48 == 1)) %>%
    group_by(trt.duration, trt.recover) %>%
    mutate(n.surv.enter = sum(status.enter == 1),
           n.surv.enter48 = sum(status.enter48 == 1),
           n.surv.return = sum(status.return == 1),
           n.surv.return48 = sum(status.return48 == 1),
           n.died.enter = sum(status.enter == 0),
           n.died.enter48 = sum(status.enter48 == 0 & status.enter == 1),
           n.died.return = sum(status.return == 0 & status.enter48 == 1),
           n.died.return48 = case_when(trt.duration == 0 ~ sum(status.return48 == 0 & status.enter48 == 1),
                                       TRUE ~ sum(status.return48 == 0 & status.return == 1))) %>%
    select(c(trt.duration, trt.recover, starts_with("n."))) %>% 
    unique() %>% ungroup()
  
  ## calc props
  survprops <- counts %>%
    select(-contains("died")) %>%
    unique() %>%
    mutate(p.surv.enter48 = n.surv.enter48/n.tot.enter,
           p.surv.return = n.surv.return/n.tot.enter48,
           p.surv.return48 = case_when(trt.duration == 0 ~ n.surv.return48/n.tot.enter48,
                                     TRUE ~ n.surv.return48/n.tot.return)) %>%
    select(c(trt.duration, trt.recover, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") 
  
  diedprops <- counts %>%
    #tmp %>%
    select(-contains("surv")) %>% 
    mutate(p.died.enter48 = n.died.enter48/n.tot.enter,
           p.died.return = n.died.return/n.tot.enter48,
           p.died.return48 = case_when(trt.duration == 0 ~ n.died.return48/n.tot.enter48,
                                       TRUE ~ n.died.return48/n.tot.return)) %>%
    select(c(trt.duration, trt.recover, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") #%>% View()
  
  ss <- merge(survprops, diedprops, all = TRUE)
  
  ss <- counts %>%
    #tmp %>%
        pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") %>%
    unique() %>% #View()
    merge(., ss, all = TRUE)
}

#testfn.surv_ss4b(testing$r3$df) %>% View()

testing$r3 <- list_assign(testing$r3, ss4b = testfn.surv_ss4b(testing$r3$df))

```


```{r}
testing$r3$ss4b %>%
  drop_na() %>%
  ggplot(aes(x = timept, y = p, 
             #fill = interaction(trt.recover, status), # this works but its a lot
             fill = as.factor(trt.recover), alpha = status
             )) +
  geom_col() +
  scale_alpha_manual(values = c(1, 0.6)) +
  facet_grid(~trt.duration) +
  theme_minimal()


testing$r3$ss4b %>%
  drop_na() %>%
  ggplot(aes(x = interaction(timept), y = p, 
             #fill = interaction(trt.recover, status), # this works but its a lot
             fill = status,
             )) +
  geom_col() +
  scale_alpha_manual(values = c(1, 0.6)) +
  facet_grid(~trt.recover) +
  #facet_grid(~trt.duration) +
  theme_minimal()
```



```{r}
## colors deaths @ trt (i.e. lumps surivors)
testfn.surv_ss4a <- function(longdata) {
  
  counts <- longdata %>%
    #tmp <- testing$test3 %>%
    group_by(trt.duration) %>%
    mutate(n.tot.enter = sum(status.enter == 1),
           n.tot.enter48 = sum(status.enter48 == 1),
           n.tot.return = sum(status.return == 1),
           n.tot.return48 = sum(status.return48 == 1),
           n.surv.enter = sum(status.enter == 1),
           n.surv.enter48 = sum(status.enter48 == 1),
           n.surv.return = sum(status.return == 1),
           n.surv.return48 = sum(status.return48 == 1)) %>%
    group_by(trt.duration, trt.recover) %>%
    mutate(n.died.enter = sum(status.enter == 0),
           n.died.enter48 = sum(status.enter48 == 0 & status.enter == 1),
           n.died.return = sum(status.return == 0 & status.enter48 == 1),
           n.died.return48 = case_when(trt.duration == 0 ~ sum(status.return48 == 0 & status.enter48 == 1),
                                       TRUE ~ sum(status.return48 == 0 & status.return == 1))) %>%
    select(c(trt.duration, trt.recover, starts_with("n."))) %>% 
    unique() %>% ungroup()
  
  ## calc props
  survprops <- counts %>%
    select(-c(trt.recover, contains("died"))) %>%
    unique() %>%
    mutate(p.surv.enter48 = n.surv.enter48/n.tot.enter,
           p.surv.return = n.surv.return/n.tot.enter48,
           p.surv.return48 = case_when(trt.duration == 0 ~ n.surv.return48/n.tot.enter48,
                                     TRUE ~ n.surv.return48/n.tot.return)) %>%
    select(c(trt.duration, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") 
  
  diedprops <- counts %>%
    #tmp %>%
    select(-contains("surv")) %>% 
    mutate(p.died.enter48 = n.died.enter48/n.tot.enter,
           p.died.return = n.died.return/n.tot.enter48,
           p.died.return48 = case_when(trt.duration == 0 ~ n.died.return48/n.tot.enter48,
                                       TRUE ~ n.died.return48/n.tot.return)) %>%
    select(c(trt.duration, trt.recover, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") #%>% View()
    
  ## merges
  counts <- counts %>%
    #tmp <- tmp %>%
    pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.")
  
  miscs <- counts %>%
    filter(status != "surv") %>%
    merge(., diedprops) %>%
    filter(status != "surv") %>%
    select(-status)
    
  survs <- counts %>%
    filter(status == "surv") %>%
    merge(., survprops, all = TRUE) %>%
    select(-trt.recover) %>%
    rename(trt.recover = "status") %>%
    unique()
  
  ss <- merge(survs, miscs, all = TRUE)
   
   return(ss)
} 

testing$r3 <- list_assign(testing$r3, ss4a = testfn.surv_ss4a(testing$r3$df))

```

```{r}
testing$r3$ss4a %>%
  filter(trt.recover != "tot") %>%
  drop_na() %>%
  ggplot(aes(x = timept, y = p, fill = trt.recover)) +
  geom_bar(stat = "identity") +
  facet_grid(~trt.duration) +
  labs(fill = "test temp (if died)", ## i.e. purple are survivors
       ) 


testing$r3$ss4a %>%
  filter(trt.recover != "tot") %>%
  drop_na() %>%
  ggplot(aes(x = as.factor(trt.duration), y = p, fill = trt.recover)) +
  geom_bar(stat = "identity") +
  facet_grid(~timept) +
  labs(fill = "test temp (if died)", ## i.e. purple are survivors
       ) 
```



```{r}
## colors survivors @ trt (i.e. lumps deaths)
testfn.surv_ss4 <- function(longdata) {
  
  counts <- longdata %>%
    #tmp <- testing$test3 %>%
    group_by(trt.duration) %>%
    mutate(n.tot.enter = sum(status.enter == 1),
           n.tot.enter48 = sum(status.enter48 == 1),
           n.tot.return = sum(status.return == 1),
           n.tot.return48 = sum(status.return48 == 1),,
           # n.tot2.enter = sum(!is.na(status.enter)),
           # n.tot2.enter48 = sum(!is.na(status.enter48)),
           # n.tot2.return = sum(!is.na(status.return)),
           # n.tot2.return48 = sum(!is.na(status.return48)),
           n.died.enter = sum(status.enter == 0),
           n.died.enter48 = sum(status.enter48 == 0 & status.enter == 1),
           n.died.return = sum(status.return == 0 & status.enter48 == 1),
           n.died.return48 = case_when(trt.duration == 0 ~ sum(status.return48 == 0 & status.enter48 == 1),
                                       TRUE ~ sum(status.return48 == 0 & status.return == 1))) %>%
    group_by(trt.duration, trt.recover) %>%
    mutate(n.surv.enter = sum(status.enter == 1),
           n.surv.enter48 = sum(status.enter48 == 1),
           n.surv.return = sum(status.return == 1),
           n.surv.return48 = sum(status.return48 == 1)) %>%
    select(c(trt.duration, trt.recover, starts_with("n."))) %>% 
    unique() %>% ungroup()
  
  ## calc props
  survprops <- counts %>%
    select(-contains("died")) %>%
    mutate(p.surv.enter48 = n.surv.enter48/n.tot.enter,
           p.surv.return = n.surv.return/n.tot.enter48,
           p.surv.return48 = case_when(trt.duration == 0 ~ n.surv.return48/n.tot.enter48,
                                     TRUE ~ n.surv.return48/n.tot.return)) %>%
    select(c(trt.duration, trt.recover, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") 
  
  diedprops <- counts %>%
    #tmp %>%
    select(-c(trt.recover, contains("surv"))) %>% 
    unique() %>% #View() %>%
      ### troubleshooting counts
    # pivot_longer(cols = starts_with(c("n.")),
    #              names_prefix = "n\\.",
    #              names_to = c(".value", "timept"), names_sep ="\\.") %>% View()
    #group_by(trt.duration) %>%
    mutate(p.died.enter48 = n.died.enter48/n.tot.enter,
           p.died.return = n.died.return/n.tot.enter48,
           p.died.return48 = case_when(trt.duration == 0 ~ n.died.return48/n.tot.enter48,
                                       TRUE ~ n.died.return48/n.tot.return)) %>%
    select(c(trt.duration, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") #%>% View()
    
  ## merges
  counts <- counts %>%
    #tmp <- tmp %>%
    pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.")
  
  survs <- counts %>%
    filter(status == "surv") %>%
    merge(., survprops) %>%
    filter(status == "surv") %>%
    select(-status)
    
  miscs <- counts %>%
    filter(status != "surv") %>%
    merge(., diedprops, all = TRUE) %>%
    select(-trt.recover) %>%
    rename(trt.recover = "status") %>%
    unique()
  
  ss <- merge(survs, miscs, all = TRUE)
   
   return(ss)
} 

testing$r3 <- list_assign(testing$r3, ss4 = testfn.surv_ss4(testing$r3$df))

```


```{r}
# filter(testing$r3$ss4,
#        trt.duration == 24 & timept %in% c("enter48", "return")) #%>% View()

testing$r3$ss4 %>%
  filter(trt.recover != "tot") %>%
  drop_na() %>%
  ggplot(aes(x = timept, y = p, fill = trt.recover)) +
  geom_bar(stat = "identity") +
  facet_grid(~trt.duration)


testing$r3$ss4 %>%
  filter(trt.recover != "tot") %>%
  drop_na() %>%
  ggplot(aes(x = as.factor(trt.duration), y = p, fill = trt.recover)) +
  geom_bar(stat = "identity") +
  facet_grid(~timept)
  
```


##### attempt 3.3
```{r eval=FALSE}
testfn.surv_ss3 <- function(longdata) {
  
  counts <- longdata %>%
    tmp <- testing$test3 %>%
    group_by(trt.duration) %>%
    mutate(n.tot.enter = sum(status.enter == 1),
           n.tot.enter48 = sum(status.enter48 == 1),
           n.tot.return = sum(status.return == 1),
           n.tot.return48 = sum(status.return48 == 1),
           n.died.enter = sum(status.enter == 0),
           n.died.enter48 = sum(status.enter48 == 0),
           n.died.return = sum(status.return == 0),
           n.died.return48 = sum(status.return48 == 0)) %>%
    ungroup() %>% group_by(trt.duration, trt.recover) %>%
    mutate(n.surv.enter = sum(status.enter == 1),
           n.surv.enter48 = sum(status.enter48 == 1),
           n.surv.return = sum(status.return == 1),
           n.surv.return48 = sum(status.return48 == 1)) %>%
    select(c(trt.duration, trt.recover, starts_with("n."))) %>% 
    ungroup() %>% unique() #%>% View()

props <- counts %>%
  tmp2 <- tmp %>%
    mutate(p.surv.enter48 = n.surv.enter48/n.tot.enter,
           p.surv.return = n.surv.return/n.tot.enter48,
           p.surv.return48 = case_when(trt.duration == 0 ~ n.surv.return48/n.tot.enter48,
                                     TRUE ~ n.surv.return48/n.tot.return)) %>%
    group_by(trt.duration) %>%
    mutate(p.died.enter48 = n.died.enter48/n.tot.enter,
           p.died.return = 1 - p.surv.return,
           p.died.return48 = 1 - p.surv.return48,
           # p.died.return = n.died.return/n.tot.enter48,
           # p.died.return48 = case_when(trt.duration == 0 ~ n.died.return48/n.tot.enter48,
           #                           TRUE ~ n.died.return48/n.tot.return),
           
           # p.died.enter48 = 1 - n.tot.enter48/n.tot.enter,
           # p.died.return = 1 - n.tot.return/n.tot.enter48,
           # p.died.return48 = case_when(trt.duration == 0 ~ 1 - n.tot.return48/n.tot.enter48,
           #                           TRUE ~ 1 - n.tot.return48/n.tot.return)
           ) %>%
    #group_by(trt.recover) %>%
    # mutate(p.surv.enter48 = n.surv.enter48/n.tot.enter,
    #        p.surv.return = n.surv.return/n.tot.enter48,
    #        p.surv.return48 = case_when(trt.duration == 0 ~ n.tot.return48/n.tot.enter48,
    #                                  TRUE ~ n.surv.return48/n.tot.return),
           # p.surv.enter48 = n.surv.enter48/n.surv.enter,
           # p.surv.return = n.surv.return/n.surv.enter48,
           # p.surv.return48 = case_when(trt.duration == 0 ~ n.surv.return48/n.surv.enter48,
           #                           TRUE ~ n.surv.return48/n.surv.return)) %>%
    ungroup() %>% unique() %>% 
    select(c(trt.duration, trt.recover, starts_with("p."))) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.") %>% View()
  
  counts <- counts %>%
  tmp <- tmp %>%
    pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "status", "timept"),
                 names_sep = "\\.")

  ### separating survival and death #s   
  survs <- merge(counts, props) %>%
    filter(status == "surv") %>%
    select(-status)
    
  miscs <- merge(counts, props, all = TRUE) %>%
    
    filter(status != "surv") %>%
    select(-trt.recover) %>%
    rename(trt.recover = "status") %>%
    unique()
  
  ss <- merge(survs, miscs, all = TRUE)
   
   return(ss)
} 

testing$r3 <- list_assign(testing$r3, ss3 = testfn.surv_ss3(testing$r3$df))

```


```{r eval=FALSE}
filter(testing$r3$ss3,
       trt.duration == 24 & timept %in% c("enter48", "return")) %>% View()

testing$r3$ss3 %>%
  filter(trt.recover != "tot") %>%
  ggplot(aes(x = timept, y = p, fill = trt.recover)) +
  geom_bar(stat = "identity") +
  facet_wrap(~trt.duration)
```



##### attempt 3.2
```{r eval=FALSE}
## figure out survivors at each time pt
testfn.surv_ss2 <- function(longdata){
  
  ## survivors
  totalsurv <- longdata %>%
    tmp <- testing$test3 %>%
    group_by(trt.duration) %>%
    mutate(tot.enter = sum(status.enter == 1),
           tot.enter48 = sum(status.enter48 == 1),
           tot.return = sum(status.return == 1),
           tot.return48 = sum(status.return48 == 1),
           ) %>%
    group_by(trt.recover) %>%
    mutate(n.enter = sum(status.enter == 1),
           n.enter48 = sum(status.enter48 == 1),
           n.return = sum(status.return == 1),
           n.return48 = sum(status.return48 == 1)) %>%
    select(c(trt.duration, trt.recover, starts_with(c("tot.", "n.")))) %>%
    unique() %>%
    ungroup() %>% View()
  
  propsurv <- totalsurv %>% 
    #### ok this is overcounting a bit cuz its counting w/in each trt.recover.... i think
    #### but i think "totalsruv" was done corerectly...
    
    
    #tmp %>%
    group_by(trt.duration, trt.recover) %>%
    summarise(p.enter48 = n.enter48/n.enter,
              p.return = n.return/n.enter48,
              p.return48 = case_when(trt.duration == 0 ~ n.return48/n.enter48,
                                  TRUE ~ n.return48/n.return),
    ) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "timept"), names_sep = "\\.",
                 values_drop_na = TRUE)%>%
    mutate(trt.recover = as.character(trt.recover)) #%>% View()
  
  ## died
  totaldied <- longdata %>%
    #tmp3 <- testing$test3 %>%
    group_by(trt.duration) %>%
    summarise(n.enter = sum(status.enter == 0),
           n.enter48 = sum(status.enter48 == 0),
           n.return = sum(status.return == 0),
           n.return48 = sum(status.return48 == 0)) #%>% View()
  
  propdied <- totalsurv %>% 
    #tmp2 <- tmp %>%
    group_by(trt.duration) %>%
    mutate(tot.enter = sum(n.enter),
              tot.enter48 = sum(n.enter48),
              tot.return = sum(n.return),
              tot.return48 = sum(n.return48)) %>% #View()
    mutate(ps.enter48 = tot.enter48/tot.enter,
              ps.return = tot.return/tot.enter48,
              ps.return48 = case_when(trt.duration == 0 ~ tot.return48/tot.enter48,
                                  TRUE ~ tot.return48/tot.return)) %>% #View()
    summarise(p.enter48 = 1 - ps.enter48,
              p.return = 1 - ps.return,
              p.return48 = 1- ps.return48) %>% #View()
        pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "timept"), names_sep = "\\.",
                 #values_drop_na = TRUE
                 ) %>%
    mutate(trt.recover = "died") #%>% View()
   
  ### pivot totals
  totalsurv <- totalsurv %>%
        pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "timept"), names_sep = "\\.") %>%
    mutate(trt.recover = as.character(trt.recover))
    
  totaldied <- totaldied %>%
    #tmp4 <- tmp3 %>%
    pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "timept"), names_sep = "\\.",
                 values_drop_na = TRUE) %>%
    mutate(trt.recover = "died") #%>% View()
  
  ### merge
  sssurv <- merge(totalsurv, propsurv,
                  by = c("trt.duration", "trt.recover", "timept"),
                  #all.x = TRUE,
                  all = TRUE
                  )
  
  ssdied <- merge(totaldied, propdied,
                  #tmp2, tmp4,
                   by = c("trt.recover", 
                          "trt.duration", "timept"),
                  #all.x = TRUE,
                  all = TRUE
                  ) #%>% View()
  
  ss <- rbind(sssurv, ssdied) %>% 
    mutate(p = case_when(is.na(p) ~ 0,
                         TRUE ~ p)) %>%
    unique()
  
  return(ss)

}

testing$r3 <- list_assign(testing$r3, ss2 = testfn.surv_ss2(testing$r3$df))
```

```{r eval=FALSE}
  # ggplot(aes(y = p, x = interaction(pop, trt), fill = stage)) +
  # geom_bar(position = "stack", 
  #          #position = position_fill(reverse = TRUE),
  #          stat = "identity") + 
  # facet_grid(~interaction(pop,year), scale = "free", 
  #            #labeller = as_labeller(interaction(pop,year) = c("field 2023", "lab 2023", "field 2024", "lab 2024"))
  #            ) +

testing$r3$ss2 %>%
  ggplot(aes(x = timept, y = p, fill = trt.recover)) +
  geom_col(stat = "identity") +
  facet_grid(~trt.duration, drop = TRUE)

filter(testing$r3$ss2,
       trt.duration == 24 & timept == "enter48") %>% View()

```


##### attempt 3.1
```{r eval=FALSE}
#### ok so pmd needs to be treated as its own "trt" lol (for the fill)
testfn.surv_ss <- function(longdata){
  
  totals <- longdata %>%
    group_by(trt, trt.duration, trt.recover) %>%
    summarise(#n.initial = n(), # dont rly need bc we mostly just care abt "enter"
              n.enter = sum(status.enter == 1),
              n.enter48 = sum(status.enter48 == 1),
              n.return = sum(status.return == 1),
              n.return48 = sum(status.return48 == 1))
            
  props <- totals %>%
    group_by(trt, trt.duration, trt.recover) %>%
    summarise(p.surv.enter48 = n.enter48/n.enter,
              p.surv.return = n.return/n.enter48,
              p.surv.return48 = case_when(trt < 100 ~ n.return48/n.enter48,
                                     trt > 100 ~ n.return48/n.return),
              
              p.died.enter48 = 1 - p.surv.enter48,
              p.died.return = 1- p.surv.return,
              p.died.return48 = 1 - p.surv.return48
              ) %>%
    pivot_longer(cols = starts_with("p."),
                 names_to = c(".value", "type", "timept"), names_sep = "\\.",
                 values_drop_na = TRUE)
  
  totals <- totals %>%
    pivot_longer(cols = starts_with("n."),
                 names_to = c(".value", "timept"), names_sep = "\\.")
  
  ss <- merge(totals, props,
              by = c("trt", "trt.duration", "trt.recover", "timept"),
              all.x = TRUE
              ) # maybe drop 40 ctrl "return" bc they dont exist
  
  return(ss)
  
}

testing$r3 <- list_assign(testing$r3, ss = testfn.surv_ss(testing$r3$df))

```

```{r eval=FALSE}
testing$r3$ss %>%
  #filter(type != "died") %>% 
  select(-n) %>% 
  ggplot(aes(y = p, x = timept, fill = interaction(as.factor(trt.recover), type))) +
  geom_col(aes(position = "fill")) +
  #ggplot(aes(y = p, x = timept, fill = as.factor(trt.recover))) +
  #geom_col() +
  facet_wrap(~trt.duration, scales = "free_x")
```






#### attempt 2
2025-06-28

```{r eval=FALSE}
# TODO WIP calc survival props per 48h group
  ## i think thisll be easier to just combine some dfs together... (i.e. calc some things separately)
testing <- list_assign(testing, r2 = list())

## attempt 2: split into trt groups after some binaries
testing$r2 <- list_assign(testing$r2, df = dfs$data %>%
                         filter(is.40C == "Y") %>%
                         mutate(# alive at the start of the next time period?
                           status.atenter = case_when(dh.exit > dh.enter ~ 1,
                                                      TRUE ~ 0),
                           status.atrecover = case_when(dh.exit > dh.recover ~ 1,
                                                        TRUE ~ 0),
                           status.atreturn = case_when(dh.exit > dh.return ~ 1,
                                                       TRUE ~ 0),
                           # 48h after the time period
                           dh.enter2 = dh.enter + 2,
                           dh.recover2 = dh.recover + 2,
                           dh.return2 = dh.return + 2,
                           
                           # alive after 48h?
                           status.atreturn2 = case_when(dh.exit > dh.return2 ~ 1,
                                                        TRUE ~ 0)
                         ))

### summary stats
testing$r2 <- list_assign(testing$r2, ss_bytrt = testing$r2$df %>%
                                            group_by(trt, trt.recover, trt.duration) %>%
                                            summarise(n.tot = n(),
                                                      n.entered = sum(status.atenter == 1),
                                                      n.enter48 = sum(status.atrecover == 1),
                                                      n.recovered = sum(status.atreturn == 1),
                                                      n.return48 = sum(status.atreturn2 == 1)
                                            ) %>%
                                            ungroup()
                                          )
testing$r2 <-
  list_assign(testing$r2, ss_bydur = testing$r2$ss_bytrt %>%
                pivot_wider(names_from = trt.recover, 
                            values_from = starts_with("n."),
                            values_fn = sum,
                            #id_cols = values_fn(~ sum(.x, na.rm = TRUE))
                ) %>%
                group_by(trt.duration) %>%
                summarise(tot.enter48 = sum(across(contains("enter48")), na.rm = TRUE),
                          tot.entered = sum(across(contains("entered")), na.rm = TRUE),
                          tot.recovered = sum(across(contains("recovered")), na.rm = TRUE),
                          tot.return48 = sum(across(contains("return48")), na.rm = TRUE),
                          p.surv.atrec = tot.enter48/tot.entered,
                          p.died.atrec = 1 - p.surv.atrec,
                          p.surv.atret = tot.recovered/tot.enter48,
                          p.died.atret = 1 - p.surv.atret,
                          p.surv.atret48 = tot.return48/tot.recovered,
                          p.died.atret48 = 1 - p.surv.atret48
                          )
              )

testing$r2 <-
  list_assign(testing$r2, ss_bydur2 = testing$r2$ss_bydur %>%
                rename(n.tot.atrec = tot.enter48,
                       n.tot.atret = tot.recovered,
                       n.tot.atret48 = tot.return48) %>%
                pivot_longer(cols = starts_with(c("p.", "n.")),
                             names_to = c(".value", "type", "time"), names_sep = "\\.")
              )
```

```{r eval=FALSE}
testing$r2$ss_bydur2 %>%
  ggplot(aes(x = time, y = p,
             fill = type)) +
  geom_col() +
  geom_text(aes(label = n)) + ## "tot" is fuckin it up. it should be surv/died
  facet_wrap(~trt.duration) +
  scale_x_discrete(labels = c("48h after\n1st HS", "after <facet> hrs\nrecovery", "48h after\n2nd HS"),
                   guide = guide_axis(angle = 45)) +
  #theme(axis.text.x = element_text(angle = 45, vjust = -1)) + 
  labs(x = "timept",
       y = "proportion")
## time to add instars...
```



attempt 1 (read-only)

```{r eval=FALSE}
## attempt 1:
### does not work to combine dfs lol
dfs$data %>%
  filter(is.40C == "Y") %>%
  mutate(status.atenter = case_when(dh.exit > dh.enter ~ 1,
                                    TRUE ~ 0),
         status.atrecover = case_when(dh.exit > dh.recover ~ 1,
                                      TRUE ~ 0),
         status.atreturn = case_when(dh.exit > dh.return ~ 1,
                                     TRUE ~ 0)) %>%
  group_by(trt.enter) %>% ## 1st df
  mutate(n.tot = n(),
         n.entered = sum(status.atenter == 1),
         # atenter.tot = n(),
         # atenter.died = sum(status.atenter == 0)
         ) %>%
  ungroup() %>%
  group_by(trt.duration, trt.recover) %>% ## 2nd df
  mutate(n.torecover = sum(status.atrecover == 1), # idk if this should go up w/ the prev..
         n.toreturn = sum(status.atreturn == 1),
         # atrecover.tot = sum(status.atrecover = 1),
         # atrecover.died = sum(status.)
         ) %>%
  summarise(prop.survtoenter = n.entered/n.tot, # group by initial 48h
            prop.survtorec = n.torecover/n.entered,
            prop.survtoret = n.toreturn/n.torecover) %>%
            # i wish there were sample sizes!!!! (prob gota do this separately)
  unique() %>%
  View()
  
  
  
  group_by(trt.enter) %>%
  mutate(n.tothatch = n(),
         n.totenter = sum(!is.na(jdate.recover))) %>% # merp doesnt work for the ctrls lol
  ungroup() %>%
  group_by(trt.duration, trt.enter) %>%
  mutate(n.totrecover = sum(!is.na(jdate.return)))

```

## 2025-06-27 kms

```{r}
# all exptal trts together
obj <- Surv(dfs$expts$exit.fromenter, dfs$expts$km.status)
fit <- survfit(obj ~ trt.recover + trt.duration, data = dfs$expts)

ggsurvplot(fit, data = dfs$expts,
           conf.int = TRUE,
           #xlim = c(45823, 45841),
           facet.by = "trt.recover",
           #linetype = "strata",
           palette = "strata"
           ) +
  geom_vline(xintercept = 2, color = "black") +
  geom_vline(xintercept = 3, color = "#F8766D") +
  geom_vline(xintercept = 4, color = "#00BFC4") +
  labs(caption = "black = entered recovery; colored = returned to 40") +
  theme(plot.caption = element_text(size = 10))

```

```{r}
# everyone together
obj <- Surv(dfs$data$exit.fromenter, dfs$data$km.status)
fit <- survfit(obj ~ trt.lab + trt.duration + trt.recover, data = dfs$data)

ggsurvplot(fit, data = dfs$data,
           #conf.int = TRUE,
           #xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #linetype = "strata",
           palette = "strata"
           ) +
  geom_vline(xintercept = 2, color = "black") +
  geom_vline(xintercept = 3, color = "#F8766D") +
  geom_vline(xintercept = 4, color = "#00BFC4")

```



## 2025-06-16 kms 

(first pass)

```{r eval=FALSE, include=FALSE}
# fit all
fit <- survfit(Surv(dh.exit, km.status) ~ trt.enter + trt.recover + trt.duration, data = dfs$data)
fit <- survfit(Surv(dh.exit, km.status) ~ trt, data = dfs$data)

           # linetype = c("1", "1", "1", "1",
           #              "2", "2", "2",
           #              "3", "3", "3"),
           palette = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")
           )  


  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_linetype_manual(values = c("1", "1", "1", "1",
                                   "2", "2", "2",
                                   "3", "3", "3"))




ggsurvplot(fit, data = dfs$data,
           #conf.int = TRUE,
           xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #linetype = "strata",
           #palette = "strata")
           )

+ 
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1",
                                "deepskyblue", "skyblue1", "goldenrod1",
                                "deepskyblue", "skyblue1", "goldenrod1")) +
  scale_linetype_manual(values = c("1", "1", "1", "1",
                                   "2", "2", "2",
                                   "3", "3", "3"))



fit <- survfit(Surv(dh.exit, km.status) ~ trt.enter + trt.recover + trt.duration, data = dfs$data)
#fit <- survfit(coxph(Surv(km.exit, km.status) ~ trt, data = km.data))


ggsurvplot(fit, data = dfs$data, 
           #conf.int = TRUE,
           xlim = c(45823, 45839),
           facet.by = "trt.duration",
           #group.by = c("trt.recover", "trt.duration"),
           #palette = "trt.recover",
           palette = "strata"
           #linetype = "trt.duration",
           ) +
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", 
                                "deepskyblue", "skyblue1", "goldenrod1", "darkorange1")) +
  scale_fill_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", 
                                "deepskyblue", "skyblue1", "goldenrod1", "darkorange1"))

  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1"),
                     labels = c(19, 26, 33, 40)) +
  scale_linetype_manual(#values = c("0" = 1, "24" = 2, "48" = 3),
                        values = c(1, 2, 3),
                        labels = c("0", "24", "48"))
  
  
  
  
  
# colorRampPalette(c(startcol, endcol))(ngroups)
  
# autoplot(fit, data)+ scale_color_manual, full
  
  
  
  
  
ggsurvplot(fit, data = dfs$data,
           #conf.int = TRUE,
           xlim = c(45823, 45841),
           facet.by = "trt.duration",
           #palette = viridis_pal(option = "H", direction = -1)(10)
           palette = hcl.colors(10, palette = "Roma")
           #linetype = "strata",
           #palette = "strata")
           ) +
  labs(caption = "2025-06-18 15:57 data; starting from hatch")
```

```{r eval=FALSE, include=FALSE}
# try subsets
subset <- dfs$data %>% filter(,
                             #trt > 100 & trt < 200,
                             is.40 == "Y",
                             #is.ctrl == "Y"
                             #trt %in% c(133, 233)
                             )

fit <- survfit(Surv(dh.exit, km.status) ~ trt + trt.duration + trt.recover, data = subset)
#fit <- survfit(Surv(km.exit, km.status) ~ trt.recover + trt.duration, data = subset)


#summary(fit)
#autoplot(fit)

ggsurvplot(fit, data = subset, 
           conf.int = TRUE,
           xlim = c(45823, 45839),
           #facet.by = "trt.recover",
           group.by = c("trt.recover", "trt.duration"),
           #palette = "trt.recover",
           #linetype = "trt.duration",
           ) +
  scale_color_manual(values = c("deepskyblue", "skyblue1", "goldenrod1", "darkorange1"),
                     labels = c(19, 26, 33, 40)) +
  scale_linetype_manual(#values = c("0" = 1, "24" = 2, "48" = 3),
                        values = c(1, 2, 3),
                        labels = c("0", "24", "48"))
  
  
  # scale_color_manual(values = c(c("19", "119", "219") = "deepskyblue", 
  #                               c("26", "126", "226") = "skyblue1",
  #                               c("33", "133", "233") = "goldenrod1",
  #                               "40" = "darkorange1")) +
  # scale_linetype_manual(values = c(c("19", "26", "33", "40") = 1,
  #                                  c("119", "126", "133") = 2,
  #                                  c("219", "226", "233") = 3))

  # scale_linetype_manual(values = c("19" = 1, "26" = 1, "33" = 1, "40" = 1,
  #                                  "119" = 2, "126" = 2, "133" = 2,
  #                                  "219" = 3, "226" = 3, "233" = 3))
```

```{r eval=FALSE, include=FALSE}
fit <- survfit(coxph(Surv(dh.exit, km.status) ~ trt, data = subset))
#library(ggplot2)
autoplot(fit)
```
