---
title: "tdt stats"
date: "2025-06-16"
---

# # preamble & loading

```{r}
# libraries
library(tidyverse)
#library(MuMIn) # for survival nested model comparisons
library(survival) # for fitting 
library(survminer) # for plotting
library(coxme) # mixed effects coxphs

# pathfinding
here::i_am("2025/tdt/analysis-stats.Rmd")
library(here)

source(here::here("set-paths.R"))
source(here(bin_paths$y25$tdt, "analysis-utils.R")) # helper functions
```

## data import + df prep
```{r}
# data import
source(here(bin_paths$y25$tdt, "wrangle_p.R"))

dfs <- bin_wrangled$km # just data for cohort B+ and onward for R2+
```

```{r}
# initialising lists
for (l in c("survObjs", "fits")) {
  assign(l, list())
}
```

## todos & notes

notes

- 2025-07-17 re: jdate vs dh
  - cohort B: after updates to cleaning script, `dh` is calculated in days (should be hours, ideally)... jdate is calculated in days (so just be careful w axes)
  

=====

# # R3

includes cohort B + C

### df prep
```{r}
# cohort C subsetting
dfs <- list_assign(dfs,
                   onlyC = dfs$data %>% filter(cohort == "C"),
                   recC = dfs$data %>% filter(cohort == "C" & trt.duration > 0))
```

## = post 250811 mtg fups

- include cohort as fixed and shelf as random

```{r}
survObjs <- list_assign(survObjs, r2 = list())
fits <- list_assign(fits, r2 = list())
```

test out stuff w/ the whole group before moving into the 24h and 48 subsets... (see what things need to be included in the models)

### - coxph rec bugs w/ 40C
```{r}
survObjs$r2 <- list_assign(survObjs$r2,
                            fromEnter = Surv(dfs$subsets40$all$exit.fromenter, dfs$subsets40$all$km.status),
                            fromRecovery = Surv(dfs$subsets40$all$exit.fromrecover, dfs$subsets40$all$km.status),
                            fromReturn = Surv(dfs$subsets40$all$exit.fromreturn, dfs$subsets40$all$km.status)
                            )
#sdf <- dfs$subsets40$all #summary() gets too ambiguous when this is omitted lol
```

```{r}
# + cohort
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxph(survObjs$r2$fromEnter ~ trt.recover*as.factor(trt.duration) + cohort, 
                                         data = dfs$subsets40$all),
                       cph.recovery = coxph(survObjs$r2$fromRecovery ~ trt.recover*as.factor(trt.duration) + cohort,
                                            data = dfs$subsets40$all),
                       cph.return = coxph(survObjs$r2$fromReturn ~ trt.recover*as.factor(trt.duration) + cohort,
                                          data = dfs$subsets40$all)
                       )
# fits$r2 creation tests lol
#lapply(survObjs$r2, coxph(~ trt.recover*as.factor(trt.duration) + cohort,  data = dfs$subsets40$all))
# lapply(survObjs$r2, 
#        \(s) coxph(eval(paste0(s, " ~ trt.recover * as.factor(trt.duration) + cohort")), data = dfs$subsets40$all))

lapply(fits$r2, summary)

```
breaks with `+ (1 | shelf)`....


### - coxph rec bugs w/o 40C
use `subsets$recs` (hot bugs w/o 40 ctrl) instead of `subsets40$all` bc stats get weird when trt.duration == 0 is included...

```{r}
survObjs$r2 <- list_assign(survObjs$r2,
                            fromEnter = Surv(dfs$subsets$recs$exit.fromenter, dfs$subsets$recs$km.status),
                            fromRecovery = Surv(dfs$subsets$recs$exit.fromrecover, dfs$subsets$recs$km.status),
                            fromReturn = Surv(dfs$subsets$recs$exit.fromreturn, dfs$subsets$recs$km.status)
                            )
```
```{r}
# + cohort
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxph(survObjs$r2$fromEnter ~ trt.recover*as.factor(trt.duration)+ cohort, 
                                         data = dfs$subsets$recs),
                       cph.recovery = coxph(survObjs$r2$fromRecovery ~ trt.recover*as.factor(trt.duration) + cohort,
                                            data = dfs$subsets$recs),
                       cph.return = coxph(survObjs$r2$fromReturn ~ trt.recover*as.factor(trt.duration) + cohort,
                                          data = dfs$subsets$recs)
                       )

lapply(fits$r2, summary)
```

```{r}
# + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ trt.recover*as.factor(trt.duration) + (1|shelf), 
                                         data = dfs$subsets$recs),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ trt.recover*as.factor(trt.duration) + (1|shelf),
                                            data = dfs$subsets$recs),
                       cph.return = coxme(survObjs$r2$fromReturn ~ trt.recover*as.factor(trt.duration) + (1|shelf),
                                          data = dfs$subsets$recs)
                       )

lapply(fits$r2, summary)
```
```{r}
# + cohort + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ 
                                           trt.recover*as.factor(trt.duration) + cohort + (1|shelf), 
                                         data = dfs$subsets$recs),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ 
                                              trt.recover*as.factor(trt.duration) + cohort + (1|shelf),
                                            data = dfs$subsets$recs),
                       cph.return = coxme(survObjs$r2$fromReturn ~ 
                                            trt.recover*as.factor(trt.duration) + cohort + (1|shelf),
                                          data = dfs$subsets$recs)
                       )

lapply(fits$r2, summary)
```



### - coxph 24h

```{r}
survObjs$r2 <- list_assign(survObjs$r2,
                            fromEnter = Surv(dfs$subsets40$rec24$exit.fromenter, dfs$subsets40$rec24$km.status),
                            fromRecovery = Surv(dfs$subsets40$rec24$exit.fromrecover, dfs$subsets40$rec24$km.status),
                            fromReturn = Surv(dfs$subsets40$rec24$exit.fromreturn, dfs$subsets40$rec24$km.status)
                            )
```
```{r}
# + cohort
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxph(survObjs$r2$fromEnter ~ trt.recover + cohort, 
                                         data = dfs$subsets40$rec24),
                       cph.recovery = coxph(survObjs$r2$fromRecovery ~ trt.recover + cohort,
                                            data = dfs$subsets40$rec24),
                       cph.return = coxph(survObjs$r2$fromReturn ~ trt.recover + cohort,
                                          data = dfs$subsets40$rec24)
                       )
lapply(fits$r2, summary)
```
```{r}
# + recT + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ trt.recover + (1|shelf), 
                                         data = dfs$subsets40$rec24),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ trt.recover + (1|shelf),
                                            data = dfs$subsets40$rec24),
                       cph.return = coxme(survObjs$r2$fromReturn ~ trt.recover + (1|shelf),
                                          data = dfs$subsets40$rec24)
                       )

lapply(fits$r2, summary)
```

```{r}
# + recT + cohort + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ trt.recover + (1|shelf) + cohort, 
                                         data = dfs$subsets40$rec24),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ trt.recover + (1|shelf) + cohort,
                                            data = dfs$subsets40$rec24),
                       cph.return = coxme(survObjs$r2$fromReturn ~ trt.recover + (1|shelf) + cohort,
                                          data = dfs$subsets40$rec24)
                       )

lapply(fits$r2, summary)
```
### - coxph 48h
```{r}
survObjs$r2 <- list_assign(survObjs$r2,
                            fromEnter = Surv(dfs$subsets40$rec48$exit.fromenter, dfs$subsets40$rec48$km.status),
                            fromRecovery = Surv(dfs$subsets40$rec48$exit.fromrecover, dfs$subsets40$rec48$km.status),
                            fromReturn = Surv(dfs$subsets40$rec48$exit.fromreturn, dfs$subsets40$rec48$km.status)
                            )
```
```{r}
# + cohort
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxph(survObjs$r2$fromEnter ~ trt.recover + cohort, 
                                         data = dfs$subsets40$rec48),
                       cph.recovery = coxph(survObjs$r2$fromRecovery ~ trt.recover + cohort,
                                            data = dfs$subsets40$rec48),
                       cph.return = coxph(survObjs$r2$fromReturn ~ trt.recover + cohort,
                                          data = dfs$subsets40$rec48)
                       )
lapply(fits$r2, summary)
```
```{r}
# + recT + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ trt.recover + (1|shelf), 
                                         data = dfs$subsets40$rec48),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ trt.recover + (1|shelf),
                                            data = dfs$subsets40$rec48),
                       cph.return = coxme(survObjs$r2$fromReturn ~ trt.recover + (1|shelf),
                                          data = dfs$subsets40$rec48)
                       )

lapply(fits$r2, summary)
```

```{r}
# + recT + cohort + (1|shelf)
fits$r2 <- list_assign(fits$r2,
                       cph.enter = coxme(survObjs$r2$fromEnter ~ trt.recover + (1|shelf) + cohort, 
                                         data = dfs$subsets40$rec48),
                       cph.recovery = coxme(survObjs$r2$fromRecovery ~ trt.recover + (1|shelf) + cohort,
                                            data = dfs$subsets40$rec48),
                       cph.return = coxme(survObjs$r2$fromReturn ~ trt.recover + (1|shelf) + cohort,
                                          data = dfs$subsets40$rec48)
                       )

lapply(fits$r2, summary)
```


### [ ] coxph masses at recovery (C, recovered bugs only)


## = 2025-08-11 pre mtg

### coxph trts (B+C)

```{r}
survObjs <- list_assign(survObjs, r1 = list())
fits <- list_assign(fits, r1 = list())
```

```{r}
# B + C
survObjs$r1 <- list_assign(survObjs$r1,
                            fromEnter = Surv(dfs$subsets$recs$exit.fromenter, dfs$subsets$recs$km.status),
                            fromRecovery = Surv(dfs$subsets$recs$exit.fromrecover, dfs$subsets$recs$km.status),
                            fromReturn = Surv(dfs$subsets$recs$exit.fromreturn, dfs$subsets$recs$km.status)
                            )

fits$r1 <- list_assign(fits$r1,
                    cph.enter = coxph(survObjs$r1$fromEnter ~ trt.recover * as.factor(trt.duration), 
                                      data = dfs$subsets$recs),
                    cph.recovery = coxph(survObjs$r1$fromRecovery ~ trt.recover * as.factor(trt.duration), 
                                         data = dfs$subsets$recs),
                    cph.return = coxph(survObjs$r1$fromReturn ~ trt.recover * as.factor(trt.duration),
                                       data = dfs$subsets$recs)
                    )

                                       
summary(fits$r1$cph.enter)
summary(fits$r1$cph.recovery)
summary(fits$r1$cph.return)
```

### coxph trts (C only)
```{r}
survObjs$r1 <- list_assign(survObjs$r1,
                            fromEnter = Surv(dfs$recC$exit.fromenter, dfs$recC$km.status),
                            fromRecovery = Surv(dfs$recC$exit.fromrecover, dfs$recC$km.status),
                            fromReturn = Surv(dfs$recC$exit.fromreturn, dfs$recC$km.status)
                            )

## duration as factor
fits$r1 <- list_assign(fits$r1,
                    cph.enter = coxph(survObjs$r1$fromEnter ~ trt.recover * as.factor(trt.duration), 
                                      data = dfs$recC),
                    cph.recovery = coxph(survObjs$r1$fromRecovery ~ trt.recover * as.factor(trt.duration), 
                                         data = dfs$recC),
                    cph.return = coxph(survObjs$r1$fromReturn ~ trt.recover * as.factor(trt.duration),
                                       data = dfs$recC)
                    )

                                       
summary(fits$r1$cph.enter)
summary(fits$r1$cph.recovery)
summary(fits$r1$cph.return)
```

```{r}
## fits as numeric
fits$r1 <- list_assign(fits$r1,
                    cph.enter = coxph(survObjs$r1$fromEnter ~ trt.recover * trt.duration, 
                                      data = dfs$recC),
                    cph.recovery = coxph(survObjs$r1$fromRecovery ~ trt.recover * trt.duration, 
                                         data = dfs$recC),
                    cph.return = coxph(survObjs$r1$fromReturn ~ trt.recover * trt.duration,
                                       data = dfs$recC)
                    )

                                       
summary(fits$r1$cph.enter)
summary(fits$r1$cph.recovery)
summary(fits$r1$cph.return)
```


======


# # R2 

## = 2025-07-16 cohort B fups
- going to rerun analyses from cohort A just focusing on cohort B data

TODO

- [ ] R2 surv props
- [ ] are there diffs in the kms?
- [ ] compare R2 vs R1
  - [ ] shelf effects?
- [ ] 33 const cohort B are all getting filtered out for some reason...

### df prep

```{r}
dfs <- bin_wrangled$km
survObjs <- list()
fits <- list()
```
  
```{r}
dfs <- list_assign(dfs, r1 = list())
survObjs <- list_assign(survObjs, r1 = list())
fits <- list_assign(fits, r1 = list())
```

```{r}
dfs$r1 <- list_assign(dfs$r1, 
                      allB = dfs$data %>% filter(cohort == "B"),
                      hotB = dfs$data %>% filter(is.40C == "Y"),
                      recB = dfs$data %>% filter(cohort == "B" & trt.duration > 0))
```

### coxph trt durations

```{r}
### use round 2 data
survObjs$r1 <- list_assign(survObjs$r1,
                            fromEnter = Surv(dfs$r1$recB$exit.fromenter, dfs$r1$recB$km.status),
                            fromRecovery = Surv(dfs$r1$recB$exit.fromrecover, dfs$r1$recB$km.status),
                            fromReturn = Surv(dfs$r1$recB$exit.fromreturn, dfs$r1$recB$km.status)
                            )

fits$r1 <- list_assign(fits$r1,
                    cph.enter = coxph(survObjs$r1$fromEnter ~ trt.recover * as.factor(trt.duration), 
                                      data = dfs$r1$recB),
                    cph.recovery = coxph(survObjs$r1$fromRecovery ~ trt.recover * as.factor(trt.duration), 
                                         data = dfs$r1$recB),
                    cph.return = coxph(survObjs$r1$fromReturn ~ trt.recover * as.factor(trt.duration),
                                       data = dfs$r1$recB)
                    )

                                       
summary(fits$r1$cph.enter)
summary(fits$r1$cph.recovery)
summary(fits$r1$cph.return)

# yay !!!
```

### coxph shelves

```{r}
fits$r1 <- list_assign(fits$r1, shelves = list())

fits$r1$shelves <- list_assign(fits$r1$shelves,
                               enterAdd = coxph(survObjs$r1$fromEnter ~ 
                                                  trt.recover * as.factor(trt.duration) + shelf,
                                                data = dfs$r1$recB),
                               enterInt = coxph(survObjs$r1$fromEnter ~ 
                                                  trt.recover * as.factor(trt.duration) * shelf,
                                                data = dfs$r1$recB),
                               recoverAdd = coxph(survObjs$r1$fromRecovery ~ 
                                                  trt.recover * as.factor(trt.duration) + shelf,
                                                data = dfs$r1$recB),
                               recoverInt = coxph(survObjs$r1$fromRecovery ~ 
                                                  trt.recover * as.factor(trt.duration) * shelf,
                                                data = dfs$r1$recB),
                               returnAdd = coxph(survObjs$r1$fromReturn ~ 
                                                  trt.recover * as.factor(trt.duration) + shelf,
                                                data = dfs$r1$recB),
                               returnInt = coxph(survObjs$r1$fromReturn ~ 
                                                  trt.recover * as.factor(trt.duration) * shelf,
                                                data = dfs$r1$recB))


summary(fits$r1$shelves$enterAdd)
summary(fits$r1$shelves$enterInt)
summary(fits$r1$shelves$recoverAdd)
summary(fits$r1$shelves$recoverInt)
summary(fits$r1$shelves$returnAdd)
summary(fits$r1$shelves$returnInt)

# ehh..
```


# # R1 archive

R1 is handled too differently so its df prep steps gets moved here

todos:

- 2025-07-16
  - [>] prob need to rerun A now that ive fixed the decimal hours calcs finally LO L L...
  - [ ] figure out when my bugs start to enter in the km viz lol 😵‍💫 
    - (idk if t0 = hatch or enter... also if the standardising is alrdy accounted for lol)
    
notes:

- cohort A df prep is separated (uses the `ogkm` df from import)

## - df prep

```{r}
# set up empty lists and dfs (different from B+ handling)
survObj <- list()
fits <- list()
testing <- list()

dfs <- bin_wrangled$kmA
```

## - 2025-06-27 redoing cox fits lmao (**)

cuz i love a fresh start!!!

- goals
  - [x] set up km
    - [/] which is the best time frame to use?
  - [x] try a few cphs
  
### create fits

```{r}
survObj <- list_assign(survObj,
                       fromEnter = Surv(dfs$expts$exit.fromenter, dfs$expts$km.status),
                       fromRecovery = Surv(dfs$expts$exit.fromrecover, dfs$expts$km.status),
                       fromReturn = Surv(dfs$expts$exit.fromreturn, dfs$expts$km.status)
                       )
```

### coxphing (trt + shelves)

```{r}
fits <- list_assign(fits, 
                    cph.enter = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration), data = dfs$expts),
                    cph.recovery = coxph(survObj$fromRecovery ~ trt.recover * as.factor(trt.duration), data = dfs$expts),
                    cph.return = coxph(survObj$fromReturn ~ trt.recover * as.factor(trt.duration), data = dfs$expts),
                    )

summary(fits$cph.enter)
summary(fits$cph.recovery)
summary(fits$cph.return)
```

tldr trt doesnt matter except except for longest time frame

```{r}
obj <- Surv(dfs$only40$exit.fromenter, dfs$only40$km.status)
obj2 <- Surv(dfs$only40$exit.fromreturn, dfs$only40$km.status)
fits <- list_assign(fits, 
                    cph.enter.addshelf = coxph(obj ~ trt.recover * as.factor(trt.duration) + shelf, data = dfs$only40),
                    cph.enter.intshelf = coxph(obj ~ trt.recover * as.factor(trt.duration) * shelf, data = dfs$only40),
                    cph.return.addshelf = coxph(obj2 ~ trt.recover * as.factor(trt.duration) + shelf, data = dfs$only40),
                    cph.return.intshelf = coxph(obj2 ~ trt.recover * as.factor(trt.duration) * shelf, data = dfs$only40),)

summary(fits$cph.enter.addshelf)
summary(fits$cph.enter.intshelf)
summary(fits$cph.return.addshelf)
summary(fits$cph.return.intshelf)
```
tldr shelf effect outweighs trt effect...



## = 2025-06-22 coxph attempts

- todo
  - [x] shelf effects: look at starting exit-fromenter...

- notes:
  - also worked on: 2025-06-23, 2025-06-24
  - 2025-06-27: 
    - lowkey the fits arent that correct bc 1) the exit dates are calced wrong and 2) the `dh.<stage>` stuff was sorta assigned wrong (i think)
    - also i calced recovery time wrong lmao (used return time...)
  


```{r}
survObj <- list_assign(survObj, 
                       #fromHatch = Surv(dfs$expts$exit.fromhatch, dfs$expts$km.status), 
                          # got yeeted from the wrangle script so would need to put back to make this work
                       fromEnter = Surv(dfs$expts$exit.fromenter, dfs$expts$km.status),
                       fromRecovery = Surv(dfs$expts$exit.fromrecover, dfs$expts$km.status),
                       fromReturn = Surv(dfs$expts$exit.fromreturn, dfs$expts$km.status)
                       )
```


### survdiffing (**)

- NOTE this is where `kmdf` is

```{r trying survdiff interactions, eval=FALSE}

## so google ai results summary says that survdiff is not intended to work w interaction terms LOL... sure!

##### trying to figure out how to get the interaction to show up but idky something about it breaks the formula
  # renaming to nicer vars doesnt work...

#survdiff(survObj$fromHatch ~ dfs$expts$trt.duration + dfs$expts$trt.recover)
# trying to pick out the columns names to get the interaction to work. see other attempts below
           # pluck(km$subsets, expts, trt.recover):pluck(km$subsets, expts, trt.duration))
           # 
           # hoist(as.data.frame(km$subsets), "trt.recover"):hoist(as.data.frame(km$subsets), "trt.duration"))
           # 
           # dfs$[expts][dfs$expts$,]:km$subsets[expts][dfs$expts$trt.recover,])


#### other attempts 

# kmdf <- dfs$expts %>%
#   rename(duration = trt.duration,
#          recover = trt.recover,
#          fromrec = exit.fromrecover,
#          stat = km.status)
# make a subset so naming is nicer, apparently u shouldnt use $ in formulas
# 
# survdiff(Surv(fromrec, stat) ~ 
#           duration + recover + interaction(duration, recover), data = kmdf) 
# this is bad (dont use interaction() cuz its ≠ to using :)
# 
# survdiff(Surv(fromrec, stat) ~ 
#           duration + recover + duration:recover, data = kmdf) # doesnt work
# 
#survdiff(formula = Surv(kmdf$fromrec, kmdf$stat) ~ duration + recover + duration:recover, data = kmdf) # doesnt work


### 3

# survdiff(survObj$fromHatch ~ poly(c(dfs$expts$trt.duration, dfs$expts$trt.recover), n = 2)) 
  # im totally using poly() wrong here but idr how to use it. but also def not what im looking for

### 4 am i going nuts

# anova(lm(fromrec ~ duration*recover, data = kmdf)) # i am not...

```


```{r eval=FALSE}
# survdiff starting from hatch time
## recovery temp + duration
# survdiff(survObj$fromHatch ~ dfs$expts$trt.duration + dfs$expts$trt.recover 
#          #+ `dfs$expts$trt.recover`:`dfs$expts$trt.duration`
#          )

#######

# survdiff starting from enter time
survdiff(survObj$fromEnter ~ dfs$expts$trt.duration + dfs$expts$trt.recover) ## recovery temp + duration


#######

# looking for diffs starting at recovery
survdiff(survObj$fromRecovery ~ dfs$expts$trt.recover) ## based on recovery temp


survdiff(survObj$fromRecovery ~ dfs$expts$trt.duration) ## based on recovery period


survdiff(survObj$fromRecovery ~ dfs$expts$trt.duration + dfs$expts$trt.recover) ## temp + duration (cant do the interaction)


########

# survdiff starting from returning time
survdiff(survObj$fromReturn ~ dfs$expts$trt.duration + dfs$expts$trt.recover) ## recovery temp + duration

```
### cphing

#### fitting cph

##### effects of recovery temp and duration
```{r}
fits <- list_assign(fits, cph.enter = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration), 
                                               data = dfs$expts))
summary(fits$cph.enter)


fits <- list_assign(fits, cph.recovery = coxph(survObj$fromRecovery ~ trt.recover * as.factor(trt.duration), 
                                               data = dfs$expts))
summary(fits$cph.recovery)


fits <- list_assign(fits, cph.return = coxph(survObj$fromReturn ~ trt.recover * as.factor(trt.duration), 
                                             data = dfs$expts))
summary(fits$cph.return)


fits <- list_assign(fits, cph.return2 = coxph(survObj$fromReturn ~ trt.recover * trt.duration, 
                                             data = dfs$expts))
summary(fits$cph.return2)
# same results as above

fits <- list_assign(fits, cph.return3 = coxph(survObj$fromReturn ~ as.factor(trt.recover) * trt.duration, 
                                             data = dfs$expts))
summary(fits$cph.return3)

```

so the cph fits only really matter starting from the enter time?

##### shelf effects

```{r}
# looking at shelf effects

fits <- list_assign(fits, cph.shelf = coxph(survObj$fromEnter ~ trt.recover * as.factor(trt.duration) * shelf, 
                                               data = dfs$expts))
summary(fits$cph.shelf)


fits <- list_assign(fits, cph.shelf2 = coxph(survObj$fromEnter ~ shelf + trt.recover * as.factor(trt.duration), 
                                               data = dfs$expts))
summary(fits$cph.shelf2)

fits <- list_assign(fits, cph.shelf3 = coxph(survObj$fromEnter ~ shelf + trt.recover * trt.duration, 
                                               data = dfs$expts))
summary(fits$cph.shelf3)
```

tldr in all cases, `trt.recover` becomes signif once u throw in the shelf LOL.. (generally p < 0.5)

#### plotting cph

```{r eval=FALSE}
# plot cphs...
# plot(x = fits$cph.recovery$linear.predictors,
#      y = fits$cph.recovery$residuals,
#      data = dfs$expts, 
#      #facet.by = trt.recovery
#      )
  # this is not how to use survival::plot.coxph lolol (cph doesnt give u a survival curve per se. its a regression)

autoplot(survfit(coxph(survObj$fromReturn ~ trt.recover * trt.duration, 
                                             data = dfs$expts)),
         facets = TRUE,
         )
```

```{r eval=FALSE}
# per google:
# check if prop hazards assumption is met (is HR betwn groups constant over time?)
  # https://stats.stackexchange.com/questions/560975/how-to-interpret-schoenfield-residual-plot
ggcoxzph(cox.zph(fits$cph.enter), ggtheme = theme_minimal())
ggcoxzph(cox.zph(fits$cph.recovery), ggtheme = theme_minimal())
ggcoxzph(cox.zph(fits$cph.return), ggtheme = theme_minimal())

# TODO idk what im doing here so... lets revisit later
```

```{r}
fit <- survfit(Surv(kmdf$exit.fromenter, kmdf$stat) ~ recover + duration, data = kmdf) 
# ok survfit super doesnt like "subsets" being in the df name lol

# kmdf_summ <- kmdf %>%
#   group_by(trt)
#   summarise()

ggsurvplot(fit, 
           data = kmdf,
           conf.int = TRUE,
           #xlim = c(45823, 45841),
           #facet.by = "trt.duration",
           facet.by = "duration",
           #linetype = "strata",
           palette = "strata") +
  
  # i think the stuff for vline needs to come from a separate df LOL...
  # also, lowkey not sure how this works considering the time depends on the hatch day... (avg them??)
  
  geom_vline(xintercept = kmdf$dh.recover - kmdf$dh.exit, #tried w dh.enter, kinda wonky
             #color = "red"
             #aes(xintercept = kmdf$dh.recover - kmdf$dh.enter, color = kmdf$recover)
             ) +
  geom_vline(xintercept = kmdf$return - kmdf$df.exit,
             color = "blue")

# TODO idk gotta think about wtf im plotting for this one a bit... draw it out
  
```

